---
title: "02_28_cosmx_TNBC_prepro_metric_report"
author: "Benedetta Banzi"
date: "2024-03-04"
output: html_document
---

<STYLE TYPE="text/css">
<!--
  td{
    font-family: Arial; 
    font-size: 10 pt;
    padding:0px;
    cellpadding="0";
    cellspacing="0"
  }
  th {
    font-family: Arial; 
    font-size: 10 pt;
    height: 20px;
    font-weight: bold;
    text-align: "center";
    background-color: #ccccff;
    padding-left: 15px;
    padding-right: 15px;
  }
  table { 
    border-spacing: 4px;
    border-collapse: collapse;
  }
--->
</STYLE>


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Library import

```{r}
library(SpatialExperiment)
library(data.table)
library(scater) # it imports scuttle
library(cowplot)
library(ggplot2)
theme_set(theme_bw())
library(matrixStats)
library(dplyr)
library(tidyr)
library(tibble)
library(arrow)
library(scales)
library(sf)
library(tmap)
library(knitr)
library(grid)
library(ggpointdensity)
library(ggpubr)
library(spdep)
library(ggExtra)
```

Read-in function

```{r}
readCosmxSPE <- function(dirname = dirname, 
                         countmatfpattern = "exprMat_file.csv", 
                         metadatafpattern = "metadata_file.csv", 
                         coord_names = c("CenterX_global_px",
                                         "CenterY_global_px")){
  
  countmat_file <- file.path(dirname, list.files(dirname, countmatfpattern))
  metadata_file <- file.path(dirname, list.files(dirname, metadatafpattern))
  
  # Read in 
  countmat <- data.table::fread(countmat_file)
  metadata <- data.table::fread(metadata_file)
  
  # Count matrix   
  counts <- merge(countmat, metadata[, c("fov", "cell_ID")])
  counts <- subset(counts, select = -c(fov, cell_ID))
  cell_codes <- rownames(counts)
  features <- colnames(counts)
  counts <- t(counts)
  
  rownames(counts) <- features
  colnames(counts) <- cell_codes
  
  # rowData (does not exist)
  
  # colData
  colData <- merge(metadata, countmat[, c("fov", "cell_ID")])
  
  spe <- SpatialExperiment::SpatialExperiment(
    assays = list(counts = counts),
    # rowData = rowData,
    colData = colData,
    spatialCoordsNames = coord_names
  )
  
  return(spe)
}
```

Setting directories

```{r}
dirname <- "/data01/Shared/Benedetta/spatial_transcriptomics/CosMx/Datasets/TNBC/UAP7_02527"
count_pattern <- "UAP7_02527_exprMat_file.csv"
meta_pattern <- "UAP7_02527_metadata_file.csv"
sample_name <- "CosMx_human_TNBC"
```

# CosMx dataset exploratory analysis for preprocessing

SpatialExperiment object creation and exploration

```{r}
spe <- readCosmxSPE(dirname = dirname, countmatfpattern = count_pattern, metadatafpattern = meta_pattern)
spe <- scuttle::addPerCellQC(spe, subsets=list(NegProb = grep("^Negative", rownames(spe)), SysCtrl = grep("^SystemControl", rownames(spe))))
names(colData(spe)@listData)
spe$control_counts <- rowSums(data.frame(colData(spe))[,grep("^subsets_.*_sum$", names(colData(spe)))])
spe$control_detected <- rowSums(data.frame(colData(spe))[,grep("^subsets_.*_detected$", names(colData(spe)))])
```

```{r}
dim(spe)
```

## Large-scale to small-scale

Approaching dataset preprocessing starting from global-scale, then fov-focused exploratory analysis

# Global sample map

To view the spatial organization of both the sample as it is and FOVs

```{r}
# Importing global coordinates from spatialCoords slot since they are not available in colData
colData(spe)$CenterX_global_px <- spatialCoords(spe)[,1]
colData(spe)$CenterY_global_px <- spatialCoords(spe)[,2]
```

```{r, out.width="600%", out.height="150%"}
fov_pattern <- "UAP7_02527_fov_positions_file.csv"
fovpos_file <- file.path(dirname, list.files(dirname, fov_pattern))

metadata <- data.table::fread(file.path(dirname, list.files(dirname, meta_pattern)))
fov_positions <- data.table::fread(fovpos_file, header = T)
fov_positions <- fov_positions[fov_positions$FOV%in%unique(metadata$fov),]

# image theme
global_map_theme <- function(back.color="black",back.border=NA,title.col="white") {
  theme(aspect.ratio = 1,
        panel.border = element_blank(),
        legend.key = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks = element_blank(),
        axis.text.y = element_blank(),
        axis.text.x = element_blank(),
        panel.grid = element_blank(),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        panel.background = element_rect(fill = "transparent",colour = NA),
        plot.title = element_text(color = title.col,hjust = 0.5,face = "bold"),
        plot.background = element_rect(fill = "transparent",colour = back.border))
}

fov_size <- 4256

ggplot() + 
  geom_point(data = metadata, mapping = aes(x = CenterX_global_px, y = CenterY_global_px), colour = "darkmagenta", fill="darkmagenta", size = 0.05, alpha = 0.2) +
  annotate("rect", xmin = fov_positions$X_px - 1/32*fov_size, 
         xmax = fov_positions$X_px + fov_size - 1/32*fov_size, 
         ymin = fov_positions$Y_px - fov_size - 1/16*fov_size, 
         ymax = fov_positions$Y_px - 1/16*fov_size,
         alpha = .2, color = "black",linewidth = 0.2) +
  geom_text(aes(x = fov_positions$X_px+11/10*fov_size/2,
                y = fov_positions$Y_px-11/10*fov_size/2,
                label = fov_positions$FOV), color="black", size = 3) +
  ggtitle(sample_name) +
  global_map_theme(back.color = "white", back.border = "white", title.col = "black")
```

# Global sample map with cell types

```{r}
anno_file <- fread(file.path(dirname, "UAP7_02527_clusterAnnot@1_simple.txt"), sep = "\t", header = T)
colnames(anno_file) <- c("id_fov", "cell_type", "type_order", "type_color")

anno_df <- subset(anno_file, select = c("id_fov", "cell_type"))

metadata$id_fov <- paste(metadata$cell_ID, metadata$fov, sep = "_")
spe$id_fov <- metadata$id_fov
metadata <- left_join(metadata, anno_df, by = join_by(id_fov))
table(table(metadata$fov) == table(colData(spe)$fov))
```

```{r}
metadata <- metadata[metadata$id_fov %in% anno_df$id_fov,]
spe <- spe[, spe$id_fov %in% metadata$id_fov]
table(table(metadata$fov) == table(colData(spe)$fov))
spe$cell_type <- metadata$cell_type
```

Some cells are missing in the annotation file. They were preprocessed because of their low-signal. Proceeding without them.

```{r, out.width="150%", out.height="300%"}
celltype_palette <- c(
  "discard"	= "gray11", 
"RBCs" = "gray47",
"Plasma_cells" =	"blue",
"Plasmablasts" =	"royalblue1",
"Mix_CTLs_Plasma_cells" =	"dodgerblue4",
"CTLs" =	"deepskyblue",
"Mix_CXCL13+Tcells_Plasma_cells" =	"cyan",
"CXCL13+Tcells" =	"lightblue1",
"Tcells" =	"darkseagreen1",
"Tcell_Monocyte_Pairs" = "darkolivegreen1",
"Monocytes" =	"limegreen",
"LipoTAMs" =	"green3",
"Mixed_LipoTAMs" =	"darkgreen",
"Mix_LipoTAMs_Plasma_cells" =	"yellow4",
"TAMs" =	"yellow",
"iTAMs" = "greenyellow",
"OPN+TAMs" =	"olivedrab1",
"DCs" =	"seagreen2",
"Mast_cells" =	"lightyellow",
"Mix_MastCells_Plasma_cells" =	"goldenrod4",
"Mix_CAFs_Plasma_cells" =	"goldenrod1",
"Mixed_CAFs" =	"lightgoldenrod1",
"CAFs" =	"lightgoldenrod4",
"Stromal_cells" =	"darkorange1",
"Pericytes" =	"plum1",
"Lymph_ECs" =	"orchid4",
"Blood_ECs" =	"orchid1",
"BC_cells" =	"lightpink",
"LC3+BC_cells" =	"lightcoral",
"MGP+BC_cells" =	"orangered1",
"Cycling_BC_cells" =	"deeppink1",
"Hypoxic_BC_cells" = "darkmagenta",
"Mixed_BC_cells" =	"red4"
)

celltype_color <- c()
for(i in 1:dim(metadata)[1]){
  celltype <- anno_df$cell_type[i]
  n <- match(celltype, names(celltype_palette))
  celltype_color[i] <- celltype_palette[n]
}

metadata$celltype_color <- as.factor(celltype_color)
spe$celltype_color <- as.factor(celltype_color)


ggplot(data = metadata, mapping = aes(x = CenterX_global_px, 
  y = CenterY_global_px, color = cell_type)) + 
  geom_point(size = 0.05) +
  scale_color_manual(values = celltype_palette) +
  ggtitle(sample_name) + 
  global_map_theme(back.color = "white", back.border = "white", 
  title.col = "black") +
  guides(colour = guide_legend(override.aes = list(size=5)))

```


# Added variables

```{r}
fov_size <- 4256

# Importing global coordinates from spatialCoords slot since they are not available in colData
colData(spe)$CenterX_global_px <- spatialCoords(spe)[,1]
colData(spe)$CenterY_global_px <- spatialCoords(spe)[,2]

# Converting global coordinates to um 
spe$CenterX_global_um <- spe$CenterX_global_px*0.18
spe$CenterY_global_um <- spe$CenterY_global_px*0.18 

# Logged total counts
spe$logtot <- log10(spe$total)

# Converting area in um
spe$um_area <- spe$Area*0.0324

# Logged width/height ratio so there are no holes in the distribution in the middle and also negative values can be appreciated
spe$log2_AspectRatio <- log2(spe$AspectRatio)

# Keeping only target counts by subtracting counts referred to control probes
spe$target_counts <- spe$total - spe$control_counts

# Keeping only detected features by subtracting negative control probes
spe$target_detected <- spe$detected - spe$control_detected

# Proportion between counts referred only to target probes and detected features
spe$tot_det_ratio <- spe$target_counts/spe$target_detected

spe$ctrl_total_ratio <- spe$control_counts/spe$total
```


# Global feature maps

To spot areas with anomalous feature values and spatial patterns on a global scale

```{r, out.width="150%", out.height="150%"}
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "total", point_size = 0.1) + ggtitle("Total counts per cell") + theme(aspect.ratio = 1)
```

```{r, out.width="150%", out.height="150%"}
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "logtot", point_size = 0.1) + ggtitle("Log-total counts per cell") + theme(aspect.ratio = 1)
```

```{r, out.width="150%", out.height="150%"}
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "detected", point_size = 0.1) + ggtitle("Detected features per cell") + theme(aspect.ratio = 1)
```

```{r, out.width="150%", out.height="150%"}
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "um_area", point_size = 0.1) + ggtitle("Area in μm per cell") + theme(aspect.ratio = 1)
```

```{r, out.width="150%", out.height="150%"}
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "log2_AspectRatio", point_size = 0.1) + ggtitle("Logged width/Height ratio per cell") + theme(aspect.ratio = 1)
```

```{r, out.width="150%", out.height="150%"}
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "tot_det_ratio", point_size = 0.1) + ggtitle("Target counts/detected features ratio per cell") + theme(aspect.ratio = 1)
```

```{r, out.width="150%", out.height="150%"}
plotColData(spe, "CenterY_global_um", "CenterX_global_um", order_by = "control_counts", colour_by = "control_counts", point_size = 0.1) +
  scale_color_gradient(low = "white", high = "red", name = "control_counts") + ggtitle("Negative control probe counts per cell") +
  theme(aspect.ratio = 1,
        panel.background = element_rect(fill = "black",color = NA),
        plot.background = element_rect(fill = "black",color = NA),
        axis.title.x = element_text(color = "white"),
        axis.title.y = element_text(color = "white"),
        axis.ticks = element_line(color = "white"),
        axis.text.y = element_text(color = "white"),
        axis.text.x = element_text(color = "white"),
        axis.line = element_line(color = "white"),
        legend.title = element_text(color = "white"),
        legend.text = element_text(color = "white"),
        plot.title = element_text(color = "white"))
```

```{r}
threshold <- 0.20
spe$flagged_tot_20 <- as.factor(ifelse(colData(spe)$total < quantile(colData(spe)$total, probs = threshold), TRUE, FALSE))

threshold <- 0.10
spe$flagged_tot_10 <- as.factor(ifelse(colData(spe)$total < quantile(colData(spe)$total, probs = threshold), TRUE, FALSE))

threshold <- 0.05
spe$flagged_tot_5 <- as.factor(ifelse(colData(spe)$total < quantile(colData(spe)$total, probs = threshold), TRUE, FALSE))

threshold <- 0.10
spe$flagged_det_10 <- as.factor(ifelse(colData(spe)$detected < quantile(colData(spe)$detected, probs = threshold), TRUE, FALSE))

threshold <- 0.05
spe$flagged_det_5 <- as.factor(ifelse(colData(spe)$detected < quantile(colData(spe)$detected, probs = threshold), TRUE, FALSE))

threshold <- 0.10
spe$flagged_umarea_10 <- as.factor(ifelse(colData(spe)$um_area < quantile(colData(spe)$um_area, probs = threshold), TRUE, FALSE))

threshold <- 0.05
spe$flagged_umarea_5 <- as.factor(ifelse(colData(spe)$um_area < quantile(colData(spe)$um_area, probs = threshold), TRUE, FALSE))

threshold <- 0.90
spe$flagged_umarea_90 <- as.factor(ifelse(colData(spe)$um_area > quantile(colData(spe)$um_area, probs = threshold), TRUE, FALSE))

threshold <- 0.95
spe$flagged_umarea_95 <- as.factor(ifelse(colData(spe)$um_area > quantile(colData(spe)$um_area, probs = threshold), TRUE, FALSE))

threshold <- 0.10
spe$flagged_aratio_10 <- as.factor(ifelse(colData(spe)$log2_AspectRatio < quantile(colData(spe)$log2_AspectRatio, probs = threshold), TRUE, FALSE))

threshold <- 0.05
spe$flagged_aratio_5 <- as.factor(ifelse(colData(spe)$log2_AspectRatio < quantile(colData(spe)$log2_AspectRatio, probs = threshold), TRUE, FALSE))

threshold <- 0.90
spe$flagged_aratio_90 <- as.factor(ifelse(colData(spe)$log2_AspectRatio > quantile(colData(spe)$log2_AspectRatio, probs = threshold), TRUE, FALSE))

threshold <- 0.95
spe$flagged_aratio_95 <- as.factor(ifelse(colData(spe)$log2_AspectRatio > quantile(colData(spe)$log2_AspectRatio, probs = threshold), TRUE, FALSE))

threshold <- 0.10
spe$flagged_tdratio_10 <- as.factor(ifelse(colData(spe)$tot_det_ratio < quantile(colData(spe)$tot_det_ratio, probs = threshold, na.rm = T), TRUE, FALSE))

threshold <- 0.05
spe$flagged_tdratio_5 <- as.factor(ifelse(colData(spe)$tot_det_ratio < quantile(colData(spe)$tot_det_ratio, probs = threshold, na.rm = T), TRUE, FALSE))

threshold <- 1
spe$flagged_tdratio_atomx <- as.factor(ifelse(colData(spe)$tot_det_ratio <= threshold, TRUE, FALSE))

threshold <- 0.50
spe$flagged_negprob_50 <- as.factor(ifelse(colData(spe)$control_counts > quantile(colData(spe)$control_counts, probs = threshold, na.rm = T), TRUE, FALSE))

threshold <- 0.75
spe$flagged_negprob_75 <- as.factor(ifelse(colData(spe)$control_counts > quantile(colData(spe)$control_counts, probs = threshold, na.rm = T), TRUE, FALSE))

spe$ctrl_total_ratio <- spe$control_counts/spe$total
threshold <- 0.10
spe$flagged_negprob_10 <- as.factor(ifelse(colData(spe)$ctrl_total_ratio > quantile(colData(spe)$ctrl_total_ratio, probs = threshold, na.rm = T), TRUE, FALSE))
```


# Feature histograms

To check feature distribution before highlighting where extreme values are located in the global sample map

```{r}
variables <- c("total", "detected", "logtot", "um_area", "log2_AspectRatio", "tot_det_ratio", "control_counts")
for (var in 1:length(variables)){
  n <- grep(paste0("^", variables[var], "$"), colnames(spe@colData))
  histo_var <- names(colData(spe))[n]
  p <- ggplot(data = data.frame(colData(spe))) +
    geom_histogram(aes(x = .data[[histo_var]]), fill = "#69b3a2") +
    ggtitle(paste0(histo_var))
  print(p)
}  
```

# Global threshold map

Cells below the selected feature threshold are highlighted in black, all the others are grey

```{r, out.width="150%", out.height="150%"}
threshold <- 0.20
spe$flagged_tot_20 <- as.factor(ifelse(colData(spe)$total < quantile(colData(spe)$total, probs = threshold), TRUE, FALSE))
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "flagged_tot_20", 
            point_size = 0.1) + scale_color_manual(values = c("TRUE" = "black", "FALSE" = "lightgrey")) + ggtitle("Counts below 20% threshold (default by AtoMx guidelines)") + theme(legend.position="none") + labs(subtitle = paste0("Flagged cells = ", table(colData(spe)$flagged_tot_20)["TRUE"], " out of ", dim(spe)[2], ", ", round(table(colData(spe)$flagged_tot_20)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"))
```

```{r, out.width="150%", out.height="150%"}
threshold <- 0.10
spe$flagged_tot_10 <- as.factor(ifelse(colData(spe)$total < quantile(colData(spe)$total, probs = threshold), TRUE, FALSE))
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "flagged_tot_10", 
            point_size = 0.1) + scale_color_manual(values = c("TRUE" = "black","FALSE" = "lightgrey")) + ggtitle("Counts below 10% threshold")  + theme(legend.position="none") + labs(subtitle = paste0("Flagged cells = ", table(colData(spe)$flagged_tot_10)["TRUE"], " out of ", dim(spe)[2], ", ",  round(table(colData(spe)$flagged_tot_10)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"))
```

```{r, out.width="150%", out.height="150%"}
threshold <- 0.05
spe$flagged_tot_5 <- as.factor(ifelse(colData(spe)$total < quantile(colData(spe)$total, probs = threshold), TRUE, FALSE))
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "flagged_tot_5", 
            point_size = 0.1) + scale_color_manual(values = c("TRUE" = "black","FALSE" = "lightgrey")) + ggtitle("Counts below 5% threshold") + theme(legend.position="none") + labs(subtitle = paste0("Flagged cells = ", table(colData(spe)$flagged_tot_5)["TRUE"], " out of ", dim(spe)[2], ", ",  round(table(colData(spe)$flagged_tot_5)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells")) 
```



```{r, out.width="300%", out.height="200%"}
threshold <- 0.10
spe$flagged_det_10 <- as.factor(ifelse(colData(spe)$detected < quantile(colData(spe)$detected, probs = threshold), TRUE, FALSE))
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "flagged_det_10", 
            point_size = 0.1) + scale_color_manual(values = c("TRUE" = "black","FALSE" = "lightgrey")) + ggtitle("Detected features below 10% threshold") + theme(legend.position="none") + labs(subtitle = paste0("Flagged cells = ", table(colData(spe)$flagged_det_10)["TRUE"], " out of ", dim(spe)[2], ", ",  round(table(colData(spe)$flagged_det_10)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"))
```

```{r, out.width="150%", out.height="150%"}
threshold <- 0.05
spe$flagged_det_5 <- as.factor(ifelse(colData(spe)$detected < quantile(colData(spe)$detected, probs = threshold), TRUE, FALSE))
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "flagged_det_5", 
            point_size = 0.1) + scale_color_manual(values = c("TRUE" = "black","FALSE" = "lightgrey")) + ggtitle("Detected features below 5% threshold")  + theme(legend.position="none") + labs(subtitle = paste0("Flagged cells = ", table(colData(spe)$flagged_det_5)["TRUE"], " out of ", dim(spe)[2], ", ",  round(table(colData(spe)$flagged_det_5)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"))
```



```{r, out.width="150%", out.height="150%"}
threshold <- 0.10
spe$flagged_umarea_10 <- as.factor(ifelse(colData(spe)$um_area < quantile(colData(spe)$um_area, probs = threshold), TRUE, FALSE))
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "flagged_umarea_10", 
            point_size = 0.1) + scale_color_manual(values = c("TRUE" = "black","FALSE" = "lightgrey")) + ggtitle("Area in μm below 10%") + theme(legend.position="none") + labs(subtitle = paste0("Flagged cells = ", table(colData(spe)$flagged_umarea_10)["TRUE"], " out of ", dim(spe)[2], ", ",  round(table(colData(spe)$flagged_umarea_10)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"))
```

```{r, out.width="150%", out.height="150%"}
threshold <- 0.05
spe$flagged_umarea_5 <- as.factor(ifelse(colData(spe)$um_area < quantile(colData(spe)$um_area, probs = threshold), TRUE, FALSE))
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "flagged_umarea_5", 
            point_size = 0.1) + scale_color_manual(values = c("TRUE" = "black","FALSE" = "lightgrey")) + ggtitle("Area in μm below 5%")  + theme(legend.position="none") + labs(subtitle = paste0("Flagged cells = ", table(colData(spe)$flagged_umarea_5)["TRUE"], " out of ", dim(spe)[2], ", ",  round(table(colData(spe)$flagged_umarea_5)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"))
```


Cells above the selected threshold are highlighted in red, all the others are grey

```{r, out.width="150%", out.height="150%"}
threshold <- 0.90
spe$flagged_umarea_90 <- as.factor(ifelse(colData(spe)$um_area > quantile(colData(spe)$um_area, probs = threshold), TRUE, FALSE))
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "flagged_umarea_90", 
            point_size = 0.1) + scale_color_manual(values = c("TRUE" = "red","FALSE" = "lightgrey")) + ggtitle("Area in μm above 90% threshold") + theme(legend.position="none") + labs(subtitle = paste0("Flagged cells = ", table(colData(spe)$flagged_umarea_90)["TRUE"], " out of ", dim(spe)[2], ", ",  round(table(colData(spe)$flagged_umarea_90)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"))
```

```{r, out.width="150%", out.height="150%"}
threshold <- 0.95
spe$flagged_umarea_95 <- as.factor(ifelse(colData(spe)$um_area > quantile(colData(spe)$um_area, probs = threshold), TRUE, FALSE))
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "flagged_umarea_95", 
            point_size = 0.1) + scale_color_manual(values = c("TRUE" = "red","FALSE" = "lightgrey")) + ggtitle("Area in μm above 95% threshold")  + theme(legend.position="none") + labs(subtitle = paste0("Flagged cells = ", table(colData(spe)$flagged_umarea_95)["TRUE"], " out of ", dim(spe)[2], ", ",  round(table(colData(spe)$flagged_umarea_95)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"))
```



```{r, out.width="150%", out.height="150%"}
threshold <- 0.10
spe$flagged_aratio_10 <- as.factor(ifelse(colData(spe)$log2_AspectRatio < quantile(colData(spe)$log2_AspectRatio, probs = threshold), TRUE, FALSE))
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "flagged_aratio_10", 
            point_size = 0.1) + scale_color_manual(values = c("TRUE" = "black","FALSE" = "lightgrey")) + ggtitle("Width/height below 10% threshold") + theme(legend.position="none") + labs(subtitle = paste0("Flagged cells = ", table(colData(spe)$flagged_aratio_10)["TRUE"], " out of ", dim(spe)[2], ", ",  round(table(colData(spe)$flagged_aratio_10)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"))
```

```{r, out.width="150%", out.height="150%"}
threshold <- 0.05
spe$flagged_aratio_5 <- as.factor(ifelse(colData(spe)$log2_AspectRatio < quantile(colData(spe)$log2_AspectRatio, probs = threshold), TRUE, FALSE))
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "flagged_aratio_5", 
            point_size = 0.1) + scale_color_manual(values = c("TRUE" = "black","FALSE" = "lightgrey")) + ggtitle("Width/height below 5% threshold")  + theme(legend.position="none") + labs(subtitle = paste0("Flagged cells = ", table(colData(spe)$flagged_aratio_5)["TRUE"], " out of ", dim(spe)[2], ", ",  round(table(colData(spe)$flagged_aratio_5)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"))
```



```{r, out.width="150%", out.height="150%"}
threshold <- 0.90
spe$flagged_aratio_90 <- as.factor(ifelse(colData(spe)$log2_AspectRatio > quantile(colData(spe)$log2_AspectRatio, probs = threshold), TRUE, FALSE))
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "flagged_aratio_90", 
            point_size = 0.1) + scale_color_manual(values = c("TRUE" = "red","FALSE" = "lightgrey")) + ggtitle("Width/height above 90% threshold") + theme(legend.position="none") + labs(subtitle = paste0("Flagged cells = ", table(colData(spe)$flagged_aratio_90)["TRUE"], " out of ", dim(spe)[2], ", ",  round(table(colData(spe)$flagged_aratio_90)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"))
```

```{r, out.width="150%", out.height="150%"}
threshold <- 0.95
spe$flagged_aratio_95 <- as.factor(ifelse(colData(spe)$log2_AspectRatio > quantile(colData(spe)$log2_AspectRatio, probs = threshold), TRUE, FALSE))
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "flagged_aratio_95", 
            point_size = 0.1) + scale_color_manual(values = c("TRUE" = "red","FALSE" = "lightgrey")) + ggtitle("Width/height above 95% threshold")  + theme(legend.position="none") + labs(subtitle = paste0("Flagged cells = ", table(colData(spe)$flagged_aratio_95)["TRUE"], " out of ", dim(spe)[2], ", ",  round(table(colData(spe)$flagged_aratio_95)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"))
```



```{r, out.width="150%", out.height="150%"}
threshold <- 0.10
spe$flagged_tdratio_10 <- as.factor(ifelse(colData(spe)$tot_det_ratio < quantile(colData(spe)$tot_det_ratio, probs = threshold, na.rm = T), TRUE, FALSE))
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "flagged_tdratio_10", 
            point_size = 0.1) + scale_color_manual(values = c("TRUE" = "black","FALSE" = "lightgrey")) + ggtitle(paste0("Total counts/detected features below 10% threshold, ratio = ", round(quantile(colData(spe)$tot_det_ratio, probs = threshold, na.rm = T), 2))) + theme(legend.position="none") + labs(subtitle = paste0("Flagged cells = ", table(colData(spe)$flagged_tdratio_10)["TRUE"], " out of ", dim(spe)[2], ", ",  round(table(colData(spe)$flagged_tdratio_10)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"))
```

```{r, out.width="150%", out.height="150%"}
threshold <- 0.05
spe$flagged_tdratio_5 <- as.factor(ifelse(colData(spe)$tot_det_ratio < quantile(colData(spe)$tot_det_ratio, probs = threshold, na.rm = T), TRUE, FALSE))
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "flagged_tdratio_5", 
            point_size = 0.1) + scale_color_manual(values = c("TRUE" = "black","FALSE" = "lightgrey")) + ggtitle(paste0("Total counts/detected features below 5% threshold, ratio = ", round(quantile(colData(spe)$tot_det_ratio, probs = threshold, na.rm = T), 2)))  + theme(legend.position="none") + labs(subtitle = paste0("Flagged cells = ", table(colData(spe)$flagged_tdratio_5)["TRUE"], " out of ", dim(spe)[2], ", ",  round(table(colData(spe)$flagged_tdratio_5)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"))
```

ATOMX GUIDELINES FOR CELL QC

Count distribution 
flag cells where (total counts) / (number of detected genes) ≤1

```{r, out.width="150%", out.height="150%"}
threshold <- 1
spe$flagged_tdratio_atomx <- as.factor(ifelse(colData(spe)$tot_det_ratio <= threshold, TRUE, FALSE))
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "flagged_tdratio_atomx", 
            point_size = 0.1) + scale_color_manual(values = c("TRUE" = "black","FALSE" = "lightgrey")) + ggtitle("Total counts/detected feature ratio below 1 (default by AtoMx guidelines), ratio = 1") + theme(legend.position="none") + labs(subtitle = paste0("Flagged cells = ", table(colData(spe)$flagged_tdratio_atomx)["TRUE"], " out of ", dim(spe)[2], ", ",  round(table(colData(spe)$flagged_tdratio_atomx)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells")) 
```


```{r, out.width="150%", out.height="150%"}
threshold <- 0.50
spe$flagged_negprob_50 <- as.factor(ifelse(colData(spe)$control_counts > quantile(colData(spe)$control_counts, probs = threshold, na.rm = T), TRUE, FALSE))
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "flagged_negprob_50", 
            point_size = 0.1) + scale_color_manual(values = c("TRUE" = "red","FALSE" = "lightgrey")) + ggtitle("Negative probe counts above 50% threshold") + theme(legend.position="none") + labs(subtitle = paste0("Flagged cells = ", table(colData(spe)$flagged_negprob_50)["TRUE"], " out of ", dim(spe)[2], ", ",  round(table(colData(spe)$flagged_negprob_50)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"))
```

```{r, out.width="150%", out.height="150%"}
threshold <- 0.75
spe$flagged_negprob_75 <- as.factor(ifelse(colData(spe)$control_counts > quantile(colData(spe)$control_counts, probs = threshold, na.rm = T), TRUE, FALSE))
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "flagged_negprob_75", 
            point_size = 0.1) + scale_color_manual(values = c("TRUE" = "red","FALSE" = "lightgrey")) + ggtitle("Negative probe counts above 75% threshold")  + theme(legend.position="none") + labs(subtitle = paste0("Flagged cells = ", table(colData(spe)$flagged_negprob_75)["TRUE"], " out of ", dim(spe)[2], ", ",  round(table(colData(spe)$flagged_negprob_75)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"))
```

ATOMX GUIDELINES FOR CELL QC

Proportion of negative counts
flag cells where > 10% (0.1; default) of the counts are negative probes

```{r, out.width="150%", out.height="150%"}
spe$ctrl_total_ratio <- spe$control_counts/spe$total
threshold <- 0.10
spe$flagged_negprob_10 <- as.factor(ifelse(colData(spe)$ctrl_total_ratio > quantile(colData(spe)$ctrl_total_ratio, probs = threshold, na.rm = T), TRUE, FALSE))
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "flagged_negprob_10", 
            point_size = 0.1) + scale_color_manual(values = c("TRUE" = "red","FALSE" = "lightgrey")) + ggtitle("Negative control probe/total counts ratio above 10% threshold (default by AtoMx guidelines)") + theme(legend.position="none") + labs(subtitle = paste0("Flagged cells = ", table(colData(spe)$flagged_negprob_10)["TRUE"], " out of ", dim(spe)[2], ", ", round(table(colData(spe)$flagged_negprob_10)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells")) 
```


# Feature heatmaps by FOV 

To compare at a glance FOV mean/median feature values and spot anomalous FOVs

```{r}
meta_df <- data.frame(colData(spe))
meta_df <- group_by(meta_df, fov) |> summarize(n_cells = n(), mean_total = mean(total), mean_detected = mean(detected), mean_umarea = mean(um_area), mean_log2_AspectRatio = mean(log2_AspectRatio), mean_tot_det_ratio = mean(tot_det_ratio), mean_control_counts = mean(control_counts), median_total = median(total), median_detected = median(detected), median_umarea = median(um_area),median_log2_AspectRatio = median(log2_AspectRatio), median_tot_det_ratio = median(tot_det_ratio), median_control_counts = median(control_counts))

fov_positions$fov <- fov_positions$FOV
fov_positions <- left_join(fov_positions, meta_df, by = join_by(fov))

fov_size <- 4256
poly_df <- transmute(fov_positions, x_1 = X_px, y_1 = Y_px, 
                     x_2 = X_px + fov_size, y_2 = Y_px, 
                     x_3 = X_px + fov_size, y_3 = Y_px + fov_size, 
                     x_4 = X_px, y_4 = Y_px + fov_size)

str(poly_df)

ls_list <- list()
for (k in 1:dim(fov_positions)[1]){
  ls_list[[k]] <- st_linestring(matrix(as.numeric(poly_df[k]), ncol = 2, byrow = T))
}

sf_data <- st_as_sf(fov_positions, coords = c("X_px", "Y_px"), geometry = st_sfc(ls_list))

sf_data <- st_cast(sf_data, "POLYGON")

# plot(sf_data)
names(sf_data)
sf_data$fov <- as.factor(sf_data$fov)
```

```{r, out.width="150%", out.height="150%"}
ggplot() +
  geom_sf(data = sf_data, aes(fill = n_cells)) + 
  scale_fill_gradient(low = "purple4", high = "#FF8247") +
  theme(axis.title.x = element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(), 
        axis.title.y = element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        panel.grid = element_blank()) +
  geom_text(data = fov_positions, aes(x = fov_positions$X_px+fov_size/2,
                y = fov_positions$Y_px+fov_size/2,
                label = fov_positions$n_cells), color="white", size = 1.8) +
    ggtitle(paste0("Number of cells by fov"))
```

```{r, out.width="150%", out.height="150%"}
heatmap_variables <- names(sf_data)[-c(1:7, length(names(sf_data)))]
for (var in 1:length(heatmap_variables)){
  heat_var <- heatmap_variables[var]
  p <- ggplot() +
  geom_sf(data = sf_data, aes(fill = .data[[heat_var]])) + 
  scale_fill_viridis_c() +
  theme(axis.title.x = element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(), 
        axis.title.y = element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        panel.grid = element_blank()) +
  geom_text(data = fov_positions, aes(x = fov_positions$X_px+fov_size/2,
                y = fov_positions$Y_px+fov_size/2,
                label = fov_positions$fov), color="white", size = 2) +
    ggtitle(paste0(heat_var, " by fov"))
  print(p)
}  
```

ATOMX GUIDELINES FOR FOV QC

The Mean method (default) flags FOVs where the total count per cell average is below a threshold (default 100)

```{r, out.width="150%", out.height="150%"}
color <- ifelse(fov_positions$mean_total <= 100, "red", "white")
fov_positions <- cbind(fov_positions, color)
ggplot() +
  geom_sf(data = sf_data, aes(fill = mean_total)) + 
  scale_fill_viridis_c() +
  theme(axis.title.x = element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(), 
        axis.title.y = element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        panel.grid = element_blank()) +
  geom_text(data = fov_positions, aes(x = fov_positions$X_px+fov_size/2,
                y = fov_positions$Y_px+fov_size/2,
                label = floor(fov_positions$mean_total)), color = color, size = 2) +
    ggtitle(paste0("Mean_total by fov"))
```

ATOMX GUIDELINES FOR FOV QC

The QC quantile must be greater than the negative probe median in order to pass QC. Default is 0, so it must be ≥ 0

```{r, out.width="150%", out.height="150%"}
threshold <- 0.90
quant_df <-  data.frame(colData(spe))
quant_df <- mutate(quant_df, target_counts = total - control_counts)
quant_df <- group_by(quant_df, fov) |> summarize(quantile_90 = quantile(target_counts, probs = threshold))
fov_positions <- left_join(fov_positions, quant_df, by = join_by(fov))

color <- ifelse(fov_positions$quantile_90 < fov_positions$median_control_counts, "red", "white")
sf_data$quantile_90 <- fov_positions$quantile_90

ggplot() +
  geom_sf(data = sf_data, aes(fill = quantile_90)) + 
  scale_fill_viridis_c() +
  theme(axis.title.x = element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(), 
        axis.title.y = element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        panel.grid = element_blank()) +
  geom_text(data = fov_positions, aes(x = fov_positions$X_px+fov_size/2,
                y = fov_positions$Y_px+fov_size/2,
                label = floor(fov_positions$quantile_90)), color = color, size = 1.8) +
    ggtitle(paste0("FOV QC based on 90th quantile"))
```


# Feature boxplot by FOV

To compare the distribution of each feature of interest among FOVs

```{r, warnings = FALSE}
variables <- c("total", "detected", "logtot", "um_area", "log2_AspectRatio", "tot_det_ratio", "control_counts")

for (var in 1:length(variables)){
  n <- grep(paste0("^", variables[var], "$"), colnames(spe@colData))
  boxplot(spe@colData@listData[[n]] ~ spe$fov, ylab = colnames(spe@colData)[n], cex = 0.2)
  title(paste0(colnames(spe@colData)[n], "~fov"))
}
```

# Cell counts barplot by FOV

To view how many cells lie inside each FOV

```{r}
fov_df <- data.frame("n_cells" = as.integer(table(spe@colData@listData$fov)), "fov" = as.factor(names(table(spe@colData@listData$fov))))
levels(fov_df$fov) <- order(levels(fov_df$fov), seq(1:length(fov_df$fov)))

ggplot(data = fov_df, aes(x = fov, y = n_cells)) +
  geom_col(fill = "#B0C4DE") + theme(legend.position ="none") +
  ggtitle("Number of cells per fov") 
```

# Final report

```{r echo=T}
data <- data.frame(colData(spe))

df <- format(t(data.frame("Total_cell_number" = dim(spe)[2], "Total_counts" = sum(spe$total), "Total_features" = sum(spe$detected), "Total_ctrl_probe_counts" = sum(spe$control_counts), "Median_counts_per_cell" = median(spe$total), "Median_features_per_cell" = median(spe$detected),  "Median_ctrl_probe_counts" = median(spe$control_counts), "Median_um_area" = median(spe$um_area), "Median_log2_aspect_ratio" = median(spe$log2_AspectRatio), "Median_total_counts/features_ratio" = median(spe$tot_det_ratio, na.rm = T),  "Median_ctrl_probe_counts/total_counts_ratio" = median(spe$ctrl_total_ratio, na.rm = T))), scientific = FALSE)

rownames(df) <- c("Total cell number", "Total counts", "Total features", "Total ctrl probe counts", "Median counts per cell", "Median features per cell", "Median ctrl probe counts", "Median um area", "Median log2 aspect ratio", "Median total counts/features ratio", "Median ctrl probe counts/total counts ratio")

df[,1] <- round(as.numeric(df[,1]), 2)

kable(df, format = "html", row.names = T, col.names = NULL, align = "c", caption = "<center>Global summary metrics</center>",
      padding = 1L)
```

#
#

```{r}
flag_variables <- c("flagged_tot_20", 
                    "flagged_tot_10",
                    "flagged_tot_5",
                    "flagged_det_10",
                    "flagged_det_5",
                    "flagged_umarea_10",
                    "flagged_umarea_5",
                    "flagged_umarea_90",
                    "flagged_umarea_95",
                    "flagged_aratio_10",
                    "flagged_aratio_5",
                    "flagged_aratio_90",
                    "flagged_aratio_95",
                    "flagged_tdratio_10",
                    "flagged_tdratio_5",
                    "flagged_tdratio_atomx",
                    "flagged_negprob_50",
                    "flagged_negprob_75",
                    "ctrl_total_ratio")

flag_data <- data.frame(Unflagged_cells = as.integer(), Flagged_cells = as.integer(), Percentage_flagged = as.numeric())
for(i in 1:length(flag_variables)){
  var <- grep(paste0("^", flag_variables[i],"$"), names(colData(spe)))
  list_i <- c(sum(colData(spe)[,var]==FALSE, na.rm = T), sum(colData(spe)[,var]==TRUE, na.rm = T), round(sum(colData(spe)[,var]==TRUE, na.rm = T)/dim(spe)[2]*100, 2))
  flag_data[i, ] <- list_i
}

row.names(flag_data) <- c("Total counts < 20%", " Total counts < 10%", "Total counts < 5%", "Features < 10%", "Features < 5%", "Area < 10%", "Area < 5%", 
                         "Area > 90%","Area > 95%", "H/w < 10%", "H/w < 5%", "H/w > 90%", "H/w > 95%", "Counts/features < 10%", "Counts/features < 5%",
                         "Counts/features < 1", "Ctrl counts > 50%", "Ctrl counts > 75%", "Ctrl/total > 10%")

kable(flag_data, format = "html", padding = 1L, align = "c", col.names = c("Unflagged cells", "Flagged cells", "% flagged cells"), caption = paste0("<center>Flagged vs unflagged cells, total cells: " , dim(spe)[2],"</center>"))
```

#
#

ATOMX GUIDELINES FOR TARGET LEVEL QC

A value of 0.5 (default) will flag probes with lower total counts than the median (50th percentile) of the negative control probes' counts

```{r}
ctrl_counts <- counts(spe)[c(grep("^Negative", rownames(spe)), grep("^SystemControl", rownames(spe))), ]
target_counts <- counts(spe)[1:(dim(counts(spe))[1] - dim(ctrl_counts)[1]), ]
temp <- rowSums(target_counts) > median(rowSums(ctrl_counts))
count_data <- rowSums(counts(spe))
color <- ifelse(count_data < median(rowSums(ctrl_counts)), "red", "black") 
color <- ifelse(match(names(color), rownames(target_counts)), color, "grey")
color <- replace_na(color, "grey") 
count_df <- data_frame(count_data, color)
count_df$log_count <- log10(count_df$count_data)
plot(count_df$log_count, col = count_df$color)
```

# Flag score development

Starting from feature scatterplots to see the relationship between possible candidate features

Total counts ~ logged width/height ratio

```{r, out.width="150%", out.height="150%"}
cell_df <- data.frame(colData(spe))

p <- ggplot(data = cell_df) +
  geom_pointdensity(aes(x = log2_AspectRatio, y = total)) +
  scale_color_viridis_c() +
  theme(legend.position = "none",
        plot.margin = margin())

ggMarginal(p, type = "density", color = "#ccccff", fill = "#ccccff") 
```

Detected features ~ logged width/height ratio

```{r, out.width="150%", out.height="150%"}
p <- ggplot(data = cell_df) +
  geom_pointdensity(aes(x = log2_AspectRatio, y = detected)) +
  scale_color_viridis_c() +
  theme(legend.position = "none",
        plot.margin = margin())

ggMarginal(p, type = "density", color = "#ccccff", fill = "#ccccff") 
```

Total counts ~ area in um

```{r, out.width="150%", out.height="150%"}
p <- ggplot(data = cell_df) +
  geom_pointdensity(aes(x = um_area, y = total)) +
  scale_color_viridis_c() +
  theme(legend.position = "none",
        plot.margin = margin())

ggMarginal(p, type = "density", color = "#ccccff", fill = "#ccccff") 
```

Logged width/height ratio ~ negative control probe counts

```{r, out.width="150%", out.height="150%"}
p <- ggplot(data = cell_df) +
  geom_pointdensity(aes(x = control_counts, y = log2_AspectRatio)) +
  scale_color_viridis_c() +
  theme(legend.position = "none",
        plot.margin = margin())

ggMarginal(p, type = "density", color = "#ccccff", fill = "#ccccff") 
```

Logged width/height ratio ~ local coordinates on X axis in px

```{r, out.width="150%", out.height="150%"}
p <- ggplot(data = cell_df) +
  geom_pointdensity(aes(x = CenterX_local_px, y = log2_AspectRatio)) +
  scale_color_viridis_c() +
  theme(legend.position = "none",
        plot.margin = margin())

ggMarginal(p, type = "density", color = "#ccccff", fill = "#ccccff") 
```

Logged width/height ratio ~ local coordinates on X axis in px

```{r, out.width="150%", out.height="150%"}
p <- ggplot(data = cell_df) +
  geom_pointdensity(aes(x = CenterY_local_px, y = log2_AspectRatio)) +
  scale_color_viridis_c() +
  theme(legend.position = "none",
        plot.margin = margin())

ggMarginal(p, type = "density", color = "#ccccff", fill = "#ccccff") 
```

# Flag score formula

```{r}
spe$flag_score <- 1/(1 + exp(-0.3*spe$target_counts/spe$um_area + 0.8*abs(spe$log2_AspectRatio)))
summary(spe$flag_score)
plot(spe$flag_score)
```

```{r}
hist(spe$flag_score)
```

## Cell boundary file loading

Loading polygon file and setting the correct format as sf object

```{r}
spat_obj <- fread(file.path(dirname, "UAP7_02527-polygons.csv"))
str(spat_obj)
```

```{r}
spat_obj$id_fov <- paste(spat_obj$cellID, spat_obj$fov, sep = "_")
spat_obj <- spat_obj[spat_obj$id_fov %in% metadata$id_fov,]
length(table(spat_obj$id_fov))
```

```{r}
table(names(table(spat_obj$fov)) == names(table(metadata$fov)))
```

```{r}
sf_global <- st_as_sf(spat_obj, coords = c("x_global_px", "y_global_px"))

polygons_whole <- sf_global %>%
  group_by(fov, cellID) %>%
  summarize(geometry = st_combine(geometry))#st_combine = combine several feature geometries into one, without unioning or resolving internal boundaries

polygons_whole <- st_cast(polygons_whole, "POLYGON")
```


```{r, out.width="150%", out.height="150%"}
polygons_whole <- polygons_whole[rownames(polygons_whole) %in% rownames(metadata),]

polygons_whole$total <- spe$total
polygons_whole$target_counts <- spe$target_counts
polygons_whole$um_area <- spe$um_area
polygons_whole$log2_AspectRatio <- spe$log2_AspectRatio
polygons_whole$flag_score <- spe$flag_score
polygons_whole$cell_type <- spe$cell_type
polygons_whole$celltype_color <- as.character(spe$celltype_color)

polygons_whole$control_counts <- spe$control_counts
```


```{r, out.width="150%", out.height="150%"}
tm_shape(polygons_whole) + 
tm_fill(col = "celltype_color") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = sample_name,
            main.title.size = 0.5,
            frame = FALSE,
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0)) +
tm_add_legend(type = "fill", labels = names(celltype_palette), col = celltype_palette)
```


```{r, out.width="150%", out.height="150%"}
tm_shape(polygons_whole) + 
  tm_fill(col= "flag_score", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE)
```

Local threshold map of flag score, target counts, cell_area and logged width/height ratio, 20th percentile

```{r, out.width="300%", out.height="900%"}
metric_color <- ifelse(polygons_whole$flag_score < quantile(spe$flag_score, probs = 0.2), "red", "grey80")
target_color <- ifelse(polygons_whole$target_counts < quantile(spe$target_counts, probs = 0.2), "red", "grey80")
area_color <- ifelse(polygons_whole$um_area < quantile(spe$um_area, probs = 0.2), "red", "grey80")
count_area_ratio_color <- ifelse((polygons_whole$target_counts/polygons_whole$um_area) < quantile((spe$target_counts/spe$um_area), probs = 0.2), "red", "grey80")
log_hw_color_lo <- ifelse(polygons_whole$log2_AspectRatio < quantile(spe$log2_AspectRatio, probs = 0.2), "red", "grey80")
log_hw_color_hi <- ifelse(polygons_whole$log2_AspectRatio > quantile(spe$log2_AspectRatio, probs = 0.85), "red", "grey80")

polygons_whole <- cbind(polygons_whole, metric_color = metric_color, target_color = target_color, area_color = area_color, count_area_ratio_color = count_area_ratio_color, log_hw_color_lo = log_hw_color_lo, log_hw_color_hi = log_hw_color_hi)

p1 <- tm_shape(polygons_whole) + 
  tm_fill(col= "metric_color", alpha = 0.5) +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Flag score",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p2 <- tm_shape(polygons_whole) + 
  tm_fill(col= "target_color", alpha = 0.5) +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Target counts",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p3 <- tm_shape(polygons_whole) + 
  tm_fill(col= "area_color", alpha = 0.5) +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "cell area",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p4 <- tm_shape(polygons_whole) + 
  tm_fill(col= "count_area_ratio_color", alpha = 0.5) +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "counts/um area",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p5 <- tm_shape(polygons_whole) + 
  tm_fill(col= "log_hw_color_lo", alpha = 0.5) +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Logged AspectRatio below",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p6 <- tm_shape(polygons_whole) + 
  tm_fill(col= "log_hw_color_hi", alpha = 0.5) +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Logged AspectRatio above",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p7 <- tm_shape(polygons_whole) + 
tm_fill(col = "celltype_color") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "celltype",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

tmap_arrange(p1, p2, p3, p4, p5, p6, p7, ncol = 2)
```

```{r, out.width="150%", out.height="150%"}
plot_list <- list(p1, p2, p3, p4, p5, p6, p7 + 
                    tm_layout(frame = FALSE,
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0)) +
tm_add_legend(type = "fill", labels = names(celltype_palette), col = celltype_palette))

for(i in 1:length(plot_list)){
  print(plot_list[[i]])
}
```


```{r, out.width="150%", out.height="150%"}

flag_value <- c(round(quantile(polygons_whole$flag_score, probs = 0.2), 2),
                round(quantile(polygons_whole$target_counts, probs = 0.2), 2),
                round(quantile(polygons_whole$um_area, probs = 0.2), 2),
                round(quantile((polygons_whole$target_counts/polygons_whole$um_area), probs = 0.2), 2),
                round(quantile(polygons_whole$log2_AspectRatio, probs = 0.2), 2),
                round(quantile(polygons_whole$log2_AspectRatio, probs = 0.85), 2))

flagged_cells <- c(sum(polygons_whole$metric_color=="red", na.rm = T),
                   sum(polygons_whole$target_color=="red", na.rm = T),
                   sum(polygons_whole$area_color=="red", na.rm = T),
                   sum(polygons_whole$count_area_ratio_color=="red", na.rm = T),
                   sum(polygons_whole$log_hw_color_lo=="red", na.rm = T),
                   sum(polygons_whole$log_hw_color_hi=="red", na.rm = T))

unflagged_cells <- c(sum(polygons_whole$metric_color=="white", na.rm = T),
                     sum(polygons_whole$target_color=="white", na.rm = T),
                     sum(polygons_whole$area_color=="white", na.rm = T),
                     sum(polygons_whole$count_area_ratio_color=="white", na.rm = T),
                     sum(polygons_whole$log_hw_color_lo=="white", na.rm = T),
                     sum(polygons_whole$log_hw_color_hi=="white", na.rm = T))

percent_flagged <- round(flagged_cells/dim(polygons_whole)[1]*100, 2)

poly <- data.frame(threshold = c(0.2, 0.2, 0.2, 0.2, 0.2, 0.85), flag_value = flag_value, unflagged_cells = unflagged_cells, flagged_cells = flagged_cells, percent_flagged = percent_flagged, row.names = c("Flag score", "Target counts", "Area in um", "Target counts/um area", "Logged aspect ratio low", "Logged aspect ratio high"))

kable(poly, format = "html", col.names = c("Threshold", "Flag value", "Unflagged cells", "Flagged cells", "% flagged cells"), align = "c", caption = paste0("<center>Flagged cells by different metrics, total cells in whole sample: ", dim(polygons_whole)[1], ", flagged cells in whole sample: ", sum(spe$flag_score < quantile(spe$flag_score, probs = 0.2, na.rm = T)), "</center>"), padding = 1L)
```

#
#

Creating sf using local coordinates for geometry

```{r}
sf_local <- st_as_sf(spat_obj, coords = c("x_local_px", "y_local_px"))

polygons <- sf_local %>%
  group_by(fov, cellID) %>%
  summarize(geometry = st_combine(geometry))#st_combine = combine several feature geometries into one, without unioning or resolving internal boundaries

polygons <- st_cast(polygons, "POLYGON")
polygons <- polygons[rownames(polygons) %in% rownames(metadata),]

polygons$total <- spe$total
polygons$target_counts <- spe$target_counts
polygons$um_area <- spe$um_area
polygons$log2_AspectRatio <- spe$log2_AspectRatio
polygons$flag_score <- spe$flag_score
polygons$cell_type <- spe$cell_type
polygons$celltype_color <- as.character(spe$celltype_color)

polygons$control_counts <- spe$control_counts
```

Local feature map of flag score, fov 91

```{r, out.width="150%", out.height="150%"}
polygons91 <- filter(polygons, fov==91) 

tm_shape(polygons91) + 
  tm_fill(col= "flag_score", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE)
```

Local threshold map of flag score, target counts, um_area and logged width/height ratio, 20th percentile

```{r, out.width="300%", out.height="600%"}
metric_color <- ifelse(polygons91$flag_score < quantile(spe$flag_score, probs = 0.2), "red", "white")
target_color <- ifelse(polygons91$target_counts < quantile(spe$target_counts, probs = 0.2), "red", "white")
area_color <- ifelse(polygons91$um_area < quantile(spe$um_area, probs = 0.2), "red", "white")
count_area_ratio_color <- ifelse((polygons91$target_counts/polygons91$um_area) < quantile((spe$target_counts/spe$um_area), probs = 0.2), "red", "white")
log_hw_color_lo <- ifelse(polygons91$log2_AspectRatio < quantile(spe$log2_AspectRatio, probs = 0.2), "red", "white")
log_hw_color_hi <- ifelse(polygons91$log2_AspectRatio > quantile(spe$log2_AspectRatio, probs = 0.85), "red", "white")

polygons91 <- cbind(polygons91, metric_color = metric_color, target_color = target_color, area_color = area_color, count_area_ratio_color = count_area_ratio_color, log_hw_color_lo = log_hw_color_lo, log_hw_color_hi = log_hw_color_hi)

p1 <- tm_shape(polygons91) + 
  tm_fill(col= "metric_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Flag score",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))


p2 <- tm_shape(polygons91) + 
  tm_fill(col= "target_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Target counts",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p3 <- tm_shape(polygons91) + 
  tm_fill(col= "area_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "um area",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p4 <- tm_shape(polygons91) + 
  tm_fill(col= "count_area_ratio_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "counts/um area",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p5 <- tm_shape(polygons91) + 
  tm_fill(col= "log_hw_color_lo", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Logged AspectRatio below",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p6 <- tm_shape(polygons91) + 
  tm_fill(col= "log_hw_color_hi", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Logged AspectRatio above",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p7 <- tm_shape(polygons91) + 
tm_fill(col = "celltype_color") +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "celltype",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

tmap_arrange(p1, p2, p3, p4, p5, p6, p7, ncol = 2)
```

```{r, out.width="150%", out.height="150%"}
plot_list <- list(p1, p2, p3, p4, p5, p6, p7 + 
                    tm_layout(frame = FALSE,
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0)) +
tm_add_legend(type = "fill", labels = names(celltype_palette), col = celltype_palette))

for(i in 1:length(plot_list)){
  print(plot_list[[i]])
}
```


```{r, out.width="150%", out.height="150%"}

flag_value <- c(round(quantile(polygons91$flag_score, probs = 0.2), 2),
                round(quantile(polygons91$target_counts, probs = 0.2), 2),
                round(quantile(polygons91$um_area, probs = 0.2), 2),
                round(quantile((polygons91$target_counts/polygons91$um_area), probs = 0.2), 2),
                round(quantile(polygons91$log2_AspectRatio, probs = 0.2), 2),
                round(quantile(polygons91$log2_AspectRatio, probs = 0.85), 2))

flagged_cells <- c(sum(polygons91$metric_color=="red", na.rm = T),
                   sum(polygons91$target_color=="red", na.rm = T),
                   sum(polygons91$area_color=="red", na.rm = T),
                   sum(polygons91$count_area_ratio_color=="red", na.rm = T),
                   sum(polygons91$log_hw_color_lo=="red", na.rm = T),
                   sum(polygons91$log_hw_color_hi=="red", na.rm = T))

unflagged_cells <- c(sum(polygons91$metric_color=="white", na.rm = T),
                     sum(polygons91$target_color=="white", na.rm = T),
                     sum(polygons91$area_color=="white", na.rm = T),
                     sum(polygons91$count_area_ratio_color=="white", na.rm = T),
                     sum(polygons91$log_hw_color_lo=="white", na.rm = T),
                     sum(polygons91$log_hw_color_hi=="white", na.rm = T))

percent_flagged <- round(flagged_cells/dim(polygons91)[1]*100, 2)

poly91 <- data.frame(threshold = c(0.2, 0.2, 0.2, 0.2, 0.2, 0.85), flag_value = flag_value, unflagged_cells = unflagged_cells, flagged_cells = flagged_cells, percent_flagged = percent_flagged, row.names = c("Flag score", "Target counts", "Area in um", "Target counts/um area", "Logged aspect ratio low", "Logged aspect ratio high"))

kable(poly91, format = "html", col.names = c("Threshold", "Flag value", "Unflagged cells", "Flagged cells", "% flagged cells"), align = "c", caption = paste0("<center>Flagged cells by different metrics, total cells in fov: ", dim(polygons91)[1], ", total cells in whole sample: ", dim(polygons)[1], ", flagged cells in whole sample: ", sum(spe$flag_score < quantile(spe$flag_score, probs = 0.2, na.rm = T)), "</center>"), padding = 1L)
```

#
#

Local threshold map of flag score, 10th percentile

```{r, out.width="300%", out.height="900%"}
metric_color <- ifelse(polygons91$flag_score < quantile(spe$flag_score, probs = 0.1), "red", "white")
target_color <- ifelse(polygons91$target_counts < quantile(spe$target_counts, probs = 0.1), "red", "white")
area_color <- ifelse(polygons91$um_area < quantile(spe$um_area, probs = 0.1), "red", "white")
count_area_ratio_color <- ifelse((polygons91$target_counts/polygons91$um_area) < quantile((spe$target_counts/spe$um_area), probs = 0.1), "red", "white")
log_hw_color_lo <- ifelse(polygons91$log2_AspectRatio < quantile(spe$log2_AspectRatio, probs = 0.1), "red", "white")
log_hw_color_hi <- ifelse(polygons91$log2_AspectRatio > quantile(spe$log2_AspectRatio, probs = 0.95), "red", "white")

polygons91 <- cbind(polygons91, metric_color = metric_color, target_color = target_color, area_color = area_color, count_area_ratio_color = count_area_ratio_color, log_hw_color_lo = log_hw_color_lo, log_hw_color_hi = log_hw_color_hi)

p1 <- tm_shape(polygons91) + 
  tm_fill(col= "metric_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Flag score",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p2 <- tm_shape(polygons91) + 
  tm_fill(col= "target_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Target counts",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p3 <- tm_shape(polygons91) + 
  tm_fill(col= "area_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "um area",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p4 <- tm_shape(polygons91) + 
  tm_fill(col= "count_area_ratio_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "counts/um area",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p5 <- tm_shape(polygons91) + 
  tm_fill(col= "log_hw_color_lo", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Logged AspectRatio below",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p6 <- tm_shape(polygons91) + 
  tm_fill(col= "log_hw_color_hi", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Logged AspectRatio above",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p7 <- tm_shape(polygons91) + 
tm_fill(col = "celltype_color") +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "celltype",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

tmap_arrange(p1, p2, p3, p4, p5, p6, p7, ncol = 2)
```

```{r, out.width="150%", out.height="150%"}
plot_list <- list(p1, p2, p3, p4, p5, p6, p7 + 
                    tm_layout(frame = FALSE,
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0)) +
tm_add_legend(type = "fill", labels = names(celltype_palette), col = celltype_palette))

for(i in 1:length(plot_list)){
  print(plot_list[[i]])
}
```

```{r, out.width="300%", out.height="300%"}

flag_value <- c(round(quantile(polygons91$flag_score, probs = 0.1), 2),
                round(quantile(polygons91$target_counts, probs = 0.1), 2),
                round(quantile(polygons91$um_area, probs = 0.1), 2),
                round(quantile((polygons91$target_counts/polygons91$um_area), probs = 0.1), 2),
                round(quantile(polygons91$log2_AspectRatio, probs = 0.1), 2),
                round(quantile(polygons91$log2_AspectRatio, probs = 0.95), 2))

flagged_cells <- c(sum(polygons91$metric_color=="red", na.rm = T),
                   sum(polygons91$target_color=="red", na.rm = T),
                   sum(polygons91$area_color=="red", na.rm = T),
                   sum(polygons91$count_area_ratio_color=="red", na.rm = T),
                   sum(polygons91$log_hw_color_lo=="red", na.rm = T),
                   sum(polygons91$log_hw_color_hi=="red", na.rm = T))

unflagged_cells <- c(sum(polygons91$metric_color=="white", na.rm = T),
                     sum(polygons91$target_color=="white", na.rm = T),
                     sum(polygons91$area_color=="white", na.rm = T),
                     sum(polygons91$count_area_ratio_color=="white", na.rm = T),
                     sum(polygons91$log_hw_color_lo=="white", na.rm = T),
                     sum(polygons91$log_hw_color_hi=="white", na.rm = T))

percent_flagged <- round(flagged_cells/dim(polygons91)[1]*100, 2)

poly91 <- data.frame(threshold = c(0.1, 0.1, 0.1, 0.1, 0.1, 0.95), flag_value = flag_value, unflagged_cells = unflagged_cells, flagged_cells = flagged_cells, percent_flagged = percent_flagged, row.names = c("Flag score", "Target counts", "Area in um", "Target counts/um area", "Logged aspect ratio low", "Logged aspect ratio high"))

kable(poly91, format = "html", col.names = c("Threshold", "Flag value", "Unflagged cells", "Flagged cells", "% flagged cells"), align = "c", caption = paste0("<center>Flagged cells by different metrics, total cells in fov: ", dim(polygons91)[1], ", total cells in whole sample: ", dim(polygons)[1], ", flagged cells in whole sample: ", sum(spe$flag_score < quantile(spe$flag_score, probs = 0.1, na.rm = T)), "</center>"), padding = 1L)
```

#
#

Local feature map of flag score, fov 56

```{r, out.width="150%", out.height="150%"}
polygons56 <- filter(polygons, fov==56)

tm_shape(polygons56) + 
  tm_fill(col= "flag_score", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE)

```

Local threshold map of flag score, 15th percentile

```{r, out.width="300%", out.height="900%"}
metric_color <- ifelse(polygons56$flag_score < quantile(spe$flag_score, probs = 0.15), "red", "white")
target_color <- ifelse(polygons56$target_counts < quantile(spe$target_counts, probs = 0.15), "red", "white")
area_color <- ifelse(polygons56$um_area < quantile(spe$um_area, probs = 0.15), "red", "white")
count_area_ratio_color <- ifelse((polygons56$target_counts/polygons56$um_area) < quantile((spe$target_counts/spe$um_area), probs = 0.15, na.rm = T), "red", "white")
log_hw_color_lo <- ifelse(polygons56$log2_AspectRatio < quantile(spe$log2_AspectRatio, probs = 0.15), "red", "white")
log_hw_color_hi <- ifelse(polygons56$log2_AspectRatio > quantile(spe$log2_AspectRatio, probs = 0.90), "red", "white")

polygons56 <- cbind(polygons56, metric_color = metric_color, target_color = target_color, area_color = area_color, count_area_ratio_color = count_area_ratio_color, log_hw_color_lo = log_hw_color_lo, log_hw_color_hi = log_hw_color_hi)

p1 <- tm_shape(polygons56) + 
  tm_fill(col= "metric_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Flag score",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p2 <- tm_shape(polygons56) + 
  tm_fill(col= "target_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Target counts",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p3 <- tm_shape(polygons56) + 
  tm_fill(col= "area_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "um area",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p4 <- tm_shape(polygons56) + 
  tm_fill(col= "count_area_ratio_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "counts/um area",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p5 <- tm_shape(polygons56) + 
  tm_fill(col= "log_hw_color_lo", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Logged AspectRatio below",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p6 <- tm_shape(polygons56) + 
  tm_fill(col= "log_hw_color_hi", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Logged AspectRatio above",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))


p7 <- tm_shape(polygons56) + 
tm_fill(col = "celltype_color") +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "celltype",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

tmap_arrange(p1, p2, p3, p4, p5, p6, p7, ncol = 2)
```

```{r, out.width="150%", out.height="150%"}
plot_list <- list(p1, p2, p3, p4, p5, p6, p7 + 
                    tm_layout(frame = FALSE,
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0)) +
tm_add_legend(type = "fill", labels = names(celltype_palette), col = celltype_palette))

for(i in 1:length(plot_list)){
  print(plot_list[[i]])
}
```

```{r, out.width="150%", out.height="150%"}

flag_value <- c(round(quantile(polygons56$flag_score, probs = 0.15), 2),
                round(quantile(polygons56$target_counts, probs = 0.15), 2),
                round(quantile(polygons56$um_area, probs = 0.15), 2),
                round(quantile((polygons56$target_counts/polygons56$um_area), probs = 0.15), 2),
                round(quantile(polygons56$log2_AspectRatio, probs = 0.15), 2),
                round(quantile(polygons56$log2_AspectRatio, probs = 0.90), 2))

flagged_cells <- c(sum(polygons56$metric_color=="red", na.rm = T),
                   sum(polygons56$target_color=="red", na.rm = T),
                   sum(polygons56$area_color=="red", na.rm = T),
                   sum(polygons56$count_area_ratio_color=="red", na.rm = T),
                   sum(polygons56$log_hw_color_lo=="red", na.rm = T),
                   sum(polygons56$log_hw_color_hi=="red", na.rm = T))

unflagged_cells <- c(sum(polygons56$metric_color=="white", na.rm = T),
                     sum(polygons56$target_color=="white", na.rm = T),
                     sum(polygons56$area_color=="white", na.rm = T),
                     sum(polygons56$count_area_ratio_color=="white", na.rm = T),
                     sum(polygons56$log_hw_color_lo=="white", na.rm = T),
                     sum(polygons56$log_hw_color_hi=="white", na.rm = T))

percent_flagged <- round(flagged_cells/dim(polygons56)[1]*100, 2)

poly56 <- data.frame(threshold = c(0.15, 0.15, 0.15, 0.15, 0.15, 0.90), flag_value = flag_value, unflagged_cells = unflagged_cells, flagged_cells = flagged_cells, 
percent_flagged = percent_flagged, row.names = c("Flag score", "Target counts", "Area in um", "Target counts/um area", "Logged aspect ratio low", "Logged aspect ratio high"))

kable(poly56, format = "html", col.names = c("Threshold", "Flag value", "Unflagged cells", "Flagged cells", "% flagged cells"), align = "c", caption = paste0("<center>Flagged cells by different metrics, total cells in fov: ", dim(polygons56)[1], ", total cells in whole sample: ", dim(polygons)[1], ", flagged cells in whole sample: ", sum(spe$flag_score < quantile(spe$flag_score, probs = 0.15, na.rm = T)), "</center>"), padding = 1L)
```

#
#

Local feature map of flag score, fov 72

```{r, out.width="150%", out.height="150%"}
polygons72 <- filter(polygons, fov==72)

tm_shape(polygons72) + 
  tm_fill(col= "flag_score", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE)

```

Local threshold map of flag score, 15th percentile

```{r, out.width="300%", out.height="900%"}
metric_color <- ifelse(polygons72$flag_score < quantile(spe$flag_score, probs = 0.15), "red", "white")
target_color <- ifelse(polygons72$target_counts < quantile(spe$target_counts, probs = 0.15), "red", "white")
area_color <- ifelse(polygons72$um_area < quantile(spe$um_area, probs = 0.15), "red", "white")
count_area_ratio_color <- ifelse((polygons72$target_counts/polygons72$um_area) < quantile((spe$target_counts/spe$um_area), probs = 0.15, na.rm = T), "red", "white")
log_hw_color_lo <- ifelse(polygons72$log2_AspectRatio < quantile(spe$log2_AspectRatio, probs = 0.15), "red", "white")
log_hw_color_hi <- ifelse(polygons72$log2_AspectRatio > quantile(spe$log2_AspectRatio, probs = 0.90), "red", "white")

polygons72 <- cbind(polygons72, metric_color = metric_color, target_color = target_color, area_color = area_color, count_area_ratio_color = count_area_ratio_color, log_hw_color_lo = log_hw_color_lo, log_hw_color_hi = log_hw_color_hi)

p1 <- tm_shape(polygons72) + 
  tm_fill(col= "metric_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Flag score",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p2 <- tm_shape(polygons72) + 
  tm_fill(col= "target_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Target counts",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p3 <- tm_shape(polygons72) + 
  tm_fill(col= "area_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "um area",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p4 <- tm_shape(polygons72) + 
  tm_fill(col= "count_area_ratio_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "counts/um area",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p5 <- tm_shape(polygons72) + 
  tm_fill(col= "log_hw_color_lo", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Logged AspectRatio below",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p6 <- tm_shape(polygons72) + 
  tm_fill(col= "log_hw_color_hi", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Logged AspectRatio above",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p7 <- tm_shape(polygons72) + 
tm_fill(col = "celltype_color") +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "celltype",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

tmap_arrange(p1, p2, p3, p4, p5, p6, p7, ncol = 2)
```

```{r, out.width="150%", out.height="150%"}
plot_list <- list(p1, p2, p3, p4, p5, p6, p7 + 
                    tm_layout(frame = FALSE,
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0)) +
tm_add_legend(type = "fill", labels = names(celltype_palette), col = celltype_palette))

for(i in 1:length(plot_list)){
  print(plot_list[[i]])
}
```

```{r, out.width="150%", out.height="150%"}

flag_value <- c(round(quantile(polygons72$flag_score, probs = 0.15), 2),
                round(quantile(polygons72$target_counts, probs = 0.15), 2),
                round(quantile(polygons72$um_area, probs = 0.15), 2),
                round(quantile((polygons72$target_counts/polygons72$um_area), probs = 0.15), 2),
                round(quantile(polygons72$log2_AspectRatio, probs = 0.15), 2),
                round(quantile(polygons72$log2_AspectRatio, probs = 0.90), 2))

flagged_cells <- c(sum(polygons72$metric_color=="red", na.rm = T),
                   sum(polygons72$target_color=="red", na.rm = T),
                   sum(polygons72$area_color=="red", na.rm = T),
                   sum(polygons72$count_area_ratio_color=="red", na.rm = T),
                   sum(polygons72$log_hw_color_lo=="red", na.rm = T),
                   sum(polygons72$log_hw_color_hi=="red", na.rm = T))

unflagged_cells <- c(sum(polygons72$metric_color=="white", na.rm = T),
                     sum(polygons72$target_color=="white", na.rm = T),
                     sum(polygons72$area_color=="white", na.rm = T),
                     sum(polygons72$count_area_ratio_color=="white", na.rm = T),
                     sum(polygons72$log_hw_color_lo=="white", na.rm = T),
                     sum(polygons72$log_hw_color_hi=="white", na.rm = T))

percent_flagged <- round(flagged_cells/dim(polygons72)[1]*100, 2)

poly72 <- data.frame(threshold = c(0.15, 0.15, 0.15, 0.15, 0.15, 0.90), flag_value = flag_value, unflagged_cells = unflagged_cells, flagged_cells = flagged_cells,
percent_flagged = percent_flagged, row.names = c("Flag score", "Target counts", "Area in um", "Target counts/um area", "Logged aspect ratio low", "Logged aspect ratio high"))

kable(poly72, format = "html", col.names = c("Threshold", "Flag value", "Unflagged cells", "Flagged cells", "% flagged cells"), align = "c", caption = paste0("<center>Flagged cells by different metrics, total cells in fov: ", dim(polygons72)[1], ", total cells in whole sample: ", dim(polygons)[1], ", flagged cells in whole sample: ", sum(spe$flag_score < quantile(spe$flag_score, probs = 0.15, na.rm = T)), "</center>"), padding = 1L)
```


#
#

Local feature map of flag score, fov 66

```{r, out.width="150%", out.height="150%"}
polygons66 <- filter(polygons, fov==66)

tm_shape(polygons66) + 
  tm_fill(col= "flag_score", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE)

```

Local threshold map of flag score, 15th percentile

```{r, out.width="300%", out.height="900%"}
metric_color <- ifelse(polygons66$flag_score < quantile(spe$flag_score, probs = 0.15), "red", "white")
target_color <- ifelse(polygons66$target_counts < quantile(spe$target_counts, probs = 0.15), "red", "white")
area_color <- ifelse(polygons66$um_area < quantile(spe$um_area, probs = 0.15), "red", "white")
count_area_ratio_color <- ifelse((polygons66$target_counts/polygons66$um_area) < quantile((spe$target_counts/spe$um_area), probs = 0.15, na.rm = T), "red", "white")
log_hw_color_lo <- ifelse(polygons66$log2_AspectRatio < quantile(spe$log2_AspectRatio, probs = 0.15), "red", "white")
log_hw_color_hi <- ifelse(polygons66$log2_AspectRatio > quantile(spe$log2_AspectRatio, probs = 0.90), "red", "white")

polygons66 <- cbind(polygons66, metric_color = metric_color, target_color = target_color, area_color = area_color, count_area_ratio_color = count_area_ratio_color, log_hw_color_lo = log_hw_color_lo, log_hw_color_hi = log_hw_color_hi)

p1 <- tm_shape(polygons66) + 
  tm_fill(col= "metric_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Flag score",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p2 <- tm_shape(polygons66) + 
  tm_fill(col= "target_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Target counts",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p3 <- tm_shape(polygons66) + 
  tm_fill(col= "area_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "um area",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p4 <- tm_shape(polygons66) + 
  tm_fill(col= "count_area_ratio_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "counts/um area",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p5 <- tm_shape(polygons66) + 
  tm_fill(col= "log_hw_color_lo", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Logged AspectRatio below",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p6 <- tm_shape(polygons66) + 
  tm_fill(col= "log_hw_color_hi", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Logged AspectRatio above",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p7 <- tm_shape(polygons66) + 
tm_fill(col = "celltype_color") +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "celltype",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

tmap_arrange(p1, p2, p3, p4, p5, p6, p7, ncol = 2)
```

```{r, out.width="150%", out.height="150%"}
plot_list <- list(p1, p2, p3, p4, p5, p6, p7 + 
                    tm_layout(frame = FALSE,
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0)) +
tm_add_legend(type = "fill", labels = names(celltype_palette), col = celltype_palette))

for(i in 1:length(plot_list)){
  print(plot_list[[i]])
}
```

```{r, out.width="150%", out.height="150%"}

flag_value <- c(round(quantile(polygons66$flag_score, probs = 0.15), 2),
                round(quantile(polygons66$target_counts, probs = 0.15), 2),
                round(quantile(polygons66$um_area, probs = 0.15), 2),
                round(quantile((polygons66$target_counts/polygons66$um_area), probs = 0.15), 2),
                round(quantile(polygons66$log2_AspectRatio, probs = 0.15), 2),
                round(quantile(polygons66$log2_AspectRatio, probs = 0.90), 2))

flagged_cells <- c(sum(polygons66$metric_color=="red", na.rm = T),
                   sum(polygons66$target_color=="red", na.rm = T),
                   sum(polygons66$area_color=="red", na.rm = T),
                   sum(polygons66$count_area_ratio_color=="red", na.rm = T),
                   sum(polygons66$log_hw_color_lo=="red", na.rm = T),
                   sum(polygons66$log_hw_color_hi=="red", na.rm = T))

unflagged_cells <- c(sum(polygons66$metric_color=="white", na.rm = T),
                     sum(polygons66$target_color=="white", na.rm = T),
                     sum(polygons66$area_color=="white", na.rm = T),
                     sum(polygons66$count_area_ratio_color=="white", na.rm = T),
                     sum(polygons66$log_hw_color_lo=="white", na.rm = T),
                     sum(polygons66$log_hw_color_hi=="white", na.rm = T))

percent_flagged <- round(flagged_cells/dim(polygons66)[1]*100, 2)

poly66 <- data.frame(threshold = c(0.15, 0.15, 0.15, 0.15, 0.15, 0.90), flag_value = flag_value, unflagged_cells = unflagged_cells, flagged_cells = flagged_cells,
percent_flagged = percent_flagged, row.names = c("Flag score", "Target counts", "Area in um", "Target counts/um area", "Logged aspect ratio low", "Logged aspect ratio high"))

kable(poly66, format = "html", col.names = c("Threshold", "Flag value", "Unflagged cells", "Flagged cells", "% flagged cells"), align = "c", caption = paste0("<center>Flagged cells by different metrics, total cells in fov: ", dim(polygons66)[1], ", total cells in whole sample: ", dim(polygons)[1], ", flagged cells in whole sample: ", sum(spe$flag_score < quantile(spe$flag_score, probs = 0.15, na.rm = T)), "</center>"), padding = 1L)
```


#
#

Local feature map of flag score, fov 146

```{r, out.width="150%", out.height="150%"}
polygons146 <- filter(polygons, fov==146)

tm_shape(polygons146) + 
  tm_fill(col= "flag_score", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE)

```

Local threshold map of flag score, 15th percentile

```{r, out.width="300%", out.height="900%"}
metric_color <- ifelse(polygons146$flag_score < quantile(spe$flag_score, probs = 0.15), "red", "white")
target_color <- ifelse(polygons146$target_counts < quantile(spe$target_counts, probs = 0.15), "red", "white")
area_color <- ifelse(polygons146$um_area < quantile(spe$um_area, probs = 0.15), "red", "white")
count_area_ratio_color <- ifelse((polygons146$target_counts/polygons146$um_area) < quantile((spe$target_counts/spe$um_area), probs = 0.15, na.rm = T), "red", "white")
log_hw_color_lo <- ifelse(polygons146$log2_AspectRatio < quantile(spe$log2_AspectRatio, probs = 0.15), "red", "white")
log_hw_color_hi <- ifelse(polygons146$log2_AspectRatio > quantile(spe$log2_AspectRatio, probs = 0.90), "red", "white")

polygons146 <- cbind(polygons146, metric_color = metric_color, target_color = target_color, area_color = area_color, count_area_ratio_color = count_area_ratio_color, log_hw_color_lo = log_hw_color_lo, log_hw_color_hi = log_hw_color_hi)

p1 <- tm_shape(polygons146) + 
  tm_fill(col= "metric_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Flag score",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p2 <- tm_shape(polygons146) + 
  tm_fill(col= "target_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Target counts",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p3 <- tm_shape(polygons146) + 
  tm_fill(col= "area_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "um area",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p4 <- tm_shape(polygons146) + 
  tm_fill(col= "count_area_ratio_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "counts/um area",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p5 <- tm_shape(polygons146) + 
  tm_fill(col= "log_hw_color_lo", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Logged AspectRatio below",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p6 <- tm_shape(polygons146) + 
  tm_fill(col= "log_hw_color_hi", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Logged AspectRatio above",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p7 <- tm_shape(polygons146) + 
tm_fill(col = "celltype_color") +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "celltype",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

tmap_arrange(p1, p2, p3, p4, p5, p6, p7, ncol = 2)
```

```{r, out.width="150%", out.height="150%"}
plot_list <- list(p1, p2, p3, p4, p5, p6, p7 + 
                    tm_layout(frame = FALSE,
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0)) +
tm_add_legend(type = "fill", labels = names(celltype_palette), col = celltype_palette))

for(i in 1:length(plot_list)){
  print(plot_list[[i]])
}
```

```{r, out.width="150%", out.height="150%"}

flag_value <- c(round(quantile(polygons146$flag_score, probs = 0.15), 2),
                round(quantile(polygons146$target_counts, probs = 0.15), 2),
                round(quantile(polygons146$um_area, probs = 0.15), 2),
                round(quantile((polygons146$target_counts/polygons146$um_area), probs = 0.15), 2),
                round(quantile(polygons146$log2_AspectRatio, probs = 0.15), 2),
                round(quantile(polygons146$log2_AspectRatio, probs = 0.90), 2))

flagged_cells <- c(sum(polygons146$metric_color=="red", na.rm = T),
                   sum(polygons146$target_color=="red", na.rm = T),
                   sum(polygons146$area_color=="red", na.rm = T),
                   sum(polygons146$count_area_ratio_color=="red", na.rm = T),
                   sum(polygons146$log_hw_color_lo=="red", na.rm = T),
                   sum(polygons146$log_hw_color_hi=="red", na.rm = T))

unflagged_cells <- c(sum(polygons146$metric_color=="white", na.rm = T),
                     sum(polygons146$target_color=="white", na.rm = T),
                     sum(polygons146$area_color=="white", na.rm = T),
                     sum(polygons146$count_area_ratio_color=="white", na.rm = T),
                     sum(polygons146$log_hw_color_lo=="white", na.rm = T),
                     sum(polygons146$log_hw_color_hi=="white", na.rm = T))

percent_flagged <- round(flagged_cells/dim(polygons146)[1]*100, 2)

poly146 <- data.frame(threshold = c(0.15, 0.15, 0.15, 0.15, 0.15, 0.90), flag_value = flag_value, unflagged_cells = unflagged_cells, flagged_cells = flagged_cells,
percent_flagged = percent_flagged, row.names = c("Flag score", "Target counts", "Area in um", "Target counts/um area", "Logged aspect ratio low", "Logged aspect ratio high"))

kable(poly146, format = "html", col.names = c("Threshold", "Flag value", "Unflagged cells", "Flagged cells", "% flagged cells"), align = "c", caption = paste0("<center>Flagged cells by different metrics, total cells in fov: ", dim(polygons146)[1], ", total cells in whole sample: ", dim(polygons)[1], ", flagged cells in whole sample: ", sum(spe$flag_score < quantile(spe$flag_score, probs = 0.15, na.rm = T)), "</center>"), padding = 1L)
```


#
#

Local feature map of flag score, fov 20

```{r, out.width="150%", out.height="150%"}
polygons20 <- filter(polygons, fov==20)

tm_shape(polygons20) + 
  tm_fill(col= "flag_score", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE)

```

Local threshold map of flag score, 15th percentile

```{r, out.width="300%", out.height="900%"}
metric_color <- ifelse(polygons20$flag_score < quantile(spe$flag_score, probs = 0.15), "red", "white")
target_color <- ifelse(polygons20$target_counts < quantile(spe$target_counts, probs = 0.15), "red", "white")
area_color <- ifelse(polygons20$um_area < quantile(spe$um_area, probs = 0.15), "red", "white")
count_area_ratio_color <- ifelse((polygons20$target_counts/polygons20$um_area) < quantile((spe$target_counts/spe$um_area), probs = 0.15, na.rm = T), "red", "white")
log_hw_color_lo <- ifelse(polygons20$log2_AspectRatio < quantile(spe$log2_AspectRatio, probs = 0.15), "red", "white")
log_hw_color_hi <- ifelse(polygons20$log2_AspectRatio > quantile(spe$log2_AspectRatio, probs = 0.90), "red", "white")

polygons20 <- cbind(polygons20, metric_color = metric_color, target_color = target_color, area_color = area_color, count_area_ratio_color = count_area_ratio_color, log_hw_color_lo = log_hw_color_lo, log_hw_color_hi = log_hw_color_hi)

p1 <- tm_shape(polygons20) + 
  tm_fill(col= "metric_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Flag score",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p2 <- tm_shape(polygons20) + 
  tm_fill(col= "target_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Target counts",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p3 <- tm_shape(polygons20) + 
  tm_fill(col= "area_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "um area",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p4 <- tm_shape(polygons20) + 
  tm_fill(col= "count_area_ratio_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "counts/um area",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p5 <- tm_shape(polygons20) + 
  tm_fill(col= "log_hw_color_lo", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Logged AspectRatio below",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p6 <- tm_shape(polygons20) + 
  tm_fill(col= "log_hw_color_hi", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Logged AspectRatio above",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p7 <- tm_shape(polygons20) + 
tm_fill(col = "celltype_color") +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "celltype",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

tmap_arrange(p1, p2, p3, p4, p5, p6, p7, ncol = 2)
```

```{r, out.width="150%", out.height="150%"}
plot_list <- list(p1, p2, p3, p4, p5, p6, p7 + 
                    tm_layout(frame = FALSE,
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0)) +
tm_add_legend(type = "fill", labels = names(celltype_palette), col = celltype_palette))

for(i in 1:length(plot_list)){
  print(plot_list[[i]])
}
```

```{r, out.width="150%", out.height="150%"}

flag_value <- c(round(quantile(polygons20$flag_score, probs = 0.15), 2),
                round(quantile(polygons20$target_counts, probs = 0.15), 2),
                round(quantile(polygons20$um_area, probs = 0.15), 2),
                round(quantile((polygons20$target_counts/polygons20$um_area), probs = 0.15), 2),
                round(quantile(polygons20$log2_AspectRatio, probs = 0.15), 2),
                round(quantile(polygons20$log2_AspectRatio, probs = 0.90), 2))

flagged_cells <- c(sum(polygons20$metric_color=="red", na.rm = T),
                   sum(polygons20$target_color=="red", na.rm = T),
                   sum(polygons20$area_color=="red", na.rm = T),
                   sum(polygons20$count_area_ratio_color=="red", na.rm = T),
                   sum(polygons20$log_hw_color_lo=="red", na.rm = T),
                   sum(polygons20$log_hw_color_hi=="red", na.rm = T))

unflagged_cells <- c(sum(polygons20$metric_color=="white", na.rm = T),
                     sum(polygons20$target_color=="white", na.rm = T),
                     sum(polygons20$area_color=="white", na.rm = T),
                     sum(polygons20$count_area_ratio_color=="white", na.rm = T),
                     sum(polygons20$log_hw_color_lo=="white", na.rm = T),
                     sum(polygons20$log_hw_color_hi=="white", na.rm = T))

percent_flagged <- round(flagged_cells/dim(polygons20)[1]*100, 2)

poly20 <- data.frame(threshold = c(0.15, 0.15, 0.15, 0.15, 0.15, 0.90), flag_value = flag_value, unflagged_cells = unflagged_cells, flagged_cells = flagged_cells,
percent_flagged = percent_flagged, row.names = c("Flag score", "Target counts", "Area in um", "Target counts/um area", "Logged aspect ratio low", "Logged aspect ratio high"))

kable(poly20, format = "html", col.names = c("Threshold", "Flag value", "Unflagged cells", "Flagged cells", "% flagged cells"), align = "c", caption = paste0("<center>Flagged cells by different metrics, total cells in fov: ", dim(polygons20)[1], ", total cells in whole sample: ", dim(polygons)[1], ", flagged cells in whole sample: ", sum(spe$flag_score < quantile(spe$flag_score, probs = 0.15, na.rm = T)), "</center>"), padding = 1L)
```

