---
title: "02_21_cosmx_mouse_hippo_preprocessing_metric_report"
author: "Benedetta Banzi"
date: "2024-02-23"
output: html_document
---

<STYLE TYPE="text/css">
<!--
  td{
    font-family: Arial; 
    font-size: 10 pt;
    padding:0px;
    cellpadding="0";
    cellspacing="0"
  }
  th {
    font-family: Arial; 
    font-size: 10 pt;
    height: 20px;
    font-weight: bold;
    text-align: "center";
    background-color: #ccccff;
    padding-left: 15px;
    padding-right: 15px;
  }
  table { 
    border-spacing: 4px;
    border-collapse: collapse;
  }
--->
</STYLE>


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Library import

```{r}
library(SpatialExperiment)
library(data.table)
library(scater) # it imports scuttle
library(cowplot)
library(ggplot2)
theme_set(theme_bw())
library(matrixStats)
library(dplyr)
library(tidyr)
library(tibble)
library(arrow)
library(scales)
library(sf)
library(tmap)
library(knitr)
library(grid)
library(ggpointdensity)
library(ggpubr)
library(spdep)
```

Read-in function

```{r}
readCosmxSPE <- function(dirname = dirname, 
                         countmatfpattern = "exprMat_file.csv", 
                         metadatafpattern = "metadata_file.csv", 
                         coord_names = c("CenterX_global_px",
                                         "CenterY_global_px")){
  
  countmat_file <- file.path(dirname, list.files(dirname, countmatfpattern))
  metadata_file <- file.path(dirname, list.files(dirname, metadatafpattern))
  
  # Read in 
  countmat <- data.table::fread(countmat_file)
  metadata <- data.table::fread(metadata_file)
  
  # Count matrix   
  counts <- merge(countmat, metadata[, c("fov", "cell_ID")])
  counts <- subset(counts, select = -c(fov, cell_ID))
  cell_codes <- rownames(counts)
  features <- colnames(counts)
  counts <- t(counts)
  
  rownames(counts) <- features
  colnames(counts) <- cell_codes
  
  # rowData (does not exist)
  
  # colData
  colData <- merge(metadata, countmat[, c("fov", "cell_ID")])
  
  spe <- SpatialExperiment::SpatialExperiment(
    assays = list(counts = counts),
    # rowData = rowData,
    colData = colData,
    spatialCoordsNames = coord_names
  )
  
  return(spe)
}
```

Setting directories

```{r}
dirname <- "/NAS06/work/Spatial_omics/CosMx/Mouse_brain/coronal_hippocampus_and_cortex"
count_pattern <- "Run5642_S3_Quarter_exprMat_file.csv"
meta_pattern <- "Run5642_S3_Quarter_metadata_file.csv"
sample_name <- "CosMx_Mouse_hippocampus_cortex"
```

# CosMx dataset exploratory analysis for preprocessing

SpatialExperiment object creation and exploration

```{r}
spe <- readCosmxSPE(dirname = dirname, countmatfpattern = count_pattern, metadatafpattern = meta_pattern)
spe <- scuttle::addPerCellQC(spe, subsets=list(NegProb = grep("^NegPrb", rownames(spe))))
names(colData(spe)@listData)
```

```{r}
dim(spe)
```

Selecting only numeric and integer variables from metadata then only variables of interest from the list

```{r}
temp <- unlist(lapply(spe@colData@listData, function(x) is.numeric(x) | is.integer(x)))
num_variables <- c()
for (i in 1:length(temp)){
  num_variables[i] <- ifelse(temp[i] == TRUE, names(temp[i]), NA)
  num_variables <- num_variables[!is.na(num_variables)]
}
num_variables
variables <- c("total", "detected")
```

## Large-scale to small-scale

Approaching dataset preprocessing starting from global-scale, then fov-focused exploratory analysis

# Global sample map

To view the spatial organization of both the sample as it is and FOVs

```{r}
# Importing global coordinates from spatialCoords slot since they are not available in colData
colData(spe)$CenterX_global_px <- spatialCoords(spe)[,1]
colData(spe)$CenterY_global_px <- spatialCoords(spe)[,2]
```

```{r, out.width="150%", out.height="150%"}
fov_pattern <- "Run5642_S3_Quarter_fov_positions_file.csv"
fovpos_file <- file.path(dirname, list.files(dirname, fov_pattern))

metadata <- data.table::fread(file.path(dirname, list.files(dirname, meta_pattern)))
fov_positions <- data.table::fread(fovpos_file, header = T)
fov_positions <- fov_positions[fov_positions$fov%in%unique(metadata$fov),]

# image theme
global_map_theme <- function(back.color="black",back.border=NA,title.col="white") {
  theme(aspect.ratio = 1,
        panel.border = element_blank(),
        legend.key = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks = element_blank(),
        axis.text.y = element_blank(),
        axis.text.x = element_blank(),
        panel.grid = element_blank(),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        panel.background = element_rect(fill = "transparent",colour = NA),
        plot.title = element_text(color = title.col,hjust = 0.5,face = "bold"),
        plot.background = element_rect(fill = "transparent",colour = back.border))
}

fov_size <- 4256

ggplot() + 
  geom_point(data = metadata, mapping = aes(x = CenterX_global_px, y = CenterY_global_px), colour = "darkmagenta", fill="darkmagenta", size = 0.05, alpha = 0.2) +
  annotate("rect", xmin = fov_positions$x_global_px, 
           xmax = fov_positions$x_global_px + fov_size, 
           ymin = fov_positions$y_global_px, 
           ymax = fov_positions$y_global_px + fov_size,
           alpha = .2, color = "black",linewidth = 0.2) +
  geom_text(aes(x = fov_positions$x_global_px+fov_size/2,
                y = fov_positions$y_global_px+fov_size/2,
                label = fov_positions$fov), color="black") +
  ggtitle(sample_name) + 
  global_map_theme(back.color = "white", back.border = "white", title.col = "black")
```

# Added variables

```{r}
fov_size <- 4256

# Importing global coordinates from spatialCoords slot since they are not available in colData
colData(spe)$CenterX_global_px <- spatialCoords(spe)[,1]
colData(spe)$CenterY_global_px <- spatialCoords(spe)[,2]

# Converting global coordinates to um 
spe$CenterX_global_um <- spe$CenterX_global_px*0.18
spe$CenterY_global_um <- spe$CenterY_global_px*0.18 

# Logged total counts
spe$logtot <- log10(spe$total)

# Converting area in um
spe$um_area <- spe$Area*0.0324

# Logged width/height ratio so there are no holes in the distribution in the middle and also negative values can be appreciated
spe$log2_AspectRatio <- log2(spe$AspectRatio)

# Keeping only target counts by subtracting counts referred to negative control probes
spe$target_counts <- spe$total - spe$subsets_NegProb_sum

# Keeping only detected features by subtracting negative control probes
spe$target_detected <- spe$detected - spe$subsets_NegProb_detected

# Proportion between counts referred only to target probes and detected features
spe$tot_det_ratio <- spe$target_counts/spe$target_detected

spe$ctrl_total_ratio <- spe$subsets_NegProb_sum/spe$total
```


# Global feature maps

To spot areas with anomalous feature values and spatial patterns on a global scale

```{r, out.width="150%", out.height="150%"}
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "total", point_size = 0.1) + ggtitle("Total counts per cell") + theme(aspect.ratio = 1)
```

```{r, out.width="150%", out.height="150%"}
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "logtot", point_size = 0.1) + ggtitle("Log-total counts per cell") + theme(aspect.ratio = 1)
```

```{r, out.width="150%", out.height="150%"}
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "detected", point_size = 0.1) + ggtitle("Detected features per cell") + theme(aspect.ratio = 1)
```

```{r, out.width="150%", out.height="150%"}
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "um_area", point_size = 0.1) + ggtitle("Area in μm per cell") + theme(aspect.ratio = 1)
```

```{r, out.width="150%", out.height="150%"}
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "log2_AspectRatio", point_size = 0.1) + ggtitle("Logged width/Height ratio per cell") + theme(aspect.ratio = 1)
```

```{r, out.width="150%", out.height="150%"}
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "tot_det_ratio", point_size = 0.1) + ggtitle("Target counts/detected features ratio per cell") + theme(aspect.ratio = 1)
```

```{r, out.width="150%", out.height="150%"}
plotColData(spe, "CenterY_global_um", "CenterX_global_um", order_by = "subsets_NegProb_sum", colour_by = "subsets_NegProb_sum", point_size = 0.1) +
  scale_color_gradient(low = "white", high = "red", name = "subsets_NegProb_sum") + ggtitle("Negative control probe counts per cell") +
  theme(aspect.ratio = 1,
        panel.background = element_rect(fill = "black",color = NA),
        plot.background = element_rect(fill = "black",color = NA),
        axis.title.x = element_text(color = "white"),
        axis.title.y = element_text(color = "white"),
        axis.ticks = element_line(color = "white"),
        axis.text.y = element_text(color = "white"),
        axis.text.x = element_text(color = "white"),
        axis.line = element_line(color = "white"),
        legend.title = element_text(color = "white"),
        legend.text = element_text(color = "white"),
        plot.title = element_text(color = "white"))
```

```{r}
threshold <- 0.20
spe$flagged_tot_20 <- as.factor(ifelse(colData(spe)$total < quantile(colData(spe)$total, probs = threshold), TRUE, FALSE))

threshold <- 0.10
spe$flagged_tot_10 <- as.factor(ifelse(colData(spe)$total < quantile(colData(spe)$total, probs = threshold), TRUE, FALSE))

threshold <- 0.05
spe$flagged_tot_5 <- as.factor(ifelse(colData(spe)$total < quantile(colData(spe)$total, probs = threshold), TRUE, FALSE))

threshold <- 0.10
spe$flagged_det_10 <- as.factor(ifelse(colData(spe)$detected < quantile(colData(spe)$detected, probs = threshold), TRUE, FALSE))

threshold <- 0.05
spe$flagged_det_5 <- as.factor(ifelse(colData(spe)$detected < quantile(colData(spe)$detected, probs = threshold), TRUE, FALSE))

threshold <- 0.10
spe$flagged_umarea_10 <- as.factor(ifelse(colData(spe)$um_area < quantile(colData(spe)$um_area, probs = threshold), TRUE, FALSE))

threshold <- 0.05
spe$flagged_umarea_5 <- as.factor(ifelse(colData(spe)$um_area < quantile(colData(spe)$um_area, probs = threshold), TRUE, FALSE))

threshold <- 0.90
spe$flagged_umarea_90 <- as.factor(ifelse(colData(spe)$um_area > quantile(colData(spe)$um_area, probs = threshold), TRUE, FALSE))

threshold <- 0.10
spe$flagged_aratio_10 <- as.factor(ifelse(colData(spe)$log2_AspectRatio < quantile(colData(spe)$log2_AspectRatio, probs = threshold), TRUE, FALSE))

threshold <- 0.05
spe$flagged_aratio_5 <- as.factor(ifelse(colData(spe)$log2_AspectRatio < quantile(colData(spe)$log2_AspectRatio, probs = threshold), TRUE, FALSE))

threshold <- 0.90
spe$flagged_aratio_90 <- as.factor(ifelse(colData(spe)$log2_AspectRatio > quantile(colData(spe)$log2_AspectRatio, probs = threshold), TRUE, FALSE))

threshold <- 0.95
spe$flagged_aratio_95 <- as.factor(ifelse(colData(spe)$log2_AspectRatio > quantile(colData(spe)$log2_AspectRatio, probs = threshold), TRUE, FALSE))

threshold <- 0.10
spe$flagged_tdratio_10 <- as.factor(ifelse(colData(spe)$tot_det_ratio < quantile(colData(spe)$tot_det_ratio, probs = threshold, na.rm = T), TRUE, FALSE))

threshold <- 0.05
spe$flagged_tdratio_5 <- as.factor(ifelse(colData(spe)$tot_det_ratio < quantile(colData(spe)$tot_det_ratio, probs = threshold, na.rm = T), TRUE, FALSE))

threshold <- 1
spe$flagged_tdratio_atomx <- as.factor(ifelse(colData(spe)$tot_det_ratio <= threshold, TRUE, FALSE))

threshold <- 0.50
spe$flagged_negprob_50 <- as.factor(ifelse(colData(spe)$subsets_NegProb_sum > quantile(colData(spe)$subsets_NegProb_sum, probs = threshold, na.rm = T), TRUE, FALSE))

threshold <- 0.75
spe$flagged_negprob_75 <- as.factor(ifelse(colData(spe)$subsets_NegProb_sum > quantile(colData(spe)$subsets_NegProb_sum, probs = threshold, na.rm = T), TRUE, FALSE))

spe$ctrl_total_ratio <- spe$subsets_NegProb_sum/spe$total
threshold <- 0.10
spe$flagged_negprob_10 <- as.factor(ifelse(colData(spe)$ctrl_total_ratio > quantile(colData(spe)$ctrl_total_ratio, probs = threshold, na.rm = T), TRUE, FALSE))
```


# Feature histograms

To check feature distribution before highlighting where extreme values are located in the global sample map

```{r}
variables <- c(variables, "logtot", "um_area", "log2_AspectRatio", "tot_det_ratio", "subsets_NegProb_sum")
for (var in 1:length(variables)){
  n <- grep(paste0("^", variables[var], "$"), colnames(spe@colData))
  histo_var <- names(colData(spe))[n]
  p <- ggplot(data = data.frame(colData(spe))) +
    geom_histogram(aes(x = .data[[histo_var]]), fill = "#69b3a2") +
    ggtitle(paste0(histo_var))
  print(p)
}  
```

# Global threshold map

Cells below the selected feature threshold are highlighted in black, all the others are grey

```{r, out.width="150%", out.height="150%"}
threshold <- 0.20
spe$flagged_tot_20 <- as.factor(ifelse(colData(spe)$total < quantile(colData(spe)$total, probs = threshold), TRUE, FALSE))
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "flagged_tot_20", 
            point_size = 0.1) + scale_color_manual(values = c("TRUE" = "black", "FALSE" = "lightgrey")) + ggtitle("Counts below 20% threshold (default by AtoMx guidelines)") + theme(legend.position="none") + labs(subtitle = paste0("Flagged cells = ", table(colData(spe)$flagged_tot_20)["TRUE"], " out of ", dim(spe)[2], ", ", round(table(colData(spe)$flagged_tot_20)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"))
```

```{r, out.width="150%", out.height="150%"}
threshold <- 0.10
spe$flagged_tot_10 <- as.factor(ifelse(colData(spe)$total < quantile(colData(spe)$total, probs = threshold), TRUE, FALSE))
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "flagged_tot_10", 
            point_size = 0.1) + scale_color_manual(values = c("TRUE" = "black","FALSE" = "lightgrey")) + ggtitle("Counts below 10% threshold")  + theme(legend.position="none") + labs(subtitle = paste0("Flagged cells = ", table(colData(spe)$flagged_tot_10)["TRUE"], " out of ", dim(spe)[2], ", ",  round(table(colData(spe)$flagged_tot_10)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"))
```

```{r, out.width="150%", out.height="150%"}
threshold <- 0.05
spe$flagged_tot_5 <- as.factor(ifelse(colData(spe)$total < quantile(colData(spe)$total, probs = threshold), TRUE, FALSE))
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "flagged_tot_5", 
            point_size = 0.1) + scale_color_manual(values = c("TRUE" = "black","FALSE" = "lightgrey")) + ggtitle("Counts below 5% threshold") + theme(legend.position="none") + labs(subtitle = paste0("Flagged cells = ", table(colData(spe)$flagged_tot_5)["TRUE"], " out of ", dim(spe)[2], ", ",  round(table(colData(spe)$flagged_tot_5)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells")) 
```



```{r, out.width="300%", out.height="200%"}
threshold <- 0.10
spe$flagged_det_10 <- as.factor(ifelse(colData(spe)$detected < quantile(colData(spe)$detected, probs = threshold), TRUE, FALSE))
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "flagged_det_10", 
            point_size = 0.1) + scale_color_manual(values = c("TRUE" = "black","FALSE" = "lightgrey")) + ggtitle("Detected features below 10% threshold") + theme(legend.position="none") + labs(subtitle = paste0("Flagged cells = ", table(colData(spe)$flagged_det_10)["TRUE"], " out of ", dim(spe)[2], ", ",  round(table(colData(spe)$flagged_det_10)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"))
```

```{r, out.width="150%", out.height="150%"}
threshold <- 0.05
spe$flagged_det_5 <- as.factor(ifelse(colData(spe)$detected < quantile(colData(spe)$detected, probs = threshold), TRUE, FALSE))
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "flagged_det_5", 
            point_size = 0.1) + scale_color_manual(values = c("TRUE" = "black","FALSE" = "lightgrey")) + ggtitle("Detected features below 5% threshold")  + theme(legend.position="none") + labs(subtitle = paste0("Flagged cells = ", table(colData(spe)$flagged_det_5)["TRUE"], " out of ", dim(spe)[2], ", ",  round(table(colData(spe)$flagged_det_5)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"))
```



```{r, out.width="150%", out.height="150%"}
threshold <- 0.10
spe$flagged_umarea_10 <- as.factor(ifelse(colData(spe)$um_area < quantile(colData(spe)$um_area, probs = threshold), TRUE, FALSE))
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "flagged_umarea_10", 
            point_size = 0.1) + scale_color_manual(values = c("TRUE" = "black","FALSE" = "lightgrey")) + ggtitle("Area in μm below 10%") + theme(legend.position="none") + labs(subtitle = paste0("Flagged cells = ", table(colData(spe)$flagged_umarea_10)["TRUE"], " out of ", dim(spe)[2], ", ",  round(table(colData(spe)$flagged_umarea_10)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"))
```

```{r, out.width="150%", out.height="150%"}
threshold <- 0.05
spe$flagged_umarea_5 <- as.factor(ifelse(colData(spe)$um_area < quantile(colData(spe)$um_area, probs = threshold), TRUE, FALSE))
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "flagged_umarea_5", 
            point_size = 0.1) + scale_color_manual(values = c("TRUE" = "black","FALSE" = "lightgrey")) + ggtitle("Area in μm below 5%")  + theme(legend.position="none") + labs(subtitle = paste0("Flagged cells = ", table(colData(spe)$flagged_umarea_5)["TRUE"], " out of ", dim(spe)[2], ", ",  round(table(colData(spe)$flagged_umarea_5)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"))
```


Cells above the selected threshold are highlighted in red, all the others are grey

```{r, out.width="150%", out.height="150%"}
threshold <- 0.90
spe$flagged_umarea_90 <- as.factor(ifelse(colData(spe)$um_area > quantile(colData(spe)$um_area, probs = threshold), TRUE, FALSE))
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "flagged_umarea_90", 
            point_size = 0.1) + scale_color_manual(values = c("TRUE" = "red","FALSE" = "lightgrey")) + ggtitle("Area in μm above 90% threshold") + theme(legend.position="none") + labs(subtitle = paste0("Flagged cells = ", table(colData(spe)$flagged_umarea_90)["TRUE"], " out of ", dim(spe)[2], ", ",  round(table(colData(spe)$flagged_umarea_90)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"))
```

```{r, out.width="150%", out.height="150%"}
threshold <- 0.95
spe$flagged_umarea_95 <- as.factor(ifelse(colData(spe)$um_area > quantile(colData(spe)$um_area, probs = threshold), TRUE, FALSE))
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "flagged_umarea_95", 
            point_size = 0.1) + scale_color_manual(values = c("TRUE" = "red","FALSE" = "lightgrey")) + ggtitle("Area in μm above 95% threshold")  + theme(legend.position="none") + labs(subtitle = paste0("Flagged cells = ", table(colData(spe)$flagged_umarea_95)["TRUE"], " out of ", dim(spe)[2], ", ",  round(table(colData(spe)$flagged_umarea_95)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"))
```



```{r, out.width="150%", out.height="150%"}
threshold <- 0.10
spe$flagged_aratio_10 <- as.factor(ifelse(colData(spe)$log2_AspectRatio < quantile(colData(spe)$log2_AspectRatio, probs = threshold), TRUE, FALSE))
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "flagged_aratio_10", 
            point_size = 0.1) + scale_color_manual(values = c("TRUE" = "black","FALSE" = "lightgrey")) + ggtitle("Width/height below 10% threshold") + theme(legend.position="none") + labs(subtitle = paste0("Flagged cells = ", table(colData(spe)$flagged_aratio_10)["TRUE"], " out of ", dim(spe)[2], ", ",  round(table(colData(spe)$flagged_aratio_10)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"))
```

```{r, out.width="150%", out.height="150%"}
threshold <- 0.05
spe$flagged_aratio_5 <- as.factor(ifelse(colData(spe)$log2_AspectRatio < quantile(colData(spe)$log2_AspectRatio, probs = threshold), TRUE, FALSE))
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "flagged_aratio_5", 
            point_size = 0.1) + scale_color_manual(values = c("TRUE" = "black","FALSE" = "lightgrey")) + ggtitle("Width/height below 5% threshold")  + theme(legend.position="none") + labs(subtitle = paste0("Flagged cells = ", table(colData(spe)$flagged_aratio_5)["TRUE"], " out of ", dim(spe)[2], ", ",  round(table(colData(spe)$flagged_aratio_5)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"))
```



```{r, out.width="150%", out.height="150%"}
threshold <- 0.90
spe$flagged_aratio_90 <- as.factor(ifelse(colData(spe)$log2_AspectRatio > quantile(colData(spe)$log2_AspectRatio, probs = threshold), TRUE, FALSE))
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "flagged_aratio_90", 
            point_size = 0.1) + scale_color_manual(values = c("TRUE" = "red","FALSE" = "lightgrey")) + ggtitle("Width/height above 90% threshold") + theme(legend.position="none") + labs(subtitle = paste0("Flagged cells = ", table(colData(spe)$flagged_aratio_90)["TRUE"], " out of ", dim(spe)[2], ", ",  round(table(colData(spe)$flagged_aratio_90)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"))
```

```{r, out.width="150%", out.height="150%"}
threshold <- 0.95
spe$flagged_aratio_95 <- as.factor(ifelse(colData(spe)$log2_AspectRatio > quantile(colData(spe)$log2_AspectRatio, probs = threshold), TRUE, FALSE))
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "flagged_aratio_95", 
            point_size = 0.1) + scale_color_manual(values = c("TRUE" = "red","FALSE" = "lightgrey")) + ggtitle("Width/height above 95% threshold")  + theme(legend.position="none") + labs(subtitle = paste0("Flagged cells = ", table(colData(spe)$flagged_aratio_95)["TRUE"], " out of ", dim(spe)[2], ", ",  round(table(colData(spe)$flagged_aratio_95)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"))
```



```{r, out.width="150%", out.height="150%"}
threshold <- 0.10
spe$flagged_tdratio_10 <- as.factor(ifelse(colData(spe)$tot_det_ratio < quantile(colData(spe)$tot_det_ratio, probs = threshold, na.rm = T), TRUE, FALSE))
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "flagged_tdratio_10", 
            point_size = 0.1) + scale_color_manual(values = c("TRUE" = "black","FALSE" = "lightgrey")) + ggtitle(paste0("Total counts/detected features below 10% threshold, ratio = ", round(quantile(colData(spe)$tot_det_ratio, probs = threshold, na.rm = T), 2))) + theme(legend.position="none") + labs(subtitle = paste0("Flagged cells = ", table(colData(spe)$flagged_tdratio_10)["TRUE"], " out of ", dim(spe)[2], ", ",  round(table(colData(spe)$flagged_tdratio_10)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"))
```

```{r, out.width="150%", out.height="150%"}
threshold <- 0.05
spe$flagged_tdratio_5 <- as.factor(ifelse(colData(spe)$tot_det_ratio < quantile(colData(spe)$tot_det_ratio, probs = threshold, na.rm = T), TRUE, FALSE))
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "flagged_tdratio_5", 
            point_size = 0.1) + scale_color_manual(values = c("TRUE" = "black","FALSE" = "lightgrey")) + ggtitle(paste0("Total counts/detected features below 5% threshold, ratio = ", round(quantile(colData(spe)$tot_det_ratio, probs = threshold, na.rm = T), 2)))  + theme(legend.position="none") + labs(subtitle = paste0("Flagged cells = ", table(colData(spe)$flagged_tdratio_5)["TRUE"], " out of ", dim(spe)[2], ", ",  round(table(colData(spe)$flagged_tdratio_5)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"))
```

ATOMX GUIDELINES FOR CELL QC

Count distribution 
flag cells where (total counts) / (number of detected genes) ≤1

```{r, out.width="150%", out.height="150%"}
threshold <- 1
spe$flagged_tdratio_atomx <- as.factor(ifelse(colData(spe)$tot_det_ratio <= threshold, TRUE, FALSE))
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "flagged_tdratio_atomx", 
            point_size = 0.1) + scale_color_manual(values = c("TRUE" = "black","FALSE" = "lightgrey")) + ggtitle("Total counts/detected feature ratio below 1 (default by AtoMx guidelines), ratio = 1") + theme(legend.position="none") + labs(subtitle = paste0("Flagged cells = ", table(colData(spe)$flagged_tdratio_atomx)["TRUE"], " out of ", dim(spe)[2], ", ",  round(table(colData(spe)$flagged_tdratio_atomx)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells")) 
```


```{r, out.width="150%", out.height="150%"}
threshold <- 0.50
spe$flagged_negprob_50 <- as.factor(ifelse(colData(spe)$subsets_NegProb_sum > quantile(colData(spe)$subsets_NegProb_sum, probs = threshold, na.rm = T), TRUE, FALSE))
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "flagged_negprob_50", 
            point_size = 0.1) + scale_color_manual(values = c("TRUE" = "red","FALSE" = "lightgrey")) + ggtitle("Negative probe counts above 50% threshold") + theme(legend.position="none") + labs(subtitle = paste0("Flagged cells = ", table(colData(spe)$flagged_negprob_50)["TRUE"], " out of ", dim(spe)[2], ", ",  round(table(colData(spe)$flagged_negprob_50)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"))
```

```{r, out.width="150%", out.height="150%"}
threshold <- 0.75
spe$flagged_negprob_75 <- as.factor(ifelse(colData(spe)$subsets_NegProb_sum > quantile(colData(spe)$subsets_NegProb_sum, probs = threshold, na.rm = T), TRUE, FALSE))
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "flagged_negprob_75", 
            point_size = 0.1) + scale_color_manual(values = c("TRUE" = "red","FALSE" = "lightgrey")) + ggtitle("Negative probe counts above 75% threshold")  + theme(legend.position="none") + labs(subtitle = paste0("Flagged cells = ", table(colData(spe)$flagged_negprob_75)["TRUE"], " out of ", dim(spe)[2], ", ",  round(table(colData(spe)$flagged_negprob_75)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"))
```

ATOMX GUIDELINES FOR CELL QC

Proportion of negative counts
flag cells where > 10% (0.1; default) of the counts are negative probes

```{r, out.width="150%", out.height="150%"}
spe$ctrl_total_ratio <- spe$subsets_NegProb_sum/spe$total
threshold <- 0.10
spe$flagged_negprob_10 <- as.factor(ifelse(colData(spe)$ctrl_total_ratio > quantile(colData(spe)$ctrl_total_ratio, probs = threshold, na.rm = T), TRUE, FALSE))
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "flagged_negprob_10", 
            point_size = 0.1) + scale_color_manual(values = c("TRUE" = "red","FALSE" = "lightgrey")) + ggtitle("Negative control probe/total counts ratio above 10% threshold (default by AtoMx guidelines)") + theme(legend.position="none") + labs(subtitle = paste0("Flagged cells = ", table(colData(spe)$flagged_negprob_10)["TRUE"], " out of ", dim(spe)[2], ", ", round(table(colData(spe)$flagged_negprob_10)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells")) 
```


# Feature heatmaps by FOV 

To compare at a glance FOV mean/median feature values and spot anomalous FOVs

```{r}
meta_df <- data.frame(colData(spe))
meta_df <- group_by(meta_df, fov) |> summarize(n_cells = n(), mean_total = mean(total), mean_detected = mean(detected), mean_umarea = mean(um_area), mean_log2_AspectRatio = mean(log2_AspectRatio), mean_tot_det_ratio = mean(tot_det_ratio), mean_subsets_NegProb_sum = mean(subsets_NegProb_sum), median_total = median(total), median_detected = median(detected), median_umarea = median(um_area),median_log2_AspectRatio = median(log2_AspectRatio), median_tot_det_ratio = median(tot_det_ratio), median_subsets_NegProb_sum = median(subsets_NegProb_sum))

fov_positions <- left_join(fov_positions, meta_df, by = join_by(fov))

fov_size <- 4256
poly_df <- transmute(fov_positions, x_1 = x_global_px, y_1 = y_global_px, 
                     x_2 = x_global_px + fov_size, y_2 = y_global_px, 
                     x_3 = x_global_px + fov_size, y_3 = y_global_px + fov_size, 
                     x_4 = x_global_px, y_4 = y_global_px + fov_size)

str(poly_df)

ls_list <- list()
for (k in 1:dim(fov_positions)[1]){
  ls_list[[k]] <- st_linestring(matrix(as.numeric(poly_df[k]), ncol = 2, byrow = T))
}

sf_data <- st_as_sf(fov_positions, coords = c("x_global_px", "y_global_px"), geometry = st_sfc(ls_list))

sf_data <- st_cast(sf_data, "POLYGON")

# plot(sf_data)
names(sf_data)
sf_data$fov <- as.factor(sf_data$fov)
```

```{r, out.width="150%", out.height="150%"}
ggplot() +
  geom_sf(data = sf_data, aes(fill = n_cells)) + 
  scale_fill_gradient(low = "purple4", high = "#FF8247") +
  theme(axis.title.x = element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(), 
        axis.title.y = element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        panel.grid = element_blank()) +
  geom_text(data = fov_positions, aes(x = fov_positions$x_global_px+fov_size/2,
                y = fov_positions$y_global_px+fov_size/2,
                label = fov_positions$n_cells), color="white", size = 1.8) +
    ggtitle(paste0("Number of cells by fov"))
```

```{r, out.width="150%", out.height="150%"}
heatmap_variables <- names(sf_data)[-c(1, 2, length(names(sf_data)), length(names(sf_data))-1)]
for (var in 1:length(heatmap_variables)){
  heat_var <- heatmap_variables[var]
  p <- ggplot() +
  geom_sf(data = sf_data, aes(fill = .data[[heat_var]])) + 
  scale_fill_viridis_c() +
  theme(axis.title.x = element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(), 
        axis.title.y = element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        panel.grid = element_blank()) +
  geom_text(data = fov_positions, aes(x = fov_positions$x_global_px+fov_size/2,
                y = fov_positions$y_global_px+fov_size/2,
                label = fov_positions$fov), color="white", size = 2) +
    ggtitle(paste0(heat_var, " by fov"))
  print(p)
}  
```

ATOMX GUIDELINES FOR FOV QC

The Mean method (default) flags FOVs where the total count per cell average is below a threshold (default 100)

```{r, out.width="150%", out.height="150%"}
color <- ifelse(fov_positions$mean_total <= 100, "red", "white")
fov_positions <- cbind(fov_positions, color)
ggplot() +
  geom_sf(data = sf_data, aes(fill = mean_total)) + 
  scale_fill_viridis_c() +
  theme(axis.title.x = element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(), 
        axis.title.y = element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        panel.grid = element_blank()) +
  geom_text(data = fov_positions, aes(x = fov_positions$x_global_px+fov_size/2,
                y = fov_positions$y_global_px+fov_size/2,
                label = floor(fov_positions$mean_total)), color = color, size = 2) +
    ggtitle(paste0("Mean_total by fov"))
```

ATOMX GUIDELINES FOR FOV QC

The QC quantile must be greater than the negative probe median in order to pass QC. Default is 0, so it must be ≥ 0

```{r, out.width="150%", out.height="150%"}
threshold <- 0.90
quant_df <-  data.frame(colData(spe))
quant_df <- mutate(quant_df, target_counts = total - subsets_NegProb_sum)
quant_df <- group_by(quant_df, fov) |> summarize(quantile_90 = quantile(target_counts, probs = threshold))
fov_positions <- left_join(fov_positions, quant_df, by = join_by(fov))

color <- ifelse(fov_positions$quantile_90 < fov_positions$median_subsets_NegProb_sum, "red", "white")
sf_data$quantile_90 <- fov_positions$quantile_90

ggplot() +
  geom_sf(data = sf_data, aes(fill = quantile_90)) + 
  scale_fill_viridis_c() +
  theme(axis.title.x = element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(), 
        axis.title.y = element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        panel.grid = element_blank()) +
  geom_text(data = fov_positions, aes(x = fov_positions$x_global_px+fov_size/2,
                y = fov_positions$y_global_px+fov_size/2,
                label = floor(fov_positions$quantile_90)), color = color, size = 1.8) +
    ggtitle(paste0("FOV QC based on 90th quantile"))
```


# Feature boxplot by FOV

To compare the distribution of each feature of interest among FOVs

```{r, warnings = FALSE}
variables <- c("total", "detected", "logtot", "um_area", "log2_AspectRatio", "tot_det_ratio", "subsets_NegProb_sum")

for (var in 1:length(variables)){
  n <- grep(paste0("^", variables[var], "$"), colnames(spe@colData))
  boxplot(spe@colData@listData[[n]] ~ spe$fov, ylab = colnames(spe@colData)[n], cex = 0.2)
  title(paste0(colnames(spe@colData)[n], "~fov"))
}
```

# Cell counts barplot by FOV

To view how many cells lie inside each FOV

```{r}
fov_df <- data.frame("n_cells" = as.integer(table(spe@colData@listData$fov)), "fov" = as.factor(names(table(spe@colData@listData$fov))))
levels(fov_df$fov) <- order(levels(fov_df$fov), seq(1:length(fov_df$fov)))

ggplot(data = fov_df, aes(x = fov, y = n_cells)) +
  geom_col(fill = "#B0C4DE") + theme(legend.position ="none") +
  ggtitle("Number of cells per fov") 
```

# Final report

```{r echo=T}
data <- data.frame(colData(spe))

df <- format(t(data.frame("Total_cell_number" = dim(spe)[2], "Total_counts" = sum(spe$total), "Total_features" = sum(spe$detected), "Total_ctrl_probe_counts" = sum(spe$subsets_NegProb_sum), "Median_counts_per_cell" = median(spe$total), "Median_features_per_cell" = median(spe$detected),  "Median_ctrl_probe_counts" = median(spe$subsets_NegProb_sum), "Median_um_area" = median(spe$um_area), "Median_log2_aspect_ratio" = median(spe$log2_AspectRatio), "Median_total_counts/features_ratio" = median(spe$tot_det_ratio, na.rm = T),  "Median_ctrl_probe_counts/total_counts_ratio" = median(spe$ctrl_total_ratio, na.rm = T))), scientific = FALSE)

rownames(df) <- c("Total cell number", "Total counts", "Total features", "Total ctrl probe counts", "Median counts per cell", "Median features per cell", "Median ctrl probe counts", "Median um area", "Median log2 aspect ratio", "Median total counts/features ratio", "Median ctrl probe counts/total counts ratio")

df[,1] <- round(as.numeric(df[,1]), 2)

kable(df, format = "html", row.names = T, col.names = NULL, align = "c", caption = "<center>Global summary metrics</center>",
      padding = 1L)
```

#
#

```{r}
flag_variables <- c("flagged_tot_20", 
                    "flagged_tot_10",
                    "flagged_tot_5",
                    "flagged_det_10",
                    "flagged_det_5",
                    "flagged_umarea_10",
                    "flagged_umarea_5",
                    "flagged_umarea_90",
                    "flagged_umarea_95",
                    "flagged_aratio_10",
                    "flagged_aratio_5",
                    "flagged_aratio_90",
                    "flagged_aratio_95",
                    "flagged_tdratio_10",
                    "flagged_tdratio_5",
                    "flagged_tdratio_atomx",
                    "flagged_negprob_50",
                    "flagged_negprob_75",
                    "ctrl_total_ratio")

flag_data <- data.frame(Unflagged_cells = as.integer(), Flagged_cells = as.integer(), Percentage_flagged = as.numeric())
for(i in 1:length(flag_variables)){
  var <- grep(paste0("^", flag_variables[i],"$"), names(colData(spe)))
  list_i <- c(sum(colData(spe)[,var]==FALSE, na.rm = T), sum(colData(spe)[,var]==TRUE, na.rm = T), round(sum(colData(spe)[,var]==TRUE, na.rm = T)/dim(spe)[2]*100, 2))
  flag_data[i, ] <- list_i
}

row.names(flag_data) <- c("Total counts < 20%", " Total counts < 10%", "Total counts < 5%", "Features < 10%", "Features < 5%", "Area < 10%", "Area < 5%", 
                         "Area > 90%","Area > 95%", "H/w < 10%", "H/w < 5%", "H/w > 90%", "H/w > 95%", "Counts/features < 10%", "Counts/features < 5%",
                         "Counts/features < 1", "Ctrl counts > 50%", "Ctrl counts > 75%", "Ctrl/total > 10%")

kable(flag_data, format = "html", padding = 1L, align = "c", col.names = c("Unflagged cells", "Flagged cells", "% flagged cells"), caption = paste0("<center>Flagged vs unflagged cells, total cells: " , dim(spe)[2],"</center>"))
```

#
#

ATOMX GUIDELINES FOR TARGET LEVEL QC

A value of 0.5 (default) will flag probes with lower total counts than the median (50th percentile) of thenegative control probes' counts

```{r}
ctrl_counts <- counts(spe)[grep("^NegPrb", rownames(spe)), ]
target_counts <- counts(spe)[1:(dim(counts(spe))[1] - length(grep("^NegPrb", rownames(spe)))), ]
temp <- rowSums(target_counts) > median(rowSums(ctrl_counts))
count_data <- rowSums(counts(spe))
color <- ifelse(count_data < median(rowSums(ctrl_counts)), "red", "black") 
color <- ifelse(match(names(color), rownames(target_counts)), color, "grey")
color <- replace_na(color, "grey") 
count_df <- data_frame(count_data, color)
count_df$log_count <- log10(count_df$count_data)
plot(count_df$log_count, col = count_df$color)
```

# Flag score development

Starting from feature scatterplots to see the relationship between possible candidate features

Total counts ~ logged width/height ratio

```{r, out.width="150%", out.height="150%"}
cell_df <- data.frame(colData(spe))

p1 <- ggplot(data = cell_df) +
  geom_pointdensity(aes(x = log2_AspectRatio, y = total)) +
  scale_color_viridis_c() +
  theme(legend.position = "none",
        plot.margin = margin())

p2 <- ggdensity(data = cell_df, x = "total", color = "#ccccff", 
                fill = "#ccccff", alpha = 0.5) + theme_classic() + 
  theme(legend.position = "none", axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line.y = element_blank(),
    axis.line.x = element_blank(),
    plot.margin = margin()) + coord_flip()

p3 <- ggdensity(data = cell_df, x = "log2_AspectRatio", color = "#ccccff", 
                fill = "#ccccff", alpha = 0.5) + theme_classic() + 
  theme(legend.position = "none", axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.line.x = element_blank(),
    axis.line.y = element_blank(),
    plot.margin = margin()) 

plot_grid(p3, NULL, p1, p2, ncol = 2, align = "hv",
          rel_widths = c(2, 1), rel_heights = c(1, 2))
```

Detected features ~ logged width/height ratio

```{r, out.width="150%", out.height="150%"}
p1 <- ggplot(data = cell_df) +
  geom_pointdensity(aes(x = log2_AspectRatio, y = detected)) +
  scale_color_viridis_c() +
  theme(legend.position = "none",
        plot.margin = margin())

p2 <- ggdensity(data = cell_df, x = "detected", color = "#ccccff", 
                fill = "#ccccff", alpha = 0.5) + theme_classic() + 
  theme(legend.position = "none", axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line.y = element_blank(),
    axis.line.x = element_blank(),
    plot.margin = margin()) + coord_flip()

p3 <- ggdensity(data = cell_df, x = "log2_AspectRatio", color = "#ccccff", 
                fill = "#ccccff", alpha = 0.5) + theme_classic() + 
  theme(legend.position = "none", axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.line.x = element_blank(),
    axis.line.y = element_blank(),
    plot.margin = margin()) 

plot_grid(p3, NULL, p1, p2, ncol = 2, align = "hv",
          rel_widths = c(2, 1), rel_heights = c(1, 2))
```

Total counts ~ area in um

```{r, out.width="150%", out.height="150%"}
p1 <- ggplot(data = cell_df) +
  geom_pointdensity(aes(x = um_area, y = total)) +
  scale_color_viridis_c() +
  theme(legend.position = "none",
        plot.margin = margin())

p2 <- ggdensity(data = cell_df, x = "total", color = "#ccccff", 
                fill = "#ccccff", alpha = 0.5) + theme_classic() + 
  theme(legend.position = "none", axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line.y = element_blank(),
    axis.line.x = element_blank(),
    plot.margin = margin()) + coord_flip()

p3 <- ggdensity(data = cell_df, x = "um_area", color = "#ccccff", 
                fill = "#ccccff", alpha = 0.5) + theme_classic() + 
  theme(legend.position = "none", axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.line.x = element_blank(),
    axis.line.y = element_blank(),
    plot.margin = margin()) 

plot_grid(p3, NULL, p1, p2, ncol = 2, align = "hv",
          rel_widths = c(2,1), rel_heights = c(1, 2))
```

Total counts ~ nucleus area

```{r, out.width="150%", out.height="150%"}
p1 <- ggplot(data = cell_df) +
  geom_pointdensity(aes(x = Area, y = total)) +
  scale_color_viridis_c() +
  theme(legend.position = "none",
        plot.margin = margin())

p2 <- ggdensity(data = cell_df, x = "total", color = "#ccccff", 
                fill = "#ccccff", alpha = 0.5) + theme_classic() + 
  theme(legend.position = "none", axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line.y = element_blank(),
    axis.line.x = element_blank(),
    plot.margin = margin()) + coord_flip()

p3 <- ggdensity(data = cell_df, x = "Area", color = "#ccccff", 
                fill = "#ccccff", alpha = 0.5) + theme_classic() + 
  theme(legend.position = "none", axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.line.x = element_blank(),
    axis.line.y = element_blank(),
    plot.margin = margin()) 

plot_grid(p3, NULL, p1, p2, ncol = 2, align = "hv",
          rel_widths = c(2,1), rel_heights = c(1, 2))
```

Logged width/height ratio ~ negative control probe counts

```{r, out.width="150%", out.height="150%"}
p1 <- ggplot(data = cell_df) +
  geom_pointdensity(aes(x = subsets_NegProb_sum, y = log2_AspectRatio)) +
  scale_color_viridis_c() +
  theme(legend.position = "none",
        plot.margin = margin())

p2 <- ggdensity(data = cell_df, x = "log2_AspectRatio", color = "#ccccff", 
                fill = "#ccccff", alpha = 0.5) + theme_classic() + 
  theme(legend.position = "none", axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line.y = element_blank(),
    axis.line.x = element_blank(),
    plot.margin = margin()) + coord_flip()

p3 <- ggdensity(data = cell_df, x = "subsets_NegProb_sum", color = "#ccccff", 
                fill = "#ccccff", alpha = 0.5) + theme_classic() + 
  theme(legend.position = "none", axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.line.x = element_blank(),
    axis.line.y = element_blank(),
    plot.margin = margin()) 

plot_grid(p3, NULL, p1, p2, ncol = 2, align = "hv",
          rel_widths = c(2,1), rel_heights = c(1, 2))
```

Logged width/height ratio ~ local coordinates on X axis in px

```{r, out.width="150%", out.height="150%"}
p1 <- ggplot(data = cell_df) +
  geom_pointdensity(aes(x = CenterX_local_px, y = log2_AspectRatio)) +
  scale_color_viridis_c() +
  theme(legend.position = "none",
        plot.margin = margin())

p2 <- ggdensity(data = cell_df, x = "log2_AspectRatio", color = "#ccccff", 
                fill = "#ccccff", alpha = 0.5) + theme_classic() + 
  theme(legend.position = "none", axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line.y = element_blank(),
    axis.line.x = element_blank(),
    plot.margin = margin()) + coord_flip()

p3 <- ggdensity(data = cell_df, x = "CenterX_local_px", color = "#ccccff", 
                fill = "#ccccff", alpha = 0.5) + theme_classic() + 
  theme(legend.position = "none", axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.line.x = element_blank(),
    axis.line.y = element_blank(),
    plot.margin = margin()) 

plot_grid(p3, NULL, p1, p2, ncol = 2, align = "hv",
          rel_widths = c(2,1), rel_heights = c(1, 2))
```

Logged width/height ratio ~ local coordinates on X axis in px

```{r, out.width="150%", out.height="150%"}
p1 <- ggplot(data = cell_df) +
  geom_pointdensity(aes(x = CenterY_local_px, y = log2_AspectRatio)) +
  scale_color_viridis_c() +
  theme(legend.position = "none",
        plot.margin = margin())

p2 <- ggdensity(data = cell_df, x = "log2_AspectRatio", color = "#ccccff", 
                fill = "#ccccff", alpha = 0.5) + theme_classic() + 
  theme(legend.position = "none", axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line.y = element_blank(),
    axis.line.x = element_blank(),
    plot.margin = margin()) + coord_flip()

p3 <- ggdensity(data = cell_df, x = "CenterY_local_px", color = "#ccccff", 
                fill = "#ccccff", alpha = 0.5) + theme_classic() + 
  theme(legend.position = "none", axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.line.x = element_blank(),
    axis.line.y = element_blank(),
    plot.margin = margin()) 

plot_grid(p3, NULL, p1, p2, ncol = 2, align = "hv",
          rel_widths = c(2,1), rel_heights = c(1, 2))
```

# Flag score formula

```{r}
spe$flag_score <- 1/(1 + exp(-0.3*spe$target_counts/spe$um_area + 0.8*abs(spe$log2_AspectRatio)))
summary(spe$flag_score)
plot(spe$flag_score)
```

```{r}
hist(spe$flag_score)
```

## Cell boundary file loading

Loading polygon file and setting the correct format as sf object

```{r}
spat_obj <- fread("/NAS06/work/Spatial_omics/CosMx/Mouse_brain/coronal_hippocampus_and_cortex/Run5642_S3_Quarter-polygons.csv")
str(spat_obj)
```

```{r}
sf_global <- st_as_sf(spat_obj, coords = c("x_global_px", "y_global_px"))

polygons_whole <- sf_global %>%
  group_by(fov, cellID) %>%
  summarize(geometry = st_combine(geometry))#st_combine = combine several feature geometries into one, without unioning or resolving internal boundaries

polygons_whole <- st_cast(polygons_whole, "POLYGON")
```


```{r, out.width="150%", out.height="150%"}
polygons_whole$total <- spe$total
polygons_whole$target_counts <- spe$target_counts
polygons_whole$um_area <- spe$um_area
polygons_whole$log2_AspectRatio <- spe$log2_AspectRatio
polygons_whole$flag_score <- spe$flag_score

polygons_whole$Mean.PanCK <- spe$Mean.PanCK
polygons_whole$Mean.CD68 <- spe$Mean.CD68
polygons_whole$Mean.MembraneStain_B2M <- spe$Mean.MembraneStain_B2M
polygons_whole$Mean.CD45 <- spe$Mean.CD45
polygons_whole$Mean.DAPI <- spe$Mean.DAPI
polygons_whole$subsets_NegProb_sum <- spe$subsets_NegProb_sum

tm_shape(polygons_whole) + 
  tm_fill(col= "flag_score", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE)

```

Local threshold map of flag score, target counts, cell_area and logged width/height ratio, 20th percentile

```{r, out.width="300%", out.height="600%"}
metric_color <- ifelse(polygons_whole$flag_score < quantile(spe$flag_score, probs = 0.2), "red", "grey80")
target_color <- ifelse(polygons_whole$target_counts < quantile(spe$target_counts, probs = 0.2), "red", "grey80")
area_color <- ifelse(polygons_whole$um_area < quantile(spe$um_area, probs = 0.2), "red", "grey80")
count_area_ratio_color <- ifelse((polygons_whole$target_counts/polygons_whole$um_area) < quantile((spe$target_counts/spe$um_area), probs = 0.2), "red", "grey80")
log_hw_color_lo <- ifelse(polygons_whole$log2_AspectRatio < quantile(spe$log2_AspectRatio, probs = 0.2), "red", "grey80")
log_hw_color_hi <- ifelse(polygons_whole$log2_AspectRatio > quantile(spe$log2_AspectRatio, probs = 0.85), "red", "grey80")

polygons_whole <- cbind(polygons_whole, metric_color = metric_color, target_color = target_color, area_color = area_color, count_area_ratio_color = count_area_ratio_color, log_hw_color_lo = log_hw_color_lo, log_hw_color_hi = log_hw_color_hi)

p1 <- tm_shape(polygons_whole) + 
  tm_fill(col= "metric_color", alpha = 0.5) +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Flag score",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p2 <- tm_shape(polygons_whole) + 
  tm_fill(col= "target_color", alpha = 0.5) +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Target counts",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p3 <- tm_shape(polygons_whole) + 
  tm_fill(col= "area_color", alpha = 0.5) +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "cell area",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p4 <- tm_shape(polygons_whole) + 
  tm_fill(col= "count_area_ratio_color", alpha = 0.5) +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "counts/um area",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p5 <- tm_shape(polygons_whole) + 
  tm_fill(col= "log_hw_color_lo", alpha = 0.5) +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Logged AspectRatio below",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p6 <- tm_shape(polygons_whole) + 
  tm_fill(col= "log_hw_color_hi", alpha = 0.5) +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Logged AspectRatio above",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

tmap_arrange(p1, p2, p3, p4, p5, p6, ncol = 2)
```

```{r, out.width="150%", out.height="150%"}
plot_list <- list(p1, p2, p3, p4, p5, p6)
for(i in 1:length(plot_list)){
  print(plot_list[[i]])
}
```


```{r, out.width="150%", out.height="150%"}

flag_value <- c(round(quantile(polygons_whole$flag_score, probs = 0.2), 2),
                round(quantile(polygons_whole$target_counts, probs = 0.2), 2),
                round(quantile(polygons_whole$um_area, probs = 0.2), 2),
                round(quantile((polygons_whole$target_counts/polygons_whole$um_area), probs = 0.2), 2),
                round(quantile(polygons_whole$log2_AspectRatio, probs = 0.2), 2),
                round(quantile(polygons_whole$log2_AspectRatio, probs = 0.85), 2))

flagged_cells <- c(sum(polygons_whole$metric_color=="red", na.rm = T),
                   sum(polygons_whole$target_color=="red", na.rm = T),
                   sum(polygons_whole$area_color=="red", na.rm = T),
                   sum(polygons_whole$count_area_ratio_color=="red", na.rm = T),
                   sum(polygons_whole$log_hw_color_lo=="red", na.rm = T),
                   sum(polygons_whole$log_hw_color_hi=="red", na.rm = T))

unflagged_cells <- c(sum(polygons_whole$metric_color=="white", na.rm = T),
                     sum(polygons_whole$target_color=="white", na.rm = T),
                     sum(polygons_whole$area_color=="white", na.rm = T),
                     sum(polygons_whole$count_area_ratio_color=="white", na.rm = T),
                     sum(polygons_whole$log_hw_color_lo=="white", na.rm = T),
                     sum(polygons_whole$log_hw_color_hi=="white", na.rm = T))

percent_flagged <- round(flagged_cells/dim(polygons_whole)[1]*100, 2)

poly <- data.frame(threshold = c(0.2, 0.2, 0.2, 0.2, 0.2, 0.85), flag_value = flag_value, unflagged_cells = unflagged_cells, flagged_cells = flagged_cells, percent_flagged = percent_flagged, row.names = c("Flag score", "Target counts", "Area in um", "Target counts/um area", "Logged aspect ratio low", "Logged aspect ratio high"))

kable(poly, format = "html", col.names = c("Threshold", "Flag value", "Unflagged cells", "Flagged cells", "% flagged cells"), align = "c", caption = paste0("<center>Flagged cells by different metrics, total cells in whole sample: ", dim(polygons_whole)[1], ", flagged cells in whole sample: ", sum(spe$flag_score < quantile(spe$flag_score, probs = 0.2, na.rm = T)), "</center>"), padding = 1L)
```

#
#

Creating sf using local coordinates for geometry

```{r}
sf_local <- st_as_sf(spat_obj, coords = c("x_local_px", "y_local_px"))

polygons <- sf_local %>%
  group_by(fov, cellID) %>%
  summarize(geometry = st_combine(geometry))#st_combine = combine several feature geometries into one, without unioning or resolving internal boundaries

polygons <- st_cast(polygons, "POLYGON")

polygons$total <- spe$total
polygons$target_counts <- spe$target_counts
polygons$um_area <- spe$um_area
polygons$log2_AspectRatio <- spe$log2_AspectRatio
polygons$flag_score <- spe$flag_score

polygons$Mean.PanCK <- spe$Mean.PanCK
polygons$Mean.CD68 <- spe$Mean.CD68
polygons$Mean.MembraneStain_B2M <- spe$Mean.MembraneStain_B2M
polygons$Mean.CD45 <- spe$Mean.CD45
polygons$Mean.DAPI <- spe$Mean.DAPI
polygons$subsets_NegProb_sum <- spe$subsets_NegProb_sum
```

Local feature map of flag score, fov 21

```{r, out.width="150%", out.height="150%"}
polygons21 <- filter(polygons, fov==21) 

tm_shape(polygons21) + 
  tm_fill(col= "flag_score", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE)
```

Local threshold map of flag score, target counts, um_area and logged width/height ratio, 20th percentile

```{r, out.width="300%", out.height="600%"}
metric_color <- ifelse(polygons21$flag_score < quantile(spe$flag_score, probs = 0.2), "red", "white")
target_color <- ifelse(polygons21$target_counts < quantile(spe$target_counts, probs = 0.2), "red", "white")
area_color <- ifelse(polygons21$um_area < quantile(spe$um_area, probs = 0.2), "red", "white")
count_area_ratio_color <- ifelse((polygons21$target_counts/polygons21$um_area) < quantile((spe$target_counts/spe$um_area), probs = 0.2), "red", "white")
log_hw_color_lo <- ifelse(polygons21$log2_AspectRatio < quantile(spe$log2_AspectRatio, probs = 0.2), "red", "white")
log_hw_color_hi <- ifelse(polygons21$log2_AspectRatio > quantile(spe$log2_AspectRatio, probs = 0.85), "red", "white")

polygons21 <- cbind(polygons21, metric_color = metric_color, target_color = target_color, area_color = area_color, count_area_ratio_color = count_area_ratio_color, log_hw_color_lo = log_hw_color_lo, log_hw_color_hi = log_hw_color_hi)

p1 <- tm_shape(polygons21) + 
  tm_fill(col= "metric_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Flag score",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))


p2 <- tm_shape(polygons21) + 
  tm_fill(col= "target_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Target counts",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p3 <- tm_shape(polygons21) + 
  tm_fill(col= "area_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "um area",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p4 <- tm_shape(polygons21) + 
  tm_fill(col= "count_area_ratio_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "counts/um area",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p5 <- tm_shape(polygons21) + 
  tm_fill(col= "log_hw_color_lo", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Logged AspectRatio below",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p6 <- tm_shape(polygons21) + 
  tm_fill(col= "log_hw_color_hi", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Logged AspectRatio above",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

tmap_arrange(p1, p2, p3, p4, p5, p6, ncol = 2)
```

```{r, out.width="150%", out.height="150%"}
plot_list <- list(p1, p2, p3, p4, p5, p6)
for(i in 1:length(plot_list)){
  print(plot_list[[i]])
}
```


```{r, out.width="150%", out.height="150%"}

flag_value <- c(round(quantile(polygons21$flag_score, probs = 0.2), 2),
                round(quantile(polygons21$target_counts, probs = 0.2), 2),
                round(quantile(polygons21$um_area, probs = 0.2), 2),
                round(quantile((polygons21$target_counts/polygons21$um_area), probs = 0.2), 2),
                round(quantile(polygons21$log2_AspectRatio, probs = 0.2), 2),
                round(quantile(polygons21$log2_AspectRatio, probs = 0.85), 2))

flagged_cells <- c(sum(polygons21$metric_color=="red", na.rm = T),
                   sum(polygons21$target_color=="red", na.rm = T),
                   sum(polygons21$area_color=="red", na.rm = T),
                   sum(polygons21$count_area_ratio_color=="red", na.rm = T),
                   sum(polygons21$log_hw_color_lo=="red", na.rm = T),
                   sum(polygons21$log_hw_color_hi=="red", na.rm = T))

unflagged_cells <- c(sum(polygons21$metric_color=="white", na.rm = T),
                     sum(polygons21$target_color=="white", na.rm = T),
                     sum(polygons21$area_color=="white", na.rm = T),
                     sum(polygons21$count_area_ratio_color=="white", na.rm = T),
                     sum(polygons21$log_hw_color_lo=="white", na.rm = T),
                     sum(polygons21$log_hw_color_hi=="white", na.rm = T))

percent_flagged <- round(flagged_cells/dim(polygons21)[1]*100, 2)

poly21 <- data.frame(threshold = c(0.2, 0.2, 0.2, 0.2, 0.2, 0.85), flag_value = flag_value, unflagged_cells = unflagged_cells, flagged_cells = flagged_cells, percent_flagged = percent_flagged, row.names = c("Flag score", "Target counts", "Area in um", "Target counts/um area", "Logged aspect ratio low", "Logged aspect ratio high"))

kable(poly21, format = "html", col.names = c("Threshold", "Flag value", "Unflagged cells", "Flagged cells", "% flagged cells"), align = "c", caption = paste0("<center>Flagged cells by different metrics, total cells in fov: ", dim(polygons21)[1], ", total cells in whole sample: ", dim(polygons)[1], ", flagged cells in whole sample: ", sum(spe$flag_score < quantile(spe$flag_score, probs = 0.2, na.rm = T)), "</center>"), padding = 1L)
```

#
#

Local threshold map of flag score, 10th percentile

```{r, out.width="300%", out.height="600%"}
metric_color <- ifelse(polygons21$flag_score < quantile(spe$flag_score, probs = 0.1), "red", "white")
target_color <- ifelse(polygons21$target_counts < quantile(spe$target_counts, probs = 0.1), "red", "white")
area_color <- ifelse(polygons21$um_area < quantile(spe$um_area, probs = 0.1), "red", "white")
count_area_ratio_color <- ifelse((polygons21$target_counts/polygons21$um_area) < quantile((spe$target_counts/spe$um_area), probs = 0.1), "red", "white")
log_hw_color_lo <- ifelse(polygons21$log2_AspectRatio < quantile(spe$log2_AspectRatio, probs = 0.1), "red", "white")
log_hw_color_hi <- ifelse(polygons21$log2_AspectRatio > quantile(spe$log2_AspectRatio, probs = 0.95), "red", "white")

polygons21 <- cbind(polygons21, metric_color = metric_color, target_color = target_color, area_color = area_color, count_area_ratio_color = count_area_ratio_color, log_hw_color_lo = log_hw_color_lo, log_hw_color_hi = log_hw_color_hi)

p1 <- tm_shape(polygons21) + 
  tm_fill(col= "metric_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Flag score",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p2 <- tm_shape(polygons21) + 
  tm_fill(col= "target_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Target counts",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p3 <- tm_shape(polygons21) + 
  tm_fill(col= "area_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "um area",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p4 <- tm_shape(polygons21) + 
  tm_fill(col= "count_area_ratio_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "counts/um area",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p5 <- tm_shape(polygons21) + 
  tm_fill(col= "log_hw_color_lo", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Logged AspectRatio below",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p6 <- tm_shape(polygons21) + 
  tm_fill(col= "log_hw_color_hi", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Logged AspectRatio above",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

tmap_arrange(p1, p2, p3, p4, p5, p6, ncol = 2)

```

```{r, out.width="150%", out.height="150%"}
plot_list <- list(p1, p2, p3, p4, p5, p6)
for(i in 1:length(plot_list)){
  print(plot_list[[i]])
}
```

```{r, out.width="300%", out.height="300%"}

flag_value <- c(round(quantile(polygons21$flag_score, probs = 0.1), 2),
                round(quantile(polygons21$target_counts, probs = 0.1), 2),
                round(quantile(polygons21$um_area, probs = 0.1), 2),
                round(quantile((polygons21$target_counts/polygons21$um_area), probs = 0.1), 2),
                round(quantile(polygons21$log2_AspectRatio, probs = 0.1), 2),
                round(quantile(polygons21$log2_AspectRatio, probs = 0.95), 2))

flagged_cells <- c(sum(polygons21$metric_color=="red", na.rm = T),
                   sum(polygons21$target_color=="red", na.rm = T),
                   sum(polygons21$area_color=="red", na.rm = T),
                   sum(polygons21$count_area_ratio_color=="red", na.rm = T),
                   sum(polygons21$log_hw_color_lo=="red", na.rm = T),
                   sum(polygons21$log_hw_color_hi=="red", na.rm = T))

unflagged_cells <- c(sum(polygons21$metric_color=="white", na.rm = T),
                     sum(polygons21$target_color=="white", na.rm = T),
                     sum(polygons21$area_color=="white", na.rm = T),
                     sum(polygons21$count_area_ratio_color=="white", na.rm = T),
                     sum(polygons21$log_hw_color_lo=="white", na.rm = T),
                     sum(polygons21$log_hw_color_hi=="white", na.rm = T))

percent_flagged <- round(flagged_cells/dim(polygons21)[1]*100, 2)

poly21 <- data.frame(threshold = c(0.1, 0.1, 0.1, 0.1, 0.1, 0.95), flag_value = flag_value, unflagged_cells = unflagged_cells, flagged_cells = flagged_cells, percent_flagged = percent_flagged, row.names = c("Flag score", "Target counts", "Area in um", "Target counts/um area", "Logged aspect ratio low", "Logged aspect ratio high"))

kable(poly21, format = "html", col.names = c("Threshold", "Flag value", "Unflagged cells", "Flagged cells", "% flagged cells"), align = "c", caption = paste0("<center>Flagged cells by different metrics, total cells in fov: ", dim(polygons21)[1], ", total cells in whole sample: ", dim(polygons)[1], ", flagged cells in whole sample: ", sum(spe$flag_score < quantile(spe$flag_score, probs = 0.1, na.rm = T)), "</center>"), padding = 1L)
```

#
#

Local feature map of flag score, fov 25

```{r, out.width="150%", out.height="150%"}
polygons25 <- filter(polygons, fov==25)

tm_shape(polygons25) + 
  tm_fill(col= "flag_score", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE)

```

Local threshold map of flag score, 15th percentile

```{r, out.width="300%", out.height="600%"}
metric_color <- ifelse(polygons25$flag_score < quantile(spe$flag_score, probs = 0.15), "red", "white")
target_color <- ifelse(polygons25$target_counts < quantile(spe$target_counts, probs = 0.15), "red", "white")
area_color <- ifelse(polygons25$um_area < quantile(spe$um_area, probs = 0.15), "red", "white")
count_area_ratio_color <- ifelse((polygons25$target_counts/polygons25$um_area) < quantile((spe$target_counts/spe$um_area), probs = 0.15, na.rm = T), "red", "white")
log_hw_color_lo <- ifelse(polygons25$log2_AspectRatio < quantile(spe$log2_AspectRatio, probs = 0.15), "red", "white")
log_hw_color_hi <- ifelse(polygons25$log2_AspectRatio > quantile(spe$log2_AspectRatio, probs = 0.90), "red", "white")

polygons25 <- cbind(polygons25, metric_color = metric_color, target_color = target_color, area_color = area_color, count_area_ratio_color = count_area_ratio_color, log_hw_color_lo = log_hw_color_lo, log_hw_color_hi = log_hw_color_hi)

p1 <- tm_shape(polygons25) + 
  tm_fill(col= "metric_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Flag score",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p2 <- tm_shape(polygons25) + 
  tm_fill(col= "target_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Target counts",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p3 <- tm_shape(polygons25) + 
  tm_fill(col= "area_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "um area",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p4 <- tm_shape(polygons25) + 
  tm_fill(col= "count_area_ratio_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "counts/um area",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p5 <- tm_shape(polygons25) + 
  tm_fill(col= "log_hw_color_lo", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Logged AspectRatio below",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p6 <- tm_shape(polygons25) + 
  tm_fill(col= "log_hw_color_hi", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Logged AspectRatio above",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))


tmap_arrange(p1, p2, p3, p4, p5, p6, ncol = 2)

```

```{r, out.width="150%", out.height="150%"}
plot_list <- list(p1, p2, p3, p4, p5, p6)
for(i in 1:length(plot_list)){
  print(plot_list[[i]])
}
```

```{r, out.width="150%", out.height="150%"}

flag_value <- c(round(quantile(polygons25$flag_score, probs = 0.15), 2),
                round(quantile(polygons25$target_counts, probs = 0.15), 2),
                round(quantile(polygons25$um_area, probs = 0.15), 2),
                round(quantile((polygons25$target_counts/polygons25$um_area), probs = 0.15), 2),
                round(quantile(polygons25$log2_AspectRatio, probs = 0.15), 2),
                round(quantile(polygons25$log2_AspectRatio, probs = 0.90), 2))

flagged_cells <- c(sum(polygons25$metric_color=="red", na.rm = T),
                   sum(polygons25$target_color=="red", na.rm = T),
                   sum(polygons25$area_color=="red", na.rm = T),
                   sum(polygons25$count_area_ratio_color=="red", na.rm = T),
                   sum(polygons25$log_hw_color_lo=="red", na.rm = T),
                   sum(polygons25$log_hw_color_hi=="red", na.rm = T))

unflagged_cells <- c(sum(polygons25$metric_color=="white", na.rm = T),
                     sum(polygons25$target_color=="white", na.rm = T),
                     sum(polygons25$area_color=="white", na.rm = T),
                     sum(polygons25$count_area_ratio_color=="white", na.rm = T),
                     sum(polygons25$log_hw_color_lo=="white", na.rm = T),
                     sum(polygons25$log_hw_color_hi=="white", na.rm = T))

percent_flagged <- round(flagged_cells/dim(polygons25)[1]*100, 2)

poly25 <- data.frame(threshold = c(0.15, 0.15, 0.15, 0.15, 0.15, 0.90), flag_value = flag_value, unflagged_cells = unflagged_cells, flagged_cells = flagged_cells, 
percent_flagged = percent_flagged, row.names = c("Flag score", "Target counts", "Area in um", "Target counts/um area", "Logged aspect ratio low", "Logged aspect ratio high"))

kable(poly25, format = "html", col.names = c("Threshold", "Flag value", "Unflagged cells", "Flagged cells", "% flagged cells"), align = "c", caption = paste0("<center>Flagged cells by different metrics, total cells in fov: ", dim(polygons25)[1], ", total cells in whole sample: ", dim(polygons)[1], ", flagged cells in whole sample: ", sum(spe$flag_score < quantile(spe$flag_score, probs = 0.15, na.rm = T)), "</center>"), padding = 1L)
```

#
#

Local feature map of flag score, fov 35

```{r, out.width="150%", out.height="150%"}
polygons35 <- filter(polygons, fov==35)

tm_shape(polygons35) + 
  tm_fill(col= "flag_score", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE)

```

Local threshold map of flag score, 15th percentile

```{r, out.width="300%", out.height="600%"}
metric_color <- ifelse(polygons35$flag_score < quantile(spe$flag_score, probs = 0.15), "red", "white")
target_color <- ifelse(polygons35$target_counts < quantile(spe$target_counts, probs = 0.15), "red", "white")
area_color <- ifelse(polygons35$um_area < quantile(spe$um_area, probs = 0.15), "red", "white")
count_area_ratio_color <- ifelse((polygons35$target_counts/polygons35$um_area) < quantile((spe$target_counts/spe$um_area), probs = 0.15, na.rm = T), "red", "white")
log_hw_color_lo <- ifelse(polygons35$log2_AspectRatio < quantile(spe$log2_AspectRatio, probs = 0.15), "red", "white")
log_hw_color_hi <- ifelse(polygons35$log2_AspectRatio > quantile(spe$log2_AspectRatio, probs = 0.90), "red", "white")

polygons35 <- cbind(polygons35, metric_color = metric_color, target_color = target_color, area_color = area_color, count_area_ratio_color = count_area_ratio_color, log_hw_color_lo = log_hw_color_lo, log_hw_color_hi = log_hw_color_hi)

p1 <- tm_shape(polygons35) + 
  tm_fill(col= "metric_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Flag score",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p2 <- tm_shape(polygons35) + 
  tm_fill(col= "target_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Target counts",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p3 <- tm_shape(polygons35) + 
  tm_fill(col= "area_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "um area",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p4 <- tm_shape(polygons35) + 
  tm_fill(col= "count_area_ratio_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "counts/um area",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p5 <- tm_shape(polygons35) + 
  tm_fill(col= "log_hw_color_lo", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Logged AspectRatio below",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p6 <- tm_shape(polygons35) + 
  tm_fill(col= "log_hw_color_hi", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Logged AspectRatio above",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

tmap_arrange(p1, p2, p3, p4, p5, p6, ncol = 2)

```

```{r, out.width="150%", out.height="150%"}
plot_list <- list(p1, p2, p3, p4, p5, p6)
for(i in 1:length(plot_list)){
  print(plot_list[[i]])
}
```

```{r, out.width="150%", out.height="150%"}

flag_value <- c(round(quantile(polygons35$flag_score, probs = 0.15), 2),
                round(quantile(polygons35$target_counts, probs = 0.15), 2),
                round(quantile(polygons35$um_area, probs = 0.15), 2),
                round(quantile((polygons35$target_counts/polygons35$um_area), probs = 0.15), 2),
                round(quantile(polygons35$log2_AspectRatio, probs = 0.15), 2),
                round(quantile(polygons35$log2_AspectRatio, probs = 0.90), 2))

flagged_cells <- c(sum(polygons35$metric_color=="red", na.rm = T),
                   sum(polygons35$target_color=="red", na.rm = T),
                   sum(polygons35$area_color=="red", na.rm = T),
                   sum(polygons35$count_area_ratio_color=="red", na.rm = T),
                   sum(polygons35$log_hw_color_lo=="red", na.rm = T),
                   sum(polygons35$log_hw_color_hi=="red", na.rm = T))

unflagged_cells <- c(sum(polygons35$metric_color=="white", na.rm = T),
                     sum(polygons35$target_color=="white", na.rm = T),
                     sum(polygons35$area_color=="white", na.rm = T),
                     sum(polygons35$count_area_ratio_color=="white", na.rm = T),
                     sum(polygons35$log_hw_color_lo=="white", na.rm = T),
                     sum(polygons35$log_hw_color_hi=="white", na.rm = T))

percent_flagged <- round(flagged_cells/dim(polygons35)[1]*100, 2)

poly35 <- data.frame(threshold = c(0.15, 0.15, 0.15, 0.15, 0.15, 0.90), flag_value = flag_value, unflagged_cells = unflagged_cells, flagged_cells = flagged_cells,
percent_flagged = percent_flagged, row.names = c("Flag score", "Target counts", "Area in um", "Target counts/um area", "Logged aspect ratio low", "Logged aspect ratio high"))

kable(poly35, format = "html", col.names = c("Threshold", "Flag value", "Unflagged cells", "Flagged cells", "% flagged cells"), align = "c", caption = paste0("<center>Flagged cells by different metrics, total cells in fov: ", dim(polygons35)[1], ", total cells in whole sample: ", dim(polygons)[1], ", flagged cells in whole sample: ", sum(spe$flag_score < quantile(spe$flag_score, probs = 0.15, na.rm = T)), "</center>"), padding = 1L)
```

#
#

Local feature map of flag score, fov 32

```{r, out.width="150%", out.height="150%"}
polygons32 <- filter(polygons, fov==32)

tm_shape(polygons32) + 
  tm_fill(col= "flag_score", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE)

```

Local threshold map of flag score, 15th percentile

```{r, out.width="300%", out.height="600%"}
metric_color <- ifelse(polygons32$flag_score < quantile(spe$flag_score, probs = 0.15), "red", "white")
target_color <- ifelse(polygons32$target_counts < quantile(spe$target_counts, probs = 0.15), "red", "white")
area_color <- ifelse(polygons32$um_area < quantile(spe$um_area, probs = 0.15), "red", "white")
count_area_ratio_color <- ifelse((polygons32$target_counts/polygons32$um_area) < quantile((spe$target_counts/spe$um_area), probs = 0.15, na.rm = T), "red", "white")
log_hw_color_lo <- ifelse(polygons32$log2_AspectRatio < quantile(spe$log2_AspectRatio, probs = 0.15), "red", "white")
log_hw_color_hi <- ifelse(polygons32$log2_AspectRatio > quantile(spe$log2_AspectRatio, probs = 0.90), "red", "white")

polygons32 <- cbind(polygons32, metric_color = metric_color, target_color = target_color, area_color = area_color, count_area_ratio_color = count_area_ratio_color, log_hw_color_lo = log_hw_color_lo, log_hw_color_hi = log_hw_color_hi)

p1 <- tm_shape(polygons32) + 
  tm_fill(col= "metric_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Flag score",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p2 <- tm_shape(polygons32) + 
  tm_fill(col= "target_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Target counts",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p3 <- tm_shape(polygons32) + 
  tm_fill(col= "area_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "um area",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p4 <- tm_shape(polygons32) + 
  tm_fill(col= "count_area_ratio_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "counts/um area",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p5 <- tm_shape(polygons32) + 
  tm_fill(col= "log_hw_color_lo", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Logged AspectRatio below",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p6 <- tm_shape(polygons32) + 
  tm_fill(col= "log_hw_color_hi", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Logged AspectRatio above",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

tmap_arrange(p1, p2, p3, p4, p5, p6, ncol = 2)

```

```{r, out.width="150%", out.height="150%"}
plot_list <- list(p1, p2, p3, p4, p5, p6)
for(i in 1:length(plot_list)){
  print(plot_list[[i]])
}
```

```{r, out.width="150%", out.height="150%"}

flag_value <- c(round(quantile(polygons32$flag_score, probs = 0.15), 2),
                round(quantile(polygons32$target_counts, probs = 0.15), 2),
                round(quantile(polygons32$um_area, probs = 0.15), 2),
                round(quantile((polygons32$target_counts/polygons32$um_area), probs = 0.15), 2),
                round(quantile(polygons32$log2_AspectRatio, probs = 0.15), 2),
                round(quantile(polygons32$log2_AspectRatio, probs = 0.90), 2))

flagged_cells <- c(sum(polygons32$metric_color=="red", na.rm = T),
                   sum(polygons32$target_color=="red", na.rm = T),
                   sum(polygons32$area_color=="red", na.rm = T),
                   sum(polygons32$count_area_ratio_color=="red", na.rm = T),
                   sum(polygons32$log_hw_color_lo=="red", na.rm = T),
                   sum(polygons32$log_hw_color_hi=="red", na.rm = T))

unflagged_cells <- c(sum(polygons32$metric_color=="white", na.rm = T),
                     sum(polygons32$target_color=="white", na.rm = T),
                     sum(polygons32$area_color=="white", na.rm = T),
                     sum(polygons32$count_area_ratio_color=="white", na.rm = T),
                     sum(polygons32$log_hw_color_lo=="white", na.rm = T),
                     sum(polygons32$log_hw_color_hi=="white", na.rm = T))

percent_flagged <- round(flagged_cells/dim(polygons32)[1]*100, 2)

poly32 <- data.frame(threshold = c(0.15, 0.15, 0.15, 0.15, 0.15, 0.90), flag_value = flag_value, unflagged_cells = unflagged_cells, flagged_cells = flagged_cells,
percent_flagged = percent_flagged, row.names = c("Flag score", "Target counts", "Area in um", "Target counts/um area", "Logged aspect ratio low", "Logged aspect ratio high"))

kable(poly32, format = "html", col.names = c("Threshold", "Flag value", "Unflagged cells", "Flagged cells", "% flagged cells"), align = "c", caption = paste0("<center>Flagged cells by different metrics, total cells in fov: ", dim(polygons32)[1], ", total cells in whole sample: ", dim(polygons)[1], ", flagged cells in whole sample: ", sum(spe$flag_score < quantile(spe$flag_score, probs = 0.15, na.rm = T)), "</center>"), padding = 1L)
```

#
#

Local feature map of flag score, fov 40

```{r, out.width="150%", out.height="150%"}
polygons40 <- filter(polygons, fov==40)

tm_shape(polygons40) + 
  tm_fill(col= "flag_score", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE)

```

Local threshold map of flag score, 15th percentile

```{r, out.width="300%", out.height="600%"}
metric_color <- ifelse(polygons40$flag_score < quantile(spe$flag_score, probs = 0.15), "red", "white")
target_color <- ifelse(polygons40$target_counts < quantile(spe$target_counts, probs = 0.15), "red", "white")
area_color <- ifelse(polygons40$um_area < quantile(spe$um_area, probs = 0.15), "red", "white")
count_area_ratio_color <- ifelse((polygons40$target_counts/polygons40$um_area) < quantile((spe$target_counts/spe$um_area), probs = 0.15, na.rm = T), "red", "white")
log_hw_color_lo <- ifelse(polygons40$log2_AspectRatio < quantile(spe$log2_AspectRatio, probs = 0.15), "red", "white")
log_hw_color_hi <- ifelse(polygons40$log2_AspectRatio > quantile(spe$log2_AspectRatio, probs = 0.90), "red", "white")

polygons40 <- cbind(polygons40, metric_color = metric_color, target_color = target_color, area_color = area_color, count_area_ratio_color = count_area_ratio_color, log_hw_color_lo = log_hw_color_lo, log_hw_color_hi = log_hw_color_hi)

p1 <- tm_shape(polygons40) + 
  tm_fill(col= "metric_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Flag score",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p2 <- tm_shape(polygons40) + 
  tm_fill(col= "target_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Target counts",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p3 <- tm_shape(polygons40) + 
  tm_fill(col= "area_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "um area",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p4 <- tm_shape(polygons40) + 
  tm_fill(col= "count_area_ratio_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "counts/um area",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p5 <- tm_shape(polygons40) + 
  tm_fill(col= "log_hw_color_lo", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Logged AspectRatio below",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p6 <- tm_shape(polygons40) + 
  tm_fill(col= "log_hw_color_hi", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Logged AspectRatio above",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

tmap_arrange(p1, p2, p3, p4, p5, p6, ncol = 2)

```

```{r, out.width="150%", out.height="150%"}
plot_list <- list(p1, p2, p3, p4, p5, p6)
for(i in 1:length(plot_list)){
  print(plot_list[[i]])
}
```

```{r, out.width="150%", out.height="150%"}

flag_value <- c(round(quantile(polygons40$flag_score, probs = 0.15), 2),
                round(quantile(polygons40$target_counts, probs = 0.15), 2),
                round(quantile(polygons40$um_area, probs = 0.15), 2),
                round(quantile((polygons40$target_counts/polygons40$um_area), probs = 0.15), 2),
                round(quantile(polygons40$log2_AspectRatio, probs = 0.15), 2),
                round(quantile(polygons40$log2_AspectRatio, probs = 0.90), 2))

flagged_cells <- c(sum(polygons40$metric_color=="red", na.rm = T),
                   sum(polygons40$target_color=="red", na.rm = T),
                   sum(polygons40$area_color=="red", na.rm = T),
                   sum(polygons40$count_area_ratio_color=="red", na.rm = T),
                   sum(polygons40$log_hw_color_lo=="red", na.rm = T),
                   sum(polygons40$log_hw_color_hi=="red", na.rm = T))

unflagged_cells <- c(sum(polygons40$metric_color=="white", na.rm = T),
                     sum(polygons40$target_color=="white", na.rm = T),
                     sum(polygons40$area_color=="white", na.rm = T),
                     sum(polygons40$count_area_ratio_color=="white", na.rm = T),
                     sum(polygons40$log_hw_color_lo=="white", na.rm = T),
                     sum(polygons40$log_hw_color_hi=="white", na.rm = T))

percent_flagged <- round(flagged_cells/dim(polygons40)[1]*100, 2)

poly40 <- data.frame(threshold = c(0.15, 0.15, 0.15, 0.15, 0.15, 0.90), flag_value = flag_value, unflagged_cells = unflagged_cells, flagged_cells = flagged_cells,
percent_flagged = percent_flagged, row.names = c("Flag score", "Target counts", "Area in um", "Target counts/um area", "Logged aspect ratio low", "Logged aspect ratio high"))

kable(poly40, format = "html", col.names = c("Threshold", "Flag value", "Unflagged cells", "Flagged cells", "% flagged cells"), align = "c", caption = paste0("<center>Flagged cells by different metrics, total cells in fov: ", dim(polygons40)[1], ", total cells in whole sample: ", dim(polygons)[1], ", flagged cells in whole sample: ", sum(spe$flag_score < quantile(spe$flag_score, probs = 0.15, na.rm = T)), "</center>"), padding = 1L)
```

#
#

Local feature map of flag score, fov 64

```{r, out.width="150%", out.height="150%"}
polygons64 <- filter(polygons, fov==64)

tm_shape(polygons64) + 
  tm_fill(col= "flag_score", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE)

```

Local threshold map of flag score, 15th percentile

```{r, out.width="300%", out.height="600%"}
metric_color <- ifelse(polygons64$flag_score < quantile(spe$flag_score, probs = 0.15), "red", "white")
target_color <- ifelse(polygons64$target_counts < quantile(spe$target_counts, probs = 0.15), "red", "white")
area_color <- ifelse(polygons64$um_area < quantile(spe$um_area, probs = 0.15), "red", "white")
count_area_ratio_color <- ifelse((polygons64$target_counts/polygons64$um_area) < quantile((spe$target_counts/spe$um_area), probs = 0.15, na.rm = T), "red", "white")
log_hw_color_lo <- ifelse(polygons64$log2_AspectRatio < quantile(spe$log2_AspectRatio, probs = 0.15), "red", "white")
log_hw_color_hi <- ifelse(polygons64$log2_AspectRatio > quantile(spe$log2_AspectRatio, probs = 0.90), "red", "white")

polygons64 <- cbind(polygons64, metric_color = metric_color, target_color = target_color, area_color = area_color, count_area_ratio_color = count_area_ratio_color, log_hw_color_lo = log_hw_color_lo, log_hw_color_hi = log_hw_color_hi)

p1 <- tm_shape(polygons64) + 
  tm_fill(col= "metric_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Flag score",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p2 <- tm_shape(polygons64) + 
  tm_fill(col= "target_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Target counts",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p3 <- tm_shape(polygons64) + 
  tm_fill(col= "area_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "um area",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p4 <- tm_shape(polygons64) + 
  tm_fill(col= "count_area_ratio_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "counts/um area",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p5 <- tm_shape(polygons64) + 
  tm_fill(col= "log_hw_color_lo", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Logged AspectRatio below",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p6 <- tm_shape(polygons64) + 
  tm_fill(col= "log_hw_color_hi", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Logged AspectRatio above",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

tmap_arrange(p1, p2, p3, p4, p5, p6, ncol = 2)

```

```{r, out.width="150%", out.height="150%"}
plot_list <- list(p1, p2, p3, p4, p5, p6)
for(i in 1:length(plot_list)){
  print(plot_list[[i]])
}
```

```{r, out.width="150%", out.height="150%"}

flag_value <- c(round(quantile(polygons64$flag_score, probs = 0.15), 2),
                round(quantile(polygons64$target_counts, probs = 0.15), 2),
                round(quantile(polygons64$um_area, probs = 0.15), 2),
                round(quantile((polygons64$target_counts/polygons64$um_area), probs = 0.15), 2),
                round(quantile(polygons64$log2_AspectRatio, probs = 0.15), 2),
                round(quantile(polygons64$log2_AspectRatio, probs = 0.90), 2))

flagged_cells <- c(sum(polygons64$metric_color=="red", na.rm = T),
                   sum(polygons64$target_color=="red", na.rm = T),
                   sum(polygons64$area_color=="red", na.rm = T),
                   sum(polygons64$count_area_ratio_color=="red", na.rm = T),
                   sum(polygons64$log_hw_color_lo=="red", na.rm = T),
                   sum(polygons64$log_hw_color_hi=="red", na.rm = T))

unflagged_cells <- c(sum(polygons64$metric_color=="white", na.rm = T),
                     sum(polygons64$target_color=="white", na.rm = T),
                     sum(polygons64$area_color=="white", na.rm = T),
                     sum(polygons64$count_area_ratio_color=="white", na.rm = T),
                     sum(polygons64$log_hw_color_lo=="white", na.rm = T),
                     sum(polygons64$log_hw_color_hi=="white", na.rm = T))

percent_flagged <- round(flagged_cells/dim(polygons64)[1]*100, 2)

poly64 <- data.frame(threshold = c(0.15, 0.15, 0.15, 0.15, 0.15, 0.90), flag_value = flag_value, unflagged_cells = unflagged_cells, flagged_cells = flagged_cells,
percent_flagged = percent_flagged, row.names = c("Flag score", "Target counts", "Area in um", "Target counts/um area", "Logged aspect ratio low", "Logged aspect ratio high"))

kable(poly64, format = "html", col.names = c("Threshold", "Flag value", "Unflagged cells", "Flagged cells", "% flagged cells"), align = "c", caption = paste0("<center>Flagged cells by different metrics, total cells in fov: ", dim(polygons64)[1], ", total cells in whole sample: ", dim(polygons)[1], ", flagged cells in whole sample: ", sum(spe$flag_score < quantile(spe$flag_score, probs = 0.15, na.rm = T)), "</center>"), padding = 1L)
```

