---
title: "02_21_xenium_BC_preprocessing_metric_report"
author: "Benedetta Banzi"
date: "2024-02-22"
output: html_document
---

<STYLE TYPE="text/css">
<!--
  td{
    font-family: Arial; 
    font-size: 10 pt;
    padding:0px;
    cellpadding="0";
    cellspacing="0"
  }
  th {
    font-family: Arial; 
    font-size: 10 pt;
    height: 20px;
    font-weight: bold;
    text-align: "center";
    background-color: #ccccff;
    padding-left: 15px;
    padding-right: 15px;
  }
  table { 
    border-spacing: 4px;
    border-collapse: collapse;
  }
--->
</STYLE>


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Library import

```{r}
library(SpatialExperiment)
library(data.table)
library(scater) # it imports scuttle
library(cowplot)
library(ggplot2)
theme_set(theme_bw())
library(matrixStats)
library(dplyr)
library(tidyr)
library(tibble)
library(arrow)
library(scales)
library(sf)
library(tmap)
library(knitr)
library(grid)
library(ggpointdensity)
library(ggpubr)
```

Read-in function

```{r}
readXeniumSPE <- function(dirname, 
                          countfname = "cell_feature_matrix.h5",
                          coordfpattern = "cells.csv.gz", 
                          coord_names = c("x_centroid", "y_centroid")){
  countfpath <- file.path(dirname, countfname)
  coord_file <- file.path(dirname, list.files(dirname, coordfpattern))
  
  # Count matrix + rowData
  sce <- DropletUtils::read10xCounts(countfpath, col.names = TRUE)
  
  # Spatial and colData
  colData <- read.csv(gzfile(coord_file), header = TRUE)
  
  # construct 'SpatialExperiment'
  spe <- SpatialExperiment::SpatialExperiment(
    assays = assays(sce),
    rowData = rowData(sce),
    colData = colData,
    spatialCoordsNames = coord_names
  )
  
  return(spe)
}
```

Setting directories

```{r}
dirname <- "/NAS06/work/Spatial_omics/Xenium/Breast_cancer_rep1/outs"
sample_name <- "Xenium_human_BC"
```

# CosMx dataset exploratory analysis for preprocessing

SpatialExperiment object creation and exploration

```{r}
spe <- readXeniumSPE(dirname = dirname)
spe <- scuttle::addPerCellQC(spe, subsets=list(NegProb = grep("^NegControlProbe_", rownames(spe)), NegCodw = grep("^NegControlCodeword_", rownames(spe)), Blank = grep("^BLANK_", rownames(spe))))
names(colData(spe)@listData)
spe$control_counts <- rowSums(data.frame(colData(spe))[,grep("^subsets_.*_sum$", names(colData(spe)))])
spe$control_detected <- rowSums(data.frame(colData(spe))[,grep("^subsets_.*_detected$", names(colData(spe)))])
```

```{r}
dim(spe)
```

Selecting only numeric and integer variables from metadata then only variables of interest from the list

```{r}
temp <- unlist(lapply(spe@colData@listData, function(x) is.numeric(x) | is.integer(x)))
num_variables <- c()
for (i in 1:length(temp)){
  num_variables[i] <- ifelse(temp[i] == TRUE, names(temp[i]), NA)
  num_variables <- num_variables[!is.na(num_variables)]
}
num_variables
variables <- c("total", "detected")
```

Importing cell boundaries file to compute AspectRatio

```{r}
cell_boundaries <- read_parquet(file.path(dirname, "cell_boundaries.parquet"))
bound_sf <- st_as_sf(cell_boundaries, coords = c("vertex_x", "vertex_y"))

polygons <- bound_sf %>%
  group_by(cell_id) %>%
  summarize(geometry = st_combine(geometry))#st_combine = combine several feature geometries into one, without unioning or resolving internal boundaries

polygons <- st_cast(polygons, "POLYGON")

AspectRatio <- c()
for(i in 1:dim(polygons)[1]){
 AspectRatio[i] <- (max(polygons$geometry[[i]][[1]][,2]) - min(polygons$geometry[[i]][[1]][,2]))/(max(polygons$geometry[[i]][[1]][,1]) -
 min(polygons$geometry[[i]][[1]][,1]))
}
```

Added variables

```{r}
# Importing global coordinates from spatialCoords slot since they are not available in colData
colData(spe)$CenterX_global_um <- spatialCoords(spe)[,1]
colData(spe)$CenterY_global_um <- spatialCoords(spe)[,2]

# Logged total counts
spe$logtot <- log10(spe$total)

# Logged width/height ratio so there are no holes in the distribution in the middle and also negative values can be appreciated
spe$AspectRatio <- AspectRatio
spe$log2_AspectRatio <- log2(spe$AspectRatio)

# Keeping only target counts by subtracting counts referred to negative control probes
spe$target_counts <- spe$total - spe$control_counts

# Keeping only detected features by subtracting negative control probes
spe$target_detected <- spe$detected - spe$control_detected

# Proportion between counts referred only to target probes and detected features
spe$tot_det_ratio <- spe$target_counts/spe$target_detected

spe$ctrl_total_ratio <- spe$control_counts/spe$total
```

Computing mock fovs starting from first identified section size

```{r}
meta_df <- data.frame(colData(spe))

# isolating only first fov
meta_df_1 <- meta_df[1:643, ]
min(meta_df_1$CenterX_global_um)
max(meta_df_1$CenterX_global_um)
min(meta_df_1$CenterY_global_um)
max(meta_df_1$CenterY_global_um)

# computing the fov size using the max and the min of the first fov coordinates
fov_size_x <- max(meta_df_1$CenterX_global_um) - min(meta_df_1$CenterX_global_um)
fov_size_y <- max(meta_df_1$CenterY_global_um) - min(meta_df_1$CenterY_global_um)

# binning coordinates using fov_size
x_bin <- cut(spe$CenterX_global_um, seq(min(spe$CenterX_global_um), max(spe$CenterX_global_um), fov_size_x), include.lowest = TRUE)
y_bin <- cut(spe$CenterY_global_um, seq(min(spe$CenterY_global_um), max(spe$CenterY_global_um), fov_size_y), include.lowest = TRUE)

# creating grid cell indexes
levels(x_bin) <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10)
levels(y_bin) <- c("A", "B", "C", "D", "E", "F", "G", "H")

# combining the indexes
mock_fov <- paste(x_bin, y_bin)

# adding the cell indexes to the Spatial object metadata
spe$mock_fov <- mock_fov
```


# Global sample map

To view the spatial organization of both the sample as it is and FOVs

```{r, out.width="600%", out.height="150%"}
# image theme
global_map_theme <- function(back.color="black",back.border=NA,title.col="white") {
  theme(aspect.ratio = 1,
        panel.border = element_blank(),
        legend.key = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks = element_blank(),
        axis.text.y = element_blank(),
        axis.text.x = element_blank(),
        panel.grid = element_blank(),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        panel.background = element_rect(fill = "transparent",colour = NA),
        plot.title = element_text(color = title.col,hjust = 0.5,face = "bold"),
        plot.background = element_rect(fill = "transparent",colour = back.border))
}

meta_df$grid_color <- rep(c("darkmagenta", "grey28"), c(14250, dim(spe)[2]-14250))

ggplot() + 
  geom_point(data = meta_df, mapping = aes(x = CenterX_global_um, y = CenterY_global_um), color = meta_df$grid_color, fill = meta_df$grid_color, size = 0.05, alpha = 0.2) +
  ggtitle(sample_name) + 
  global_map_theme(back.color = "white", back.border = "white", title.col = "black")
```


# Global sample map with cell types

```{r, out.width="150%", out.height="300%"}
anno_file <- fread(file.path(dirname, "Xenium_BC_rep1_annotation.txt"))

colnames(anno_file) <- c("cell_id", "celltype_annotation")
meta_df <- left_join(meta_df, anno_file, by = join_by(cell_id))

table(meta_df$celltype_annotation)

celltype_palette <- c(
  "B_Cells" = "blueviolet",
  "CD4+_T_Cells" = "#43e109",
  "CD8+_T_Cells" = "#94e1cf",
  "DCIS_1" = "#4b98af",
  "DCIS_2" = "#ff6a00", 
  "Endothelial" = "#47c8ff",
  "Invasive_Tumor" = "red",
  "IRF7+_DCs" = "#eb37b2",
  "LAMP3+_DCs" = "#6665DD",
  "Macrophages_1" = "#8abd00",
  "Macrophages_2" = "#AF4D98",
  "Mast_Cells" = "#3936D3",
  "Myoepi_ACTA2+" = "#ffae57",
  "Myoepi_KRT15+" = "#b80040",
  "Perivascular-Like" = "#9B9ECE",
  'Prolif_Invasive_Tumor' = "firebrick",
  "Stromal" = "#086fc9",
  "Stromal_&_T_Cell_Hybrid" = "#134502",
  "T_Cell_&_Tumor_Hybrid" = "#dcf500",
  "Unlabeled" = "grey80"
)

celltype_color <- c()
for(i in 1:dim(spe)[2]){
  celltype <- meta_df$celltype_annotation[i]
  n <- match(celltype, names(celltype_palette))
  celltype_color[i] <- celltype_palette[n]
}

meta_df$celltype_color <- celltype_color
polygons$celltype_color <- celltype_color

tm_shape(polygons) + 
tm_fill(col = "celltype_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = sample_name,
            main.title.size = 0.5,
            frame = FALSE,
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0)) +
tm_add_legend(type = "fill", labels = names(celltype_palette), col = celltype_palette)

```

# Global feature maps

To spot areas with anomalous feature values and spatial patterns on a global scale

```{r, out.width="300%", out.height="150%"}
 plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "total", point_size = 0.1) + ggtitle("Total counts per cell") + theme(aspect.ratio = 1)
```

```{r, out.width="300%", out.height="150%"}
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "logtot", point_size = 0.1) + ggtitle("Log-total counts per cell") + theme(aspect.ratio = 1)
```

```{r, out.width="300%", out.height="150%"}
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "detected", point_size = 0.1) + ggtitle("Detected features per cell") + theme(aspect.ratio = 1)
```

```{r, out.width="300%", out.height="150%"}
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "cell_area", point_size = 0.1) + ggtitle("Area in μm per cell") + theme(aspect.ratio = 1)
```

```{r, out.width="300%", out.height="150%"}
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "nucleus_area", point_size = 0.1) + ggtitle("Nucleus area per cell") + theme(aspect.ratio = 1)
```

```{r, out.width="300%", out.height="150%"}
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "log2_AspectRatio", point_size = 0.1) + ggtitle("Logged width/Height ratio per cell") + theme(aspect.ratio = 1)
```

```{r, out.width="300%", out.height="150%"}
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "tot_det_ratio", point_size = 0.1) + ggtitle("Target counts/detected features ratio per cell") + theme(aspect.ratio = 1)
```

```{r, out.width="300%", out.height="150%"}
plotColData(spe, "CenterY_global_um", "CenterX_global_um", order_by = "control_counts", colour_by = "control_counts", point_size = 0.1) +
  scale_color_gradient(low = "white", high = "red", name = "control_counts") + ggtitle("Control counts per cell") +
  theme(aspect.ratio = 1,
        panel.background = element_rect(fill = "black",color = NA),
        plot.background = element_rect(fill = "black",color = NA),
        axis.title.x = element_text(color = "white"),
        axis.title.y = element_text(color = "white"),
        axis.ticks = element_line(color = "white"),
        axis.text.y = element_text(color = "white"),
        axis.text.x = element_text(color = "white"),
        axis.line = element_line(color = "white"),
        legend.title = element_text(color = "white"),
        legend.text = element_text(color = "white"),
        plot.title = element_text(color = "white"))
```

```{r}
threshold <- 0.20
spe$flagged_tot_20 <- as.factor(ifelse(colData(spe)$total < quantile(colData(spe)$total, probs = threshold), TRUE, FALSE))

threshold <- 0.10
spe$flagged_tot_10 <- as.factor(ifelse(colData(spe)$total < quantile(colData(spe)$total, probs = threshold), TRUE, FALSE))

threshold <- 0.05
spe$flagged_tot_5 <- as.factor(ifelse(colData(spe)$total < quantile(colData(spe)$total, probs = threshold), TRUE, FALSE))

threshold <- 0.10
spe$flagged_det_10 <- as.factor(ifelse(colData(spe)$detected < quantile(colData(spe)$detected, probs = threshold), TRUE, FALSE))

threshold <- 0.05
spe$flagged_det_5 <- as.factor(ifelse(colData(spe)$detected < quantile(colData(spe)$detected, probs = threshold), TRUE, FALSE))

threshold <- 0.10
spe$flagged_carea_10 <- as.factor(ifelse(colData(spe)$cell_area < quantile(colData(spe)$cell_area, probs = threshold), TRUE, FALSE))

threshold <- 0.05
spe$flagged_carea_5 <- as.factor(ifelse(colData(spe)$cell_area < quantile(colData(spe)$cell_area, probs = threshold), TRUE, FALSE))

threshold <- 0.90
spe$flagged_carea_90 <- as.factor(ifelse(colData(spe)$cell_area > quantile(colData(spe)$cell_area, probs = threshold), TRUE, FALSE))

threshold <- 0.95
spe$flagged_carea_95 <- as.factor(ifelse(colData(spe)$cell_area > quantile(colData(spe)$cell_area, probs = threshold), TRUE, FALSE))

threshold <- 0.10
spe$flagged_narea_10 <- as.factor(ifelse(colData(spe)$nucleus_area < quantile(colData(spe)$nucleus_area, probs = threshold), TRUE, FALSE))

threshold <- 0.05
spe$flagged_narea_5 <- as.factor(ifelse(colData(spe)$nucleus_area < quantile(colData(spe)$nucleus_area, probs = threshold), TRUE, FALSE))

threshold <- 0.90
spe$flagged_narea_90 <- as.factor(ifelse(colData(spe)$nucleus_area > quantile(colData(spe)$nucleus_area, probs = threshold), TRUE, FALSE))

threshold <- 0.95
spe$flagged_narea_95 <- as.factor(ifelse(colData(spe)$nucleus_area > quantile(colData(spe)$nucleus_area, probs = threshold), TRUE, FALSE))

threshold <- 0.10
spe$flagged_aratio_10 <- as.factor(ifelse(colData(spe)$log2_AspectRatio < quantile(colData(spe)$log2_AspectRatio, probs = threshold), TRUE, FALSE))

threshold <- 0.05
spe$flagged_aratio_5 <- as.factor(ifelse(colData(spe)$log2_AspectRatio < quantile(colData(spe)$log2_AspectRatio, probs = threshold), TRUE, FALSE))

threshold <- 0.90
spe$flagged_aratio_90 <- as.factor(ifelse(colData(spe)$log2_AspectRatio > quantile(colData(spe)$log2_AspectRatio, probs = threshold), TRUE, FALSE))

threshold <- 0.95
spe$flagged_aratio_95 <- as.factor(ifelse(colData(spe)$log2_AspectRatio > quantile(colData(spe)$log2_AspectRatio, probs = threshold), TRUE, FALSE))

threshold <- 0.10
spe$flagged_tdratio_10 <- as.factor(ifelse(colData(spe)$tot_det_ratio < quantile(colData(spe)$tot_det_ratio, probs = threshold, na.rm = T), TRUE, FALSE))

threshold <- 0.05
spe$flagged_tdratio_5 <- as.factor(ifelse(colData(spe)$tot_det_ratio < quantile(colData(spe)$tot_det_ratio, probs = threshold, na.rm = T), TRUE, FALSE))

threshold <- 1
spe$flagged_tdratio_atomx <- as.factor(ifelse(colData(spe)$tot_det_ratio <= threshold, TRUE, FALSE))

threshold <- 0.50
spe$flagged_negprob_50 <- as.factor(ifelse(colData(spe)$subsets_NegProb_sum > quantile(colData(spe)$subsets_NegProb_sum, probs = threshold, na.rm = T), TRUE, FALSE))

threshold <- 0.75
spe$flagged_negprob_75 <- as.factor(ifelse(colData(spe)$subsets_NegProb_sum > quantile(colData(spe)$subsets_NegProb_sum, probs = threshold, na.rm = T), TRUE, FALSE))

spe$ctrl_total_ratio <- spe$subsets_NegProb_sum/spe$total
threshold <- 0.10
spe$flagged_negprob_10 <- as.factor(ifelse(colData(spe)$ctrl_total_ratio > quantile(colData(spe)$ctrl_total_ratio, probs = threshold, na.rm = T), TRUE, FALSE))
```


# Feature histograms

To check feature distribution before highlighting where extreme values are located in the global sample map

```{r}
variables <- c(variables, "logtot", "cell_area", "nucleus_area", "log2_AspectRatio", "tot_det_ratio", "control_counts")
for (var in 1:length(variables)){
  n <- grep(paste0("^", variables[var], "$"), colnames(spe@colData))
  histo_var <- names(colData(spe))[n]
  p <- ggplot(data = data.frame(colData(spe))) +
    geom_histogram(aes(x = .data[[histo_var]]), fill = "#69b3a2") +
    ggtitle(paste0(histo_var))
  print(p)
}  
```

# Global threshold map

Cells below the selected feature threshold are highlighted in black, all the others are grey

```{r, out.width="150%", out.height="150%"}
threshold <- 0.20
spe$flagged_tot_20 <- as.factor(ifelse(colData(spe)$total < quantile(colData(spe)$total, probs = threshold), TRUE, FALSE))
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "flagged_tot_20", 
            point_size = 0.1) + scale_color_manual(values = c("TRUE" = "black", "FALSE" = "lightgrey")) + ggtitle("Counts below 20% threshold (default by AtoMx guidelines)") + theme(legend.position="none") + labs(subtitle = paste0("Flagged cells = ", table(colData(spe)$flagged_tot_20)["TRUE"], " out of ", dim(spe)[2], ", ", round(table(colData(spe)$flagged_tot_20)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"))
```

```{r, out.width="150%", out.height="150%"}
threshold <- 0.10
spe$flagged_tot_10 <- as.factor(ifelse(colData(spe)$total < quantile(colData(spe)$total, probs = threshold), TRUE, FALSE))
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "flagged_tot_10", 
            point_size = 0.1) + scale_color_manual(values = c("TRUE" = "black","FALSE" = "lightgrey")) + ggtitle("Counts below 10% threshold")  + theme(legend.position="none") + labs(subtitle = paste0("Flagged cells = ", table(colData(spe)$flagged_tot_10)["TRUE"], " out of ", dim(spe)[2], ", ",  round(table(colData(spe)$flagged_tot_10)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"))
```

```{r, out.width="150%", out.height="150%"}
threshold <- 0.05
spe$flagged_tot_5 <- as.factor(ifelse(colData(spe)$total < quantile(colData(spe)$total, probs = threshold), TRUE, FALSE))
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "flagged_tot_5", 
            point_size = 0.1) + scale_color_manual(values = c("TRUE" = "black","FALSE" = "lightgrey")) + ggtitle("Counts below 5% threshold") + theme(legend.position="none") + labs(subtitle = paste0("Flagged cells = ", table(colData(spe)$flagged_tot_5)["TRUE"], " out of ", dim(spe)[2], ", ",  round(table(colData(spe)$flagged_tot_5)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells")) 
```



```{r, out.width="300%", out.height="200%"}
threshold <- 0.10
spe$flagged_det_10 <- as.factor(ifelse(colData(spe)$detected < quantile(colData(spe)$detected, probs = threshold), TRUE, FALSE))
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "flagged_det_10", 
            point_size = 0.1) + scale_color_manual(values = c("TRUE" = "black","FALSE" = "lightgrey")) + ggtitle("Detected features below 10% threshold") + theme(legend.position="none") + labs(subtitle = paste0("Flagged cells = ", table(colData(spe)$flagged_det_10)["TRUE"], " out of ", dim(spe)[2], ", ",  round(table(colData(spe)$flagged_det_10)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"))
```

```{r, out.width="150%", out.height="150%"}
threshold <- 0.05
spe$flagged_det_5 <- as.factor(ifelse(colData(spe)$detected < quantile(colData(spe)$detected, probs = threshold), TRUE, FALSE))
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "flagged_det_5", 
            point_size = 0.1) + scale_color_manual(values = c("TRUE" = "black","FALSE" = "lightgrey")) + ggtitle("Detected features below 5% threshold")  + theme(legend.position="none") + labs(subtitle = paste0("Flagged cells = ", table(colData(spe)$flagged_det_5)["TRUE"], " out of ", dim(spe)[2], ", ",  round(table(colData(spe)$flagged_det_5)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"))
```



```{r, out.width="150%", out.height="150%"}
threshold <- 0.10
spe$flagged_carea_10 <- as.factor(ifelse(colData(spe)$cell_area < quantile(colData(spe)$cell_area, probs = threshold), TRUE, FALSE))
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "flagged_carea_10", 
            point_size = 0.1) + scale_color_manual(values = c("TRUE" = "black","FALSE" = "lightgrey")) + ggtitle("Area in μm below 10%") + theme(legend.position="none") + labs(subtitle = paste0("Flagged cells = ", table(colData(spe)$flagged_carea_10)["TRUE"], " out of ", dim(spe)[2], ", ",  round(table(colData(spe)$flagged_carea_10)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"))
```

```{r, out.width="150%", out.height="150%"}
threshold <- 0.05
spe$flagged_carea_5 <- as.factor(ifelse(colData(spe)$cell_area < quantile(colData(spe)$cell_area, probs = threshold), TRUE, FALSE))
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "flagged_carea_5", 
            point_size = 0.1) + scale_color_manual(values = c("TRUE" = "black","FALSE" = "lightgrey")) + ggtitle("Area in μm below 5%")  + theme(legend.position="none") + labs(subtitle = paste0("Flagged cells = ", table(colData(spe)$flagged_carea_5)["TRUE"], " out of ", dim(spe)[2], ", ",  round(table(colData(spe)$flagged_carea_5)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"))
```


Cells above the selected threshold are highlighted in red, all the others are grey

```{r, out.width="150%", out.height="150%"}
threshold <- 0.90
spe$flagged_carea_90 <- as.factor(ifelse(colData(spe)$cell_area > quantile(colData(spe)$cell_area, probs = threshold), TRUE, FALSE))
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "flagged_carea_90", 
            point_size = 0.1) + scale_color_manual(values = c("TRUE" = "red","FALSE" = "lightgrey")) + ggtitle("Area in μm above 90% threshold") + theme(legend.position="none") + labs(subtitle = paste0("Flagged cells = ", table(colData(spe)$flagged_carea_90)["TRUE"], " out of ", dim(spe)[2], ", ",  round(table(colData(spe)$flagged_carea_90)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"))
```

```{r, out.width="150%", out.height="150%"}
threshold <- 0.95
spe$flagged_carea_95 <- as.factor(ifelse(colData(spe)$cell_area > quantile(colData(spe)$cell_area, probs = threshold), TRUE, FALSE))
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "flagged_carea_95", 
            point_size = 0.1) + scale_color_manual(values = c("TRUE" = "red","FALSE" = "lightgrey")) + ggtitle("Area in μm above 95% threshold")  + theme(legend.position="none") + labs(subtitle = paste0("Flagged cells = ", table(colData(spe)$flagged_carea_95)["TRUE"], " out of ", dim(spe)[2], ", ",  round(table(colData(spe)$flagged_carea_95)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"))
```



```{r, out.width="150%", out.height="150%"}
threshold <- 0.10
spe$flagged_narea_10 <- as.factor(ifelse(colData(spe)$nucleus_area < quantile(colData(spe)$nucleus_area, probs = threshold), TRUE, FALSE))
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "flagged_narea_10", 
            point_size = 0.1) + scale_color_manual(values = c("TRUE" = "black","FALSE" = "lightgrey")) + ggtitle("Area in μm below 10%") + theme(legend.position="none") + labs(subtitle = paste0("Flagged cells = ", table(colData(spe)$flagged_narea_10)["TRUE"], " out of ", dim(spe)[2], ", ",  round(table(colData(spe)$flagged_carea_10)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"))
```

```{r, out.width="150%", out.height="150%"}
threshold <- 0.05
spe$flagged_narea_5 <- as.factor(ifelse(colData(spe)$nucleus_area < quantile(colData(spe)$nucleus_area, probs = threshold), TRUE, FALSE))
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "flagged_narea_5", 
            point_size = 0.1) + scale_color_manual(values = c("TRUE" = "black","FALSE" = "lightgrey")) + ggtitle("Area in μm below 5%")  + theme(legend.position="none") + labs(subtitle = paste0("Flagged cells = ", table(colData(spe)$flagged_narea_5)["TRUE"], " out of ", dim(spe)[2], ", ",  round(table(colData(spe)$flagged_narea_5)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"))
```



```{r, out.width="150%", out.height="150%"}
threshold <- 0.90
spe$flagged_narea_90 <- as.factor(ifelse(colData(spe)$nucleus_area > quantile(colData(spe)$nucleus_area, probs = threshold), TRUE, FALSE))
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "flagged_narea_90", 
            point_size = 0.1) + scale_color_manual(values = c("TRUE" = "red","FALSE" = "lightgrey")) + ggtitle("Area in μm above 90% threshold") + theme(legend.position="none") + labs(subtitle = paste0("Flagged cells = ", table(colData(spe)$flagged_narea_90)["TRUE"], " out of ", dim(spe)[2], ", ",  round(table(colData(spe)$flagged_narea_90)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"))
```

```{r, out.width="150%", out.height="150%"}
threshold <- 0.95
spe$flagged_narea_95 <- as.factor(ifelse(colData(spe)$nucleus_area > quantile(colData(spe)$nucleus_area, probs = threshold), TRUE, FALSE))
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "flagged_narea_95", 
            point_size = 0.1) + scale_color_manual(values = c("TRUE" = "red","FALSE" = "lightgrey")) + ggtitle("Area in μm above 95% threshold")  + theme(legend.position="none") + labs(subtitle = paste0("Flagged cells = ", table(colData(spe)$flagged_narea_95)["TRUE"], " out of ", dim(spe)[2], ", ",  round(table(colData(spe)$flagged_narea_95)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"))
```



```{r, out.width="150%", out.height="150%"}
threshold <- 0.10
spe$flagged_aratio_10 <- as.factor(ifelse(colData(spe)$log2_AspectRatio < quantile(colData(spe)$log2_AspectRatio, probs = threshold), TRUE, FALSE))
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "flagged_aratio_10", 
            point_size = 0.1) + scale_color_manual(values = c("TRUE" = "black","FALSE" = "lightgrey")) + ggtitle("Width/height below 10% threshold") + theme(legend.position="none") + labs(subtitle = paste0("Flagged cells = ", table(colData(spe)$flagged_aratio_10)["TRUE"], " out of ", dim(spe)[2], ", ",  round(table(colData(spe)$flagged_aratio_10)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"))
```

```{r, out.width="150%", out.height="150%"}
threshold <- 0.05
spe$flagged_aratio_5 <- as.factor(ifelse(colData(spe)$log2_AspectRatio < quantile(colData(spe)$log2_AspectRatio, probs = threshold), TRUE, FALSE))
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "flagged_aratio_5", 
            point_size = 0.1) + scale_color_manual(values = c("TRUE" = "black","FALSE" = "lightgrey")) + ggtitle("Width/height below 5% threshold")  + theme(legend.position="none") + labs(subtitle = paste0("Flagged cells = ", table(colData(spe)$flagged_aratio_5)["TRUE"], " out of ", dim(spe)[2], ", ",  round(table(colData(spe)$flagged_aratio_5)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"))
```



```{r, out.width="150%", out.height="150%"}
threshold <- 0.90
spe$flagged_aratio_90 <- as.factor(ifelse(colData(spe)$log2_AspectRatio > quantile(colData(spe)$log2_AspectRatio, probs = threshold), TRUE, FALSE))
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "flagged_aratio_90", 
            point_size = 0.1) + scale_color_manual(values = c("TRUE" = "red","FALSE" = "lightgrey")) + ggtitle("Width/height above 90% threshold") + theme(legend.position="none") + labs(subtitle = paste0("Flagged cells = ", table(colData(spe)$flagged_aratio_90)["TRUE"], " out of ", dim(spe)[2], ", ",  round(table(colData(spe)$flagged_aratio_90)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"))
```

```{r, out.width="150%", out.height="150%"}
threshold <- 0.95
spe$flagged_aratio_95 <- as.factor(ifelse(colData(spe)$log2_AspectRatio > quantile(colData(spe)$log2_AspectRatio, probs = threshold), TRUE, FALSE))
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "flagged_aratio_95", 
            point_size = 0.1) + scale_color_manual(values = c("TRUE" = "red","FALSE" = "lightgrey")) + ggtitle("Width/height above 95% threshold")  + theme(legend.position="none") + labs(subtitle = paste0("Flagged cells = ", table(colData(spe)$flagged_aratio_95)["TRUE"], " out of ", dim(spe)[2], ", ",  round(table(colData(spe)$flagged_aratio_95)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"))
```



```{r, out.width="150%", out.height="150%"}
threshold <- 0.10
spe$flagged_tdratio_10 <- as.factor(ifelse(colData(spe)$tot_det_ratio < quantile(colData(spe)$tot_det_ratio, probs = threshold, na.rm = T), TRUE, FALSE))
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "flagged_tdratio_10", 
            point_size = 0.1) + scale_color_manual(values = c("TRUE" = "black","FALSE" = "lightgrey")) + ggtitle(paste0("Total counts/detected features below 10% threshold, ratio = ", round(quantile(colData(spe)$tot_det_ratio, probs = threshold, na.rm = T), 2))) + theme(legend.position="none") + labs(subtitle = paste0("Flagged cells = ", table(colData(spe)$flagged_tdratio_10)["TRUE"], " out of ", dim(spe)[2], ", ",  round(table(colData(spe)$flagged_tdratio_10)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"))
```

```{r, out.width="150%", out.height="150%"}
threshold <- 0.05
spe$flagged_tdratio_5 <- as.factor(ifelse(colData(spe)$tot_det_ratio < quantile(colData(spe)$tot_det_ratio, probs = threshold, na.rm = T), TRUE, FALSE))
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "flagged_tdratio_5", 
            point_size = 0.1) + scale_color_manual(values = c("TRUE" = "black","FALSE" = "lightgrey")) + ggtitle(paste0("Total counts/detected features below 5% threshold, ratio = ", round(quantile(colData(spe)$tot_det_ratio, probs = threshold, na.rm = T), 2)))  + theme(legend.position="none") + labs(subtitle = paste0("Flagged cells = ", table(colData(spe)$flagged_tdratio_5)["TRUE"], " out of ", dim(spe)[2], ", ",  round(table(colData(spe)$flagged_tdratio_5)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"))
```

ATOMX GUIDELINES FOR CELL QC

Count distribution 
flag cells where (total counts) / (number of detected genes) ≤1

```{r, out.width="150%", out.height="150%"}
threshold <- 1
spe$flagged_tdratio_atomx <- as.factor(ifelse(colData(spe)$tot_det_ratio <= threshold, TRUE, FALSE))
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "flagged_tdratio_atomx", 
            point_size = 0.1) + scale_color_manual(values = c("TRUE" = "black","FALSE" = "lightgrey")) + ggtitle("Total counts/detected feature ratio below 1 (default by AtoMx guidelines), ratio = 1") + theme(legend.position="none") + labs(subtitle = paste0("Flagged cells = ", table(colData(spe)$flagged_tdratio_atomx)["TRUE"], " out of ", dim(spe)[2], ", ",  round(table(colData(spe)$flagged_tdratio_atomx)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells")) 
```


```{r, out.width="150%", out.height="150%"}
threshold <- 0.50
spe$flagged_negprob_50 <- as.factor(ifelse(colData(spe)$subsets_NegProb_sum > quantile(colData(spe)$subsets_NegProb_sum, probs = threshold, na.rm = T), TRUE, FALSE))
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "flagged_negprob_50", 
            point_size = 0.1) + scale_color_manual(values = c("TRUE" = "red","FALSE" = "lightgrey")) + ggtitle("Negative probe counts above 50% threshold") + theme(legend.position="none") + labs(subtitle = paste0("Flagged cells = ", table(colData(spe)$flagged_negprob_50)["TRUE"], " out of ", dim(spe)[2], ", ",  round(table(colData(spe)$flagged_negprob_50)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"))
```

```{r, out.width="150%", out.height="150%"}
threshold <- 0.75
spe$flagged_negprob_75 <- as.factor(ifelse(colData(spe)$subsets_NegProb_sum > quantile(colData(spe)$subsets_NegProb_sum, probs = threshold, na.rm = T), TRUE, FALSE))
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "flagged_negprob_75", 
            point_size = 0.1) + scale_color_manual(values = c("TRUE" = "red","FALSE" = "lightgrey")) + ggtitle("Negative probe counts above 75% threshold")  + theme(legend.position="none") + labs(subtitle = paste0("Flagged cells = ", table(colData(spe)$flagged_negprob_75)["TRUE"], " out of ", dim(spe)[2], ", ",  round(table(colData(spe)$flagged_negprob_75)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"))
```

ATOMX GUIDELINES FOR CELL QC

Proportion of negative counts
flag cells where > 10% (0.1; default) of the counts are negative probes

```{r, out.width="150%", out.height="150%"}
spe$ctrl_total_ratio <- spe$subsets_NegProb_sum/spe$total
threshold <- 0.10
spe$flagged_negprob_10 <- as.factor(ifelse(colData(spe)$ctrl_total_ratio > quantile(colData(spe)$ctrl_total_ratio, probs = threshold, na.rm = T), TRUE, FALSE))
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "flagged_negprob_10", 
            point_size = 0.1) + scale_color_manual(values = c("TRUE" = "red","FALSE" = "lightgrey")) + ggtitle("Negative control probe/total counts ratio above 10% threshold (default by AtoMx guidelines)") + theme(legend.position="none") + labs(subtitle = paste0("Flagged cells = ", table(colData(spe)$flagged_negprob_10)["TRUE"], " out of ", dim(spe)[2], ", ", round(table(colData(spe)$flagged_negprob_10)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells")) 
```


# Feature heatmaps by FOV 

To compare at a glance FOV mean/median feature values and spot anomalous FOVs

```{r}
meta_df <- data.frame(colData(spe))
meta_df <- group_by(meta_df, mock_fov) |> summarize(n_cells = n(), mean_total = mean(total), mean_detected = mean(detected), mean_carea = mean(cell_area), mean_narea = mean(nucleus_area), mean_log2_AspectRatio = mean(log2_AspectRatio), mean_tot_det_ratio = mean(tot_det_ratio), mean_control_counts = mean(control_counts), median_total = median(total), median_detected = median(detected), median_carea = median(cell_area), median_narea = median(nucleus_area), median_log2_AspectRatio = median(log2_AspectRatio), median_tot_det_ratio = median(tot_det_ratio), median_control_counts = median(control_counts))

#fov_positions <- left_join(fov_positions, meta_df, by = join_by(fov))

#poly_df <- transmute(fov_positions, x_1 = x_global_px, y_1 = y_global_px, 
                     # x_2 = x_global_px + fov_size, y_2 = y_global_px, 
                     # x_3 = x_global_px + fov_size, y_3 = y_global_px + fov_size, 
                     # x_4 = x_global_px, y_4 = y_global_px + fov_size)

#str(poly_df)

#ls_list <- list()
#for (k in 1:dim(fov_positions)[1]){
  #ls_list[[k]] <- st_linestring(matrix(as.numeric(poly_df[k]), ncol = 2, byrow = T))
#}

#sf_data <- st_as_sf(fov_positions, coords = c("x_global_px", "y_global_px"), geometry = st_sfc(ls_list))

#sf_data <- st_cast(sf_data, "POLYGON")

# plot(sf_data)
#names(sf_data)
#sf_data$fov <- as.factor(sf_data$fov)
```

```{r, out.width="150%", out.height="150%"}
#ggplot() +
  #geom_sf(data = sf_data, aes(fill = n_cells)) + 
  #scale_fill_gradient(low = "purple4", high = "#FF8247") +
  #theme(axis.title.x = element_blank(),
        #axis.text.x=element_blank(),
        #axis.ticks.x=element_blank(), 
        #axis.title.y = element_blank(),
        #axis.text.y=element_blank(),
        #axis.ticks.y=element_blank(),
        #panel.grid = element_blank()) +
  #geom_text(data = fov_positions, aes(x = fov_positions$x_global_px+fov_size/2,
                #y = fov_positions$y_global_px+fov_size/2,
                #label = fov_positions$n_cells), color="white", size = 1.8) +
    #ggtitle(paste0("Number of cells by fov"))
```

```{r, out.width="150%", out.height="150%"}
#heatmap_variables <- names(sf_data)[-c(1, 2, length(names(sf_data)), length(names(sf_data))-1)]
#for (var in 1:length(heatmap_variables)){
  #heat_var <- heatmap_variables[var]
  #p <- ggplot() +
  #geom_sf(data = sf_data, aes(fill = .data[[heat_var]])) + 
  #scale_fill_viridis_c() +
  #theme(axis.title.x = element_blank(),
        #axis.text.x=element_blank(),
        #axis.ticks.x=element_blank(), 
        #axis.title.y = element_blank(),
        #axis.text.y=element_blank(),
        #axis.ticks.y=element_blank(),
        #panel.grid = element_blank()) +
  #geom_text(data = fov_positions, aes(x = fov_positions$x_global_px+fov_size/2,
                #y = fov_positions$y_global_px+fov_size/2,
                #label = fov_positions$fov), color="white", size = 2) +
    #ggtitle(paste0(heat_var, " by fov"))
  #print(p)
#}  
```

ATOMX GUIDELINES FOR FOV QC

The Mean method (default) flags FOVs where the total count per cell average is below a threshold (default 100)

```{r, out.width="150%", out.height="150%"}
#color <- ifelse(fov_positions$mean_total <= 100, "red", "white")
#fov_positions <- cbind(fov_positions, color)
#ggplot() +
  #geom_sf(data = sf_data, aes(fill = mean_total)) + 
  #scale_fill_viridis_c() +
  #theme(axis.title.x = element_blank(),
        #axis.text.x=element_blank(),
        #axis.ticks.x=element_blank(), 
        #axis.title.y = element_blank(),
        #axis.text.y=element_blank(),
        #axis.ticks.y=element_blank(),
        #panel.grid = element_blank()) +
  #geom_text(data = fov_positions, aes(x = fov_positions$x_global_px+fov_size/2,
                #y = fov_positions$y_global_px+fov_size/2,
                #label = floor(fov_positions$mean_total)), color = color, size = 2) +
    #ggtitle(paste0("Mean_total by fov"))
```

ATOMX GUIDELINES FOR FOV QC

The QC quantile must be greater than the negative probe median in order to pass QC. Default is 0, so it must be ≥ 0

```{r, out.width="150%", out.height="150%"}
# threshold <- 0.90
# quant_df <-  data.frame(colData(spe))
# quant_df <- mutate(quant_df, target_counts = total - subsets_NegProb_sum)
# quant_df <- group_by(quant_df, fov) |> summarize(quantile_90 = quantile(target_counts, probs = threshold))
# fov_positions <- left_join(fov_positions, quant_df, by = join_by(fov))

# color <- ifelse(fov_positions$quantile_90 < fov_positions$median_subsets_NegProb_sum, "red", "white")
# sf_data$quantile_90 <- fov_positions$quantile_90

# ggplot() +
#   geom_sf(data = sf_data, aes(fill = quantile_90)) + 
  # scale_fill_viridis_c() +
  # theme(axis.title.x = element_blank(),
  #       axis.text.x=element_blank(),
  #       axis.ticks.x=element_blank(), 
  #       axis.title.y = element_blank(),
  #       axis.text.y=element_blank(),
  #       axis.ticks.y=element_blank(),
  #       panel.grid = element_blank()) +
  # geom_text(data = fov_positions, aes(x = fov_positions$x_global_px+fov_size/2,
  #               y = fov_positions$y_global_px+fov_size/2,
  #               label = floor(fov_positions$quantile_90)), color = color, size = 1.8) +
  # ggtitle(paste0("FOV QC based on 90th quantile"))
```


# Feature boxplot

```{r, warnings = FALSE}
meta_df <- data.frame(colData(spe))
variables <- c("total", "detected", "logtot", "cell_area", "nucleus_area", "log2_AspectRatio", "tot_det_ratio", "control_counts")

for (var in 1:length(variables)){
  n <- grep(paste0("^", variables[var], "$"), names(colData(spe)))
  boxplot(meta_df[,n] ~ spe$mock_fov, ylab = names(colData(spe))[n], cex = 0.2)
  title(paste0(colnames(spe@colData)[n]))
}
```

# Cell counts barplot by FOV

To view how many cells lie inside each FOV

```{r}
#fov_df <- data.frame("n_cells" = as.integer(table(spe@colData@listData$fov)), "fov" = as.factor(names(table(spe@colData@listData$fov))))
#levels(fov_df$fov) <- order(levels(fov_df$fov), seq(1:length(fov_df$fov)))

#ggplot(data = fov_df, aes(x = fov, y = n_cells)) +
  #geom_col(fill = "#B0C4DE") + theme(legend.position ="none") +
  #ggtitle("Number of cells per fov") 
```

# Final report

```{r echo=T}
data <- data.frame(colData(spe))

df <- format(t(data.frame("Total_cell_number" = dim(spe)[2], "Total_counts" = sum(spe$total), "Total_features" = sum(spe$detected), "Total_ctrl_probe_counts" = sum(spe$control_counts), "Median_counts_per_cell" = median(spe$total), "Median_features_per_cell" = median(spe$detected),  "Median_ctrl_probe_counts" = median(spe$control_counts), "Median_um_cell_area" = median(spe$cell_area), "Median_um_nucleus_area" = median(spe$nucleus_area), "Median_log2_aspect_ratio" = median(spe$log2_AspectRatio), "Median_total_counts/features_ratio" = median(spe$tot_det_ratio, na.rm = T),  "Median_ctrl_probe_counts/total_counts_ratio" = median(spe$ctrl_total_ratio, na.rm = T))), scientific = FALSE)

rownames(df) <- c("Total cell number", "Total counts", "Total features", "Total ctrl probe counts", "Median counts per cell", "Median features per cell", "Median ctrl probe counts", "Median um cell area", "Median um nucleus area", "Median log2 aspect ratio", "Median total counts/features ratio", "Median ctrl probe counts/total counts ratio")

df[,1] <- round(as.numeric(df[,1]), 2)

kable(df, format = "html", row.names = T, col.names = NULL, align = "c", caption = "<center>Global summary metrics</center>",
      padding = 1L)
```

#
#

```{r}
flag_variables <- c("flagged_tot_20", 
                    "flagged_tot_10",
                    "flagged_tot_5",
                    "flagged_det_10",
                    "flagged_det_5",
                    "flagged_carea_10",
                    "flagged_carea_5",
                    "flagged_carea_90",
                    "flagged_carea_95",
                    "flagged_narea_10",
                    "flagged_narea_5",
                    "flagged_narea_90",
                    "flagged_narea_95",
                    "flagged_aratio_10",
                    "flagged_aratio_5",
                    "flagged_aratio_90",
                    "flagged_aratio_95",
                    "flagged_tdratio_10",
                    "flagged_tdratio_5",
                    "flagged_tdratio_atomx",
                    "flagged_negprob_50",
                    "flagged_negprob_75",
                    "ctrl_total_ratio")

flag_data <- data.frame(Unflagged_cells = as.integer(), Flagged_cells = as.integer(), Percentage_flagged = as.numeric())
for(i in 1:length(flag_variables)){
  var <- grep(paste0("^", flag_variables[i],"$"), names(colData(spe)))
  list_i <- c(sum(colData(spe)[,var]==FALSE, na.rm = T), sum(colData(spe)[,var]==TRUE, na.rm = T), round(sum(colData(spe)[,var]==TRUE, na.rm = T)/dim(spe)[2]*100, 2))
  flag_data[i, ] <- list_i
}

row.names(flag_data) <- c("Total counts < 20%", " Total counts < 10%", "Total counts < 5%", "Features < 10%", "Features < 5%", "Cell area < 10%", "Cell area < 5%", 
                         "Cell area > 90%","Cell area > 95%", "Nucleus area < 10%", "Nucleus area < 5%", "Nucleus area > 90%","Nucleus area > 95%", "H/w < 10%", 
                         "H/w < 5%", "H/w > 90%", "H/w > 95%", "Counts/features < 10%", "Counts/features < 5%",
                         "Counts/features < 1", "Ctrl counts > 50%", "Ctrl counts > 75%", "Ctrl/total > 10%")

kable(flag_data, format = "html", padding = 1L, align = "c", col.names = c("Unflagged cells", "Flagged cells", "% flagged cells"), caption = paste0("<center>Flagged vs unflagged cells, total cells: " , dim(spe)[2],"</center>"))
```

#
#

ATOMX GUIDELINES FOR TARGET LEVEL QC

A value of 0.5 (default) will flag probes with lower total counts than the median (50th percentile) of thenegative control probes' counts

```{r}
# target_counts <- counts(spe)[]
# count_data <- rowSums(counts(spe))
# color <- ifelse(count_data < median(spe$control_counts), "red", "black") 
# color <- ifelse(match(names(color), rownames(target_counts)), color, "grey")
# color <- replace_na(color, "grey") 
# count_df <- data_frame(count_data, color)
# count_df$log_count <- log10(count_df$count_data)
# plot(count_df$log_count, col = count_df$color)
```


# Flag score development

Starting from feature scatterplots to see the relationship between possible candidate features

Total counts ~ logged width/height ratio

```{r, out.width="150%", out.height="150%"}
cell_df <- data.frame(colData(spe))

p1 <- ggplot(data = cell_df) +
  geom_pointdensity(aes(x = log2_AspectRatio, y = total)) +
  scale_color_viridis_c() +
  theme(legend.position = "none",
        plot.margin = margin())

p2 <- ggdensity(data = cell_df, x = "total", color = "#ccccff", 
                fill = "#ccccff", alpha = 0.5) + theme_classic() + 
  theme(legend.position = "none", axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line.y = element_blank(),
    axis.line.x = element_blank(),
    plot.margin = margin()) + coord_flip()

p3 <- ggdensity(data = cell_df, x = "log2_AspectRatio", color = "#ccccff", 
                fill = "#ccccff", alpha = 0.5) + theme_classic() + 
  theme(legend.position = "none", axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.line.x = element_blank(),
    axis.line.y = element_blank(),
    plot.margin = margin()) 

plot_grid(p3, NULL, p1, p2, ncol = 2, align = "hv",
          rel_widths = c(2, 1), rel_heights = c(1, 2))
```

Detected features ~ logged width/height ratio

```{r, out.width="150%", out.height="150%"}
p1 <- ggplot(data = cell_df) +
  geom_pointdensity(aes(x = log2_AspectRatio, y = detected)) +
  scale_color_viridis_c() +
  theme(legend.position = "none",
        plot.margin = margin())

p2 <- ggdensity(data = cell_df, x = "detected", color = "#ccccff", 
                fill = "#ccccff", alpha = 0.5) + theme_classic() + 
  theme(legend.position = "none", axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line.y = element_blank(),
    axis.line.x = element_blank(),
    plot.margin = margin()) + coord_flip()

p3 <- ggdensity(data = cell_df, x = "log2_AspectRatio", color = "#ccccff", 
                fill = "#ccccff", alpha = 0.5) + theme_classic() + 
  theme(legend.position = "none", axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.line.x = element_blank(),
    axis.line.y = element_blank(),
    plot.margin = margin()) 

plot_grid(p3, NULL, p1, p2, ncol = 2, align = "hv",
          rel_widths = c(2, 1), rel_heights = c(1, 2))
```

Total counts ~ area in um

```{r, out.width="150%", out.height="150%"}
p1 <- ggplot(data = cell_df) +
  geom_pointdensity(aes(x = cell_area, y = total)) +
  scale_color_viridis_c() +
  theme(legend.position = "none",
        plot.margin = margin())

p2 <- ggdensity(data = cell_df, x = "total", color = "#ccccff", 
                fill = "#ccccff", alpha = 0.5) + theme_classic() + 
  theme(legend.position = "none", axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line.y = element_blank(),
    axis.line.x = element_blank(),
    plot.margin = margin()) + coord_flip()

p3 <- ggdensity(data = cell_df, x = "cell_area", color = "#ccccff", 
                fill = "#ccccff", alpha = 0.5) + theme_classic() + 
  theme(legend.position = "none", axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.line.x = element_blank(),
    axis.line.y = element_blank(),
    plot.margin = margin()) 

plot_grid(p3, NULL, p1, p2, ncol = 2, align = "hv",
          rel_widths = c(2,1), rel_heights = c(1, 2))
```

Total counts ~ nucleus area

```{r, out.width="150%", out.height="150%"}
p1 <- ggplot(data = cell_df) +
  geom_pointdensity(aes(x = nucleus_area, y = total)) +
  scale_color_viridis_c() +
  theme(legend.position = "none",
        plot.margin = margin())

p2 <- ggdensity(data = cell_df, x = "total", color = "#ccccff", 
                fill = "#ccccff", alpha = 0.5) + theme_classic() + 
  theme(legend.position = "none", axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line.y = element_blank(),
    axis.line.x = element_blank(),
    plot.margin = margin()) + coord_flip()

p3 <- ggdensity(data = cell_df, x = "nucleus_area", color = "#ccccff", 
                fill = "#ccccff", alpha = 0.5) + theme_classic() + 
  theme(legend.position = "none", axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.line.x = element_blank(),
    axis.line.y = element_blank(),
    plot.margin = margin()) 

plot_grid(p3, NULL, p1, p2, ncol = 2, align = "hv",
          rel_widths = c(2,1), rel_heights = c(1, 2))
```

Logged width/height ratio ~ negative control probe counts

```{r, out.width="150%", out.height="150%"}
p1 <- ggplot(data = cell_df) +
  geom_pointdensity(aes(x = control_counts, y = log2_AspectRatio)) +
  scale_color_viridis_c() +
  theme(legend.position = "none",
        plot.margin = margin())

p2 <- ggdensity(data = cell_df, x = "log2_AspectRatio", color = "#ccccff", 
                fill = "#ccccff", alpha = 0.5) + theme_classic() + 
  theme(legend.position = "none", axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line.y = element_blank(),
    axis.line.x = element_blank(),
    plot.margin = margin()) + coord_flip()

p3 <- ggdensity(data = cell_df, x = "control_counts", color = "#ccccff", 
                fill = "#ccccff", alpha = 0.5) + theme_classic() + 
  theme(legend.position = "none", axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.line.x = element_blank(),
    axis.line.y = element_blank(),
    plot.margin = margin()) 

plot_grid(p3, NULL, p1, p2, ncol = 2, align = "hv",
          rel_widths = c(2,1), rel_heights = c(1, 2))
```

Logged width/height ratio ~ local coordinates on X axis in px

```{r, out.width="150%", out.height="150%"}
p1 <- ggplot(data = cell_df) +
  geom_pointdensity(aes(x = CenterX_global_um, y = log2_AspectRatio)) +
  scale_color_viridis_c() +
  theme(legend.position = "none",
        plot.margin = margin())

p2 <- ggdensity(data = cell_df, x = "log2_AspectRatio", color = "#ccccff", 
                fill = "#ccccff", alpha = 0.5) + theme_classic() + 
  theme(legend.position = "none", axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line.y = element_blank(),
    axis.line.x = element_blank(),
    plot.margin = margin()) + coord_flip()

p3 <- ggdensity(data = cell_df, x = "CenterX_global_um", color = "#ccccff", 
                fill = "#ccccff", alpha = 0.5) + theme_classic() + 
  theme(legend.position = "none", axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.line.x = element_blank(),
    axis.line.y = element_blank(),
    plot.margin = margin()) 

plot_grid(p3, NULL, p1, p2, ncol = 2, align = "hv",
          rel_widths = c(2,1), rel_heights = c(1, 2))
```

Logged width/height ratio ~ local coordinates on X axis in px

```{r, out.width="150%", out.height="150%"}
p1 <- ggplot(data = cell_df) +
  geom_pointdensity(aes(x = CenterY_global_um, y = log2_AspectRatio)) +
  scale_color_viridis_c() +
  theme(legend.position = "none",
        plot.margin = margin())

p2 <- ggdensity(data = cell_df, x = "log2_AspectRatio", color = "#ccccff", 
                fill = "#ccccff", alpha = 0.5) + theme_classic() + 
  theme(legend.position = "none", axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line.y = element_blank(),
    axis.line.x = element_blank(),
    plot.margin = margin()) + coord_flip()

p3 <- ggdensity(data = cell_df, x = "CenterY_global_um", color = "#ccccff", 
                fill = "#ccccff", alpha = 0.5) + theme_classic() + 
  theme(legend.position = "none", axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.line.x = element_blank(),
    axis.line.y = element_blank(),
    plot.margin = margin()) 

plot_grid(p3, NULL, p1, p2, ncol = 2, align = "hv",
          rel_widths = c(2,1), rel_heights = c(1, 2))
```

# Flag score formula

```{r}
spe$flag_score <- 1/(1 + exp(-0.3*spe$target_counts/spe$nucleus_area + 0.8*abs(spe$log2_AspectRatio)))
summary(spe$flag_score)
plot(spe$flag_score)
```

```{r}
hist(spe$flag_score)
```


```{r, out.width="300%", out.height="150%"}
polygons$total <- spe$total
polygons$target_counts <- spe$target_counts
polygons$cell_area <- spe$cell_area
polygons$nucleus_area <- spe$nucleus_area
polygons$log2_AspectRatio <- spe$log2_AspectRatio
polygons$flag_score <- spe$flag_score
polygons$mock_fov <- spe$mock_fov

tm_shape(polygons) + 
  tm_fill(col= "flag_score", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE)
```

Local threshold map of flag score, target counts, cell_area and logged width/height ratio, 20th percentile

```{r, out.width="300%", out.height="300%"}
metric_color <- ifelse(polygons$flag_score < quantile(spe$flag_score, probs = 0.15), "red", "white")
target_color <- ifelse(polygons$target_counts < quantile(spe$target_counts, probs = 0.15), "red", "white")
area_color <- ifelse(polygons$cell_area < quantile(spe$cell_area, probs = 0.15), "red", "white")
nucleus_color <- ifelse(polygons$nucleus_area < quantile(spe$nucleus_area, probs = 0.15), "red", "white")
count_area_ratio_color <- ifelse((polygons$target_counts/polygons$cell_area) < quantile((spe$target_counts/spe$cell_area), probs = 0.15), "red", "white")
count_nucleus_ratio_color <- ifelse((polygons$target_counts/polygons$nucleus_area) < quantile((spe$target_counts/spe$nucleus_area), probs = 0.15), "red", "white")
log_hw_color_lo <- ifelse(polygons$log2_AspectRatio < quantile(spe$log2_AspectRatio, probs = 0.15), "red", "white")
log_hw_color_hi <- ifelse(polygons$log2_AspectRatio > quantile(spe$log2_AspectRatio, probs = 0.85), "red", "white")

polygons_whole <- cbind(polygons, metric_color = metric_color, target_color = target_color, area_color = area_color, nucleus_color = nucleus_color, count_area_ratio_color = count_area_ratio_color, count_nucleus_ratio_color = count_nucleus_ratio_color, log_hw_color_lo = log_hw_color_lo, log_hw_color_hi = log_hw_color_hi)

p1 <- tm_shape(polygons_whole) + 
  tm_fill(col= "metric_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Flag score",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))


p2 <- tm_shape(polygons_whole) + 
  tm_fill(col= "target_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Target counts",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p3 <- tm_shape(polygons_whole) + 
  tm_fill(col= "area_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "cell area",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p4 <- tm_shape(polygons_whole) + 
  tm_fill(col= "nucleus_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "nucleus area",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p5 <- tm_shape(polygons_whole) + 
  tm_fill(col= "count_area_ratio_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "counts/um area",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p6 <- tm_shape(polygons_whole) + 
  tm_fill(col= "count_nucleus_ratio_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "counts/nucleus area",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p7 <- tm_shape(polygons_whole) + 
  tm_fill(col= "log_hw_color_lo", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Logged AspectRatio below",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p8 <- tm_shape(polygons_whole) + 
  tm_fill(col= "log_hw_color_hi", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Logged AspectRatio above",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p9 <- tm_shape(polygons_whole) + 
tm_fill(col = "celltype_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "celltype",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))


tmap_arrange(p1, p2, p3, p4, p5, p6, p7, p8, p9, ncol = 3)
```


```{r, out.width="150%", out.height="150%"}
plot_list <- list(p1, p2, p3, p4, p5, p6, p7, p8, p9 + 
                    tm_layout(frame = FALSE,
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0)) +
tm_add_legend(type = "fill", labels = names(celltype_palette), col = celltype_palette))

for(i in 1:length(plot_list)){
  print(plot_list[[i]])
}
```


```{r, out.width="150%", out.height="150%"}

flag_value <- c(round(quantile(polygons_whole$flag_score, probs = 0.15), 2),
                round(quantile(polygons_whole$target_counts, probs = 0.15), 2),
                round(quantile(polygons_whole$cell_area, probs = 0.15), 2),
                round(quantile(polygons_whole$nucleus_area, probs = 0.15), 2),
                round(quantile((polygons_whole$target_counts/polygons$cell_area), probs = 0.15), 2),
                round(quantile((polygons_whole$target_counts/polygons$nucleus_area), probs = 0.15), 2),
                round(quantile(polygons_whole$log2_AspectRatio, probs = 0.15), 2),
                round(quantile(polygons_whole$log2_AspectRatio, probs = 0.85), 2))

flagged_cells <- c(sum(polygons_whole$metric_color=="red", na.rm = T),
                   sum(polygons_whole$target_color=="red", na.rm = T),
                   sum(polygons_whole$area_color=="red", na.rm = T),
                   sum(polygons_whole$nucleus_color=="red", na.rm = T),
                   sum(polygons_whole$count_area_ratio_color=="red", na.rm = T),
                   sum(polygons_whole$count_nucleus_ratio_color=="red", na.rm = T),
                   sum(polygons_whole$log_hw_color_lo=="red", na.rm = T),
                   sum(polygons_whole$log_hw_color_hi=="red", na.rm = T))

unflagged_cells <- c(sum(polygons_whole$metric_color=="white", na.rm = T),
                     sum(polygons_whole$target_color=="white", na.rm = T),
                     sum(polygons_whole$area_color=="white", na.rm = T),
                     sum(polygons_whole$nucleus_color=="white", na.rm = T),
                     sum(polygons_whole$count_area_ratio_color=="white", na.rm = T),
                     sum(polygons_whole$count_nucleus_ratio_color=="white", na.rm = T),
                     sum(polygons_whole$log_hw_color_lo=="white", na.rm = T),
                     sum(polygons_whole$log_hw_color_hi=="white", na.rm = T))

percent_flagged <- round(flagged_cells/dim(polygons_whole)[1]*100, 2)

poly <- data.frame(threshold = c(0.15, 0.15, 0.15, 0.15, 0.15, 0.15, 0.15, 0.85), flag_value = flag_value, unflagged_cells = unflagged_cells, flagged_cells = flagged_cells, percent_flagged = percent_flagged, row.names = c("Flag score", "Target counts", "Cell area in um", "Nucleus area in um", "Target counts/cell area", "Target counts/nucleus area", "Logged aspect ratio low", "Logged aspect ratio high"))

kable(poly, format = "html", col.names = c("Threshold", "Flag value", "Unflagged cells", "Flagged cells", "% flagged cells"), align = "c", caption = paste0("<center>Flagged cells by different metrics, total cells in fov: ", dim(polygons_whole)[1], ", total cells in whole sample: ", dim(polygons_whole)[1], ", flagged cells in whole sample: ", sum(spe$flag_score < quantile(spe$flag_score, probs = 0.15, na.rm = T)), "</center>"), padding = 1L)
```

# Cells on borders 

```{r}
polygons_1 <- polygons[1:14250,]
tm_shape(polygons_1) + 
  tm_fill(col= "flag_score", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE)
```

Local threshold map of flag score, target counts, cell_area and logged width/height ratio, 20th percentile

```{r, out.width="300%", out.height="300%"}
metric_color <- ifelse(polygons_1$flag_score < quantile(spe$flag_score, probs = 0.15), "red", "white")
target_color <- ifelse(polygons_1$target_counts < quantile(spe$target_counts, probs = 0.15), "red", "white")
area_color <- ifelse(polygons_1$cell_area < quantile(spe$cell_area, probs = 0.15), "red", "white")
nucleus_color <- ifelse(polygons_1$nucleus_area < quantile(spe$nucleus_area, probs = 0.15), "red", "white")
count_area_ratio_color <- ifelse((polygons_1$target_counts/polygons_1$cell_area) < quantile((spe$target_counts/spe$cell_area), probs = 0.15), "red", "white")
count_nucleus_ratio_color <- ifelse((polygons_1$target_counts/polygons_1$nucleus_area) < quantile((spe$target_counts/spe$nucleus_area), probs = 0.15), "red", "white")
log_hw_color_lo <- ifelse(polygons_1$log2_AspectRatio < quantile(spe$log2_AspectRatio, probs = 0.15), "red", "white")
log_hw_color_hi <- ifelse(polygons_1$log2_AspectRatio > quantile(spe$log2_AspectRatio, probs = 0.85), "red", "white")

polygons_1 <- cbind(polygons_1, metric_color = metric_color, target_color = target_color, area_color = area_color, nucleus_color = nucleus_color, count_area_ratio_color = count_area_ratio_color, count_nucleus_ratio_color = count_nucleus_ratio_color, log_hw_color_lo = log_hw_color_lo, log_hw_color_hi = log_hw_color_hi)

p1 <- tm_shape(polygons_1) + 
  tm_fill(col= "metric_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Flag score",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))


p2 <- tm_shape(polygons_1) + 
  tm_fill(col= "target_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Target counts",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p3 <- tm_shape(polygons_1) + 
  tm_fill(col= "area_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "cell area",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p4 <- tm_shape(polygons_1) + 
  tm_fill(col= "nucleus_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "nucleus area",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p5 <- tm_shape(polygons_1) + 
  tm_fill(col= "count_area_ratio_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "counts/cell area",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p6 <- tm_shape(polygons_1) + 
  tm_fill(col= "count_nucleus_ratio_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "counts/nucleus area",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p7 <- tm_shape(polygons_1) + 
  tm_fill(col= "log_hw_color_lo", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Logged AspectRatio below",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p8 <- tm_shape(polygons_1) + 
  tm_fill(col= "log_hw_color_hi", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Logged AspectRatio above",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p9 <- tm_shape(polygons_1) + 
tm_fill(col = "celltype_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "celltype",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))


tmap_arrange(p1, p2, p3, p4, p5, p6, p7, p8, p9, ncol = 3)
```

```{r, out.width="150%", out.height="150%"}
plot_list <- list(p1, p2, p3, p4, p5, p6, p7, p8, p9 + 
                    tm_layout(frame = FALSE,
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0)) +
tm_add_legend(type = "fill", labels = names(celltype_palette), col = celltype_palette))

for(i in 1:length(plot_list)){
  print(plot_list[[i]])
}
```

```{r, out.width="150%", out.height="150%"}

flag_value <- c(round(quantile(polygons_1$flag_score, probs = 0.15), 2),
                round(quantile(polygons_1$target_counts, probs = 0.15), 2),
                round(quantile(polygons_1$cell_area, probs = 0.15), 2),
                round(quantile(polygons_1$nucleus_area, probs = 0.15), 2),
                round(quantile((polygons_1$target_counts/polygons_1$cell_area), probs = 0.15), 2),
                round(quantile((polygons_1$target_counts/polygons_1$nucleus_area), probs = 0.15), 2),
                round(quantile(polygons_1$log2_AspectRatio, probs = 0.15), 2),
                round(quantile(polygons_1$log2_AspectRatio, probs = 0.85), 2))

flagged_cells <- c(sum(polygons_1$metric_color=="red", na.rm = T),
                   sum(polygons_1$target_color=="red", na.rm = T),
                   sum(polygons_1$area_color=="red", na.rm = T),
                   sum(polygons_1$nucleus_color=="red", na.rm = T),
                   sum(polygons_1$count_area_ratio_color=="red", na.rm = T),
                   sum(polygons_1$count_nucleus_ratio_color=="red", na.rm = T),
                   sum(polygons_1$log_hw_color_lo=="red", na.rm = T),
                   sum(polygons_1$log_hw_color_hi=="red", na.rm = T))

unflagged_cells <- c(sum(polygons_1$metric_color=="white", na.rm = T),
                     sum(polygons_1$target_color=="white", na.rm = T),
                     sum(polygons_1$area_color=="white", na.rm = T),
                     sum(polygons_1$nucleus_color=="white", na.rm = T),
                     sum(polygons_1$count_area_ratio_color=="white", na.rm = T),
                     sum(polygons_1$count_nucleus_ratio_color=="white", na.rm = T),
                     sum(polygons_1$log_hw_color_lo=="white", na.rm = T),
                     sum(polygons_1$log_hw_color_hi=="white", na.rm = T))

percent_flagged <- round(flagged_cells/dim(polygons_1)[1]*100, 2)

poly_1 <- data.frame(threshold = c(0.15, 0.15, 0.15, 0.15, 0.15, 0.15, 0.15, 0.85), flag_value = flag_value, unflagged_cells = unflagged_cells, flagged_cells = flagged_cells, percent_flagged = percent_flagged, row.names = c("Flag score", "Target counts", "Cell area in um", "Nucleus area in um", "Target counts/cell area", "Target counts/nucleus area", "Logged aspect ratio low", "Logged aspect ratio high"))

kable(poly_1, format = "html", col.names = c("Threshold", "Flag value", "Unflagged cells", "Flagged cells", "% flagged cells"), align = "c", caption = paste0("<center>Flagged cells by different metrics, total cells in fov: ", dim(polygons_1)[1], ", total cells in whole sample: ", dim(polygons_1)[1], ", flagged cells in whole sample: ", sum(spe$flag_score < quantile(spe$flag_score, probs = 0.15, na.rm = T)), "</center>"), padding = 1L)
```

# Cells of mock fov 1A

```{r}
polygons_1A <- subset(polygons, mock_fov == "1 A")

tm_shape(polygons_1A) + 
  tm_fill(col= "flag_score", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE)
```

Local threshold map of flag score, target counts, cell_area and logged width/height ratio, 20th percentile

```{r, out.width="300%", out.height="300%"}
metric_color <- ifelse(polygons_1A$flag_score < quantile(spe$flag_score, probs = 0.15), "red", "white")
target_color <- ifelse(polygons_1A$target_counts < quantile(spe$target_counts, probs = 0.15), "red", "white")
area_color <- ifelse(polygons_1A$cell_area < quantile(spe$cell_area, probs = 0.15), "red", "white")
nucleus_color <- ifelse(polygons_1A$nucleus_area < quantile(spe$nucleus_area, probs = 0.15), "red", "white")
count_area_ratio_color <- ifelse((polygons_1A$target_counts/polygons_1A$cell_area) < quantile((spe$target_counts/spe$cell_area), probs = 0.15), "red", "white")
count_nucleus_ratio_color <- ifelse((polygons_1A$target_counts/polygons_1A$nucleus_area) < quantile((spe$target_counts/spe$nucleus_area), probs = 0.15), "red", "white")
log_hw_color_lo <- ifelse(polygons_1A$log2_AspectRatio < quantile(spe$log2_AspectRatio, probs = 0.15), "red", "white")
log_hw_color_hi <- ifelse(polygons_1A$log2_AspectRatio > quantile(spe$log2_AspectRatio, probs = 0.85), "red", "white")

polygons_1A <- cbind(polygons_1A, metric_color = metric_color, target_color = target_color, area_color = area_color, nucleus_color = nucleus_color, count_area_ratio_color = count_area_ratio_color, count_nucleus_ratio_color = count_nucleus_ratio_color, log_hw_color_lo = log_hw_color_lo, log_hw_color_hi = log_hw_color_hi)

p1 <- tm_shape(polygons_1A) + 
  tm_fill(col= "metric_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Flag score",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))


p2 <- tm_shape(polygons_1A) + 
  tm_fill(col= "target_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Target counts",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p3 <- tm_shape(polygons_1A) + 
  tm_fill(col= "area_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "cell area",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p4 <- tm_shape(polygons_1A) + 
  tm_fill(col= "nucleus_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "nucleus area",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p5 <- tm_shape(polygons_1A) + 
  tm_fill(col= "count_area_ratio_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "counts/cell area",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p6 <- tm_shape(polygons_1A) + 
  tm_fill(col= "count_nucleus_ratio_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "counts/nucleus area",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p7 <- tm_shape(polygons_1A) + 
  tm_fill(col= "log_hw_color_lo", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Logged AspectRatio below",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p8 <- tm_shape(polygons_1A) + 
  tm_fill(col= "log_hw_color_hi", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Logged AspectRatio above",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p9 <- tm_shape(polygons_1A) + 
tm_fill(col = "celltype_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "celltype",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))


tmap_arrange(p1, p2, p3, p4, p5, p6, p7, p8, p9, ncol = 3)
```

```{r, out.width="150%", out.height="150%"}
plot_list <- list(p1, p2, p3, p4, p5, p6, p7, p8, p9 + 
                    tm_layout(frame = FALSE,
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0)) +
tm_add_legend(type = "fill", labels = names(celltype_palette), col = celltype_palette))

for(i in 1:length(plot_list)){
  print(plot_list[[i]])
}
```


```{r, out.width="150%", out.height="150%"}

flag_value <- c(round(quantile(polygons_1A$flag_score, probs = 0.15), 2),
                round(quantile(polygons_1A$target_counts, probs = 0.15), 2),
                round(quantile(polygons_1A$cell_area, probs = 0.15), 2),
                round(quantile(polygons_1A$nucleus_area, probs = 0.15), 2),
                round(quantile((polygons_1A$target_counts/polygons_1A$cell_area), probs = 0.15), 2),
                round(quantile((polygons_1A$target_counts/polygons_1A$nucleus_area), probs = 0.15), 2),
                round(quantile(polygons_1A$log2_AspectRatio, probs = 0.15), 2),
                round(quantile(polygons_1A$log2_AspectRatio, probs = 0.85), 2))

flagged_cells <- c(sum(polygons_1A$metric_color=="red", na.rm = T),
                   sum(polygons_1A$target_color=="red", na.rm = T),
                   sum(polygons_1A$area_color=="red", na.rm = T),
                   sum(polygons_1A$nucleus_color=="red", na.rm = T),
                   sum(polygons_1A$count_area_ratio_color=="red", na.rm = T),
                   sum(polygons_1A$count_nucleus_ratio_color=="red", na.rm = T),
                   sum(polygons_1A$log_hw_color_lo=="red", na.rm = T),
                   sum(polygons_1A$log_hw_color_hi=="red", na.rm = T))

unflagged_cells <- c(sum(polygons_1A$metric_color=="white", na.rm = T),
                     sum(polygons_1A$target_color=="white", na.rm = T),
                     sum(polygons_1A$area_color=="white", na.rm = T),
                     sum(polygons_1A$nucleus_color=="white", na.rm = T),
                     sum(polygons_1A$count_area_ratio_color=="white", na.rm = T),
                     sum(polygons_1A$count_nucleus_ratio_color=="white", na.rm = T),
                     sum(polygons_1A$log_hw_color_lo=="white", na.rm = T),
                     sum(polygons_1A$log_hw_color_hi=="white", na.rm = T))

percent_flagged <- round(flagged_cells/dim(polygons_1A)[1]*100, 2)

poly_1A <- data.frame(threshold = c(0.15, 0.15, 0.15, 0.15, 0.15, 0.15, 0.15, 0.85), flag_value = flag_value, unflagged_cells = unflagged_cells, flagged_cells = flagged_cells, percent_flagged = percent_flagged, row.names = c("Flag score", "Target counts", "Cell area in um", "Nucleus area in um", "Target counts/cell area", "Target counts/nucleus area", "Logged aspect ratio low", "Logged aspect ratio high"))

kable(poly_1A, format = "html", col.names = c("Threshold", "Flag value", "Unflagged cells", "Flagged cells", "% flagged cells"), align = "c", caption = paste0("<center>Flagged cells by different metrics, total cells in fov: ", dim(polygons_1A)[1], ", total cells in whole sample: ", dim(polygons_1A)[1], ", flagged cells in whole sample: ", sum(spe$flag_score < quantile(spe$flag_score, probs = 0.15, na.rm = T)), "</center>"), padding = 1L)
```

# Cells of mock fov 8F

```{r}
polygons_8F <- subset(polygons, mock_fov == "8 F")

tm_shape(polygons_8F) + 
  tm_fill(col= "flag_score", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE)
```

Local threshold map of flag score, target counts, cell_area and logged width/height ratio, 20th percentile

```{r, out.width="300%", out.height="300%"}
metric_color <- ifelse(polygons_8F$flag_score < quantile(spe$flag_score, probs = 0.15), "red", "white")
target_color <- ifelse(polygons_8F$target_counts < quantile(spe$target_counts, probs = 0.15), "red", "white")
area_color <- ifelse(polygons_8F$cell_area < quantile(spe$cell_area, probs = 0.15), "red", "white")
nucleus_color <- ifelse(polygons_8F$nucleus_area < quantile(spe$nucleus_area, probs = 0.15), "red", "white")
count_area_ratio_color <- ifelse((polygons_8F$target_counts/polygons_8F$cell_area) < quantile((spe$target_counts/spe$cell_area), probs = 0.15), "red", "white")
count_nucleus_ratio_color <- ifelse((polygons_8F$target_counts/polygons_8F$nucleus_area) < quantile((spe$target_counts/spe$nucleus_area), probs = 0.15), "red", "white")
log_hw_color_lo <- ifelse(polygons_8F$log2_AspectRatio < quantile(spe$log2_AspectRatio, probs = 0.15), "red", "white")
log_hw_color_hi <- ifelse(polygons_8F$log2_AspectRatio > quantile(spe$log2_AspectRatio, probs = 0.85), "red", "white")

polygons_8F <- cbind(polygons_8F, metric_color = metric_color, target_color = target_color, area_color = area_color, nucleus_color = nucleus_color, count_area_ratio_color = count_area_ratio_color, count_nucleus_ratio_color = count_nucleus_ratio_color, log_hw_color_lo = log_hw_color_lo, log_hw_color_hi = log_hw_color_hi)

p1 <- tm_shape(polygons_8F) + 
  tm_fill(col= "metric_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Flag score",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))


p2 <- tm_shape(polygons_8F) + 
  tm_fill(col= "target_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Target counts",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p3 <- tm_shape(polygons_8F) + 
  tm_fill(col= "area_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "cell area",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p4 <- tm_shape(polygons_8F) + 
  tm_fill(col= "nucleus_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "nucleus area",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p5 <- tm_shape(polygons_8F) + 
  tm_fill(col= "count_area_ratio_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "counts/cell area",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p6 <- tm_shape(polygons_8F) + 
  tm_fill(col= "count_nucleus_ratio_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "counts/nucleus area",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p7 <- tm_shape(polygons_8F) + 
  tm_fill(col= "log_hw_color_lo", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Logged AspectRatio below",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p8 <- tm_shape(polygons_8F) + 
  tm_fill(col= "log_hw_color_hi", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Logged AspectRatio above",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p9 <- tm_shape(polygons_8F) + 
tm_fill(col = "celltype_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "celltype",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))


tmap_arrange(p1, p2, p3, p4, p5, p6, p7, p8, p9, ncol = 3)
```

```{r, out.width="150%", out.height="150%"}
plot_list <- list(p1, p2, p3, p4, p5, p6, p7, p8, p9 + 
                    tm_layout(frame = FALSE,
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0)) +
tm_add_legend(type = "fill", labels = names(celltype_palette), col = celltype_palette))

for(i in 1:length(plot_list)){
  print(plot_list[[i]])
}
```


```{r, out.width="150%", out.height="150%"}

flag_value <- c(round(quantile(polygons_8F$flag_score, probs = 0.15), 2),
                round(quantile(polygons_8F$target_counts, probs = 0.15), 2),
                round(quantile(polygons_8F$cell_area, probs = 0.15), 2),
                round(quantile(polygons_8F$nucleus_area, probs = 0.15), 2),
                round(quantile((polygons_8F$target_counts/polygons_8F$cell_area), probs = 0.15), 2),
                round(quantile((polygons_8F$target_counts/polygons_8F$nucleus_area), probs = 0.15), 2),
                round(quantile(polygons_8F$log2_AspectRatio, probs = 0.15), 2),
                round(quantile(polygons_8F$log2_AspectRatio, probs = 0.85), 2))

flagged_cells <- c(sum(polygons_8F$metric_color=="red", na.rm = T),
                   sum(polygons_8F$target_color=="red", na.rm = T),
                   sum(polygons_8F$area_color=="red", na.rm = T),
                   sum(polygons_8F$nucleus_color=="red", na.rm = T),
                   sum(polygons_8F$count_area_ratio_color=="red", na.rm = T),
                   sum(polygons_8F$count_nucleus_ratio_color=="red", na.rm = T),
                   sum(polygons_8F$log_hw_color_lo=="red", na.rm = T),
                   sum(polygons_8F$log_hw_color_hi=="red", na.rm = T))

unflagged_cells <- c(sum(polygons_8F$metric_color=="white", na.rm = T),
                     sum(polygons_8F$target_color=="white", na.rm = T),
                     sum(polygons_8F$area_color=="white", na.rm = T),
                     sum(polygons_8F$nucleus_color=="white", na.rm = T),
                     sum(polygons_8F$count_area_ratio_color=="white", na.rm = T),
                     sum(polygons_8F$count_nucleus_ratio_color=="white", na.rm = T),
                     sum(polygons_8F$log_hw_color_lo=="white", na.rm = T),
                     sum(polygons_8F$log_hw_color_hi=="white", na.rm = T))

percent_flagged <- round(flagged_cells/dim(polygons_8F)[1]*100, 2)

poly_8F <- data.frame(threshold = c(0.15, 0.15, 0.15, 0.15, 0.15, 0.15, 0.15, 0.85), flag_value = flag_value, unflagged_cells = unflagged_cells, flagged_cells = flagged_cells, percent_flagged = percent_flagged, row.names = c("Flag score", "Target counts", "Cell area in um", "Nucleus area in um", "Target counts/cell area", "Target counts/nucleus area", "Logged aspect ratio low", "Logged aspect ratio high"))

kable(poly_8F, format = "html", col.names = c("Threshold", "Flag value", "Unflagged cells", "Flagged cells", "% flagged cells"), align = "c", caption = paste0("<center>Flagged cells by different metrics, total cells in fov: ", dim(polygons_8F)[1], ", total cells in whole sample: ", dim(polygons_8F)[1], ", flagged cells in whole sample: ", sum(spe$flag_score < quantile(spe$flag_score, probs = 0.15, na.rm = T)), "</center>"), padding = 1L)
```


# Cells of mock fov 2E

```{r}
polygons_2E <- subset(polygons, mock_fov == "2 E")

tm_shape(polygons_2E) + 
  tm_fill(col= "flag_score", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE)
```

Local threshold map of flag score, target counts, cell_area and logged width/height ratio, 20th percentile

```{r, out.width="300%", out.height="300%"}
metric_color <- ifelse(polygons_2E$flag_score < quantile(spe$flag_score, probs = 0.15), "red", "white")
target_color <- ifelse(polygons_2E$target_counts < quantile(spe$target_counts, probs = 0.15), "red", "white")
area_color <- ifelse(polygons_2E$cell_area < quantile(spe$cell_area, probs = 0.15), "red", "white")
nucleus_color <- ifelse(polygons_2E$nucleus_area < quantile(spe$nucleus_area, probs = 0.15), "red", "white")
count_area_ratio_color <- ifelse((polygons_2E$target_counts/polygons_2E$cell_area) < quantile((spe$target_counts/spe$cell_area), probs = 0.15), "red", "white")
count_nucleus_ratio_color <- ifelse((polygons_2E$target_counts/polygons_2E$nucleus_area) < quantile((spe$target_counts/spe$nucleus_area), probs = 0.15), "red", "white")
log_hw_color_lo <- ifelse(polygons_2E$log2_AspectRatio < quantile(spe$log2_AspectRatio, probs = 0.15), "red", "white")
log_hw_color_hi <- ifelse(polygons_2E$log2_AspectRatio > quantile(spe$log2_AspectRatio, probs = 0.85), "red", "white")

polygons_2E <- cbind(polygons_2E, metric_color = metric_color, target_color = target_color, area_color = area_color, nucleus_color = nucleus_color, count_area_ratio_color = count_area_ratio_color, count_nucleus_ratio_color = count_nucleus_ratio_color, log_hw_color_lo = log_hw_color_lo, log_hw_color_hi = log_hw_color_hi)

p1 <- tm_shape(polygons_2E) + 
  tm_fill(col= "metric_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Flag score",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))


p2 <- tm_shape(polygons_2E) + 
  tm_fill(col= "target_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Target counts",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p3 <- tm_shape(polygons_2E) + 
  tm_fill(col= "area_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "cell area",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p4 <- tm_shape(polygons_2E) + 
  tm_fill(col= "nucleus_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "nucleus area",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p5 <- tm_shape(polygons_2E) + 
  tm_fill(col= "count_area_ratio_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "counts/cell area",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p6 <- tm_shape(polygons_2E) + 
  tm_fill(col= "count_nucleus_ratio_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "counts/nucleus area",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p7 <- tm_shape(polygons_2E) + 
  tm_fill(col= "log_hw_color_lo", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Logged AspectRatio below",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p8 <- tm_shape(polygons_2E) + 
  tm_fill(col= "log_hw_color_hi", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Logged AspectRatio above",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p9 <- tm_shape(polygons_2E) + 
tm_fill(col = "celltype_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "celltype",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))


tmap_arrange(p1, p2, p3, p4, p5, p6, p7, p8, p9, ncol = 3)
```

```{r, out.width="150%", out.height="150%"}
plot_list <- list(p1, p2, p3, p4, p5, p6, p7, p8, p9 + 
                    tm_layout(frame = FALSE,
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0)) +
tm_add_legend(type = "fill", labels = names(celltype_palette), col = celltype_palette))

for(i in 1:length(plot_list)){
  print(plot_list[[i]])
}
```

```{r, out.width="150%", out.height="150%"}

flag_value <- c(round(quantile(polygons_2E$flag_score, probs = 0.15), 2),
                round(quantile(polygons_2E$target_counts, probs = 0.15), 2),
                round(quantile(polygons_2E$cell_area, probs = 0.15), 2),
                round(quantile(polygons_2E$nucleus_area, probs = 0.15), 2),
                round(quantile((polygons_2E$target_counts/polygons_2E$cell_area), probs = 0.15), 2),
                round(quantile((polygons_2E$target_counts/polygons_2E$nucleus_area), probs = 0.15), 2),
                round(quantile(polygons_2E$log2_AspectRatio, probs = 0.15), 2),
                round(quantile(polygons_2E$log2_AspectRatio, probs = 0.85), 2))

flagged_cells <- c(sum(polygons_2E$metric_color=="red", na.rm = T),
                   sum(polygons_2E$target_color=="red", na.rm = T),
                   sum(polygons_2E$area_color=="red", na.rm = T),
                   sum(polygons_2E$nucleus_color=="red", na.rm = T),
                   sum(polygons_2E$count_area_ratio_color=="red", na.rm = T),
                   sum(polygons_2E$count_nucleus_ratio_color=="red", na.rm = T),
                   sum(polygons_2E$log_hw_color_lo=="red", na.rm = T),
                   sum(polygons_2E$log_hw_color_hi=="red", na.rm = T))

unflagged_cells <- c(sum(polygons_2E$metric_color=="white", na.rm = T),
                     sum(polygons_2E$target_color=="white", na.rm = T),
                     sum(polygons_2E$area_color=="white", na.rm = T),
                     sum(polygons_2E$nucleus_color=="white", na.rm = T),
                     sum(polygons_2E$count_area_ratio_color=="white", na.rm = T),
                     sum(polygons_2E$count_nucleus_ratio_color=="white", na.rm = T),
                     sum(polygons_2E$log_hw_color_lo=="white", na.rm = T),
                     sum(polygons_2E$log_hw_color_hi=="white", na.rm = T))

percent_flagged <- round(flagged_cells/dim(polygons_2E)[1]*100, 2)

poly_2E <- data.frame(threshold = c(0.15, 0.15, 0.15, 0.15, 0.15, 0.15, 0.15, 0.85), flag_value = flag_value, unflagged_cells = unflagged_cells, flagged_cells = flagged_cells, percent_flagged = percent_flagged, row.names = c("Flag score", "Target counts", "Cell area in um", "Nucleus area in um", "Target counts/cell area", "Target counts/nucleus area", "Logged aspect ratio low", "Logged aspect ratio high"))

kable(poly_2E, format = "html", col.names = c("Threshold", "Flag value", "Unflagged cells", "Flagged cells", "% flagged cells"), align = "c", caption = paste0("<center>Flagged cells by different metrics, total cells in fov: ", dim(polygons_2E)[1], ", total cells in whole sample: ", dim(polygons_2E)[1], ", flagged cells in whole sample: ", sum(spe$flag_score < quantile(spe$flag_score, probs = 0.15, na.rm = T)), "</center>"), padding = 1L)
```

```{r}
tx <- read_parquet("/NAS06/work/Spatial_omics/Xenium/Breast_cancer_rep1/outs/transcripts.parquet")

polygons_1A <- subset(polygons, mock_fov == "1 A")

tx1 <- tx[tx$cell_id %in% polygons_1A$cell_id,] |> st_as_sf(coords=c("x_location", "y_location"))

tm_shape(tx1) +
    tm_dots(col = "firebrick") +
    tm_shape(polygons_1A) + 
    tm_fill(col="grey90", alpha=0.5) +
    tm_borders(lwd = 0.1, col = "grey50") +
    tm_layout(legend.outside = TRUE)

```










