---
title: "03_27_xenium_human_lung_cancer_report"
author: "Benedetta Banzi"
date: "2024-03-27"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---
<STYLE TYPE="text/css">
<!--
  td{
    font-family: Arial; 
    font-size: 10 pt;
    padding:0px;
    cellpadding="0";
    cellspacing="0"
  }
  th {
    font-family: Arial; 
    font-size: 10 pt;
    height: 20px;
    font-weight: bold;
    text-align: "center";
    background-color: #ccccff;
    padding-left: 15px;
    padding-right: 15px;
  }
  table { 
    border-spacing: 4px;
    border-collapse: collapse;
  }
--->
</STYLE>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Preliminary steps

Library import

```{r}
library(SpatialExperiment)
library(data.table)
library(scater) # it imports scuttle
library(cowplot)
library(ggplot2)
theme_set(theme_bw())
library(matrixStats)
library(dplyr)
library(tidyr)
library(tibble)
library(arrow)
library(scales)
library(sf)
library(tmap)
library(knitr)
library(grid)
library(ggpointdensity)
library(ggpubr)
library(spdep)
library(ggExtra)
library(DT)
library(e1071)
library(robustbase)
```

Read-in function

```{r}
readXeniumSPE <- function(dirname, 
                          countfname = "cell_feature_matrix.h5",
                          coordfpattern = "cells.csv.gz", 
                          coord_names = c("x_centroid", "y_centroid")){
  countfpath <- file.path(dirname, countfname)
  coord_file <- file.path(dirname, list.files(dirname, coordfpattern))
  
  # Count matrix + rowData
  sce <- DropletUtils::read10xCounts(countfpath, col.names = TRUE)
  
  # Spatial and colData
  colData <- read.csv(gzfile(coord_file), header = TRUE)
  
  # construct 'SpatialExperiment'
  spe <- SpatialExperiment::SpatialExperiment(
    assays = assays(sce),
    rowData = rowData(sce),
    colData = colData,
    spatialCoordsNames = coord_names
  )
  
  return(spe)
}
```

Setting directories

```{r}
dirname <- "/NAS06/work/Spatial_omics/Xenium/Human_lung_cancer"
sample_name <- "Xenium_human_lung_cancer"
```

# Xenium dataset exploratory analysis for preprocessing {.tabset .tabset-fade .tabset-pills}

SpatialExperiment object creation and exploration

```{r}
spe <- readXeniumSPE(dirname = dirname)
spe <- scuttle::addPerCellQC(spe, subsets=list(NegProb = grep("^NegControlProbe_", rownames(spe)), NegCodw = grep("^NegControlCodeword_", rownames(spe)), UnasCodw = grep("^UnassignedCodeword_", rownames(spe))))
names(colData(spe)@listData)
spe$control_counts <- rowSums(data.frame(colData(spe))[,grep("^subsets_.*_sum$", names(colData(spe)))])
spe$control_detected <- rowSums(data.frame(colData(spe))[,grep("^subsets_.*_detected$", names(colData(spe)))])
```
Dataset dimensions

```{r}
dim(spe)
```

Adding variables

```{r}
cell_boundaries <- read_parquet(file.path(dirname, "cell_boundaries.parquet"))
bound_sf <- st_as_sf(cell_boundaries, coords = c("vertex_x", "vertex_y"))

polygons_whole <- bound_sf %>%
  group_by(cell_id) %>%
  summarize(geometry = st_combine(geometry))#st_combine = combine several feature geometries into one, without unioning or resolving internal boundaries

polygons_whole <- st_cast(polygons_whole, "POLYGON")

AspectRatio <- c()
for(i in 1:dim(polygons_whole)[1]){
 AspectRatio[i] <- (max(polygons_whole$geometry[[i]][[1]][,2]) - min(polygons_whole$geometry[[i]][[1]][,2]))/(max(polygons_whole$geometry[[i]][[1]][,1]) -
 min(polygons_whole$geometry[[i]][[1]][,1]))
}

polygons_whole$log2_AspectRatio <- log2(AspectRatio) 
```

Checking if the computed polygons are valid and removing elements that could result in geometry errors (Voyager)

```{r, class.source = 'fold-show'}
valid_poly <- st_is_valid(polygons_whole)
table(valid_poly)
```

Computing buffer (a circumscribing space) around geometries to handle invalid geometries using 0 distance, so keeping it adherent to the original space
occupied by the cell. (Voyager)

```{r, class.source = 'fold-show'}
polygons_whole <- st_buffer(polygons_whole, dist = 0)

# In st_cast.MULTIPOLYGON(X[[i]], ...) : polygon from first part only
# 32 are still multipolygon

table(st_geometry_type(polygons_whole))

cellids <- c() 
for(i in 1:dim(polygons_whole)[1]){
	cellids[i] <- attr(polygons_whole$geometry[[i]], "class")[2]
	cellids[i] <- cellids[i] == "MULTIPOLYGON"
	
}

polygons_whole <- polygons_whole[cellids == FALSE, ]
table(st_geometry_type(polygons_whole))

spe <- spe[,spe$cell_id %in% polygons_whole$cell_id]
```


Added variables

```{r}
spe <- spe[, is.na(spe$nucleus_area) == FALSE]
polygons_whole <- polygons_whole[polygons_whole$cell_id %in% spe$cell_id, ]

# Importing global coordinates from spatialCoords slot since they are not available in colData
colData(spe)$CenterX_global_um <- spatialCoords(spe)[,1]
colData(spe)$CenterY_global_um <- spatialCoords(spe)[,2]

# Logged total counts
spe$logtot <- log10(spe$total)

# Logged width/height ratio so there are no holes in the distribution in the middle and also negative values can be appreciated
spe$log2_AspectRatio <- polygons_whole$log2_AspectRatio

# Keeping only target counts by subtracting counts referred to negative control probes
spe$target_counts <- spe$total - spe$control_counts

# Keeping only detected features by subtracting negative control probes
spe$target_detected <- spe$detected - spe$control_detected

# Proportion between counts referred only to target probes and detected features
spe$tot_det_ratio <- spe$target_counts/spe$target_detected

# Control probe counts/total counts
ctrl_total_ratio <- spe$control_counts/spe$total
ctrl_total_ratio <- replace_na(ctrl_total_ratio, 0)
spe$ctrl_total_ratio <- ctrl_total_ratio

# Target area ratio
spe$target_area_ratio <- spe$target_counts/spe$cell_area
```

# Large-scale to small-scale {.tabset .tabset-fade .tabset-pills}

Approaching dataset preprocessing starting from global-scale, then fov-focused exploratory analysis

## Global sample map

To view the spatial organization of both the sample as it is and FOVs

```{r, out.width="600%", out.height="150%"}
metadata <- data.frame(colData(spe))

# image theme
global_map_theme <- function(back.color="black",back.border=NA,title.col="white") {
  theme(aspect.ratio = 1,
        panel.border = element_blank(),
        legend.key = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks = element_blank(),
        axis.text.y = element_blank(),
        axis.text.x = element_blank(),
        panel.grid = element_blank(),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        panel.background = element_rect(fill = "transparent",colour = NA),
        plot.title = element_text(color = title.col,hjust = 0.5,face = "bold"),
        plot.background = element_rect(fill = "transparent",colour = back.border))
}

tm_shape(polygons_whole)+
  tm_borders(lwd = 0.01, col = "grey50")+
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = sample_name,
            main.title.size = 0.5,
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))
```

## Global feature maps

To spot areas with anomalous feature values and spatial patterns on a global scale

```{r, out.width="150%", out.height="150%"}
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "total", point_size = 0.1) + ggtitle("Total counts per cell") + theme(aspect.ratio = 1)
```

```{r, out.width="150%", out.height="150%"}
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "logtot", point_size = 0.1) + ggtitle("Log-total counts per cell") + theme(aspect.ratio = 1)
```

```{r, out.width="150%", out.height="150%"}
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "detected", point_size = 0.1) + ggtitle("Detected features per cell") + theme(aspect.ratio = 1)
```

```{r, out.width="150%", out.height="150%"}
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "nucleus_area", point_size = 0.1) + ggtitle("Cell area in μm per cell") + theme(aspect.ratio = 1)
```

```{r, out.width="150%", out.height="150%"}
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "nucleus_area", point_size = 0.1) + ggtitle("Nucleus area in μm per cell") + theme(aspect.ratio = 1)
```

```{r, out.width="150%", out.height="150%"}
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "log2_AspectRatio", point_size = 0.1) + ggtitle("Logged width/height ratio per cell") + theme(aspect.ratio = 1)
```

```{r, out.width="150%", out.height="150%"}
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "tot_det_ratio", point_size = 0.1) + ggtitle("Target counts/detected features ratio per cell") + theme(aspect.ratio = 1)
```

```{r, out.width="150%", out.height="150%"}
plotColData(spe, "CenterY_global_um", "CenterX_global_um", order_by = "control_counts", colour_by = "control_counts", point_size = 0.1) +
  scale_color_gradient(low = "white", high = "red", name = "control_counts") + ggtitle("Negative control probe counts per cell") +
  theme(aspect.ratio = 1,
        panel.background = element_rect(fill = "black",color = NA),
        plot.background = element_rect(fill = "black",color = NA),
        axis.title.x = element_text(color = "white"),
        axis.title.y = element_text(color = "white"),
        axis.ticks = element_line(color = "white"),
        axis.text.y = element_text(color = "white"),
        axis.text.x = element_text(color = "white"),
        axis.line = element_line(color = "white"),
        legend.title = element_text(color = "white"),
        legend.text = element_text(color = "white"),
        plot.title = element_text(color = "white"))
```

```{r, out.width="150%", out.height="150%"}
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "ctrl_total_ratio", point_size = 0.1) + ggtitle("Control counts/total counts ratio per cell") + theme(aspect.ratio = 1)
```


```{r}
threshold <- 0.20
spe$flagged_tot_20 <- as.factor(ifelse(colData(spe)$total < quantile(colData(spe)$total, probs = threshold), TRUE, FALSE))

threshold <- 0.10
spe$flagged_tot_10 <- as.factor(ifelse(colData(spe)$total < quantile(colData(spe)$total, probs = threshold), TRUE, FALSE))

threshold <- 0.05
spe$flagged_tot_5 <- as.factor(ifelse(colData(spe)$total < quantile(colData(spe)$total, probs = threshold), TRUE, FALSE))

threshold <- 0.10
spe$flagged_det_10 <- as.factor(ifelse(colData(spe)$detected < quantile(colData(spe)$detected, probs = threshold), TRUE, FALSE))

threshold <- 0.05
spe$flagged_det_5 <- as.factor(ifelse(colData(spe)$detected < quantile(colData(spe)$detected, probs = threshold), TRUE, FALSE))

threshold <- 0.10
spe$flagged_cellarea_10 <- as.factor(ifelse(colData(spe)$cell_area < quantile(colData(spe)$cell_area, probs = threshold), TRUE, FALSE))

threshold <- 0.05
spe$flagged_cellarea_5 <- as.factor(ifelse(colData(spe)$cell_area < quantile(colData(spe)$cell_area, probs = threshold), TRUE, FALSE))

threshold <- 0.90
spe$flagged_cellarea_90 <- as.factor(ifelse(colData(spe)$cell_area > quantile(colData(spe)$cell_area, probs = threshold), TRUE, FALSE))

threshold <- 0.95
spe$flagged_cellarea_95 <- as.factor(ifelse(colData(spe)$cell_area > quantile(colData(spe)$cell_area, probs = threshold), TRUE, FALSE))

threshold <- 0.10
spe$flagged_nucleusarea_10 <- as.factor(ifelse(colData(spe)$nucleus_area < quantile(colData(spe)$nucleus_area, probs = threshold), TRUE, FALSE))

threshold <- 0.05
spe$flagged_nucleusarea_5 <- as.factor(ifelse(colData(spe)$nucleus_area < quantile(colData(spe)$nucleus_area, probs = threshold), TRUE, FALSE))

threshold <- 0.90
spe$flagged_nucleusarea_90 <- as.factor(ifelse(colData(spe)$nucleus_area > quantile(colData(spe)$nucleus_area, probs = threshold), TRUE, FALSE))

threshold <- 0.95
spe$flagged_nucleusarea_95 <- as.factor(ifelse(colData(spe)$nucleus_area > quantile(colData(spe)$nucleus_area, probs = threshold), TRUE, FALSE))


threshold <- 0.10
spe$flagged_aratio_10 <- as.factor(ifelse(colData(spe)$log2_AspectRatio < quantile(colData(spe)$log2_AspectRatio, probs = threshold), TRUE, FALSE))

threshold <- 0.05
spe$flagged_aratio_5 <- as.factor(ifelse(colData(spe)$log2_AspectRatio < quantile(colData(spe)$log2_AspectRatio, probs = threshold), TRUE, FALSE))

threshold <- 0.90
spe$flagged_aratio_90 <- as.factor(ifelse(colData(spe)$log2_AspectRatio > quantile(colData(spe)$log2_AspectRatio, probs = threshold), TRUE, FALSE))

threshold <- 0.95
spe$flagged_aratio_95 <- as.factor(ifelse(colData(spe)$log2_AspectRatio > quantile(colData(spe)$log2_AspectRatio, probs = threshold), TRUE, FALSE))

threshold <- 0.10
spe$flagged_tdratio_10 <- as.factor(ifelse(colData(spe)$tot_det_ratio < quantile(colData(spe)$tot_det_ratio, probs = threshold, na.rm = T), TRUE, FALSE))

threshold <- 0.05
spe$flagged_tdratio_5 <- as.factor(ifelse(colData(spe)$tot_det_ratio < quantile(colData(spe)$tot_det_ratio, probs = threshold, na.rm = T), TRUE, FALSE))

threshold <- 1
spe$flagged_tdratio_atomx <- as.factor(ifelse(colData(spe)$tot_det_ratio <= threshold, TRUE, FALSE))

threshold <- 0.50
spe$flagged_negprob_50 <- as.factor(ifelse(colData(spe)$control_counts > quantile(colData(spe)$control_counts, probs = threshold, na.rm = T), TRUE, FALSE))

threshold <- 0.75
spe$flagged_negprob_75 <- as.factor(ifelse(colData(spe)$control_counts > quantile(colData(spe)$control_counts, probs = threshold, na.rm = T), TRUE, FALSE))

threshold <- 0.10
spe$flagged_negprob_10 <- as.factor(ifelse(colData(spe)$ctrl_total_ratio > quantile(colData(spe)$ctrl_total_ratio, probs = threshold, na.rm = T), TRUE, FALSE))

threshold <- 0.95
spe$flagged_negprob_95 <- as.factor(ifelse(colData(spe)$control_counts > quantile(colData(spe)$control_counts, probs = threshold, na.rm = T), TRUE, FALSE))

threshold <- 0.99
spe$flagged_negprob_99 <- as.factor(ifelse(colData(spe)$control_counts > quantile(colData(spe)$control_counts, probs = threshold, na.rm = T), TRUE, FALSE))

threshold <- 0.01
spe$flagged_cellarea_1 <- as.factor(ifelse(colData(spe)$cell_area < quantile(colData(spe)$cell_area, probs = threshold), TRUE, FALSE))

threshold <- 0.99
spe$flagged_cellarea_99 <- as.factor(ifelse(colData(spe)$cell_area > quantile(colData(spe)$cell_area, probs = threshold), TRUE, FALSE))

threshold <- 0.01
spe$flagged_nucleusarea_1 <- as.factor(ifelse(colData(spe)$nucleus_area < quantile(colData(spe)$nucleus_area, probs = threshold), TRUE, FALSE))

threshold <- 0.05
spe$flagged_nucleusarea_5 <- as.factor(ifelse(colData(spe)$nucleus_area < quantile(colData(spe)$nucleus_area, probs = threshold), TRUE, FALSE))

threshold <- 0.95
spe$flagged_nucleusarea_95 <- as.factor(ifelse(colData(spe)$nucleus_area > quantile(colData(spe)$nucleus_area, probs = threshold), TRUE, FALSE))

threshold <- 0.99
spe$flagged_nucleusarea_99 <- as.factor(ifelse(colData(spe)$nucleus_area > quantile(colData(spe)$nucleus_area, probs = threshold), TRUE, FALSE))

threshold <- 0.01
spe$flagged_target_1 <- as.factor(ifelse(colData(spe)$target_counts < quantile(colData(spe)$target_counts, probs = threshold), TRUE, FALSE))

threshold <- 0.05
spe$flagged_target_5 <- as.factor(ifelse(colData(spe)$target_counts < quantile(colData(spe)$target_counts, probs = threshold), TRUE, FALSE))
```

# Feature histograms 

To check feature distribution before highlighting where extreme values are located in the global sample map

```{r}
variables <- c("total", "detected", "logtot", "cell_area", "nucleus_area",  "log2_AspectRatio", "tot_det_ratio", "control_counts", "ctrl_total_ratio")
for (var in 1:length(variables)){
  n <- grep(paste0("^", variables[var], "$"), colnames(spe@colData))
  histo_var <- names(colData(spe))[n]
  p <- ggplot(data = data.frame(colData(spe))) +
    geom_histogram(aes(x = .data[[histo_var]]), fill = "#69b3a2") +
    ggtitle(paste0(histo_var))
  print(p)
}  
```

# First cell skimming with quantile thresholds {.tabset .tabset-fade .tabset-pills}

Coarse-grain cell filtering to remove gross segmentation errors.

## Control probes

```{r}
summary(spe$control_counts)
```

```{r, out.width="150%", out.height="150%"}
hist(spe$control_counts)
```

```{r, out.width="150%", out.height="150%"}
boxplot(spe$control_counts)
```


```{r, out.width="150%", out.height="150%"}
threshold <- 0.99
spe$flagged_negprob_99 <- as.factor(ifelse(colData(spe)$control_counts > quantile(colData(spe)$control_counts, probs = threshold, na.rm = T), TRUE, FALSE))
polygons_whole$flagged_negprob_99 <- ifelse(spe$flagged_negprob_99 == TRUE, "red", "grey80")

tm_shape(polygons_whole) + 
  tm_fill(col= "flagged_negprob_99") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = paste0("Negative probe counts above 99% threshold, value = ", 
                                round(quantile(colData(spe)$control_counts, probs = threshold, na.rm = T), 2), "\nFlagged cells = ",
                                table(colData(spe)$flagged_negprob_99)["TRUE"], " out of ", dim(spe)[2], ", ",
                                round(table(colData(spe)$flagged_negprob_99)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"),
            main.title.fontface = 2,
            main.title.size = 1, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))
```

```{r}
negprob_spe <- spe[,spe$flagged_negprob_99 == FALSE]
summary(negprob_spe$control_counts)
```

```{r, out.width="150%", out.height="150%"}
hist(negprob_spe$control_counts)
```

```{r, out.width="150%", out.height="150%"}
boxplot(negprob_spe$control_counts)
```

## Control count / total count

```{r}
summary(spe$ctrl_total_ratio)
```

```{r, out.width="150%", out.height="150%"}
hist(spe$ctrl_total_ratio)
```

```{r, out.width="150%", out.height="150%"}
boxplot(spe$ctrl_total_ratio)
```


```{r, out.width="150%", out.height="150%"}
threshold <- 0.99
spe$flagged_ctrl_tot_ratio_99 <- as.factor(ifelse(colData(spe)$ctrl_total_ratio > quantile(colData(spe)$ctrl_total_ratio, probs = threshold, na.rm = T), TRUE, FALSE))
polygons_whole$flagged_ctrl_tot_ratio_99 <- ifelse(spe$flagged_ctrl_tot_ratio_99 == TRUE, "red", "grey80")

tm_shape(polygons_whole) + 
  tm_fill(col= "flagged_ctrl_tot_ratio_99") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = paste0("Control / total counts ratio above 99% threshold, value = ", 
                                round(quantile(colData(spe)$ctrl_total_ratio, probs = threshold, na.rm = T), 2), "\nFlagged cells = ",
                                table(colData(spe)$flagged_ctrl_tot_ratio_99)["TRUE"], " out of ", dim(spe)[2], ", ",
                                round(table(colData(spe)$flagged_ctrl_tot_ratio_99)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"),
            main.title.fontface = 2,
            main.title.size = 1, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))
```

```{r}
ctrltotratio_spe <- spe[,spe$flagged_ctrl_tot_ratio_99 == FALSE]
summary(ctrltotratio_spe$ctrl_total_ratio)
```

```{r, out.width="150%", out.height="150%"}
hist(ctrltotratio_spe$ctrl_total_ratio)
```

```{r, out.width="150%", out.height="150%"}
boxplot(ctrltotratio_spe$ctrl_total_ratio)
```


## Cell area, low threshold

```{r}
summary(spe$cell_area)
```

```{r, out.width="150%", out.height="150%"}
hist(spe$cell_area)
```

```{r, out.width="150%", out.height="150%"}
boxplot(spe$cell_area)
```

```{r, out.width="150%", out.height="150%"}
threshold <- 0.01
spe$flagged_cellarea_1 <- as.factor(ifelse(colData(spe)$cell_area < quantile(colData(spe)$cell_area, probs = threshold), TRUE, FALSE))
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "flagged_cellarea_1", 
            point_size = 0.1) + scale_color_manual(values = c("TRUE" = "black","FALSE" = "lightgrey")) + ggtitle(paste0("Cell area in μm below 1%, threshold value = ", round(quantile(colData(spe)$cell_area, probs = threshold, na.rm = T), 2)))  + theme(legend.position="none") + labs(subtitle = paste0("Flagged cells = ", table(colData(spe)$flagged_cellarea_1)["TRUE"], " out of ", dim(spe)[2], ", ",  round(table(colData(spe)$flagged_cellarea_1)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"))
```


```{r, out.width="150%", out.height="150%"}
threshold <- 0.01
spe$flagged_target_1 <- as.factor(ifelse(colData(spe)$target_counts < quantile(colData(spe)$target_counts, probs = threshold), TRUE, FALSE))

plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "flagged_target_1", 
            point_size = 0.1) + scale_color_manual(values = c("TRUE" = "black","FALSE" = "lightgrey")) + ggtitle(paste0("Target counts below 1%, threshold value = ", round(quantile(colData(spe)$target_counts, probs = threshold, na.rm = T), 2)))  + theme(legend.position="none") + labs(subtitle = paste0("Flagged cells = ", table(colData(spe)$flagged_target_1)["TRUE"], " out of ", dim(spe)[2], ", ",  round(table(colData(spe)$flagged_target_1)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"))
```

```{r}
threshold <- 0.01
spe$flagged_target_1 <- as.factor(ifelse(colData(spe)$target_counts < quantile(colData(spe)$target_counts, probs = threshold), TRUE, FALSE))
# to remove small cells that also have a number of target counts below 1%
table(spe$flagged_cellarea_1, spe$flagged_target_1, dnn = c("flagged_cellarea_1", "flagged_target_1"))
```

## Cell area, high threshold

```{r, out.width="150%", out.height="150%"}
threshold <- 0.99
spe$flagged_cellarea_99 <- as.factor(ifelse(colData(spe)$cell_area > quantile(colData(spe)$cell_area, probs = threshold), TRUE, FALSE))
polygons_whole$flagged_cellarea_99 <- ifelse(spe$flagged_cellarea_99 == TRUE, "red", "grey80")

tm_shape(polygons_whole) + 
  tm_fill(col= "flagged_cellarea_99") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = paste0("Cell area in um above 99% threshold, threshold value = ", 
                                round(quantile(colData(spe)$cell_area, probs = threshold, na.rm = T), 2), "\nFlagged cells = ",
                                table(colData(spe)$flagged_cellarea_99)["TRUE"], " out of ", dim(spe)[2], ", ",
                                round(table(colData(spe)$flagged_cellarea_99)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"),
            main.title.fontface = 2,
            main.title.size = 1, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))
```

```{r}
# Big problematic cells we noticed in polygons map can be captured using both negative probe counts and area above 90%
table(spe$flagged_negprob_99, spe$flagged_cellarea_99, dnn = c("flagged_negprob_99", "flagged_cellarea_99,"))
```

```{r}
area_spe <- spe[, spe$flagged_cellarea_99 == FALSE & spe$flagged_cellarea_1 == FALSE]
summary(area_spe$cell_area)
```

```{r, out.width="150%", out.height="150%"}
hist(area_spe$cell_area)
```

```{r, out.width="150%", out.height="150%"}
boxplot(area_spe$cell_area)
```

## Nucleus area, low threshold

```{r}
summary(spe$nucleus_area)
```

```{r, out.width="150%", out.height="150%"}
hist(spe$nucleus_area)
```

```{r, out.width="150%", out.height="150%"}
boxplot(spe$nucleus_area)
```

```{r, out.width="150%", out.height="150%"}
threshold <- 0.01
spe$flagged_nucleusarea_1 <- as.factor(ifelse(colData(spe)$nucleus_area < quantile(colData(spe)$nucleus_area, probs = threshold), TRUE, FALSE))
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "flagged_nucleusarea_1", 
            point_size = 0.1) + scale_color_manual(values = c("TRUE" = "black","FALSE" = "lightgrey")) + ggtitle(paste0("Nucleus area in μm below 1%, threshold value = ", round(quantile(colData(spe)$nucleus_area, probs = threshold, na.rm = T), 2)))  + theme(legend.position="none") + labs(subtitle = paste0("Flagged cells = ", table(colData(spe)$flagged_nucleusarea_1)["TRUE"], " out of ", dim(spe)[2], ", ",  round(table(colData(spe)$flagged_nucleusarea_1)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"))
```

```{r, out.width="150%", out.height="150%"}
threshold <- 0.01
spe$flagged_target_1 <- as.factor(ifelse(colData(spe)$target_counts < quantile(colData(spe)$target_counts, probs = threshold), TRUE, FALSE))

plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "flagged_target_1", 
            point_size = 0.1) + scale_color_manual(values = c("TRUE" = "black","FALSE" = "lightgrey")) + ggtitle(paste0("Target counts below 1%, threshold value = ", round(quantile(colData(spe)$target_counts, probs = threshold, na.rm = T), 2)))  + theme(legend.position="none") + labs(subtitle = paste0("Flagged cells = ", table(colData(spe)$flagged_target_1)["TRUE"], " out of ", dim(spe)[2], ", ",  round(table(colData(spe)$flagged_target_1)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"))
```

## Nucleus area, high threshold

```{r}
table(is.na(spe$nucleus_area))
```


```{r, out.width="150%", out.height="150%"}
threshold <- 0.99
spe$flagged_nucleusarea_99 <- as.factor(ifelse(spe$nucleus_area > quantile(spe$nucleus_area, probs = threshold, na.rm = T), TRUE, FALSE))
polygons_whole$flagged_nucleusarea_99 <- ifelse(spe$flagged_nucleusarea_99 == TRUE, "red", "grey80")

tm_shape(polygons_whole) + 
  tm_fill(col= "flagged_nucleusarea_99") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = paste0("Nucleus area in um above 99% threshold, threshold value = ", 
                                round(quantile(colData(spe)$nucleus_area, probs = threshold, na.rm = T), 2), "\nFlagged cells = ",
                                table(colData(spe)$flagged_nucleusarea_99)["TRUE"], " out of ", dim(spe)[2], ", ",
                                round(table(colData(spe)$flagged_nucleusarea_99)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"),
            main.title.fontface = 2,
            main.title.size = 1, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))
```

```{r}
nucleus_spe <- spe[, spe$flagged_nucleusarea_99 == FALSE & spe$flagged_nucleusarea_1 == FALSE]
summary(nucleus_spe$nucleus_area)
```

```{r, out.width="150%", out.height="150%"}
hist(nucleus_spe$nucleus_area)
```

```{r, out.width="150%", out.height="150%"}
boxplot(nucleus_spe$nucleus_area)
```

## Logged AspectRatio, low threshold

```{r}
summary(spe$log2_AspectRatio)
```

```{r, out.width="150%", out.height="150%"}
hist(spe$log2_AspectRatio)
```

```{r, out.width="150%", out.height="150%"}
boxplot(spe$log2_AspectRatio)
```

```{r, out.width="150%", out.height="150%"}
threshold <- 0.01
spe$flagged_logAspectRatio_1 <- as.factor(ifelse(spe$log2_AspectRatio < quantile(spe$log2_AspectRatio, probs = threshold, na.rm = T), TRUE, FALSE))
polygons_whole$flagged_logAspectRatio_1 <- ifelse(spe$flagged_logAspectRatio_1 == TRUE, "black", "grey80")

tm_shape(polygons_whole) + 
  tm_fill(col= "flagged_logAspectRatio_1") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = paste0("Logged AspectRatio below 1% threshold, threshold value = ", 
                                round(quantile(colData(spe)$log2_AspectRatio, probs = threshold, na.rm = T), 2), "\nFlagged cells = ",
                                table(colData(spe)$flagged_logAspectRatio_1)["TRUE"], " out of ", dim(spe)[2], ", ",
                                round(table(colData(spe)$flagged_logAspectRatio_1)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"),
            main.title.fontface = 2,
            main.title.size = 1, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))
```

## Logged AspectRatio, high threshold

```{r, out.width="150%", out.height="150%"}
threshold <- 0.99
spe$flagged_logAspectRatio_99 <- as.factor(ifelse(spe$log2_AspectRatio > quantile(spe$log2_AspectRatio, probs = threshold, na.rm = T), TRUE, FALSE))
polygons_whole$flagged_logAspectRatio_99 <- ifelse(spe$flagged_logAspectRatio_99 == TRUE, "red", "grey80")

tm_shape(polygons_whole) + 
  tm_fill(col= "flagged_logAspectRatio_99") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = paste0("Logged AspectRatio above 99% threshold, threshold value = ", 
                                round(quantile(colData(spe)$log2_AspectRatio, probs = threshold, na.rm = T), 2), "\nFlagged cells = ",
                                table(colData(spe)$flagged_logAspectRatio_99)["TRUE"], " out of ", dim(spe)[2], ", ",
                                round(table(colData(spe)$flagged_logAspectRatio_99)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"),
            main.title.fontface = 2,
            main.title.size = 1, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))
```

```{r}
aspectratio_spe <- spe[, spe$flagged_logAspectRatio_99 == FALSE & spe$flagged_logAspectRatio_1 == FALSE]
summary(aspectratio_spe$log2_AspectRatio)
```

```{r, out.width="150%", out.height="150%"}
hist(aspectratio_spe$log2_AspectRatio)
```

```{r, out.width="150%", out.height="150%"}
boxplot(aspectratio_spe$log2_AspectRatio)
```


# First cell skimming with scuttle::isOutlier (Hampel filter) {.tabset .tabset-fade .tabset-pills}

Introducing scuttle::isOutlier function to move to a different filter beyond threshold
Outliers are defined when their value is higher or lower than the median, plus or minus 3 median absolute deviations (MAD) respectively.
However, when using MAD a symmetric distribution is assumed

## Control counts

```{r}
summary(spe$control_counts)
```

```{r, out.width="150%", out.height="150%"}
hist(spe$control_counts)
```

```{r, out.width="150%", out.height="150%"}
boxplot(spe$control_counts)
```

```{r, out.width="150%", out.height="150%"}
spe$outlier_negprob <- scuttle::isOutlier(spe$control_counts, type = "higher")
polygons_whole$outlier_negprob <- ifelse(spe$outlier_negprob == TRUE, "red", "grey80")

tm_shape(polygons_whole) + 
  tm_fill(col= "outlier_negprob") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = paste0("Outlier cells for negative probe counts", 
                                "\n lower thr. = " , round(attr(spe$outlier_negprob,"thresholds")["lower"], 2), ", higher thr. = ",
                                round(attr(spe$outlier_negprob,"thresholds")["higher"], 2),
                                "\nFlagged cells = ",
                                table(colData(spe)$outlier_negprob)["TRUE"], " out of ", dim(spe)[2], ", ",
                                round(table(colData(spe)$outlier_negprob)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"),
            main.title.fontface = 2,
            main.title.size = 1, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

```

```{r}
negprob_spe <- spe[,spe$outlier_negprob == FALSE]
summary(negprob_spe$control_counts)
```

```{r, out.width="150%", out.height="150%"}
hist(negprob_spe$control_counts)
```

```{r, out.width="150%", out.height="150%"}
boxplot(negprob_spe$control_counts)
```

## Control count / total counts

```{r}
summary(spe$ctrl_total_ratio)
```

```{r, out.width="150%", out.height="150%"}
hist(spe$ctrl_total_ratio)
```

```{r, out.width="150%", out.height="150%"}
boxplot(spe$ctrl_total_ratio)
```

```{r, out.width="150%", out.height="150%"}
spe$outlier_ctrl_tot <- scuttle::isOutlier(spe$ctrl_total_ratio, type = "higher")
polygons_whole$outlier_ctrl_tot <- ifelse(spe$outlier_ctrl_tot == TRUE, "red", "grey80")

tm_shape(polygons_whole) + 
  tm_fill(col= "outlier_ctrl_tot") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = paste0("Outlier cells for control / total counts ratio", 
                                "\n lower thr. = " , round(attr(spe$outlier_ctrl_tot,"thresholds")["lower"], 2), ", higher thr. = ",
                                round(attr(spe$outlier_ctrl_tot,"thresholds")["higher"], 2),
                                "\nFlagged cells = ",
                                table(colData(spe)$outlier_ctrl_tot)["TRUE"], " out of ", dim(spe)[2], ", ",
                                round(table(colData(spe)$outlier_ctrl_tot)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"),
            main.title.fontface = 2,
            main.title.size = 1, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

```

```{r}
ctrltotratio_spe <- spe[,spe$outlier_ctrl_tot == FALSE]
summary(ctrltotratio_spe$ctrl_total_ratio)
```

```{r, out.width="150%", out.height="150%"}
hist(ctrltotratio_spe$ctrl_total_ratio)
```

```{r, out.width="150%", out.height="150%"}
boxplot(ctrltotratio_spe$ctrl_total_ratio)
```

## Cell area

```{r}
summary(spe$cell_area)
```

```{r, out.width="150%", out.height="150%"}
hist(spe$cell_area)
```

```{r, out.width="150%", out.height="150%"}
boxplot(spe$cell_area)
```

```{r, out.width="150%", out.height="150%"}
spe$outlier_cellarea <- scuttle::isOutlier(colData(spe)$cell_area, type = "both")
polygons_whole$outlier_cellarea <- ifelse(spe$outlier_cellarea == TRUE, "red", "grey80")

tm_shape(polygons_whole) + 
  tm_fill(col= "outlier_cellarea") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = paste0("Outlier cells for cell area", 
                                "\n lower thr. = " , round(attr(spe$outlier_cellarea,"thresholds")["lower"], 2), ", higher thr. = ",
                                round(attr(spe$outlier_cellarea,"thresholds")["higher"], 2),
                                "\nFlagged cells = ",
                                table(colData(spe)$outlier_cellarea)["TRUE"], " out of ", dim(spe)[2], ", ",
                                round(table(colData(spe)$outlier_cellarea)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"),
            main.title.fontface = 2,
            main.title.size = 1, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))
```

```{r}
area_spe <- spe[, spe$outlier_cellarea == FALSE]
summary(area_spe$cell_area)
```

```{r, out.width="150%", out.height="150%"}
hist(area_spe$cell_area)
```

```{r, out.width="150%", out.height="150%"}
boxplot(area_spe$cell_area)
```

## Nucleus area

```{r}
summary(spe$nucleus_area)
```

```{r, out.width="150%", out.height="150%"}
hist(spe$nucleus_area)
```

```{r, out.width="150%", out.height="150%"}
boxplot(spe$nucleus_area)
```

```{r, out.width="150%", out.height="150%"}
spe$outlier_nucleusarea <- scuttle::isOutlier(spe$nucleus_area, type = "both")
polygons_whole$outlier_nucleusarea<- ifelse(spe$outlier_nucleusarea == TRUE, "red", "grey80")
polygons_whole$outlier_nucleusarea <- replace_na(polygons_whole$outlier_nucleusarea, "grey80")

tm_shape(polygons_whole) + 
  tm_fill(col= "outlier_nucleusarea") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = paste0("Outlier cells for nucleus area signal",
                                "\n lower thr. = " , round(attr(spe$outlier_nucleusarea,"thresholds")["lower"],2), ", higher thr. = ",
                                round(attr(spe$outlier_nucleusarea,"thresholds")["higher"], 2),
                                "\nFlagged cells = ",
                                table(colData(spe)$outlier_nucleusarea)["TRUE"], " out of ", dim(spe)[2], ", ",
                                round(table(colData(spe)$outlier_nucleusarea)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"),
            main.title.fontface = 2,
            main.title.size = 1, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))
```

```{r}
nucleusarea_spe <- spe[, spe$outlier_nucleusarea == FALSE]
summary(nucleusarea_spe$nucleus_area)
```

```{r, out.width="150%", out.height="150%"}
hist(nucleusarea_spe$nucleus_area)
```

```{r, out.width="150%", out.height="150%"}
boxplot(nucleusarea_spe$nucleus_area)
```

## Log2 AspectRatio

```{r}
summary(spe$log2_AspectRatio)
```

```{r, out.width="150%", out.height="150%"}
hist(spe$log2_AspectRatio)
```

```{r, out.width="150%", out.height="150%"}
boxplot(spe$log2_AspectRatio)
```

```{r, out.width="150%", out.height="150%"}
spe$outlier_loghw <- scuttle::isOutlier(spe$log2_AspectRatio, type = "both")
polygons_whole$outlier_loghw<- ifelse(spe$outlier_loghw == TRUE, "red", "grey80")

tm_shape(polygons_whole) + 
  tm_fill(col= "outlier_loghw") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = paste0("Outlier cells for logged aspect ratio",
                                "\n lower thr. = " , round(attr(spe$outlier_loghw,"thresholds")["lower"],2), ", higher thr. = ",
                                round(attr(spe$outlier_loghw,"thresholds")["higher"], 2),
                                "\nFlagged cells = ",
                                table(colData(spe)$outlier_loghw)["TRUE"], " out of ", dim(spe)[2], ", ",
                                round(table(colData(spe)$outlier_loghw)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"),
            main.title.fontface = 2,
            main.title.size = 1, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))
```

```{r}
attr(spe$outlier_loghw, "threshold")
```

```{r}
loghw_spe <- spe[, spe$outlier_loghw == FALSE]
summary(loghw_spe$log2_AspectRatio)
```

```{r, out.width="150%", out.height="150%"}
hist(loghw_spe$log2_AspectRatio)
```

```{r, out.width="150%", out.height="150%"}
boxplot(loghw_spe$log2_AspectRatio)
```


# First cell skimming with skewness adjustment to boxplot {.tabset .tabset-fade .tabset-pills}

Introducing medcouple (MC) from Hubert, M. and Vandervieren, E. (2008). Boxplot whiskers are defined as:
1st quartile - 1.5 * hl(MC) * IQR 
3rd quartile + 1.5 * hu(MC) * IQR
If MC >= 0, hl = exp(-4 * MC), hu = exp(3 * MC) 
If MC < 0, hl = exp(-3 * MC), hu = exp(4 * MC)

```{r}
polygons_whole$control_counts <- spe$control_counts
polygons_whole$cell_area <- spe$cell_area
polygons_whole$nucleus_area <- spe$nucleus_area
```

## Control counts

```{r}
summary(spe$control_counts)
```

```{r, out.width="150%", out.height="150%"}
hist(spe$control_counts)
```

A normal distribution should have a skewness of 0.

```{r}
#it should be 0
library(e1071)
skewness(spe$control_counts, na.rm = T)[1]
```

−0.6 ≤ MC ≤ 0.6 is requested by Hubert and Vandervieren (2008) for their adjusted boxplot.

```{r}
library(robustbase)
mc(spe$control_counts, na.rm = T)[1] 
```

## Control / total counts ratio

```{r}
summary(spe$ctrl_total_ratio)
```

```{r, out.width="150%", out.height="150%"}
hist(spe$ctrl_total_ratio)
```

```{r}
skewness(spe$ctrl_total_ratio, na.rm = T)[1]
```

```{r}
mc(spe$ctrl_total_ratio, na.rm = T)[1] 
```


## Cell area

```{r}
summary(spe$cell_area)
```

```{r, out.width="150%", out.height="150%"}
hist(spe$cell_area)
```

Skewness

```{r}
skewness(spe$cell_area)[1]
```

Medcouple

```{r}
mc(spe$cell_area)[1]
```

How many outliers were found?

```{r}
names(spe$cell_area) <- colnames(spe)
out <- adjbox(spe$cell_area)
cellarea_fence <- out$fence
length(out$out)
```

Distribution of outlier values

```{r}
summary(out$out)
```

```{r, out.width="150%", out.height="150%"}
outlier_color <- colnames(spe)
for(i in 1:out$n){
  outlier_color[i] <- ifelse(outlier_color[i] %in% names(out$out), "red", "grey80")
}

polygons_whole$cellarea_outlier <- outlier_color

tm_shape(polygons_whole) + 
  tm_fill(col= "cellarea_outlier") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = paste0("Outlier cells for cell_area", 
                                "\n fence = " , round(out$fence, 2),
                                "\nFlagged cells = ",
                                table(polygons_whole$cellarea_outlier)["red"], " out of ", dim(spe)[2], ", ",
                                round(table(polygons_whole$cellarea_outlier)["red"]/dim(spe)[2]*100, 2), "% of total cells"),
            main.title.fontface = 2,
            main.title.size = 1, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))
```

```{r}
area_poly <- polygons_whole[polygons_whole$cellarea_outlier == "grey80",]
hist(area_poly$cell_area)
```

```{r, out.width="150%", out.height="150%"}
boxplot(area_poly$cell_area)
```

```{r}
summary(area_poly$cell_area)
```


## Nucleus area

```{r}
summary(spe$nucleus_area)
```

```{r, out.width="150%", out.height="150%"}
hist(spe$nucleus_area)
```

Skewness

```{r}
skewness(spe$nucleus_area, na.rm = TRUE)[1]
```

Medcouple

```{r}
mc(spe$nucleus_area, na.rm = TRUE)[1]
```

How many outliers were found?

```{r}
names(spe$nucleus_area) <- colnames(spe)
out <- adjbox(spe$nucleus_area)
nucleusarea_fence <- out$fence
length(out$out)
```

Distribution of outlier values

```{r}
summary(out$out)
```

```{r, out.width="150%", out.height="150%"}
outlier_color <- colnames(spe)
for(i in 1:dim(spe)[2])
  outlier_color[i] <- ifelse(outlier_color[i] %in% names(out$out), "red", "grey80")

polygons_whole$nucleusarea_outlier <- outlier_color

tm_shape(polygons_whole) + 
  tm_fill(col= "nucleusarea_outlier") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = paste0("Outlier cells for nucleus area", 
                                "\n fence = " , round(out$fence, 2),
                                "\nFlagged cells = ",
                                table(polygons_whole$nucleusarea_outlier)["red"], " out of ", dim(spe)[2], ", ",
                                round(table(polygons_whole$nucleusarea_outlier)["red"]/dim(spe)[2]*100, 2), "% of total cells"),
            main.title.fontface = 2,
            main.title.size = 1, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))
```

```{r}
table(spe$outlier_negprob, spe$outlier_cellarea, dnn = c("outlier_negprob", "outlier_cellarea"))
```

```{r}
nucleusarea_poly <- polygons_whole[polygons_whole$nucleusarea_outlier == "grey80",]
hist(nucleusarea_poly$nucleus_area)
```

```{r, out.width="150%", out.height="150%"}
boxplot(nucleusarea_poly$nucleus_area)
```

```{r}
summary(nucleusarea_poly$nucleus_area)
```

## Log2 AspectRatio

```{r}
summary(spe$log2_AspectRatio)
```

```{r, out.width="150%", out.height="150%"}
hist(spe$log2_AspectRatio)
```

Skewness

```{r}
skewness(spe$log2_AspectRatio, na.rm = TRUE)[1]
```

Medcouple

```{r}
mc(spe$log2_AspectRatio, na.rm = TRUE)[1]
```

How many outliers were found?

```{r}
names(spe$log2_AspectRatio) <- colnames(spe)
out <- adjbox(spe$log2_AspectRatio)
loghw_fence <- out$fence 
length(out$out)
```

Distribution of outlier values

```{r}
summary(out$out)
```

```{r, out.width="150%", out.height="150%"}
outlier_color <- colnames(spe)
for(i in 1:dim(spe)[2])
  outlier_color[i] <- ifelse(outlier_color[i] %in% names(out$out), "red", "grey80")

polygons_whole$loghw_outlier <- outlier_color

tm_shape(polygons_whole) + 
  tm_fill(col= "loghw_outlier") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = paste0("Outlier cells for logged aspect ratio", 
                                "\n fence = " , round(out$fence, 2),
                                "\nFlagged cells = ",
                                table(polygons_whole$loghw_outlier)["red"], " out of ", dim(spe)[2], ", ",
                                round(table(polygons_whole$loghw_outlier)["red"]/dim(spe)[2]*100, 2), "% of total cells"),
            main.title.fontface = 2,
            main.title.size = 1, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))
```

```{r}
loghw_poly <- polygons_whole[polygons_whole$loghw_outlier == "grey80",]
hist(loghw_poly$log2_AspectRatio)
```

```{r, out.width="150%", out.height="150%"}
boxplot(loghw_poly$log2_AspectRatio)
```

```{r}
summary(loghw_poly$log2_AspectRatio)
```



# Flag score development 

## Exploratory scatterplots {.tabset .tabset-fade .tabset-pills}

### Total counts ~ cell area

```{r, out.width="150%", out.height="150%"}
cell_df <- data.frame(colData(spe))

p <- ggplot(data = cell_df) +
  geom_pointdensity(aes(x = cell_area, y = total)) +
  scale_color_viridis_c() +
  theme(legend.position = "none",
        plot.margin = margin())

ggMarginal(p, type = "density", color = "#ccccff", fill = "#ccccff") 
```

### Total counts ~ nucleus area

```{r, out.width="150%", out.height="150%"}

p <- ggplot(data = cell_df) +
  geom_pointdensity(aes(x = nucleus_area, y = total)) +
  scale_color_viridis_c() +
  theme(legend.position = "none",
        plot.margin = margin())

ggMarginal(p, type = "density", color = "#ccccff", fill = "#ccccff") 
```

### Cell area ~ global coordinates on X axis in px

```{r, out.width="150%", out.height="150%"}
p <- ggplot(data = cell_df) +
  geom_pointdensity(aes(x = CenterX_global_um, y = cell_area)) +
  scale_color_viridis_c() +
  theme(legend.position = "none",
        plot.margin = margin())

ggMarginal(p, type = "density", color = "#ccccff", fill = "#ccccff") 
```

### Cell area ~ global coordinates on Y axis in px

```{r, out.width="150%", out.height="150%"}
p <- ggplot(data = cell_df) +
  geom_pointdensity(aes(x = CenterY_global_um, y = cell_area)) +
  scale_color_viridis_c() +
  theme(legend.position = "none",
        plot.margin = margin())

ggMarginal(p, type = "density", color = "#ccccff", fill = "#ccccff") 
```

### Nucleus area ~ global coordinates on X axis in px

```{r, out.width="150%", out.height="150%"}
p <- ggplot(data = cell_df) +
  geom_pointdensity(aes(x = CenterX_global_um, y = nucleus_area)) +
  scale_color_viridis_c() +
  theme(legend.position = "none",
        plot.margin = margin())

ggMarginal(p, type = "density", color = "#ccccff", fill = "#ccccff") 
```

### Nucleus area ~ global coordinates on Y axis in px

```{r, out.width="150%", out.height="150%"}
p <- ggplot(data = cell_df) +
  geom_pointdensity(aes(x = CenterY_global_um, y = nucleus_area)) +
  scale_color_viridis_c() +
  theme(legend.position = "none",
        plot.margin = margin())

ggMarginal(p, type = "density", color = "#ccccff", fill = "#ccccff") 
```

### Cell area ~ control counts

```{r, out.width="150%", out.height="150%"}
p <- ggplot(data = cell_df) +
  geom_pointdensity(aes(x = control_counts, y = cell_area)) +
  scale_color_viridis_c() +
  theme(legend.position = "none",
        plot.margin = margin())

ggMarginal(p, type = "density", color = "#ccccff", fill = "#ccccff") 
```

### Nucleus area ~ control counts

```{r, out.width="150%", out.height="150%"}
p <- ggplot(data = cell_df) +
  geom_pointdensity(aes(x = control_counts, y = nucleus_area)) +
  scale_color_viridis_c() +
  theme(legend.position = "none",
        plot.margin = margin())

ggMarginal(p, type = "density", color = "#ccccff", fill = "#ccccff") 
```

### Global coordinates on X axis ~ logged AspectRatio

```{r, out.width="150%", out.height="150%"}
p <- ggplot(data = cell_df) +
  geom_pointdensity(aes(x = CenterX_global_um, y = log2_AspectRatio)) +
  scale_color_viridis_c() +
  theme(legend.position = "none",
        plot.margin = margin())

ggMarginal(p, type = "density", color = "#ccccff", fill = "#ccccff") 
```

### Global coordinates on Y axis ~ logged AspectRatio

```{r, out.width="150%", out.height="150%"}
p <- ggplot(data = cell_df) +
  geom_pointdensity(aes(x = CenterY_global_um, y = log2_AspectRatio)) +
  scale_color_viridis_c() +
  theme(legend.position = "none",
        plot.margin = margin())

ggMarginal(p, type = "density", color = "#ccccff", fill = "#ccccff") 
```


```{r, out.width="150%", out.height="150%"}
polygons_whole$total <- spe$total
polygons_whole$target_counts <- spe$target_counts
polygons_whole$control_counts <- spe$control_counts
polygons_whole$target_area_ratio <- spe$target_area_ratio
polygons_whole$cell_area <- spe$cell_area
polygons_whole$nucleus_area <- spe$nucleus_area
```

```{r, out.width="150%", out.height="150%"}
spe$flag_score_hw <- 1/(1 + exp(-0.4*spe$target_counts/spe$cell_area + 0.4*abs(spe$log2_AspectRatio)))
summary(spe$flag_score_hw)

polygons_whole$flag_score_hw <- spe$flag_score_hw

tm_shape(polygons_whole) + 
  tm_fill(col = "flag_score_hw") +
  tm_layout(legend.outside = TRUE)
```

# Flag threshold map with AspectRatio in flag score formula and scuttle outliers {.tabset .tabset-fade .tabset-pills}

```{r}
spe$outlier_target <- scuttle::isOutlier(spe$target_counts, type = "lower", nmad = 1)
polygons_whole$outlier_target <- ifelse(spe$outlier_target == TRUE, "red", "grey80")

spe$outlier_targetarea <- scuttle::isOutlier(spe$target_area_ratio, type = "lower", nmad = 1)
polygons_whole$outlier_targetarea <- ifelse(spe$outlier_targetarea == TRUE, "red", "grey80")

spe$outlier_cellarea <- scuttle::isOutlier(colData(spe)$cell_area, type = "both", nmad = 1)
polygons_whole$outlier_cellarea <- ifelse(spe$outlier_cellarea == TRUE, "red", "grey80")

spe$outlier_nucleusarea <- scuttle::isOutlier(spe$nucleus_area, type = "both", nmad = 1)
polygons_whole$outlier_nucleusarea<- ifelse(spe$outlier_nucleusarea == TRUE, "red", "grey80")
polygons_whole$outlier_nucleusarea <- replace_na(polygons_whole$outlier_nucleusarea, "grey80")

spe$outlier_loghw <- scuttle::isOutlier(spe$log2_AspectRatio, type = "both", nmad = 1)
polygons_whole$outlier_loghw<- ifelse(spe$outlier_loghw == TRUE, "red", "grey80")

spe$outlier_flagscorehw <- scuttle::isOutlier(spe$flag_score_hw, type = "lower", nmad = 1)
polygons_whole$outlier_flagscorehw <- ifelse(spe$outlier_flagscorehw == TRUE, "red", "grey80")
```


```{r, out.width="300%", out.height="900%"}
p1 <- tm_shape(polygons_whole) + 
  tm_fill(col= "outlier_flagscorehw", alpha = 0.5) +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Flag score outliers",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p2 <- tm_shape(polygons_whole) + 
  tm_fill(col= "outlier_target", alpha = 0.5) +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Target counts outliers",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p3 <- tm_shape(polygons_whole) + 
  tm_fill(col= "outlier_targetarea", alpha = 0.5) +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Target counts/cell area outliers",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p4 <- tm_shape(polygons_whole) + 
  tm_fill(col= "outlier_cellarea", alpha = 0.5) +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Cell area outliers",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p5 <- tm_shape(polygons_whole) + 
  tm_fill(col= "outlier_nucleusarea", alpha = 0.5) +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Nucleus area outliers",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p6 <- tm_shape(polygons_whole) + 
  tm_fill(col= "outlier_loghw", alpha = 0.5) +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Logged AspectRatio outliers",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

tmap_arrange(p1, p2, p3, p4, p5, p6, ncol = 2)
```

## Flagged cell stats

```{r, out.width="150%", out.height="150%"}

out_fence <- c(paste(round(attr(spe$outlier_flagscorehw,"thresholds")["lower"], 2), round(attr(spe$outlier_flagscorehw,"thresholds")["higher"], 2)),
               paste(round(attr(spe$outlier_target,"thresholds")["lower"], 2), round(attr(spe$outlier_target,"thresholds")["higher"], 2)),
               paste(round(attr(spe$outlier_targetarea,"thresholds")["lower"], 2), round(attr(spe$outlier_targetarea,"thresholds")["higher"], 2)),
               paste(round(attr(spe$outlier_cellarea,"thresholds")["lower"], 2), round(attr(spe$outlier_cellarea,"thresholds")["higher"], 2)),
               paste(round(attr(spe$outlier_nucleusarea,"thresholds")["lower"], 2), round(attr(spe$outlier_nucleusarea,"thresholds")["higher"], 2)),
               paste(round(attr(spe$outlier_loghw,"thresholds")["lower"], 2), round(attr(spe$outlier_loghw,"thresholds")["higher"], 2)))

flagged_cells <- c(sum(polygons_whole$outlier_flagscorehw=="red", na.rm = T),
                   sum(polygons_whole$outlier_target=="red", na.rm = T),
                   sum(polygons_whole$outlier_targetarea=="red", na.rm = T),
                   sum(polygons_whole$outlier_cellarea=="red",na.rm = T),
                   sum(polygons_whole$outlier_nucleusarea=="red",na.rm = T),
                   sum(polygons_whole$outlier_loghw=="red",na.rm = T))

unflagged_cells <- c(sum(polygons_whole$outlier_flagscorehw=="grey80", na.rm = T),
                   sum(polygons_whole$outlier_target=="grey80", na.rm = T),
                   sum(polygons_whole$outlier_targetarea=="grey80", na.rm = T),
                   sum(polygons_whole$outlier_cellarea=="grey80",na.rm = T),
                   sum(polygons_whole$outlier_nucleusarea=="grey80",na.rm = T),
                   sum(polygons_whole$outlier_loghw=="grey80",na.rm = T))

percent_flagged <- round(flagged_cells/dim(polygons_whole)[1]*100, 2)

poly <- data.frame(out_fence = out_fence, unflagged_cells = unflagged_cells, flagged_cells = flagged_cells, percent_flagged = percent_flagged, row.names = c("Flag score", "Target counts", "Target counts/cell area", "Cell area in um", "Nucleus area in um", "Logged aspect ratio"))

kable(poly, format = "html", col.names = c("Outlier thresholds", "Unflagged cells", "Flagged cells", "% flagged cells"), align = "c", caption = paste0("<center>Flagged cells by different metrics, total cells in whole sample: ", dim(polygons_whole)[1], ", flagged cells in whole sample: ", sum(polygons_whole$outlier_flagscorehw=="red"), "</center>"), padding = 1L)
```
#
#

## Big flag threshold map 

```{r, out.width="150%", out.height="150%"}
plot_list <- list(p1, p2, p3, p4, p5, p6)

for(i in 1:length(plot_list)){
  print(plot_list[[i]])
}
```

# Flag threshold map with AspectRatio in flag score formula and thresholds {.tabset .tabset-fade .tabset-pills}

```{r, out.width="300%", out.height="900%"}
metric_color <- ifelse(polygons_whole$flag_score_hw < quantile(spe$flag_score_hw, probs = 0.01), "red", "grey80")
target_color <- ifelse(polygons_whole$target_counts < quantile(spe$target_counts, probs = 0.15), "red", "grey80")
count_area_ratio_color <- ifelse(polygons_whole$target_area_ratio < quantile(spe$target_area_ratio, probs = 0.15), "red", "grey80")
area_color_lo <- ifelse(polygons_whole$cell_area < quantile(spe$cell_area, probs = 0.15), "red", "grey80")
area_color_hi <- ifelse(polygons_whole$cell_area > quantile(spe$cell_area, probs = 0.85), "red", "grey80")
nucleus_color_lo <- ifelse(polygons_whole$nucleus_area < quantile(spe$nucleus_area, probs = 0.15), "red", "grey80")
nucleus_color_hi <- ifelse(polygons_whole$nucleus_area > quantile(spe$nucleus_area, probs = 0.85), "red", "grey80")
log_hw_color_lo <- ifelse(polygons_whole$log2_AspectRatio < quantile(spe$log2_AspectRatio, probs = 0.15), "red", "grey80")
log_hw_color_hi <- ifelse(polygons_whole$log2_AspectRatio > quantile(spe$log2_AspectRatio, probs = 0.85), "red", "grey80")

polygons_whole <- cbind(polygons_whole, metric_color = metric_color, target_color = target_color, count_area_ratio_color = count_area_ratio_color, area_color_lo = area_color_lo, area_color_hi = area_color_hi, nucleus_color_lo = nucleus_color_lo, nucleus_color_hi = nucleus_color_hi, log_hw_color_lo = log_hw_color_lo, log_hw_color_hi = log_hw_color_hi)

p1 <- tm_shape(polygons_whole) + 
  tm_fill(col= "metric_color", alpha = 0.5) +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Flag score below 1st percentile",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p2 <- tm_shape(polygons_whole) + 
  tm_fill(col= "target_color", alpha = 0.5) +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Target counts below 15th percentile",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p3 <- tm_shape(polygons_whole) + 
  tm_fill(col= "count_area_ratio_color", alpha = 0.5) +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Target counts/custom um area below 15th percentile",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p4 <- tm_shape(polygons_whole) + 
  tm_fill(col= "area_color_lo", alpha = 0.5) +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Cell area below 15th percentile",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p5 <- tm_shape(polygons_whole) + 
  tm_fill(col= "area_color_hi", alpha = 0.5) +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Cell area above 85th percentile",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p6 <- tm_shape(polygons_whole) + 
tm_fill(col = "nucleus_color_lo", alpha = 0.5) +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Nucleus area below 15th percentile",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p7 <- tm_shape(polygons_whole) + 
tm_fill(col = "nucleus_color_hi", alpha = 0.5) +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Nucleus area above 85th percentile",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p8 <- tm_shape(polygons_whole) + 
  tm_fill(col= "log_hw_color_lo", alpha = 0.5) +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Logged AspectRatio below 15th percentile",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p9 <- tm_shape(polygons_whole) + 
  tm_fill(col= "log_hw_color_hi", alpha = 0.5) +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Logged AspectRatio above 85th percentile",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

tmap_arrange(p1, p2, p3, p4, p5, p6, p7, p8, p9, ncol = 2)
```

```{r, out.width="150%", out.height="150%"}

flag_value <- c(round(quantile(polygons_whole$flag_score_hw, probs = 0.01), 2),
                round(quantile(polygons_whole$target_counts, probs = 0.15), 2),
                round(quantile((polygons_whole$target_area_ratio), probs = 0.15), 2),
                round(quantile(polygons_whole$cell_area, probs = 0.15), 2),
                round(quantile(polygons_whole$cell_area, probs = 0.85), 2),
                round(quantile(polygons_whole$nucleus_area, probs = 0.15), 2),
                round(quantile(polygons_whole$nucleus_area, probs = 0.85), 2),
                round(quantile(polygons_whole$log2_AspectRatio, probs = 0.15), 2),
                round(quantile(polygons_whole$log2_AspectRatio, probs = 0.85), 2))

flagged_cells <- c(sum(polygons_whole$metric_color=="red", na.rm = T),
                   sum(polygons_whole$target_color=="red", na.rm = T),
                   sum(polygons_whole$count_area_ratio_color=="red", na.rm = T),
                   sum(polygons_whole$area_color_lo=="red", na.rm = T),
                   sum(polygons_whole$area_color_hi=="red", na.rm = T),
                   sum(polygons_whole$nucleus_color_lo=="red", na.rm = T),
                   sum(polygons_whole$nucleus_color_hi=="red", na.rm = T),
                   sum(polygons_whole$log_hw_color_lo=="red", na.rm = T),
                   sum(polygons_whole$log_hw_color_hi=="red", na.rm = T))

unflagged_cells <- c(sum(polygons_whole$metric_color=="grey80", na.rm = T),
                   sum(polygons_whole$target_color=="grey80", na.rm = T),
                   sum(polygons_whole$count_area_ratio_color=="grey80", na.rm = T),
                   sum(polygons_whole$area_color_lo=="grey80", na.rm = T),
                   sum(polygons_whole$area_color_hi=="grey80", na.rm = T),
                   sum(polygons_whole$nucleus_color_lo=="grey80", na.rm = T),
                   sum(polygons_whole$nucleus_color_hi=="grey80", na.rm = T),
                   sum(polygons_whole$log_hw_color_lo=="grey80", na.rm = T),
                   sum(polygons_whole$log_hw_color_hi=="grey80", na.rm = T))

percent_flagged <- round(flagged_cells/dim(polygons_whole)[1]*100, 2)

poly <- data.frame(threshold = c(0.01, 0.15, 0.15, 0.15, 0.85, 0.15, 0.85, 0.15, 0.85), flag_value = flag_value, unflagged_cells = unflagged_cells, flagged_cells = flagged_cells, percent_flagged = percent_flagged, row.names = c("Flag score", "Target counts", "Target counts/cell area", "Cell area in um low", "Cell area in um high", "Nucleus area in um low", "Nucleus area in um high", "Logged aspect ratio low", "Logged aspect ratio high"))

kable(poly, format = "html", col.names = c("Threshold", "Flag value", "Unflagged cells", "Flagged cells", "% flagged cells"), align = "c", caption = paste0("<center>Flagged cells by different metrics, total cells in whole sample: ", dim(polygons_whole)[1], ", flagged cells in whole sample: ", sum(spe$flag_score < quantile(spe$flag_score, probs = 0.15, na.rm = T)), "</center>"), padding = 1L)
```

#
#

## Big flag threshold map 

```{r, out.width="150%", out.height="150%"}
plot_list <- list(p1, p2, p3, p4, p5, p6, p7, p8, p9)

for(i in 1:length(plot_list)){
  print(plot_list[[i]])
}
```



Computing mock fovs starting from first identified section size {.tabset .tabset-fade .tabset-pills}

```{r}
meta_df <- data.frame(colData(spe))

# computing the fov size using the max and the min of the first fov coordinates
fov_size_x <- 3520*0.12
fov_size_y <- 2960*0.12

# binning coordinates using fov_size
x_bin <- cut(spe$CenterX_global_um, seq(min(spe$CenterX_global_um), max(spe$CenterX_global_um), fov_size_x), include.lowest = TRUE)
y_bin <- cut(spe$CenterY_global_um, seq(min(spe$CenterY_global_um), max(spe$CenterY_global_um), fov_size_y), include.lowest = TRUE)

# creating grid cell indexes
levels(x_bin) <- c(1:25)
levels(y_bin) <- LETTERS[1:9]

# combining the indexes
mock_fov <- paste(x_bin, y_bin)

# adding the cell indexes to the Spatial object metadata
spe$mock_fov <- mock_fov
polygons_whole$mock_fov <- spe$mock_fov
```

# Mock fov 12E with AspectRatio and scuttle outliers {.tabset .tabset-fade .tabset-pills}

```{r, out.width="150%", out.height="150%"}
polygons12E <- filter(polygons_whole, mock_fov == "12 E")

tm_shape(polygons12E) + 
  tm_fill(col= "flag_score_hw", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE)

```

## Flag threshold map

```{r, out.width="300%", out.height="900%"}

p1 <- tm_shape(polygons12E) + 
  tm_fill(col= "outlier_flagscorehw", alpha = 0.5) +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Flag score outliers",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p2 <- tm_shape(polygons12E) + 
  tm_fill(col= "outlier_target", alpha = 0.5) +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Target counts outliers",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p3 <- tm_shape(polygons12E) + 
  tm_fill(col= "outlier_targetarea", alpha = 0.5) +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Target counts/cell area outliers",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p4 <- tm_shape(polygons12E) + 
  tm_fill(col= "outlier_cellarea", alpha = 0.5) +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Cell area outliers",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p5 <- tm_shape(polygons12E) + 
  tm_fill(col= "outlier_nucleusarea", alpha = 0.5) +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Nucleus area outliers",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p6 <- tm_shape(polygons12E) + 
  tm_fill(col= "outlier_loghw", alpha = 0.5) +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Logged AspectRatio outliers",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

tmap_arrange(p1, p2, p3, p4, p5, p6, ncol = 2)
```

## Flagged cell stats

```{r, out.width="150%", out.height="150%"}
out_fence <- c(paste(round(attr(spe$outlier_flagscorehw,"thresholds")["lower"], 2), round(attr(spe$outlier_flagscorehw,"thresholds")["higher"], 2)),
               paste(round(attr(spe$outlier_target,"thresholds")["lower"], 2), round(attr(spe$outlier_target,"thresholds")["higher"], 2)),
               paste(round(attr(spe$outlier_targetarea,"thresholds")["lower"], 2), round(attr(spe$outlier_targetarea,"thresholds")["higher"], 2)),
               paste(round(attr(spe$outlier_cellarea,"thresholds")["lower"], 2), round(attr(spe$outlier_cellarea,"thresholds")["higher"], 2)),
               paste(round(attr(spe$outlier_nucleusarea,"thresholds")["lower"], 2), round(attr(spe$outlier_nucleusarea,"thresholds")["higher"], 2)),
               paste(round(attr(spe$outlier_loghw,"thresholds")["lower"], 2), round(attr(spe$outlier_loghw,"thresholds")["higher"], 2)))

flagged_cells <- c(sum(polygons12E$outlier_flagscorehw=="red", na.rm = T),
                   sum(polygons12E$outlier_target=="red", na.rm = T),
                   sum(polygons12E$outlier_targetarea=="red", na.rm = T),
                   sum(polygons12E$outlier_cellarea=="red",na.rm = T),
                   sum(polygons12E$outlier_nucleusarea=="red",na.rm = T),
                   sum(polygons12E$outlier_loghw=="red",na.rm = T))

unflagged_cells <- c(sum(polygons12E$outlier_flagscorehw=="grey80", na.rm = T),
                   sum(polygons12E$outlier_target=="grey80", na.rm = T),
                   sum(polygons12E$outlier_targetarea=="grey80", na.rm = T),
                   sum(polygons12E$outlier_cellarea=="grey80",na.rm = T),
                   sum(polygons12E$outlier_nucleusarea=="grey80",na.rm = T),
                   sum(polygons12E$outlier_loghw=="grey80",na.rm = T))

percent_flagged <- round(flagged_cells/dim(polygons12E)[1]*100, 2)

poly12E <- data.frame(out_fence = out_fence, unflagged_cells = unflagged_cells, flagged_cells = flagged_cells, percent_flagged = percent_flagged, row.names = c("Flag score", "Target counts", "Target counts/cell area", "Cell area in um", "Nucleus area in um", "Logged aspect ratio"))

kable(poly12E, format = "html", col.names = c("Outlier thresholds", "Unflagged cells", "Flagged cells", "% flagged cells"), align = "c", caption = paste0("<center>Flagged cells by different metrics, total cells in whole sample: ", dim(polygons_whole)[1], ", flagged cells in whole sample: ", sum(polygons_whole$outlier_flagscorehw=="red"), " , total cells in mock fov: ", dim(polygons12E)[1],"</center>"), padding = 1L)
```

#
#

## Big flag threshold map

```{r, out.width="150%", out.height="150%"}
plot_list <- list(p1, p2, p3, p4, p5, p6)

for(i in 1:length(plot_list)){
  print(plot_list[[i]])
}
```

# Mock fov 12E with AspectRatio and percentile outliers {.tabset .tabset-fade .tabset-pills}

```{r, out.width="300%", out.height="900%"}
p1 <- tm_shape(polygons12E) + 
  tm_fill(col= "metric_color", alpha = 0.5) +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Flag score below 1st percentile",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p2 <- tm_shape(polygons12E) + 
  tm_fill(col= "target_color", alpha = 0.5) +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Target counts below 15th percentile",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p3 <- tm_shape(polygons12E) + 
  tm_fill(col= "count_area_ratio_color", alpha = 0.5) +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Target counts/custom um area below 15th percentile",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p4 <- tm_shape(polygons12E) + 
  tm_fill(col= "area_color_lo", alpha = 0.5) +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Cell area below 15th percentile",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p5 <- tm_shape(polygons12E) + 
  tm_fill(col= "area_color_hi", alpha = 0.5) +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Cell area above 85th percentile",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p6 <- tm_shape(polygons12E) + 
tm_fill(col = "nucleus_color_lo", alpha = 0.5) +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Nucleus area below 15th percentile",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p7 <- tm_shape(polygons12E) + 
tm_fill(col = "nucleus_color_hi", alpha = 0.5) +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Nucleus area above 85th percentile",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p8 <- tm_shape(polygons12E) + 
  tm_fill(col= "log_hw_color_lo", alpha = 0.5) +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Logged AspectRatio below 15th percentile",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p9 <- tm_shape(polygons12E) + 
  tm_fill(col= "log_hw_color_hi", alpha = 0.5) +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Logged AspectRatio above 85th percentile",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

tmap_arrange(p1, p2, p3, p4, p5, p6, p7, p8, p9, ncol = 2)
```


```{r, out.width="150%", out.height="150%"}

flag_value <- c(round(quantile(polygons12E$flag_score_hw, probs = 0.01), 2),
                round(quantile(polygons12E$target_counts, probs = 0.15), 2),
                round(quantile((polygons12E$target_area_ratio), probs = 0.15), 2),
                round(quantile(polygons12E$cell_area, probs = 0.15), 2),
                round(quantile(polygons12E$cell_area, probs = 0.85), 2),
                round(quantile(polygons12E$nucleus_area, probs = 0.15), 2),
                round(quantile(polygons12E$nucleus_area, probs = 0.85), 2),
                round(quantile(polygons12E$log2_AspectRatio, probs = 0.15), 2),
                round(quantile(polygons12E$log2_AspectRatio, probs = 0.85), 2))

flagged_cells <- c(sum(polygons12E$metric_color=="red", na.rm = T),
                   sum(polygons12E$target_color=="red", na.rm = T),
                   sum(polygons12E$count_area_ratio_color=="red", na.rm = T),
                   sum(polygons12E$area_color_lo=="red", na.rm = T),
                   sum(polygons12E$area_color_hi=="red", na.rm = T),
                   sum(polygons12E$nucleus_color_lo=="red", na.rm = T),
                   sum(polygons12E$nucleus_color_hi=="red", na.rm = T),
                   sum(polygons12E$log_hw_color_lo=="red", na.rm = T),
                   sum(polygons12E$log_hw_color_hi=="red", na.rm = T))

unflagged_cells <- c(sum(polygons12E$metric_color=="grey80", na.rm = T),
                   sum(polygons12E$target_color=="grey80", na.rm = T),
                   sum(polygons12E$count_area_ratio_color=="grey80", na.rm = T),
                   sum(polygons12E$area_color_lo=="grey80", na.rm = T),
                   sum(polygons12E$area_color_hi=="grey80", na.rm = T),
                   sum(polygons12E$nucleus_color_lo=="grey80", na.rm = T),
                   sum(polygons12E$nucleus_color_hi=="grey80", na.rm = T),
                   sum(polygons12E$log_hw_color_lo=="grey80", na.rm = T),
                   sum(polygons12E$log_hw_color_hi=="grey80", na.rm = T))

percent_flagged <- round(flagged_cells/dim(polygons12E)[1]*100, 2)

poly12E <- data.frame(threshold = c(0.01, 0.15, 0.15, 0.15, 0.85, 0.15, 0.85, 0.15, 0.85), flag_value = flag_value, unflagged_cells = unflagged_cells, flagged_cells = flagged_cells, percent_flagged = percent_flagged, row.names = c("Flag score", "Target counts", "Target counts/cell area", "Cell area in um low", "Cell area in um high", "Nucleus area in um low", "Nucleus area in um high", "Logged aspect ratio low", "Logged aspect ratio high"))

kable(poly12E, format = "html", col.names = c("Threshold", "Flag value", "Unflagged cells", "Flagged cells", "% flagged cells"), align = "c", caption = paste0("<center>Flagged cells by different metrics, total cells in whole sample: ", dim(polygons_whole)[1], ", flagged cells in whole sample: ", sum(spe$flag_score < quantile(spe$flag_score, probs = 0.15, na.rm = T)), " , total cells in mock fov: ", dim(polygons12E)[1],"</center>"), padding = 1L)
```

#
#

## Big flag threshold map 

```{r, out.width="150%", out.height="150%"}
plot_list <- list(p1, p2, p3, p4, p5, p6, p7, p8, p9)

for(i in 1:length(plot_list)){
  print(plot_list[[i]])
}
```
