---
title: "04_03_cosmx_human_tnbc_25027_prepro_metric_report"
author: "Benedetta Banzi"
date: "2024-04-03"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---


<STYLE TYPE="text/css">
<!--
  td{
    font-family: Arial; 
    font-size: 10 pt;
    padding:0px;
    cellpadding="0";
    cellspacing="0"
  }
  th {
    font-family: Arial; 
    font-size: 10 pt;
    height: 20px;
    font-weight: bold;
    text-align: "center";
    background-color: #ccccff;
    padding-left: 15px;
    padding-right: 15px;
  }
  table { 
    border-spacing: 4px;
    border-collapse: collapse;
  }
--->
</STYLE>


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Preliminary steps

Library import

```{r}
library(SpatialExperiment)
library(data.table)
library(scater) # it imports scuttle
library(cowplot)
library(ggplot2)
theme_set(theme_bw())
library(matrixStats)
library(dplyr)
library(tidyr)
library(tibble)
library(arrow)
library(scales)
library(sf)
library(tmap)
library(knitr)
library(grid)
library(ggpointdensity)
library(ggpubr)
library(spdep)
library(ggExtra)
library(DT)
library(e1071)
library(robustbase)
```

Read-in function

```{r}
readCosmxSPE <- function(dirname = dirname, 
                         countmatfpattern = "exprMat_file.csv", 
                         metadatafpattern = "metadata_file.csv", 
                         coord_names = c("CenterX_global_px",
                                         "CenterY_global_px")){
  
  countmat_file <- file.path(dirname, list.files(dirname, countmatfpattern))
  metadata_file <- file.path(dirname, list.files(dirname, metadatafpattern))
  
  # Read in 
  countmat <- data.table::fread(countmat_file)
  metadata <- data.table::fread(metadata_file)
  
  # Count matrix   
  counts <- merge(countmat, metadata[, c("fov", "cell_ID")])
  counts <- subset(counts, select = -c(fov, cell_ID))
  cell_codes <- rownames(counts)
  features <- colnames(counts)
  counts <- t(counts)
  
  rownames(counts) <- features
  colnames(counts) <- cell_codes
  
  # rowData (does not exist)
  
  # colData
  colData <- merge(metadata, countmat[, c("fov", "cell_ID")])
  
  spe <- SpatialExperiment::SpatialExperiment(
    assays = list(counts = counts),
    # rowData = rowData,
    colData = colData,
    spatialCoordsNames = coord_names
  )
  
  return(spe)
}
```

Setting directories

```{r, class.source = 'fold-show'}
dirname <- "/data01/Shared/Benedetta/spatial_transcriptomics/CosMx/Datasets/TNBC/UAP7_02527"
count_pattern <- "UAP7_02527_exprMat_file.csv"
meta_pattern <- "UAP7_02527_metadata_file.csv"
sample_name <- "CosMx_human_TNBC_02527"
```

# CosMx dataset preprocessing {.tabset .tabset-fade .tabset-pills}

Data import and SpatialExperiment object creation

```{r}
spe <- readCosmxSPE(dirname = dirname, countmatfpattern = count_pattern, metadatafpattern = meta_pattern)
spe <- scuttle::addPerCellQC(spe, subsets=list(NegProb = grep("^Negative", rownames(spe)), SysCtrl = grep("^SystemControl", rownames(spe))))
names(colData(spe)@listData)
spe$control_counts <- rowSums(data.frame(colData(spe))[,grep("^subsets_.*_sum$", names(colData(spe)))])
spe$control_detected <- rowSums(data.frame(colData(spe))[,grep("^subsets_.*_detected$", names(colData(spe)))])
```

Dataset dimensions

```{r}
dim(spe)
```

Adding variables

```{r, class.source = 'fold-show'}
fov_size <- 4256

names(colData(spe))[names(colData(spe)) == "cell_ID"] <- "cell_id"

spe$id_fov <- paste(spe$cell_id, spe$fov, sep = "_")

# Importing global coordinates from spatialCoords slot since they are not available in colData
spe$CenterX_global_px <- spatialCoords(spe)[,1]
spe$CenterY_global_px <- spatialCoords(spe)[,2]

# Converting global coordinates to um 
spe$CenterX_global_um <- spe$CenterX_global_px*0.12
spe$CenterY_global_um <- spe$CenterY_global_px*0.12 

# Logged total counts
spe$logtot <- log10(spe$total)

# Converting area in um
spe$um_area <- spe$Area*0.0144

# Logged width/height ratio so there are no holes in the distribution in the middle and also negative values can be appreciated
spe$log2_AspectRatio <- log2(spe$AspectRatio)

# Keeping only target counts by subtracting counts referred to control probes
spe$target_counts <- spe$total - spe$control_counts

# Keeping only detected features by subtracting negative control probes
spe$target_detected <- spe$detected - spe$control_detected

# Proportion between counts referred only to target probes and detected features
spe$tot_det_ratio <- spe$target_counts/spe$target_detected

# Control probe counts/total counts
ctrl_total_ratio <- spe$control_counts/spe$total
ctrl_total_ratio <- replace_na(ctrl_total_ratio, 0)
spe$ctrl_total_ratio <- ctrl_total_ratio
```

Annotation file loading to get cell types

```{r}
metadata <- data.frame(colData(spe))
anno_file <- fread(file.path(dirname, "UAP7_02527_clusterAnnot@1_simple.txt"), sep = "\t", header = T)
colnames(anno_file) <- c("id_fov", "cell_type", "type_order", "type_color")

anno_df <- subset(anno_file, select = c("id_fov", "cell_type"))

metadata$id_fov <- paste(metadata$cell_id, metadata$fov, sep = "_")
metadata <- left_join(metadata, anno_df, by = join_by(id_fov))
table(table(metadata$fov) == table(colData(spe)$fov))
```

```{r}
metadata <- metadata[metadata$id_fov %in% anno_df$id_fov,]
spe <- spe[, spe$id_fov %in% metadata$id_fov]
dim(spe)
```

Checking whether the same fov were kept after subsetting and the number of remnant cells

```{r}
table(table(metadata$fov) == table(colData(spe)$fov))
spe$cell_type <- metadata$cell_type
```

Some cells are missing in the annotation file. They were preprocessed because of their low-signal. Proceeding without them.

```{r, out.width="150%", out.height="300%"}
celltype_palette <- c(
  "discard"	= "gray11", 
"RBCs" = "gray47",
"Plasma_cells" =	"blue",
"Plasmablasts" =	"royalblue1",
"Mix_CTLs_Plasma_cells" =	"dodgerblue4",
"CTLs" =	"deepskyblue",
"Mix_CXCL13+Tcells_Plasma_cells" =	"cyan",
"CXCL13+Tcells" =	"lightblue1",
"Tcells" =	"darkseagreen1",
"Tcell_Monocyte_Pairs" = "darkolivegreen1",
"Monocytes" =	"limegreen",
"LipoTAMs" =	"green3",
"Mixed_LipoTAMs" =	"darkgreen",
"Mix_LipoTAMs_Plasma_cells" =	"yellow4",
"TAMs" =	"yellow",
"iTAMs" = "greenyellow",
"OPN+TAMs" =	"olivedrab1",
"DCs" =	"seagreen2",
"Mast_cells" =	"lightyellow",
"Mix_MastCells_Plasma_cells" =	"goldenrod4",
"Mix_CAFs_Plasma_cells" =	"goldenrod1",
"Mixed_CAFs" =	"lightgoldenrod1",
"CAFs" =	"lightgoldenrod4",
"Stromal_cells" =	"darkorange1",
"Pericytes" =	"plum1",
"Lymph_ECs" =	"orchid4",
"Blood_ECs" =	"orchid1",
"BC_cells" =	"lightpink",
"LC3+BC_cells" =	"lightcoral",
"MGP+BC_cells" =	"orangered1",
"Cycling_BC_cells" =	"deeppink1",
"Hypoxic_BC_cells" = "darkmagenta",
"Mixed_BC_cells" =	"red4"
)

celltype_color <- c()
for(i in 1:dim(metadata)[1]){
  celltype <- anno_df$cell_type[i]
  n <- match(celltype, names(celltype_palette))
  celltype_color[i] <- celltype_palette[n]
}

metadata$celltype_color <- as.factor(celltype_color)
spe$celltype_color <- as.factor(celltype_color)
```

Custom area calculation from custom polygons

Loading custom polygon file and setting the correct format as sf object

```{r}
spat_obj <- fread(file.path(dirname, "UAP7_02527-polygons.csv"))
names(spat_obj)[names(spat_obj) == "cellID"] <- "cell_id"
str(spat_obj)
```

```{r}
spat_obj$id_fov <- paste(spat_obj$cell_id, spat_obj$fov, sep = "_")
length(table(spat_obj$id_fov))
```

```{r}
table(names(table(spat_obj$fov)) == names(table(metadata$fov)))
```

```{r}
sf_global <- st_as_sf(spat_obj, coords = c("x_global_px", "y_global_px"))

polygons_whole <- sf_global %>%
  group_by(fov, cell_id) %>%
  summarize(geometry = st_combine(geometry))#st_combine = combine several feature geometries into one, without unioning or resolving internal boundaries

polygons_whole <- st_cast(polygons_whole, "POLYGON")
```

Checking if the computed polygons are valid and removing elements that could result in geometry errors (Voyager)

```{r, class.source = 'fold-show'}
valid_poly <- st_is_valid(polygons_whole)
table(valid_poly)
```

Computing buffer (a circumscribing space) around geometries to handle invalid geometries using 0 distance, so keeping it adherent to the original space
occupied by the cell. (Voyager)

```{r, class.source = 'fold-show'}
polygons_whole <- st_buffer(polygons_whole, dist = 0)
table(st_geometry_type(polygons_whole$geometry))
```

```{r}
cellids <- c() 
for(i in 1:dim(polygons_whole)[1]){
	cellids[i] <- attr(polygons_whole$geometry[[i]], "class")[2]
	cellids[i] <- cellids[i] == "MULTIPOLYGON"
	
}

polygons_whole <- polygons_whole[cellids == FALSE, ]
table(st_geometry_type(polygons_whole))

spe <- spe[,spe$cell_id %in% polygons_whole$cell_id]
```

Adding custom polygons to cell metadata

```{r}
polygons_whole$custom_area <- st_area(polygons_whole$geometry)
polygons_whole$custom_um_area <- polygons_whole$custom_area*0.0144
polygons_whole$id_fov <- paste(polygons_whole$cell_id, polygons_whole$fov, sep = "_")

metadata <- metadata[metadata$id_fov %in% polygons_whole$id_fov,]
spe <- spe[, spe$id_fov %in% metadata$id_fov]
dim(spe)

metadata <- left_join(metadata, data.frame(id_fov = polygons_whole$id_fov, custom_area = polygons_whole$custom_area, custom_um_area = polygons_whole$custom_um_area), by = join_by(id_fov))
spe$custom_area <- metadata$custom_area
spe$custom_um_area <- metadata$custom_um_area

# Target area ratio
spe$target_area_ratio <- spe$target_counts/spe$custom_um_area
polygons_whole$target_area_ratio <- spe$target_area_ratio

polygons_whole$ctrl_total_ratio <- spe$ctrl_total_ratio

# log2 AspectRatio
polygons_whole$log2_AspectRatio <- spe$log2_AspectRatio
```


# Large-scale to small-scale {.tabset .tabset-fade .tabset-pills}

Approaching dataset preprocessing starting from global-scale, then fov-focused exploratory analysis

## Global sample map

To view the spatial organization of both the sample as it is and FOVs

```{r, out.width="600%", out.height="150%"}
fov_pattern <- "UAP7_02527_fov_positions_file.csv"
fovpos_file <- file.path(dirname, list.files(dirname, fov_pattern))

fov_positions <- data.table::fread(fovpos_file, header = T)
fov_positions <- rename(fov_positions, fov = FOV)
fov_positions <- fov_positions[fov_positions$fov %in% unique(metadata$fov),]

# image theme
global_map_theme <- function(back.color="black",back.border=NA,title.col="white") {
  theme(aspect.ratio = 1,
        panel.border = element_blank(),
        legend.key = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks = element_blank(),
        axis.text.y = element_blank(),
        axis.text.x = element_blank(),
        panel.grid = element_blank(),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        panel.background = element_rect(fill = "transparent",colour = NA),
        plot.title = element_text(color = title.col,hjust = 0.5,face = "bold"),
        plot.background = element_rect(fill = "transparent",colour = back.border))
}

fov_size <- 4256

ggplot() + 
  geom_point(data = metadata, mapping = aes(x = CenterX_global_px, y = CenterY_global_px), colour = "darkmagenta", fill="darkmagenta", size = 0.05, alpha = 0.2) +
  annotate("rect", xmin = fov_positions$X_px - 1/32*fov_size, 
         xmax = fov_positions$X_px + fov_size - 1/32*fov_size, 
         ymin = fov_positions$Y_px - fov_size - 1/16*fov_size, 
         ymax = fov_positions$Y_px - 1/16*fov_size,
         alpha = .2, color = "black",linewidth = 0.2) +
  geom_text(aes(x = fov_positions$X_px+11/10*fov_size/2,
                y = fov_positions$Y_px-11/10*fov_size/2,
                label = fov_positions$fov), color="black", size = 3) +
  ggtitle(sample_name) +
  global_map_theme(back.color = "white", back.border = "white", title.col = "black")
```

## Global sample map with cell types

```{r, out.width="150%", out.height="300%"}
ggplot(data = metadata, mapping = aes(x = CenterX_global_px, 
  y = CenterY_global_px, color = cell_type)) + 
  geom_point(size = 0.05) +
  scale_color_manual(values = celltype_palette) +
  ggtitle(sample_name) + 
  global_map_theme(back.color = "white", back.border = "white", 
  title.col = "black") +
  guides(colour = guide_legend(override.aes = list(size=5)))

```

## Global feature maps

To spot areas with anomalous feature values and spatial patterns on a global scale

```{r, out.width="150%", out.height="150%"}
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "total", point_size = 0.1) + ggtitle("Total counts per cell") + theme(aspect.ratio = 1)
```

```{r, out.width="150%", out.height="150%"}
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "logtot", point_size = 0.1) + ggtitle("Log-total counts per cell") + theme(aspect.ratio = 1)
```

```{r, out.width="150%", out.height="150%"}
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "detected", point_size = 0.1) + ggtitle("Detected features per cell") + theme(aspect.ratio = 1)
```

```{r, out.width="150%", out.height="150%"}
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "um_area", point_size = 0.1) + ggtitle("Area in μm per cell") + theme(aspect.ratio = 1)
```

```{r, out.width="150%", out.height="150%"}
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "custom_um_area", point_size = 0.1) + ggtitle("Custom area in μm per cell") + theme(aspect.ratio = 1)
```

```{r, out.width="150%", out.height="150%"}
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "log2_AspectRatio", point_size = 0.1) + ggtitle("Logged width/height ratio per cell") + theme(aspect.ratio = 1)
```

```{r, out.width="150%", out.height="150%"}
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "tot_det_ratio", point_size = 0.1) + ggtitle("Target counts/detected features ratio per cell") + theme(aspect.ratio = 1)
```

```{r, out.width="150%", out.height="150%"}
plotColData(spe, "CenterY_global_um", "CenterX_global_um", order_by = "control_counts", colour_by = "control_counts", point_size = 0.1) +
  scale_color_gradient(low = "white", high = "red", name = "control_counts") + ggtitle("Negative control probe counts per cell") +
  theme(aspect.ratio = 1,
        panel.background = element_rect(fill = "black",color = NA),
        plot.background = element_rect(fill = "black",color = NA),
        axis.title.x = element_text(color = "white"),
        axis.title.y = element_text(color = "white"),
        axis.ticks = element_line(color = "white"),
        axis.text.y = element_text(color = "white"),
        axis.text.x = element_text(color = "white"),
        axis.line = element_line(color = "white"),
        legend.title = element_text(color = "white"),
        legend.text = element_text(color = "white"),
        plot.title = element_text(color = "white"))
```

```{r, out.width="150%", out.height="150%"}
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "ctrl_total_ratio", point_size = 0.1) + ggtitle("Control counts/total counts ratio per cell") + theme(aspect.ratio = 1)
```

```{r}
threshold <- 0.20
spe$flagged_tot_20 <- as.factor(ifelse(colData(spe)$total < quantile(colData(spe)$total, probs = threshold), TRUE, FALSE))

threshold <- 0.10
spe$flagged_tot_10 <- as.factor(ifelse(colData(spe)$total < quantile(colData(spe)$total, probs = threshold), TRUE, FALSE))

threshold <- 0.05
spe$flagged_tot_5 <- as.factor(ifelse(colData(spe)$total < quantile(colData(spe)$total, probs = threshold), TRUE, FALSE))

threshold <- 0.10
spe$flagged_det_10 <- as.factor(ifelse(colData(spe)$detected < quantile(colData(spe)$detected, probs = threshold), TRUE, FALSE))

threshold <- 0.05
spe$flagged_det_5 <- as.factor(ifelse(colData(spe)$detected < quantile(colData(spe)$detected, probs = threshold), TRUE, FALSE))

threshold <- 0.10
spe$flagged_umarea_10 <- as.factor(ifelse(colData(spe)$um_area < quantile(colData(spe)$um_area, probs = threshold), TRUE, FALSE))

threshold <- 0.05
spe$flagged_umarea_5 <- as.factor(ifelse(colData(spe)$um_area < quantile(colData(spe)$um_area, probs = threshold), TRUE, FALSE))

threshold <- 0.90
spe$flagged_umarea_90 <- as.factor(ifelse(colData(spe)$um_area > quantile(colData(spe)$um_area, probs = threshold), TRUE, FALSE))

threshold <- 0.95
spe$flagged_umarea_95 <- as.factor(ifelse(colData(spe)$um_area > quantile(colData(spe)$um_area, probs = threshold), TRUE, FALSE))

threshold <- 0.10
spe$flagged_cumarea_10 <- as.factor(ifelse(colData(spe)$custom_um_area < quantile(colData(spe)$custom_um_area, probs = threshold), TRUE, FALSE))

threshold <- 0.05
spe$flagged_cumarea_5 <- as.factor(ifelse(colData(spe)$custom_um_area < quantile(colData(spe)$custom_um_area, probs = threshold), TRUE, FALSE))

threshold <- 0.90
spe$flagged_cumarea_90 <- as.factor(ifelse(colData(spe)$custom_um_area > quantile(colData(spe)$custom_um_area, probs = threshold), TRUE, FALSE))

threshold <- 0.95
spe$flagged_cumarea_95 <- as.factor(ifelse(colData(spe)$custom_um_area > quantile(colData(spe)$custom_um_area, probs = threshold), TRUE, FALSE))

threshold <- 0.10
spe$flagged_aratio_10 <- as.factor(ifelse(colData(spe)$log2_AspectRatio < quantile(colData(spe)$log2_AspectRatio, probs = threshold), TRUE, FALSE))

threshold <- 0.05
spe$flagged_aratio_5 <- as.factor(ifelse(colData(spe)$log2_AspectRatio < quantile(colData(spe)$log2_AspectRatio, probs = threshold), TRUE, FALSE))

threshold <- 0.90
spe$flagged_aratio_90 <- as.factor(ifelse(colData(spe)$log2_AspectRatio > quantile(colData(spe)$log2_AspectRatio, probs = threshold), TRUE, FALSE))

threshold <- 0.95
spe$flagged_aratio_95 <- as.factor(ifelse(colData(spe)$log2_AspectRatio > quantile(colData(spe)$log2_AspectRatio, probs = threshold), TRUE, FALSE))

threshold <- 0.10
spe$flagged_tdratio_10 <- as.factor(ifelse(colData(spe)$tot_det_ratio < quantile(colData(spe)$tot_det_ratio, probs = threshold, na.rm = T), TRUE, FALSE))

threshold <- 0.05
spe$flagged_tdratio_5 <- as.factor(ifelse(colData(spe)$tot_det_ratio < quantile(colData(spe)$tot_det_ratio, probs = threshold, na.rm = T), TRUE, FALSE))

threshold <- 1
spe$flagged_tdratio_atomx <- as.factor(ifelse(colData(spe)$tot_det_ratio <= threshold, TRUE, FALSE))

threshold <- 0.50
spe$flagged_negprob_50 <- as.factor(ifelse(colData(spe)$control_counts > quantile(colData(spe)$control_counts, probs = threshold, na.rm = T), TRUE, FALSE))

threshold <- 0.75
spe$flagged_negprob_75 <- as.factor(ifelse(colData(spe)$control_counts > quantile(colData(spe)$control_counts, probs = threshold, na.rm = T), TRUE, FALSE))

threshold <- 0.10
spe$flagged_negprob_10 <- as.factor(ifelse(colData(spe)$ctrl_total_ratio > quantile(colData(spe)$ctrl_total_ratio, probs = threshold, na.rm = T), TRUE, FALSE))

threshold <- 0.95
spe$flagged_negprob_95 <- as.factor(ifelse(colData(spe)$control_counts > quantile(colData(spe)$control_counts, probs = threshold, na.rm = T), TRUE, FALSE))

threshold <- 0.99
spe$flagged_negprob_99 <- as.factor(ifelse(colData(spe)$control_counts > quantile(colData(spe)$control_counts, probs = threshold, na.rm = T), TRUE, FALSE))

threshold <- 0.01
spe$flagged_cumarea_1 <- as.factor(ifelse(colData(spe)$custom_um_area < quantile(colData(spe)$custom_um_area, probs = threshold), TRUE, FALSE))

threshold <- 0.99
spe$flagged_cumarea_99 <- as.factor(ifelse(colData(spe)$custom_um_area > quantile(colData(spe)$custom_um_area, probs = threshold), TRUE, FALSE))

threshold <- 0.01
spe$flagged_dapi_1 <- as.factor(ifelse(colData(spe)$Mean.DAPI < quantile(colData(spe)$Mean.DAPI, probs = threshold), TRUE, FALSE))

threshold <- 0.05
spe$flagged_dapi_5 <- as.factor(ifelse(colData(spe)$Mean.DAPI < quantile(colData(spe)$Mean.DAPI, probs = threshold), TRUE, FALSE))

threshold <- 0.95
spe$flagged_dapi_95 <- as.factor(ifelse(colData(spe)$Mean.DAPI > quantile(colData(spe)$Mean.DAPI, probs = threshold), TRUE, FALSE))

threshold <- 0.99
spe$flagged_dapi_99 <- as.factor(ifelse(colData(spe)$Mean.DAPI > quantile(colData(spe)$Mean.DAPI, probs = threshold), TRUE, FALSE))

threshold <- 0.01
spe$flagged_target_1 <- as.factor(ifelse(colData(spe)$target_counts < quantile(colData(spe)$target_counts, probs = threshold), TRUE, FALSE))

threshold <- 0.05
spe$flagged_target_5 <- as.factor(ifelse(colData(spe)$target_counts < quantile(colData(spe)$target_counts, probs = threshold), TRUE, FALSE))
```

# Feature histograms 

To check feature distribution before highlighting where extreme values are located in the global sample map

```{r}
variables <- c("total", "detected", "logtot", "um_area", "log2_AspectRatio", "tot_det_ratio", "control_counts", "custom_um_area", "Mean.DAPI", "ctrl_total_ratio")
for (var in 1:length(variables)){
  n <- grep(paste0("^", variables[var], "$"), colnames(spe@colData))
  histo_var <- names(colData(spe))[n]
  p <- ggplot(data = data.frame(colData(spe))) +
    geom_histogram(aes(x = .data[[histo_var]]), fill = "#69b3a2") +
    ggtitle(paste0(histo_var))
  print(p)
}  
```


# First cell skimming with quantile thresholds {.tabset .tabset-fade .tabset-pills}

Coarse-grain cell filtering to remove gross segmentation errors.

## Control probes

```{r}
summary(spe$control_counts)
```

```{r, out.width="150%", out.height="150%"}
hist(spe$control_counts)
```

```{r, out.width="150%", out.height="150%"}
boxplot(spe$control_counts)
```


```{r, out.width="150%", out.height="150%"}
threshold <- 0.99
spe$flagged_negprob_99 <- as.factor(ifelse(colData(spe)$control_counts > quantile(colData(spe)$control_counts, probs = threshold, na.rm = T), TRUE, FALSE))
polygons_whole$flagged_negprob_99 <- ifelse(spe$flagged_negprob_99 == TRUE, "red", "grey80")

tm_shape(polygons_whole) + 
  tm_fill(col= "flagged_negprob_99") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = paste0("Negative probe counts above 99% threshold, value = ", 
                                round(quantile(colData(spe)$control_counts, probs = threshold, na.rm = T), 2), "\nFlagged cells = ",
                                table(colData(spe)$flagged_negprob_99)["TRUE"], " out of ", dim(spe)[2], ", ",
                                round(table(colData(spe)$flagged_negprob_99)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"),
            main.title.fontface = 2,
            main.title.size = 1, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))
```

```{r}
negprob_spe <- spe[,spe$flagged_negprob_99 == FALSE]
summary(negprob_spe$control_counts)
```

```{r, out.width="150%", out.height="150%"}
hist(negprob_spe$control_counts)
```

```{r, out.width="150%", out.height="150%"}
boxplot(negprob_spe$control_counts)
```

## Control count / total count

```{r}
summary(spe$ctrl_total_ratio)
```

```{r, out.width="150%", out.height="150%"}
hist(spe$ctrl_total_ratio)
```

```{r, out.width="150%", out.height="150%"}
boxplot(spe$ctrl_total_ratio)
```


```{r, out.width="150%", out.height="150%"}
threshold <- 0.99
spe$flagged_ctrl_tot_ratio_99 <- as.factor(ifelse(colData(spe)$ctrl_total_ratio > quantile(colData(spe)$ctrl_total_ratio, probs = threshold, na.rm = T), TRUE, FALSE))
polygons_whole$flagged_ctrl_tot_ratio_99 <- ifelse(spe$flagged_ctrl_tot_ratio_99 == TRUE, "red", "grey80")

tm_shape(polygons_whole) + 
  tm_fill(col= "flagged_ctrl_tot_ratio_99") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = paste0("Control / total counts ratio above 99% threshold, value = ", 
                                round(quantile(colData(spe)$ctrl_total_ratio, probs = threshold, na.rm = T), 2), "\nFlagged cells = ",
                                table(colData(spe)$flagged_ctrl_tot_ratio_99)["TRUE"], " out of ", dim(spe)[2], ", ",
                                round(table(colData(spe)$flagged_ctrl_tot_ratio_99)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"),
            main.title.fontface = 2,
            main.title.size = 1, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))
```

```{r}
ctrltotratio_spe <- spe[,spe$flagged_ctrl_tot_ratio_99 == FALSE]
summary(ctrltotratio_spe$ctrl_total_ratio)
```

```{r, out.width="150%", out.height="150%"}
hist(ctrltotratio_spe$ctrl_total_ratio)
```

```{r, out.width="150%", out.height="150%"}
boxplot(ctrltotratio_spe$ctrl_total_ratio)
```


## Custom area, low threshold

```{r}
summary(spe$custom_um_area)
```

```{r, out.width="150%", out.height="150%"}
hist(spe$custom_um_area)
```

```{r, out.width="150%", out.height="150%"}
boxplot(spe$custom_um_area)
```

```{r, out.width="150%", out.height="150%"}
threshold <- 0.01
spe$flagged_cumarea_1 <- as.factor(ifelse(colData(spe)$custom_um_area < quantile(colData(spe)$custom_um_area, probs = threshold), TRUE, FALSE))
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "flagged_cumarea_1", 
            point_size = 0.1) + scale_color_manual(values = c("TRUE" = "black","FALSE" = "lightgrey")) + ggtitle(paste0("Custom area in μm below 1%, threshold value = ", round(quantile(colData(spe)$custom_um_area, probs = threshold, na.rm = T), 2)))  + theme(legend.position="none") + labs(subtitle = paste0("Flagged cells = ", table(colData(spe)$flagged_cumarea_1)["TRUE"], " out of ", dim(spe)[2], ", ",  round(table(colData(spe)$flagged_cumarea_1)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"))
```

```{r, out.width="150%", out.height="150%"}
threshold <- 0.01
spe$flagged_target_1 <- as.factor(ifelse(colData(spe)$target_counts < quantile(colData(spe)$target_counts, probs = threshold), TRUE, FALSE))

plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "flagged_target_1", 
            point_size = 0.1) + scale_color_manual(values = c("TRUE" = "black","FALSE" = "lightgrey")) + ggtitle(paste0("Target counts below 1%, threshold value = ", round(quantile(colData(spe)$target_counts, probs = threshold, na.rm = T), 2)))  + theme(legend.position="none") + labs(subtitle = paste0("Flagged cells = ", table(colData(spe)$flagged_target_1)["TRUE"], " out of ", dim(spe)[2], ", ",  round(table(colData(spe)$flagged_target_1)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"))
```

```{r}
# to remove small cells that also have a number of target counts below 1%
table(spe$flagged_cumarea_1, spe$flagged_target_1, dnn = c("flagged_cumarea_1", "flagged_target_1"))
```

## Custom area, high threshold

```{r, out.width="150%", out.height="150%"}
threshold <- 0.99
spe$flagged_cumarea_99 <- as.factor(ifelse(colData(spe)$custom_um_area > quantile(colData(spe)$custom_um_area, probs = threshold), TRUE, FALSE))
polygons_whole$flagged_cumarea_99 <- ifelse(spe$flagged_cumarea_99 == TRUE, "red", "grey80")

tm_shape(polygons_whole) + 
  tm_fill(col= "flagged_cumarea_99") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = paste0("Custom area in um above 99% threshold, threshold value = ", 
                                round(quantile(colData(spe)$custom_um_area, probs = threshold, na.rm = T), 2), "\nFlagged cells = ",
                                table(colData(spe)$flagged_cumarea_99)["TRUE"], " out of ", dim(spe)[2], ", ",
                                round(table(colData(spe)$flagged_cumarea_99)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"),
            main.title.fontface = 2,
            main.title.size = 1, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))
```

```{r}
# Big problematic cells we noticed in polygons map can be captured using both negative probe counts and area above 90%
table(spe$flagged_negprob_99, spe$flagged_cumarea_99, dnn = c("flagged_negprob_99", "flagged_cumarea_99,"))
```

```{r}
area_spe <- spe[, spe$flagged_cumarea_99 == FALSE & spe$flagged_cumarea_1 == FALSE]
summary(area_spe$custom_um_area)
```

```{r, out.width="150%", out.height="150%"}
hist(area_spe$custom_um_area)
```

```{r, out.width="150%", out.height="150%"}
boxplot(area_spe$custom_um_area)
```

## DAPI, low threshold

```{r}
summary(spe$Mean.DAPI)
```

```{r, out.width="150%", out.height="150%"}
hist(spe$Mean.DAPI)
```

```{r, out.width="150%", out.height="150%"}
boxplot(spe$Mean.DAPI)
```

```{r, out.width="150%", out.height="150%"}
threshold <- 0.01
spe$flagged_dapi_1 <- as.factor(ifelse(colData(spe)$Mean.DAPI < quantile(colData(spe)$Mean.DAPI, probs = threshold), TRUE, FALSE))
polygons_whole$flagged_dapi_1 <- ifelse(spe$flagged_dapi_1 == TRUE, "black", "grey80")

tm_shape(polygons_whole) + 
  tm_fill(col= "flagged_dapi_1") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = paste0("DAPI signal below 1% threshold, threshold value = ", 
                                round(quantile(colData(spe)$Mean.DAPI, probs = threshold, na.rm = T), 2), "\nFlagged cells = ",
                                table(colData(spe)$flagged_dapi_1)["TRUE"], " out of ", dim(spe)[2], ", ",
                                round(table(colData(spe)$flagged_dapi_1)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"),
            main.title.fontface = 2,
            main.title.size = 1, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))
```


```{r}
#Flagged cells for custom area in um below 1% are completely different compared to flagged cells for DAPI signal below 1%
table(spe$flagged_cumarea_1, spe$flagged_dapi_1, dnn = c("flagged_cumarea_1", "flagged_dapi_1"))
```

```{r}
#Flagged cells for DAPI signal below 1% show a big area
table(spe$flagged_cumarea_99, spe$flagged_dapi_1, dnn = c("flagged_cumarea_99", "flagged_dapi_1"))
```


## DAPI, high threshold

```{r, out.width="150%", out.height="150%"}
threshold <- 0.99
spe$flagged_dapi_99 <- as.factor(ifelse(colData(spe)$Mean.DAPI > quantile(colData(spe)$Mean.DAPI, probs = threshold), TRUE, FALSE))
polygons_whole$flagged_dapi_99 <- ifelse(spe$flagged_dapi_99 == TRUE, "red", "grey80")

tm_shape(polygons_whole) + 
  tm_fill(col= "flagged_dapi_99") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = paste0("DAPI signal above 99% threshold, threshold value = ", 
                                round(quantile(colData(spe)$Mean.DAPI, probs = threshold, na.rm = T), 2), "\nFlagged cells = ",
                                table(colData(spe)$flagged_dapi_99)["TRUE"], " out of ", dim(spe)[2], ", ",
                                round(table(colData(spe)$flagged_dapi_99)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"),
            main.title.fontface = 2,
            main.title.size = 1, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))
```


```{r}
#small cells with high DAPI signal?
table(spe$flagged_cumarea_1, spe$flagged_dapi_99, dnn = c("flagged_cumarea_1", "flagged_dapi_99"))
```

```{r}
dapi_spe <- spe[, spe$flagged_dapi_99 == FALSE & spe$flagged_dapi_1 == FALSE]
summary(dapi_spe$Mean.DAPI)
```

```{r, out.width="150%", out.height="150%"}
hist(dapi_spe$Mean.DAPI)
```

```{r, out.width="150%", out.height="150%"}
boxplot(dapi_spe$Mean.DAPI)
```

## Logged AspectRatio, low threshold

```{r}
summary(spe$log2_AspectRatio)
```

```{r, out.width="150%", out.height="150%"}
hist(spe$log2_AspectRatio)
```

```{r, out.width="150%", out.height="150%"}
boxplot(spe$log2_AspectRatio)
```

```{r, out.width="150%", out.height="150%"}
threshold <- 0.01
spe$flagged_logAspectRatio_1 <- as.factor(ifelse(spe$log2_AspectRatio < quantile(spe$log2_AspectRatio, probs = threshold, na.rm = T), TRUE, FALSE))
polygons_whole$flagged_logAspectRatio_1 <- ifelse(spe$flagged_logAspectRatio_1 == TRUE, "black", "grey80")

tm_shape(polygons_whole) + 
  tm_fill(col= "flagged_logAspectRatio_1") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = paste0("Logged AspectRatio below 1% threshold, threshold value = ", 
                                round(quantile(colData(spe)$log2_AspectRatio, probs = threshold, na.rm = T), 2), "\nFlagged cells = ",
                                table(colData(spe)$flagged_logAspectRatio_1)["TRUE"], " out of ", dim(spe)[2], ", ",
                                round(table(colData(spe)$flagged_logAspectRatio_1)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"),
            main.title.fontface = 2,
            main.title.size = 1, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))
```

## Logged AspectRatio, high threshold

```{r, out.width="150%", out.height="150%"}
threshold <- 0.99
spe$flagged_logAspectRatio_99 <- as.factor(ifelse(spe$log2_AspectRatio > quantile(spe$log2_AspectRatio, probs = threshold, na.rm = T), TRUE, FALSE))
polygons_whole$flagged_logAspectRatio_99 <- ifelse(spe$flagged_logAspectRatio_99 == TRUE, "red", "grey80")

tm_shape(polygons_whole) + 
  tm_fill(col= "flagged_logAspectRatio_99") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = paste0("Logged AspectRatio above 99% threshold, threshold value = ", 
                                round(quantile(colData(spe)$log2_AspectRatio, probs = threshold, na.rm = T), 2), "\nFlagged cells = ",
                                table(colData(spe)$flagged_logAspectRatio_99)["TRUE"], " out of ", dim(spe)[2], ", ",
                                round(table(colData(spe)$flagged_logAspectRatio_99)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"),
            main.title.fontface = 2,
            main.title.size = 1, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))
```

```{r}
aspectratio_spe <- spe[, spe$flagged_logAspectRatio_99 == FALSE & spe$flagged_logAspectRatio_1 == FALSE]
summary(aspectratio_spe$log2_AspectRatio)
```

```{r, out.width="150%", out.height="150%"}
hist(aspectratio_spe$log2_AspectRatio)
```

```{r, out.width="150%", out.height="150%"}
boxplot(aspectratio_spe$log2_AspectRatio)
```


# First cell skimming with scuttle::isOutlier (Hampel filter) {.tabset .tabset-fade .tabset-pills}

Introducing scuttle::isOutlier function to move to a different filter beyond threshold
Outliers are defined when their value is higher or lower than the median, plus or minus 3 median absolute deviations (MAD) respectively.
However, when using MAD a symmetric distribution is assumed

## Control counts

```{r}
summary(spe$control_counts)
```

```{r, out.width="150%", out.height="150%"}
hist(spe$control_counts)
```

```{r, out.width="150%", out.height="150%"}
boxplot(spe$control_counts)
```

```{r, out.width="150%", out.height="150%"}
spe$outlier_negprob <- scuttle::isOutlier(colData(spe)$control_counts, type = "higher")
polygons_whole$outlier_negprob <- ifelse(spe$outlier_negprob == TRUE, "red", "grey80")

tm_shape(polygons_whole) + 
  tm_fill(col= "outlier_negprob") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = paste0("Outlier cells for negative probe counts", 
                                "\n lower thr. = " , attr(spe$outlier_negprob,"thresholds")["lower"], ", higher thr. = ",
                                attr(spe$outlier_negprob,"thresholds")["higher"],
                                "\nFlagged cells = ",
                                table(colData(spe)$outlier_negprob)["TRUE"], " out of ", dim(spe)[2], ", ",
                                round(table(colData(spe)$outlier_negprob)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"),
            main.title.fontface = 2,
            main.title.size = 1, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

```

```{r}
negprob_spe <- spe[,spe$outlier_negprob == FALSE]
summary(negprob_spe$control_counts)
```

```{r, out.width="150%", out.height="150%"}
hist(negprob_spe$control_counts)
```

```{r, out.width="150%", out.height="150%"}
boxplot(negprob_spe$control_counts)
```

## Control count / total counts

```{r}
summary(spe$ctrl_total_ratio)
```

```{r, out.width="150%", out.height="150%"}
hist(spe$ctrl_total_ratio)
```

```{r, out.width="150%", out.height="150%"}
boxplot(spe$ctrl_total_ratio)
```

```{r, out.width="150%", out.height="150%"}
spe$outlier_ctrl_tot <- scuttle::isOutlier(spe$ctrl_total_ratio, type = "higher")
polygons_whole$outlier_ctrl_tot <- ifelse(spe$outlier_ctrl_tot == TRUE, "red", "grey80")

tm_shape(polygons_whole) + 
  tm_fill(col= "outlier_ctrl_tot") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = paste0("Outlier cells for control / total counts ratio", 
                                "\n lower thr. = " , round(attr(spe$outlier_ctrl_tot,"thresholds")["lower"], 2), ", higher thr. = ",
                                round(attr(spe$outlier_ctrl_tot,"thresholds")["higher"], 2),
                                "\nFlagged cells = ",
                                table(colData(spe)$outlier_ctrl_tot)["TRUE"], " out of ", dim(spe)[2], ", ",
                                round(table(colData(spe)$outlier_ctrl_tot)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"),
            main.title.fontface = 2,
            main.title.size = 1, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

```

```{r}
ctrltotratio_spe <- spe[,spe$outlier_ctrl_tot == FALSE]
summary(ctrltotratio_spe$ctrl_total_ratio)
```

```{r, out.width="150%", out.height="150%"}
hist(ctrltotratio_spe$ctrl_total_ratio)
```

```{r, out.width="150%", out.height="150%"}
boxplot(ctrltotratio_spe$ctrl_total_ratio)
```

## Custom area

```{r}
summary(spe$custom_um_area)
```

```{r, out.width="150%", out.height="150%"}
hist(spe$custom_um_area)
```

```{r, out.width="150%", out.height="150%"}
boxplot(spe$custom_um_area)
```

```{r, out.width="150%", out.height="150%"}
spe$outlier_cumarea <- scuttle::isOutlier(colData(spe)$custom_um_area, type = "both")
polygons_whole$outlier_cumarea <- ifelse(spe$outlier_cumarea == TRUE, "red", "grey80")

tm_shape(polygons_whole) + 
  tm_fill(col= "outlier_cumarea") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = paste0("Outlier cells for custom um area", 
                                "\n lower thr. = " , attr(spe$outlier_cumarea,"thresholds")["lower"], ", higher thr. = ",
                                attr(spe$outlier_cumarea,"thresholds")["higher"],
                                "\nFlagged cells = ",
                                table(colData(spe)$outlier_cumarea)["TRUE"], " out of ", dim(spe)[2], ", ",
                                round(table(colData(spe)$outlier_cumarea)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"),
            main.title.fontface = 2,
            main.title.size = 1, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))
```

```{r}
table(spe$outlier_negprob, spe$outlier_cumarea, dnn = c("outlier_negprob", "outlier_cumarea"))
```

```{r}
area_spe <- spe[, spe$outlier_cumarea == FALSE]
summary(area_spe$custom_um_area)
```

```{r, out.width="150%", out.height="150%"}
hist(area_spe$custom_um_area)
```

```{r, out.width="150%", out.height="150%"}
boxplot(area_spe$custom_um_area)
```

## DAPI signal

```{r}
summary(spe$Mean.DAPI)
```

```{r, out.width="150%", out.height="150%"}
hist(spe$Mean.DAPI)
```

```{r, out.width="150%", out.height="150%"}
boxplot(spe$Mean.DAPI)
```

```{r, out.width="150%", out.height="150%"}
spe$outlier_dapi <- scuttle::isOutlier(colData(spe)$Mean.DAPI, type = "both")
polygons_whole$outlier_dapi<- ifelse(spe$outlier_dapi == TRUE, "red", "grey80")

tm_shape(polygons_whole) + 
  tm_fill(col= "outlier_dapi") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = paste0("Outlier cells for DAPI signal",
                                "\n lower thr. = " , attr(spe$outlier_dapi,"thresholds")["lower"], ", higher thr. = ",
                                attr(spe$outlier_dapi,"thresholds")["higher"],
                                "\nFlagged cells = ",
                                table(colData(spe)$outlier_dapi)["TRUE"], " out of ", dim(spe)[2], ", ",
                                round(table(colData(spe)$outlier_dapi)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"),
            main.title.fontface = 2,
            main.title.size = 1, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))
```

```{r}
table(spe$outlier_negprob, spe$outlier_cumarea, dnn = c("outlier_negprob", "outlier_cumarea"))
```

```{r}
dapi_spe <- spe[, spe$outlier_dapi == FALSE]
summary(dapi_spe$Mean.DAPI)
```

```{r, out.width="150%", out.height="150%"}
hist(dapi_spe$Mean.DAPI)
```

```{r, out.width="150%", out.height="150%"}
boxplot(dapi_spe$Mean.DAPI)
```


# First cell skimming with skewness adjustment to boxplot {.tabset .tabset-fade .tabset-pills}

Introducing medcouple (MC) from Hubert, M. and Vandervieren, E. (2008). Boxplot whiskers are defined as:
1st quartile - 1.5 * hl(MC) * IQR 
3rd quartile + 1.5 * hu(MC) * IQR
If MC >= 0, hl = exp(-4 * MC), hu = exp(3 * MC) 
If MC < 0, hl = exp(-3 * MC), hu = exp(4 * MC)

```{r}
polygons_whole <- polygons_whole[polygons_whole$id_fov %in% metadata$id_fov,]
polygons_whole$control_counts <- spe$control_counts
polygons_whole$Mean.DAPI <- spe$Mean.DAPI
```

## Control counts

```{r}
summary(spe$control_counts)
```

```{r, out.width="150%", out.height="150%"}
hist(spe$control_counts)
```

A normal distribution should have a skewness of 0.

```{r}
#it should be 0
library(e1071)
skewness(spe$control_counts)[1]
```

−0.6 ≤ MC ≤ 0.6 is requested by Hubert and Vandervieren (2008) for their adjusted boxplot.

```{r}
library(robustbase)
mc(spe$control_counts)[1] #0.5
```

How many outliers were found?

```{r}
out <- adjbox(spe$control_counts)
length(out$out)
```

Distribution of outlier values

```{r}
summary(out$out)
```

```{r, out.width="150%", out.height="150%"}
outlier_color <- colnames(spe)
for(i in 1:out$n){
  outlier_color[i] <- ifelse(outlier_color[i] %in% names(out$out), "red", "grey80")
}

polygons_whole$negprob_outlier <- outlier_color

tm_shape(polygons_whole) + 
  tm_fill(col= "negprob_outlier") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = paste0("Outlier cells for negative probes", 
                                "\n fence = " , out$fence,
                                "\nFlagged cells = ",
                                table(polygons_whole$negprob_outlier)["red"], " out of ", dim(spe)[2], ", ",
                                round(table(polygons_whole$negprob_outlier)["red"]/dim(spe)[2]*100, 2), "% of total cells"),
            main.title.fontface = 2,
            main.title.size = 1, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))
```

```{r, out.width="150%", out.height="150%"}
negprob_poly <- polygons_whole[polygons_whole$negprob_outlier == "grey80",]
hist(negprob_poly$control_counts)
```

```{r, out.width="150%", out.height="150%"}
boxplot(negprob_poly$control_counts)
```

```{r}
summary(negprob_poly$control_counts)
```

## Control / total counts ratio

```{r}
summary(spe$ctrl_total_ratio)
```

```{r, out.width="150%", out.height="150%"}
hist(spe$ctrl_total_ratio)
```

Skewness

```{r}
skewness(spe$ctrl_total_ratio, na.rm = T)[1]
```

Medcouple

```{r}
mc(spe$ctrl_total_ratio, na.rm = T)[1] 
```

How many outliers were found?

```{r}
out <- adjbox(spe$ctrl_total_ratio)
length(out$out)
```

Distribution of outlier values

```{r}
summary(out$out)
```

```{r, out.width="150%", out.height="150%"}
outlier_color <- colnames(spe)
for(i in 1:out$n){
  outlier_color[i] <- ifelse(outlier_color[i] %in% names(out$out), "red", "grey80")
}

polygons_whole$ctrltotal_outlier <- outlier_color

tm_shape(polygons_whole) + 
  tm_fill(col = "ctrltotal_outlier") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = paste0("Outlier cells for control/total outliers", 
                                "\n fence = " , out$fence,
                                "\nFlagged cells = ",
                                table(polygons_whole$ctrltotal_outlier)["red"], " out of ", dim(spe)[2], ", ",
                                round(table(polygons_whole$ctrltotal_outlier)["red"]/dim(spe)[2]*100, 2), "% of total cells"),
            main.title.fontface = 2,
            main.title.size = 1, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))
```

```{r, out.width="150%", out.height="150%"}
ctrltotal_poly <- polygons_whole[polygons_whole$ctrltotal_outlier == "grey80",]
hist(ctrltotal_poly$ctrl_total_ratio)
```

```{r, out.width="150%", out.height="150%"}
boxplot(ctrltotal_poly$ctrl_total_ratio)
```

```{r}
summary(ctrltotal_poly$ctrl_total_ratio)
```

## Custom um area

```{r}
summary(spe$custom_um_area)
```

```{r, out.width="150%", out.height="150%"}
hist(spe$custom_um_area)
```

Skewness

```{r}
skewness(spe$custom_um_area)[1]
```

Medcouple

```{r}
mc(spe$custom_um_area)[1]
```

How many outliers were found?

```{r}
names(spe$custom_um_area) <- colnames(spe)
out <- adjbox(spe$custom_um_area)
length(out$out)
```

Distribution of outlier values

```{r}
summary(out$out)
```

```{r, out.width="150%", out.height="150%"}
outlier_color <- colnames(spe)
for(i in 1:out$n){
  outlier_color[i] <- ifelse(outlier_color[i] %in% names(out$out), "red", "grey80")
}

polygons_whole$cumarea_outlier <- outlier_color

tm_shape(polygons_whole) + 
  tm_fill(col= "cumarea_outlier") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = paste0("Outlier cells for custom_um_area", 
                                "\n fence = " , out$fence,
                                "\nFlagged cells = ",
                                table(polygons_whole$cumarea_outlier)["red"], " out of ", dim(spe)[2], ", ",
                                round(table(polygons_whole$cumarea_outlier)["red"]/dim(spe)[2]*100, 2), "% of total cells"),
            main.title.fontface = 2,
            main.title.size = 1, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))
```

```{r}
area_poly <- polygons_whole[polygons_whole$cumarea_outlier == "grey80",]
hist(negprob_poly$custom_um_area)
```

```{r, out.width="150%", out.height="150%"}
boxplot(area_poly$custom_um_area)
```

```{r}
summary(area_poly$custom_um_area)
```

## Mean DAPI signal

```{r}
summary(spe$Mean.DAPI)
```

```{r, out.width="150%", out.height="150%"}
hist(spe$Mean.DAPI)
```

Skewness

```{r}
skewness(spe$Mean.DAPI)[1]
```

Medcouple

```{r}
mc(spe$Mean.DAPI)[1]
```

How many outliers were found?

```{r}
names(spe$Mean.DAPI) <- colnames(spe)
out <- adjbox(spe$Mean.DAPI)
length(out$out)
```

Distribution of outlier values

```{r}
summary(out$out)
```

```{r, out.width="150%", out.height="150%"}
outlier_color <- colnames(spe)
for(i in 1:out$n)
  outlier_color[i] <- ifelse(outlier_color[i] %in% names(out$out), "red", "grey80")

polygons_whole$dapi_outlier <- outlier_color

tm_shape(polygons_whole) + 
  tm_fill(col= "dapi_outlier") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = paste0("Outlier cells for dapi", 
                                "\n fence = " , out$fence,
                                "\nFlagged cells = ",
                                table(polygons_whole$dapi_outlier)["red"], " out of ", dim(spe)[2], ", ",
                                round(table(polygons_whole$dapi_outlier)["red"]/dim(spe)[2]*100, 2), "% of total cells"),
            main.title.fontface = 2,
            main.title.size = 1, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))
```

```{r}
table(spe$outlier_negprob, spe$outlier_cumarea, dnn = c("outlier_negprob", "outlier_cumarea"))
```

```{r}
dapi_poly <- polygons_whole[polygons_whole$dapi_outlier == "grey80",]
hist(dapi_poly$Mean.DAPI)
```

```{r, out.width="150%", out.height="150%"}
boxplot(dapi_poly$Mean.DAPI)
```

```{r}
summary(dapi_poly$Mean.DAPI)
```

## Log2 AspectRatio

```{r}
summary(spe$log2_AspectRatio)
```

```{r, out.width="150%", out.height="150%"}
hist(spe$log2_AspectRatio)
```

Skewness

```{r}
skewness(spe$log2_AspectRatio, na.rm = TRUE)[1]
```

Medcouple

```{r}
mc(spe$log2_AspectRatio, na.rm = TRUE)[1]
```

How many outliers were found?

```{r}
names(spe$log2_AspectRatio) <- colnames(spe)
out <- adjbox(spe$log2_AspectRatio)
loghw_fence <- out$fence 
length(out$out)
```

Distribution of outlier values

```{r}
summary(out$out)
```

```{r, out.width="150%", out.height="150%"}
outlier_color <- colnames(spe)
for(i in 1:dim(spe)[2])
  outlier_color[i] <- ifelse(outlier_color[i] %in% names(out$out), "red", "grey80")

polygons_whole$loghw_outlier <- outlier_color

tm_shape(polygons_whole) + 
  tm_fill(col= "loghw_outlier") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = paste0("Outlier cells for logged aspect ratio", 
                                "\n fence = " , round(out$fence, 2),
                                "\nFlagged cells = ",
                                table(polygons_whole$loghw_outlier)["red"], " out of ", dim(spe)[2], ", ",
                                round(table(polygons_whole$loghw_outlier)["red"]/dim(spe)[2]*100, 2), "% of total cells"),
            main.title.fontface = 2,
            main.title.size = 1, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))
```

```{r}
loghw_poly <- polygons_whole[polygons_whole$loghw_outlier == "grey80",]
hist(loghw_poly$log2_AspectRatio)
```

```{r, out.width="150%", out.height="150%"}
boxplot(loghw_poly$log2_AspectRatio)
```

```{r}
summary(loghw_poly$log2_AspectRatio)
```


# Flag score development 

## Exploratory scatterplots {.tabset .tabset-fade .tabset-pills}

### Total counts ~ logged width/height ratio

```{r, out.width="150%", out.height="150%"}
cell_df <- data.frame(colData(spe))

p <- ggplot(data = cell_df) +
  geom_pointdensity(aes(x = log2_AspectRatio, y = total)) +
  scale_color_viridis_c() +
  theme(legend.position = "none",
        plot.margin = margin())

ggMarginal(p, type = "density", color = "#ccccff", fill = "#ccccff") 
```

### Detected features ~ logged width/height ratio

```{r, out.width="150%", out.height="150%"}
p <- ggplot(data = cell_df) +
  geom_pointdensity(aes(x = log2_AspectRatio, y = detected)) +
  scale_color_viridis_c() +
  theme(legend.position = "none",
        plot.margin = margin())

ggMarginal(p, type = "density", color = "#ccccff", fill = "#ccccff") 
```

### Total counts ~ custom area in um

```{r, out.width="150%", out.height="150%"}
p <- ggplot(data = cell_df) +
  geom_pointdensity(aes(x = custom_um_area, y = total)) +
  scale_color_viridis_c() +
  theme(legend.position = "none",
        plot.margin = margin())

ggMarginal(p, type = "density", color = "#ccccff", fill = "#ccccff") 
```

### Logged width/height ratio ~ negative control probe counts

```{r, out.width="150%", out.height="150%"}
p <- ggplot(data = cell_df) +
  geom_pointdensity(aes(x = control_counts, y = log2_AspectRatio)) +
  scale_color_viridis_c() +
  theme(legend.position = "none",
        plot.margin = margin())

ggMarginal(p, type = "density", color = "#ccccff", fill = "#ccccff") 
```

### Logged width/height ratio ~ local coordinates on X axis in px

```{r, out.width="150%", out.height="150%"}
p <- ggplot(data = cell_df) +
  geom_pointdensity(aes(x = CenterX_local_px, y = log2_AspectRatio)) +
  scale_color_viridis_c() +
  theme(legend.position = "none",
        plot.margin = margin())

ggMarginal(p, type = "density", color = "#ccccff", fill = "#ccccff") 
```

### Logged width/height ratio ~ local coordinates on X axis in px

```{r, out.width="150%", out.height="150%"}
p <- ggplot(data = cell_df) +
  geom_pointdensity(aes(x = CenterY_local_px, y = log2_AspectRatio)) +
  scale_color_viridis_c() +
  theme(legend.position = "none",
        plot.margin = margin())

ggMarginal(p, type = "density", color = "#ccccff", fill = "#ccccff") 
```

### Control counts ~ custom um area

```{r, out.width="150%", out.height="150%"}
p <- ggplot(data = cell_df) +
  geom_pointdensity(aes(x = custom_um_area, y = control_counts)) +
  scale_color_viridis_c() +
  theme(legend.position = "none",
        plot.margin = margin())

ggMarginal(p, type = "density", color = "#ccccff", fill = "#ccccff") 
```

### DAPI ~ custom um area

```{r, out.width="150%", out.height="150%"}
p <- ggplot(data = cell_df) +
  geom_pointdensity(aes(x = custom_um_area, y = Mean.DAPI)) +
  scale_color_viridis_c() +
  theme(legend.position = "none",
        plot.margin = margin())

ggMarginal(p, type = "density", color = "#ccccff", fill = "#ccccff") 
```

### DAPI ~ control counts

```{r, out.width="150%", out.height="150%"}
p <- ggplot(data = cell_df) +
    geom_pointdensity(aes(x = control_counts, y = Mean.DAPI)) +
  scale_color_viridis_c() +
  theme(legend.position = "none",
        plot.margin = margin())

ggMarginal(p, type = "density", color = "#ccccff", fill = "#ccccff") 
```


## Flag score formula

```{r class.source = 'fold-show'}
spe$flag_score_hw <- 1/(1 + exp(-0.3*spe$target_counts/spe$custom_um_area + 0.8*abs(spe$log2_AspectRatio)))
summary(spe$flag_score_hw)
plot(spe$flag_score_hw)
```

```{r}
hist(spe$flag_score_hw)
```

Adding variables to polygon sf object

```{r, out.width="150%", out.height="150%"}
polygons_whole <- polygons_whole[polygons_whole$id_fov %in% metadata$id_fov,]

polygons_whole$total <- spe$total
polygons_whole$target_counts <- spe$target_counts
polygons_whole$log2_AspectRatio <- spe$log2_AspectRatio
polygons_whole$flag_score_hw <- spe$flag_score_hw
polygons_whole$cell_type <- spe$cell_type
polygons_whole$celltype_color <- as.character(spe$celltype_color)
polygons_whole$control_counts <- spe$control_counts
polygons_whole$target_area_ratio <- spe$target_area_ratio
```

# Whole polygon dataset {.tabset .tabset-fade .tabset-pills}

```{r, out.width="150%", out.height="150%"}
tm_shape(polygons_whole) + 
tm_fill(col = "celltype_color") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = sample_name,
            main.title.size = 0.5,
            frame = FALSE,
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0)) +
tm_add_legend(type = "fill", labels = names(celltype_palette), col = celltype_palette)
```

## Flag score gradient map

```{r, out.width="150%", out.height="150%"}
tm_shape(polygons_whole) + 
  tm_fill(col = "flag_score_hw") +
  tm_layout(legend.outside = TRUE)
```

# Flag threshold map with AspectRatio in flag score formula and scuttle outliers {.tabset .tabset-fade .tabset-pills}

```{r}
spe$outlier_target <- scuttle::isOutlier(spe$target_counts, type = "lower", nmad = 1)
polygons_whole$outlier_target <- ifelse(spe$outlier_target == TRUE, "red", "grey80")

spe$outlier_targetarea <- scuttle::isOutlier(spe$target_area_ratio, type = "lower", nmad = 1)
polygons_whole$outlier_targetarea <- ifelse(spe$outlier_targetarea == TRUE, "red", "grey80")

spe$outlier_cellarea <- scuttle::isOutlier(colData(spe)$custom_um_area, type = "both", nmad = 1)
polygons_whole$outlier_cellarea <- ifelse(spe$outlier_cellarea == TRUE, "red", "grey80")

spe$outlier_loghw <- scuttle::isOutlier(spe$log2_AspectRatio, type = "both", nmad = 1)
polygons_whole$outlier_loghw<- ifelse(spe$outlier_loghw == TRUE, "red", "grey80")

spe$outlier_flagscorehw <- scuttle::isOutlier(spe$flag_score_hw, type = "lower", nmad = 1)
polygons_whole$outlier_flagscorehw <- ifelse(spe$outlier_flagscorehw == TRUE, "red", "grey80")
```

```{r, out.width="300%", out.height="900%"}
p1 <- tm_shape(polygons_whole) + 
  tm_fill(col= "outlier_flagscorehw", alpha = 0.5) +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Flag score outliers",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p2 <- tm_shape(polygons_whole) + 
  tm_fill(col = "celltype_color") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Flag score outliers",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p3 <- tm_shape(polygons_whole) + 
  tm_fill(col= "outlier_target", alpha = 0.5) +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Target counts outliers",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p4 <- tm_shape(polygons_whole) + 
  tm_fill(col= "outlier_targetarea", alpha = 0.5) +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Target counts/cell area outliers",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p5 <- tm_shape(polygons_whole) + 
  tm_fill(col= "outlier_cellarea", alpha = 0.5) +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Cell area outliers",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p6 <- tm_shape(polygons_whole) + 
  tm_fill(col= "outlier_loghw", alpha = 0.5) +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Logged AspectRatio outliers",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

tmap_arrange(p1, p2, p3, p4, p5, p6, ncol = 2)
```

## Flagged cell stats

```{r, out.width="150%", out.height="150%"}

out_fence <- c(paste(round(attr(spe$outlier_flagscorehw,"thresholds")["lower"], 2), round(attr(spe$outlier_flagscorehw,"thresholds")["higher"], 2)),
               paste(round(attr(spe$outlier_target,"thresholds")["lower"], 2), round(attr(spe$outlier_target,"thresholds")["higher"], 2)),
               paste(round(attr(spe$outlier_targetarea,"thresholds")["lower"], 2), round(attr(spe$outlier_targetarea,"thresholds")["higher"], 2)),
               paste(round(attr(spe$outlier_cellarea,"thresholds")["lower"], 2), round(attr(spe$outlier_cellarea,"thresholds")["higher"], 2)),
               paste(round(attr(spe$outlier_loghw,"thresholds")["lower"], 2), round(attr(spe$outlier_loghw,"thresholds")["higher"], 2)))

flagged_cells <- c(sum(polygons_whole$outlier_flagscorehw=="red", na.rm = T),
                   sum(polygons_whole$outlier_target=="red", na.rm = T),
                   sum(polygons_whole$outlier_targetarea=="red", na.rm = T),
                   sum(polygons_whole$outlier_cellarea=="red",na.rm = T),
                   sum(polygons_whole$outlier_loghw=="red",na.rm = T))

unflagged_cells <- c(sum(polygons_whole$outlier_flagscorehw=="grey80", na.rm = T),
                   sum(polygons_whole$outlier_target=="grey80", na.rm = T),
                   sum(polygons_whole$outlier_targetarea=="grey80", na.rm = T),
                   sum(polygons_whole$outlier_cellarea=="grey80",na.rm = T),
                   sum(polygons_whole$outlier_loghw=="grey80",na.rm = T))

percent_flagged <- round(flagged_cells/dim(polygons_whole)[1]*100, 2)

poly <- data.frame(out_fence = out_fence, unflagged_cells = unflagged_cells, flagged_cells = flagged_cells, percent_flagged = percent_flagged, row.names = c("Flag score", "Target counts", "Target counts/cell area", "Cell area in um", "Logged aspect ratio"))

kable(poly, format = "html", col.names = c("Outlier thresholds", "Unflagged cells", "Flagged cells", "% flagged cells"), align = "c", caption = paste0("<center>Flagged cells by different metrics, total cells in whole sample: ", dim(polygons_whole)[1], ", flagged cells in whole sample: ", sum(polygons_whole$outlier_flagscorehw=="red"), "</center>"), padding = 1L)
```
#
#

## Big flag threshold map 

```{r, out.width="150%", out.height="150%"}
plot_list <- list(p1, p2 +
tm_add_legend(type = "fill", labels = names(celltype_palette), col = celltype_palette),
                  p3, p4, p5, p6)

for(i in 1:length(plot_list)){
  print(plot_list[[i]])
}
```


# Flag threshold map with AspectRatio in flag score formula and percentile outliers {.tabset .tabset-fade .tabset-pills}

```{r, out.width="300%", out.height="900%"}
metric_color <- ifelse(polygons_whole$flag_score_hw < quantile(spe$flag_score_hw, probs = 0.15), "red", "grey80")
target_color <- ifelse(polygons_whole$target_counts < quantile(spe$target_counts, probs = 0.15), "red", "grey80")
count_area_ratio_color <- ifelse((polygons_whole$target_counts/polygons_whole$custom_um_area) < quantile((spe$target_counts/spe$custom_um_area), probs = 0.15), "red", "grey80")
area_color_lo <- ifelse(polygons_whole$custom_um_area < quantile(spe$custom_um_area, probs = 0.15), "red", "grey80")
area_color_hi <- ifelse(polygons_whole$custom_um_area > quantile(spe$custom_um_area, probs = 0.85), "red", "grey80")
log_hw_color_lo <- ifelse(polygons_whole$log2_AspectRatio < quantile(spe$log2_AspectRatio, probs = 0.15), "red", "grey80")
log_hw_color_hi <- ifelse(polygons_whole$log2_AspectRatio > quantile(spe$log2_AspectRatio, probs = 0.85), "red", "grey80")

polygons_whole <- cbind(polygons_whole, metric_color = metric_color, target_color = target_color, count_area_ratio_color = count_area_ratio_color, area_color_lo = area_color_lo, area_color_hi = area_color_hi, log_hw_color_lo = log_hw_color_lo, log_hw_color_hi = log_hw_color_hi)

p1 <- tm_shape(polygons_whole) + 
  tm_fill(col= "metric_color", alpha = 0.5) +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Flag score below 15th percentile",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p2 <- tm_shape(polygons_whole) + 
tm_fill(col = "celltype_color") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Celltype",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p3 <- tm_shape(polygons_whole) + 
  tm_fill(col= "target_color", alpha = 0.5) +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Target counts below 15th percentile",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p4 <- tm_shape(polygons_whole) + 
  tm_fill(col= "count_area_ratio_color", alpha = 0.5) +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Target counts/custom um area below 15th percentile",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p5 <- tm_shape(polygons_whole) + 
  tm_fill(col= "area_color_lo", alpha = 0.5) +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Custom cell area below 15th percentile",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p6 <- tm_shape(polygons_whole) + 
  tm_fill(col= "area_color_hi", alpha = 0.5) +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Custom cell area above 85th percentile",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p7 <- tm_shape(polygons_whole) + 
  tm_fill(col= "log_hw_color_lo", alpha = 0.5) +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Logged AspectRatio below 15th percentile",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p8 <- tm_shape(polygons_whole) + 
  tm_fill(col= "log_hw_color_hi", alpha = 0.5) +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Logged AspectRatio above 85th percentile",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

tmap_arrange(p1, p2, p3, p4, p5, p6, p7, p8, ncol = 2)
```

## Flagged cell stats

```{r, out.width="150%", out.height="150%"}

flag_value <- c(round(quantile(polygons_whole$flag_score_hw, probs = 0.15), 2),
                round(quantile(polygons_whole$target_counts, probs = 0.15), 2),
                round(quantile((polygons_whole$target_area_ratio), probs = 0.15), 2),
                round(quantile(polygons_whole$custom_um_area, probs = 0.15), 2),
                round(quantile(polygons_whole$custom_um_area, probs = 0.85), 2),
                round(quantile(polygons_whole$log2_AspectRatio, probs = 0.15), 2),
                round(quantile(polygons_whole$log2_AspectRatio, probs = 0.85), 2))

flagged_cells <- c(sum(polygons_whole$metric_color=="red", na.rm = T),
                   sum(polygons_whole$target_color=="red", na.rm = T),
                   sum(polygons_whole$count_area_ratio_color=="red", na.rm = T),
                   sum(polygons_whole$area_color_lo=="red", na.rm = T),
                   sum(polygons_whole$area_color_hi=="red", na.rm = T),
                   sum(polygons_whole$log_hw_color_lo=="red", na.rm = T),
                   sum(polygons_whole$log_hw_color_hi=="red", na.rm = T))

unflagged_cells <- c(sum(polygons_whole$metric_color=="grey80", na.rm = T),
                     sum(polygons_whole$target_color=="grey80", na.rm = T),
                     sum(polygons_whole$count_area_ratio_color=="grey80", na.rm = T),
                     sum(polygons_whole$area_color_lo=="grey80", na.rm = T),
                     sum(polygons_whole$area_color_hi=="grey80", na.rm = T),
                     sum(polygons_whole$log_hw_color_lo=="grey80", na.rm = T),
                     sum(polygons_whole$log_hw_color_hi=="grey80", na.rm = T))

percent_flagged <- round(flagged_cells/dim(polygons_whole)[1]*100, 2)

poly <- data.frame(threshold = c(0.15, 0.15, 0.15, 0.15, 0.85, 0.15, 0.85), flag_value = flag_value, unflagged_cells = unflagged_cells, flagged_cells = flagged_cells, percent_flagged = percent_flagged, row.names = c("Flag score", "Target counts", "Target counts/um area", "Custom area in um low", "Custom area in um high", "Logged aspect ratio low", "Logged aspect ratio high"))

kable(poly, format = "html", col.names = c("Threshold", "Flag value", "Unflagged cells", "Flagged cells", "% flagged cells"), align = "c", caption = paste0("<center>Flagged cells by different metrics, total cells in whole sample: ", dim(polygons_whole)[1], ", flagged cells in whole sample: ", sum(spe$flag_score_hw < quantile(spe$flag_score_hw, probs = 0.15, na.rm = T)), "</center>"), padding = 1L)
```

#
#

## Big flag threshold map 

```{r, out.width="150%", out.height="150%"}
plot_list <- list(p1, p2 +
tm_add_legend(type = "fill", labels = names(celltype_palette), col = celltype_palette),
p3, p4, p5, p6, p7, p8)

for(i in 1:length(plot_list)){
  print(plot_list[[i]])
}
```

# Creating sf using local coordinates for geometry

```{r}
sf_local <- st_as_sf(spat_obj, coords = c("x_local_px", "y_local_px"))
names(sf_local)[names(sf_local) == "cellID"] <- "cell_id"

polygons <- sf_local %>%
  group_by(fov, cell_id) %>%
  summarize(geometry = st_combine(geometry))#st_combine = combine several feature geometries into one, without unioning or resolving internal boundaries

polygons <- st_cast(polygons, "POLYGON")

polygons$id_fov <- paste(polygons$cell_id, polygons$fov, sep = "_") 

polygons <- polygons[polygons$id_fov %in% metadata$id_fov,]
```

```{r}
valid_poly <- st_is_valid(polygons)
table(valid_poly)
```

Computing buffer (a circumscribing space) around geometries to handle invalid geometries using 0 distance, so keeping it adherent to the original space
occupied by the cell. (Voyager)

```{r}
polygons <- st_buffer(polygons, dist = 0)
table(st_geometry_type(polygons))
```

```{r}
cellids <- c() 
for(i in 1:dim(polygons)[1]){
	cellids[i] <- attr(polygons$geometry[[i]], "class")[2]
	cellids[i] <- cellids[i] == "MULTIPOLYGON"
	
}

polygons <- polygons[cellids == FALSE, ]
table(st_geometry_type(polygons$geometry))

spe <- spe[,spe$cell_id %in% polygons$cell_id]
```


```{r}
polygons$total <- spe$total
polygons$target_counts <- spe$target_counts
polygons$custom_um_area <- spe$custom_um_area
polygons$log2_AspectRatio <- spe$log2_AspectRatio
polygons$flag_score_hw <- spe$flag_score_hw
polygons$cell_type <- spe$cell_type
polygons$celltype_color <- as.character(spe$celltype_color)
polygons$control_counts <- spe$control_counts
polygons$target_area_ratio <- spe$target_area_ratio
```

```{r}
polygons$outlier_target <- ifelse(spe$outlier_target == TRUE, "red", "grey80")

polygons$outlier_targetarea <- ifelse(spe$outlier_targetarea == TRUE, "red", "grey80")

polygons$outlier_cellarea <- ifelse(spe$outlier_cellarea == TRUE, "red", "grey80")

polygons$outlier_loghw <- ifelse(spe$outlier_loghw == TRUE, "red", "grey80")

polygons$outlier_flagscorehw <- ifelse(spe$outlier_flagscorehw== TRUE, "red", "grey80")
```

# Fov 91 {.tabset .tabset-fade .tabset-pills}

```{r, out.width="150%", out.height="150%"}
polygons91 <- filter(polygons, fov == 91) 

tm_shape(polygons91) + 
  tm_fill(col= "flag_score_hw", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE)
```

```{r, out.width="300%", out.height="900%"}
p1 <- tm_shape(polygons91) + 
  tm_fill(col= "outlier_flagscorehw", alpha = 0.5) +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Flag score outliers",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p2 <- tm_shape(polygons91) + 
  tm_fill(col = "celltype_color") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Flag score outliers",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p3 <- tm_shape(polygons91) + 
  tm_fill(col= "outlier_target", alpha = 0.5) +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Target counts outliers",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p4 <- tm_shape(polygons91) + 
  tm_fill(col= "outlier_targetarea", alpha = 0.5) +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Target counts/cell area outliers",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p5 <- tm_shape(polygons91) + 
  tm_fill(col= "outlier_cellarea", alpha = 0.5) +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Cell area outliers",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p6 <- tm_shape(polygons91) + 
  tm_fill(col= "outlier_loghw", alpha = 0.5) +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Logged AspectRatio outliers",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

tmap_arrange(p1, p2, p3, p4, p5, p6, ncol = 2)
```

## Flagged cell stats

```{r, out.width="150%", out.height="150%"}

out_fence <- c(paste(round(attr(spe$outlier_flagscorehw,"thresholds")["lower"], 2), round(attr(spe$outlier_flagscorehw,"thresholds")["higher"], 2)),
               paste(round(attr(spe$outlier_target,"thresholds")["lower"], 2), round(attr(spe$outlier_target,"thresholds")["higher"], 2)),
               paste(round(attr(spe$outlier_targetarea,"thresholds")["lower"], 2), round(attr(spe$outlier_targetarea,"thresholds")["higher"], 2)),
               paste(round(attr(spe$outlier_cellarea,"thresholds")["lower"], 2), round(attr(spe$outlier_cellarea,"thresholds")["higher"], 2)),
               paste(round(attr(spe$outlier_loghw,"thresholds")["lower"], 2), round(attr(spe$outlier_loghw,"thresholds")["higher"], 2)))

flagged_cells <- c(sum(polygons91$outlier_flagscorehw=="red", na.rm = T),
                   sum(polygons91$outlier_target=="red", na.rm = T),
                   sum(polygons91$outlier_targetarea=="red", na.rm = T),
                   sum(polygons91$outlier_cellarea=="red",na.rm = T),
                   sum(polygons91$outlier_loghw=="red",na.rm = T))

unflagged_cells <- c(sum(polygons91$outlier_flagscorehw=="grey80", na.rm = T),
                   sum(polygons91$outlier_target=="grey80", na.rm = T),
                   sum(polygons91$outlier_targetarea=="grey80", na.rm = T),
                   sum(polygons91$outlier_cellarea=="grey80",na.rm = T),
                   sum(polygons91$outlier_loghw=="grey80",na.rm = T))

percent_flagged <- round(flagged_cells/dim(polygons91)[1]*100, 2)

poly91 <- data.frame(out_fence = out_fence, unflagged_cells = unflagged_cells, flagged_cells = flagged_cells, percent_flagged = percent_flagged, row.names = c("Flag score", "Target counts", "Target counts/cell area", "Cell area in um", "Logged aspect ratio"))

kable(poly91, format = "html", col.names = c("Outlier thresholds", "Unflagged cells", "Flagged cells", "% flagged cells"), align = "c", caption = paste0("<center>Flagged cells by different metrics, total cells in whole sample: ", dim(polygons_whole)[1], ", flagged cells in whole sample: ", sum(polygons_whole$outlier_flagscorehw=="red"), " , total cells in mock fov: ", dim(polygons91)[1], "</center>"), padding = 1L)
```
#
#

## Big flag threshold map 

```{r, out.width="150%", out.height="150%"}
plot_list <- list(p1, p2 +
tm_add_legend(type = "fill", labels = names(celltype_palette), col = celltype_palette), p3, p4, p5, p6)

for(i in 1:length(plot_list)){
  print(plot_list[[i]])
}
```


## Flag threshold map

```{r, out.width="300%", out.height="900%"}
metric_color <- ifelse(polygons91$flag_score_hw < quantile(spe$flag_score_hw, probs = 0.15), "red", "white")
target_color <- ifelse(polygons91$target_counts < quantile(spe$target_counts, probs = 0.15), "red", "white")
count_area_ratio_color <- ifelse((polygons91$target_area_ratio) < quantile((spe$target_area_ratio), probs = 0.15), "red", "white")
area_color_lo <- ifelse(polygons91$custom_um_area < quantile(spe$custom_um_area, probs = 0.15), "red", "white")
area_color_hi <- ifelse(polygons91$custom_um_area > quantile(spe$custom_um_area, probs = 0.85), "red", "white")
log_hw_color_lo <- ifelse(polygons91$log2_AspectRatio < quantile(spe$log2_AspectRatio, probs = 0.15), "red", "white")
log_hw_color_hi <- ifelse(polygons91$log2_AspectRatio > quantile(spe$log2_AspectRatio, probs = 0.85), "red", "white")

polygons91 <- cbind(polygons91, metric_color = metric_color, target_color = target_color, count_area_ratio_color = count_area_ratio_color, area_color_lo = area_color_lo, area_color_hi = area_color_hi, log_hw_color_lo = log_hw_color_lo, log_hw_color_hi = log_hw_color_hi)

p1 <- tm_shape(polygons91) + 
  tm_fill(col= "metric_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Flag score below 15th percentile",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p2 <- tm_shape(polygons91) + 
tm_fill(col = "celltype_color") +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Celltype",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p3 <- tm_shape(polygons91) + 
  tm_fill(col= "target_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Target counts below 15th percentile",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p4 <- tm_shape(polygons91) + 
  tm_fill(col= "count_area_ratio_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Target counts/custom um area below 15th percentile",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p5 <- tm_shape(polygons91) + 
  tm_fill(col= "area_color_lo", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Custom cell area below 15th percentile",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p6 <- tm_shape(polygons91) + 
  tm_fill(col= "area_color_hi", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Custom cell area above 85th percentile",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p7 <- tm_shape(polygons91) + 
  tm_fill(col= "log_hw_color_lo", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Logged AspectRatio below 15th percentile",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p8 <- tm_shape(polygons91) + 
  tm_fill(col= "log_hw_color_hi", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Logged AspectRatio above 85th percentile",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

tmap_arrange(p1, p2, p3, p4, p5, p6, p7, p8, ncol = 2)
```

## Flagged cell stats 

```{r, out.width="300%", out.height="300%"}

flagged_cells <- c(sum(polygons91$metric_color=="red", na.rm = T),
                   sum(polygons91$target_color=="red", na.rm = T),
                   sum(polygons91$count_area_ratio_color=="red", na.rm = T),
                   sum(polygons91$area_color_lo=="red", na.rm = T),
                   sum(polygons91$area_color_hi=="red", na.rm = T),
                   sum(polygons91$log_hw_color_lo=="red", na.rm = T),
                   sum(polygons91$log_hw_color_hi=="red", na.rm = T))

unflagged_cells <- c(sum(polygons91$metric_color=="white", na.rm = T),
                     sum(polygons91$target_color=="white", na.rm = T),
                     sum(polygons91$count_area_ratio_color=="white", na.rm = T),
                     sum(polygons91$area_color_lo=="white", na.rm = T),
                     sum(polygons91$area_color_hi=="white", na.rm = T),
                     sum(polygons91$log_hw_color_lo=="white", na.rm = T),
                     sum(polygons91$log_hw_color_hi=="white", na.rm = T))

percent_flagged <- round(flagged_cells/dim(polygons91)[1]*100, 2)

poly91 <- data.frame(threshold = c(0.15, 0.15, 0.15, 0.15, 0.85, 0.15, 0.85), flag_value = flag_value, unflagged_cells = unflagged_cells, flagged_cells = flagged_cells, percent_flagged = percent_flagged, row.names = c("Flag score", "Target counts", "Target counts/um area", "Custom area in um low", "Custom area in um high", "Logged aspect ratio low", "Logged aspect ratio high"))

kable(poly91, format = "html", col.names = c("Threshold", "Flag value", "Unflagged cells", "Flagged cells", "% flagged cells"), align = "c", caption = paste0("<center>Flagged cells by different metrics, total cells in fov: ", dim(polygons91)[1], ", total cells in whole sample: ", dim(polygons_whole)[1], ", flagged cells in whole sample: ", sum(spe$flag_score_hw < quantile(spe$flag_score_hw, probs = 0.15, na.rm = T)), "</center>"), padding = 1L)
```

#
#

## Big threshold map

```{r, out.width="150%", out.height="150%"}
plot_list <- list(p1, p2 +
tm_add_legend(type = "fill", labels = names(celltype_palette), col = celltype_palette),
p3, p4, p5, p6, p7, p8)

for(i in 1:length(plot_list)){
  print(plot_list[[i]])
}
```


# Fov 56 {.tabset .tabset-fade .tabset-pills}

```{r, out.width="150%", out.height="150%"}
polygons56 <- filter(polygons, fov==56)

tm_shape(polygons56) + 
  tm_fill(col= "flag_score_hw", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE)

```

```{r, out.width="300%", out.height="900%"}
p1 <- tm_shape(polygons56) + 
  tm_fill(col= "outlier_flagscorehw", alpha = 0.5) +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Flag score outliers",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p2 <- tm_shape(polygons56) + 
  tm_fill(col = "celltype_color") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Flag score outliers",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p3 <- tm_shape(polygons56) + 
  tm_fill(col= "outlier_target", alpha = 0.5) +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Target counts outliers",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p4 <- tm_shape(polygons56) + 
  tm_fill(col= "outlier_targetarea", alpha = 0.5) +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Target counts/cell area outliers",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p5 <- tm_shape(polygons56) + 
  tm_fill(col= "outlier_cellarea", alpha = 0.5) +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Cell area outliers",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p6 <- tm_shape(polygons56) + 
  tm_fill(col= "outlier_loghw", alpha = 0.5) +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Logged AspectRatio outliers",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

tmap_arrange(p1, p2, p3, p4, p5, p6, ncol = 2)
```

## Flagged cell stats

```{r, out.width="150%", out.height="150%"}

out_fence <- c(paste(round(attr(spe$outlier_flagscorehw,"thresholds")["lower"], 2), round(attr(spe$outlier_flagscorehw,"thresholds")["higher"], 2)),
               paste(round(attr(spe$outlier_target,"thresholds")["lower"], 2), round(attr(spe$outlier_target,"thresholds")["higher"], 2)),
               paste(round(attr(spe$outlier_targetarea,"thresholds")["lower"], 2), round(attr(spe$outlier_targetarea,"thresholds")["higher"], 2)),
               paste(round(attr(spe$outlier_cellarea,"thresholds")["lower"], 2), round(attr(spe$outlier_cellarea,"thresholds")["higher"], 2)),
               paste(round(attr(spe$outlier_loghw,"thresholds")["lower"], 2), round(attr(spe$outlier_loghw,"thresholds")["higher"], 2)))

flagged_cells <- c(sum(polygons56$outlier_flagscorehw=="red", na.rm = T),
                   sum(polygons56$outlier_target=="red", na.rm = T),
                   sum(polygons56$outlier_targetarea=="red", na.rm = T),
                   sum(polygons56$outlier_cellarea=="red",na.rm = T),
                   sum(polygons56$outlier_loghw=="red",na.rm = T))

unflagged_cells <- c(sum(polygons56$outlier_flagscorehw=="grey80", na.rm = T),
                   sum(polygons56$outlier_target=="grey80", na.rm = T),
                   sum(polygons56$outlier_targetarea=="grey80", na.rm = T),
                   sum(polygons56$outlier_cellarea=="grey80",na.rm = T),
                   sum(polygons56$outlier_loghw=="grey80",na.rm = T))

percent_flagged <- round(flagged_cells/dim(polygons56)[1]*100, 2)

poly56 <- data.frame(out_fence = out_fence, unflagged_cells = unflagged_cells, flagged_cells = flagged_cells, percent_flagged = percent_flagged, row.names = c("Flag score", "Target counts", "Target counts/cell area", "Cell area in um", "Logged aspect ratio"))

kable(poly56, format = "html", col.names = c("Outlier thresholds", "Unflagged cells", "Flagged cells", "% flagged cells"), align = "c", caption = paste0("<center>Flagged cells by different metrics, total cells in whole sample: ", dim(polygons_whole)[1], ", flagged cells in whole sample: ", sum(polygons_whole$outlier_flagscorehw=="red"), " , total cells in mock fov: ", dim(polygons56)[1], "</center>"), padding = 1L)
```
#
#

## Big flag threshold map 

```{r, out.width="150%", out.height="150%"}
plot_list <- list(p1, p2 +
tm_add_legend(type = "fill", labels = names(celltype_palette), col = celltype_palette), p3, p4, p5, p6)

for(i in 1:length(plot_list)){
  print(plot_list[[i]])
}
```

## Flag threshold map

```{r, out.width="300%", out.height="900%"}
metric_color <- ifelse(polygons56$flag_score_hw < quantile(spe$flag_score_hw, probs = 0.15), "red", "white")
target_color <- ifelse(polygons56$target_counts < quantile(spe$target_counts, probs = 0.15), "red", "white")
count_area_ratio_color <- ifelse((polygons56$target_counts/polygons56$custom_um_area) < quantile((spe$target_counts/spe$custom_um_area), probs = 0.15), "red", "white")
area_color_lo <- ifelse(polygons56$custom_um_area < quantile(spe$custom_um_area, probs = 0.15), "red", "white")
area_color_hi <- ifelse(polygons56$custom_um_area > quantile(spe$custom_um_area, probs = 0.85), "red", "white")
log_hw_color_lo <- ifelse(polygons56$log2_AspectRatio < quantile(spe$log2_AspectRatio, probs = 0.15), "red", "white")
log_hw_color_hi <- ifelse(polygons56$log2_AspectRatio > quantile(spe$log2_AspectRatio, probs = 0.90), "red", "white")

polygons56 <- cbind(polygons56, metric_color = metric_color, target_color = target_color, count_area_ratio_color = count_area_ratio_color, area_color_lo = area_color_lo, area_color_hi = area_color_hi, log_hw_color_lo = log_hw_color_lo, log_hw_color_hi = log_hw_color_hi)

p1 <- tm_shape(polygons56) + 
  tm_fill(col= "metric_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Flag score below 15th percentile",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p2 <- tm_shape(polygons56) + 
tm_fill(col = "celltype_color") +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Celltype",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p3 <- tm_shape(polygons56) + 
  tm_fill(col= "target_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Target counts below 15th percentile",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p4 <- tm_shape(polygons56) + 
  tm_fill(col= "count_area_ratio_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Target counts/custom um area below 15th percentile",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p5 <- tm_shape(polygons56) + 
  tm_fill(col= "area_color_lo", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Custom cell area below 15th percentile",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p6 <- tm_shape(polygons56) + 
  tm_fill(col= "area_color_hi", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Custom cell area above 85th percentile",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p7 <- tm_shape(polygons56) + 
  tm_fill(col= "log_hw_color_lo", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Logged AspectRatio below 15th percentile",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p8 <- tm_shape(polygons56) + 
  tm_fill(col= "log_hw_color_hi", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Logged AspectRatio above 85th percentile",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

tmap_arrange(p1, p2, p3, p4, p5, p6, p7, p8, ncol = 2)
```

## Flagged cell stats

```{r, out.width="150%", out.height="150%"}

flagged_cells <- c(sum(polygons56$metric_color=="red", na.rm = T),
                   sum(polygons56$target_color=="red", na.rm = T),
                   sum(polygons56$count_area_ratio_color=="red", na.rm = T),
                   sum(polygons56$area_color_lo=="red", na.rm = T),
                   sum(polygons56$area_color_hi=="red", na.rm = T),
                   sum(polygons56$log_hw_color_lo=="red", na.rm = T),
                   sum(polygons56$log_hw_color_hi=="red", na.rm = T))

unflagged_cells <- c(sum(polygons56$metric_color=="white", na.rm = T),
                     sum(polygons56$target_color=="white", na.rm = T),
                     sum(polygons56$count_area_ratio_color=="white", na.rm = T),
                     sum(polygons56$area_color_lo=="white", na.rm = T),
                     sum(polygons56$area_color_hi=="white", na.rm = T),
                     sum(polygons56$log_hw_color_lo=="white", na.rm = T),
                     sum(polygons56$log_hw_color_hi=="white", na.rm = T))

percent_flagged <- round(flagged_cells/dim(polygons56)[1]*100, 2)

poly56 <- data.frame(threshold = c(0.15, 0.15, 0.15, 0.15, 0.85, 0.15, 0.85), flag_value = flag_value, unflagged_cells = unflagged_cells, flagged_cells = flagged_cells, percent_flagged = percent_flagged, row.names = c("Flag score", "Target counts", "Target counts/um area", "Custom area in um low", "Custom area in um high", "Logged aspect ratio low", "Logged aspect ratio high"))


kable(poly56, format = "html", col.names = c("Threshold", "Flag value", "Unflagged cells", "Flagged cells", "% flagged cells"), align = "c", caption = paste0("<center>Flagged cells by different metrics, total cells in fov: ", dim(polygons56)[1], ", total cells in whole sample: ", dim(polygons_whole)[1], ", flagged cells in whole sample: ", sum(spe$flag_score_hw < quantile(spe$flag_score_hw, probs = 0.15, na.rm = T)), "</center>"), padding = 1L)
```

#
#

## Big flag threshold map

```{r, out.width="150%", out.height="150%"}
plot_list <- list(p1, p2 +
tm_add_legend(type = "fill", labels = names(celltype_palette), col = celltype_palette),
p3, p4, p5, p6, p7, p8)

for(i in 1:length(plot_list)){
  print(plot_list[[i]])
}
```


