---
title: "05_08_cosmx_human_tnbc_14603_first_flag_mc"
author: "Benedetta Banzi"
date: "2024-05-17"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

<STYLE TYPE="text/css">
<!--
  td{
    font-family: Arial; 
    font-size: 10 pt;
    padding:0px;
    cellpadding="0";
    cellspacing="0"
  }
  th {
    font-family: Arial; 
    font-size: 10 pt;
    height: 20px;
    font-weight: bold;
    text-align: "center";
    background-color: #ccccff;
    padding-left: 15px;
    padding-right: 15px;
  }
  table { 
    border-spacing: 4px;
    border-collapse: collapse;
  }
--->
</STYLE>


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Library import

```{r}
library(SpatialExperiment)
library(data.table)
library(scater) # it imports scuttle
library(cowplot)
library(ggplot2)
theme_set(theme_bw())
library(matrixStats)
library(dplyr)
library(tidyr)
library(tibble)
library(arrow)
library(scales)
library(sf)
library(tmap)
library(knitr)
library(grid)
library(ggpointdensity)
library(ggpubr)
library(spdep)
library(ggExtra)
library(terra)
library(DT)
library(e1071)
library(robustbase)
library(InSituType)
library(UpSetR)
```

Read-in function

```{r}
readCosmxSPE <- function(dirname = dirname, 
                         countmatfpattern = "exprMat_file.csv", 
                         metadatafpattern = "metadata_file.csv", 
                         coord_names = c("CenterX_global_px",
                                         "CenterY_global_px")){
  
  countmat_file <- file.path(dirname, list.files(dirname, countmatfpattern))
  metadata_file <- file.path(dirname, list.files(dirname, metadatafpattern))
  
  # Read in 
  countmat <- t(data.table::fread(countmat_file))
  metadata <- data.table::fread(metadata_file)
  
  metadata$fov_id <- paste(metadata$fov, metadata$cell_ID, sep = "_")
  colnames(countmat) <- paste(countmat["fov",], countmat["cell_ID",], sep = "_")
  metadata <- metadata[base::order(metadata$fov, metadata$cell_ID), ]
  countmat <- countmat[, base::order(countmat["fov",], countmat["cell_ID",])]
  
  # Count matrix 
  counts <- countmat[,metadata$fov_id]
  counts <- counts[!rownames(counts) %in% c("fov", "cell_ID"), ]
  
  # rowData (does not exist)
  
  # colData
  colData <- metadata
  
  spe <- SpatialExperiment::SpatialExperiment(
    assays = list(counts = counts),
    # rowData = rowData,
    colData = colData,
    spatialCoordsNames = coord_names
  )
  
  return(spe)
}
```

Setting directories

```{r}
dirname <- "/data01/Shared/Benedetta/spatial_transcriptomics/CosMx/Datasets/TNBC/UAP7_14603"
count_pattern <- "UAP7_14603_exprMat_file.csv"
meta_pattern <- "UAP7_14603_metadata_file.csv"
sample_name <- "CosMx_human_TNBC_14603"
```

# CosMx dataset preprocessing {.tabset .tabset-fade .tabset-pills}

Data import and SpatialExperiment object creation

```{r}
spe <- readCosmxSPE(dirname = dirname, countmatfpattern = count_pattern, metadatafpattern = meta_pattern)
spe <- scuttle::addPerCellQC(spe, subsets=list(NegProb = grep("^Negative", rownames(spe)), SysCtrl = grep("^SystemControl", rownames(spe))))
names(colData(spe)@listData)
spe$control_counts <- rowSums(data.frame(colData(spe))[,grep("^subsets_.*_sum$", names(colData(spe)))])
spe$control_detected <- rowSums(data.frame(colData(spe))[,grep("^subsets_.*_detected$", names(colData(spe)))])
```

Dataset dimensions

```{r}
dim(spe)
```

Adding variables

```{r, class.source = 'fold-show'}
fov_size <- 4256

names(colData(spe))[names(colData(spe)) == "cell_id"] <- "cell_fov"
names(colData(spe))[names(colData(spe)) == "cell_ID"] <- "cell_id"

# Importing global coordinates from spatialCoords slot since they are not available in colData
# coordinates are retrieved from the slot spe@int_colData@listData$spatialCoords
spe$CenterX_global_px <- spatialCoords(spe)[,1]
spe$CenterY_global_px <- spatialCoords(spe)[,2]

# Converting global coordinates to um 
spe$CenterX_global_um <- spe$CenterX_global_px*0.12
spe$CenterY_global_um <- spe$CenterY_global_px*0.12 

# Logged total counts
spe$logtot <- log10(spe$total)

# Converting area in um
spe$um_area <- spe$Area*0.0144

# Keeping only target counts by subtracting counts referred to control probes
spe$target_counts <- spe$total - spe$control_counts

# Target area ratio
spe$target_area_ratio <- spe$target_counts/spe$um_area

# Logged width/height ratio so there are no holes in the distribution in the middle and also negative values can be appreciated
spe$log2_AspectRatio <- log2(spe$AspectRatio)

# Keeping only detected features by subtracting negative control probes
spe$target_detected <- spe$detected - spe$control_detected

# Proportion between counts referred only to target probes and detected features
spe$tot_det_ratio <- spe$target_counts/spe$target_detected

# Control probe counts/total counts
ctrl_total_ratio <- spe$control_counts/spe$total
ctrl_total_ratio <- replace_na(ctrl_total_ratio, 0)
spe$ctrl_total_ratio <- ctrl_total_ratio
```

Loading custom polygon file and setting the correct format as sf object

Creating custom polygon file to add 

```{r}
fov_pattern <- "UAP7_14603_fov_positions_file.csv"
fovpos_file <- file.path(dirname, list.files(dirname, fov_pattern))

metadata <- data.frame(colData(spe))
fov_positions <- data.table::fread(fovpos_file, header = T)
fov_positions <- rename(fov_positions, fov = FOV)
fov_positions <- fov_positions[order(fov_positions$fov),]
fov_positions <- fov_positions[fov_positions$fov %in% unique(metadata$fov),]

spat_obj <- fread(file.path(dirname, "UAP7_14603-polygons.csv"))
colnames(spat_obj)[colnames(spat_obj) == "cellID"] <- "cell_id"

sf_global <- st_as_sf(spat_obj, coords = c("x_global_px", "y_global_px"))

polygons_whole_custom <- sf_global %>%
  group_by(fov, cell_id) %>%
  summarize(geometry = st_combine(geometry))#st_combine = combine several feature geometries into one, without unioning or resolving internal boundaries

polygons_whole_custom <- st_cast(polygons_whole_custom, "POLYGON")  

polygons_whole_custom <- polygons_whole_custom[base::order(polygons_whole_custom$fov, polygons_whole_custom$cell_id),]
rownames(polygons_whole_custom) <- paste(polygons_whole_custom$fov, polygons_whole_custom$cell_id, sep = "_")

if(attr(table(rownames(polygons_whole_custom) == colnames(spe)), "dimnames")[[1]][1] == FALSE){
  spe <- spe[, base::order(spe$fov, spe$cell_id)]
  colnames(spe) <- paste(spe$fov, spe$cell_id, sep = "_")
  if(dim(spe)[2] > dim(polygons_whole_custom)[1]){
    spe <- spe[, colnames(spe) %in% rownames(polygons_whole_custom)]
  }
  else{
    polygons_whole_custom <- polygons_whole_custom[rownames(polygons_whole_custom) %in% colnames(spe),]
  }
}

spe$polygons_whole_custom <- polygons_whole_custom
```

# Creating sf using local coordinates for geometry

```{r}
sf_local <- st_as_sf(spat_obj, coords = c("x_local_px", "y_local_px"))
names(sf_local)[names(sf_local) == "cellID"] <- "cell_id"

polygons <- sf_local %>%
  group_by(fov, cell_id) %>%
  summarize(geometry = st_combine(geometry))#st_combine = combine several feature geometries into one, without unioning or resolving internal boundaries

polygons <- st_cast(polygons, "POLYGON")

polygons <- polygons[base::order(polygons$fov, polygons$cell_id),]
rownames(polygons) <- paste(polygons$fov, polygons$cell_id, sep = "_") 

if(attr(table(rownames(polygons) == colnames(spe)), "dimnames")[[1]][1] == FALSE){
  spe <- spe[, base::order(spe$fov, spe$cell_id)]
  colnames(spe) <- paste(spe$fov, spe$cell_id, sep = "_")
  if(dim(spe)[2] > dim(polygons)[1]){
    spe <- spe[, colnames(spe) %in% rownames(polygons)]
  }
  else{
    polygons <- polygons[rownames(polygons) %in% colnames(spe),]
  }
}

spe$polygons <- polygons
```

Checking if the computed polygons are valid and removing elements that could result in geometry errors (Voyager)

```{r}
valid_poly <- st_is_valid(spe$polygons_whole_custom)
table(valid_poly)
```

```{r}
if(attr(table(valid_poly), "dimnames")[[1]][1] == FALSE){
  spe$polygons_whole_custom <- st_buffer(spe$polygons_whole_custom, dist = 0)
}
valid_poly <- st_is_valid(spe$polygons_whole_custom)
table(valid_poly)
```

Checking if all the geometries are polygons

```{r}
table(st_geometry_type(spe$polygons_whole_custom$geometry))
```

Subsetting to remove non-polygons

```{r}
cellids <- c() 
for(i in 1:dim(spe$polygons_whole_custom)[1]){
	cellids[i] <- attr(spe$polygons_whole_custom$geometry[[i]], "class")[2]
	cellids[i] <- cellids[i] == "MULTIPOLYGON"
	
}

spe <- spe[, cellids == FALSE]
table(st_geometry_type(spe$polygons_whole_custom))
```

```{r}
custom_area <- st_area(spe$polygons_whole_custom)
spe$um_area <- custom_area*0.0144
spe$polygons_custom_whole$um_area <- spe$um_area 

AspectRatio <- c()
for(i in 1:dim(spe$polygons_whole_custom)[1]){
 AspectRatio[i] <- (max(spe$polygons_whole_custom$geometry[[i]][[1]][,2]) -     min(spe$polygons_whole_custom$geometry[[i]][[1]][,2]))/(max(spe$polygons_whole_custom$geometry[[i]][[1]][,1]) - min(spe$polygons_whole_custom$geometry[[i]][[1]][,1]))
}

spe$polygons_whole_custom$log2_AspectRatio <- log2(AspectRatio) 
spe$log2_AspectRatio <- log2(AspectRatio) 
```

Checking if local polygons are valid

```{r}
valid_poly <- st_is_valid(spe$polygons)
table(valid_poly)
```

```{r}
if(attr(table(valid_poly), "dimnames")[[1]][1] == FALSE){
  spe$polygons <- st_buffer(spe$polygons, dist = 0)
}

valid_poly <- st_is_valid(spe$polygons)
table(valid_poly)
```

Checking if all the geometries are polygons

```{r}
table(st_geometry_type(spe$polygons$geometry))
```

# Large-scale to small-scale {.tabset .tabset-fade .tabset-pills}

Approaching dataset preprocessing starting from global-scale, then fov-focused exploratory analysis

## Global sample map

To view the spatial organization of both the sample as it is and FOVs

```{r, out.width="600%", out.height="150%"}
# image theme
global_map_theme <- function(back.color="black",back.border=NA,title.col="white") {
  theme(aspect.ratio = 1,
        panel.border = element_blank(),
        legend.key = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks = element_blank(),
        axis.text.y = element_blank(),
        axis.text.x = element_blank(),
        panel.grid = element_blank(),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        panel.background = element_rect(fill = "transparent",colour = NA),
        plot.title = element_text(color = title.col,hjust = 0.5,face = "bold"),
        plot.background = element_rect(fill = "transparent",colour = back.border))
}

ggplot() + 
  geom_point(data = metadata, mapping = aes(x = CenterX_global_px, y = CenterY_global_px), colour = "darkmagenta", fill="darkmagenta", size = 0.05, alpha = 0.2) +
  annotate("rect", xmin = fov_positions$X_px - 1/32*fov_size, 
         xmax = fov_positions$X_px + fov_size - 1/32*fov_size, 
         ymin = fov_positions$Y_px - fov_size - 1/16*fov_size, 
         ymax = fov_positions$Y_px - 1/16*fov_size,
         alpha = .2, color = "black",linewidth = 0.2) +
  geom_text(aes(x = fov_positions$X_px+11/10*fov_size/2,
                y = fov_positions$Y_px-11/10*fov_size/2,
                label = fov_positions$fov), color="black", size = 3) +
  ggtitle(sample_name) +
  global_map_theme(back.color = "white", back.border = "white", title.col = "black")
```

## Global polygon map

```{r, out.width="600%", out.height="900%"}
tm_shape(spe$polygons_whole_custom) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = sample_name,
            main.title.size = 0.5,
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))
```

## Global feature maps

To spot areas with anomalous feature values and spatial patterns on a global scale

```{r, out.width="150%", out.height="150%"}
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "total", point_size = 0.1) + ggtitle("Total counts per cell") + theme(aspect.ratio = 1)
```

```{r, out.width="150%", out.height="150%"}
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "logtot", point_size = 0.1) + ggtitle("Log-total counts per cell") + theme(aspect.ratio = 1)
```

```{r, out.width="150%", out.height="150%"}
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "detected", point_size = 0.1) + ggtitle("Detected features per cell") + theme(aspect.ratio = 1)
```

```{r, out.width="150%", out.height="150%"}
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "um_area", point_size = 0.1) + ggtitle("Area in μm per cell") + theme(aspect.ratio = 1)
```

```{r, out.width="150%", out.height="150%"}
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "log2_AspectRatio", point_size = 0.1) + ggtitle("Logged width/height ratio per cell") + theme(aspect.ratio = 1)
```

```{r, out.width="150%", out.height="150%"}
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "tot_det_ratio", point_size = 0.1) + ggtitle("Target counts/detected features ratio per cell") + theme(aspect.ratio = 1)
```

```{r, out.width="150%", out.height="150%"}
plotColData(spe, "CenterY_global_um", "CenterX_global_um", order_by = "control_counts", colour_by = "control_counts", point_size = 0.1) +
  scale_color_gradient(low = "white", high = "red", name = "control_counts") + ggtitle("Negative control probe counts per cell") +
  theme(aspect.ratio = 1,
        panel.background = element_rect(fill = "black",color = NA),
        plot.background = element_rect(fill = "black",color = NA),
        axis.title.x = element_text(color = "white"),
        axis.title.y = element_text(color = "white"),
        axis.ticks = element_line(color = "white"),
        axis.text.y = element_text(color = "white"),
        axis.text.x = element_text(color = "white"),
        axis.line = element_line(color = "white"),
        legend.title = element_text(color = "white"),
        legend.text = element_text(color = "white"),
        plot.title = element_text(color = "white"))
```

```{r, out.width="150%", out.height="150%"}
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "ctrl_total_ratio", point_size = 0.1) + ggtitle("Control counts/total counts ratio per cell") + theme(aspect.ratio = 1)
```

# Feature histograms 

To check feature distribution before highlighting where extreme values are located in the global sample map

```{r}
variables <- c("total", "detected", "logtot", "um_area", "log2_AspectRatio", "tot_det_ratio", "control_counts", "Mean.DAPI", "ctrl_total_ratio")
for (var in 1:length(variables)){
  n <- grep(paste0("^", variables[var], "$"), colnames(spe@colData))
  histo_var <- names(colData(spe))[n]
  p <- ggplot(data = data.frame(colData(spe))) +
    geom_histogram(aes(x = .data[[histo_var]]), fill = "#69b3a2") +
    ggtitle(paste0(histo_var))
  print(p)
}  
```

# First cell skimming with constant values {.tabset .tabset-fade .tabset-pills}

Coarse-grain cell filtering to remove gross segmentation errors.

## Total counts

```{r}
summary(spe$total)
```

```{r, out.width="150%", out.height="150%"}
hist(spe$total)
```

```{r, out.width="150%", out.height="150%"}
boxplot(spe$total)
```

```{r, out.width="150%", out.height="150%"}
threshold <- 0
spe$flagged_0total <- as.factor(ifelse(colData(spe)$total == 0, TRUE, FALSE))
spe$polygons_whole_custom$flagged_0total <- ifelse(spe$flagged_0total == TRUE, "red", "grey80")

tm_shape(spe$polygons_whole_custom) + 
  tm_fill(col = "flagged_0total") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = paste0("Cells having 0 total counts", "\nFlagged cells = ",
                                table(colData(spe)$flagged_0total)["TRUE"], " out of ", dim(spe)[2], ", ",
                                round(table(colData(spe)$flagged_0total)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"),
            main.title.fontface = 2,
            main.title.size = 1, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))
```

```{r}
total0_spe <- spe[,spe$flagged_0total == FALSE]
dim(total0_spe)
```

## Control count / total counts

```{r}
summary(spe$ctrl_total_ratio)
```

```{r, out.width="150%", out.height="150%"}
hist(spe$ctrl_total_ratio)
```

```{r, out.width="150%", out.height="150%"}
boxplot(spe$ctrl_total_ratio)
```

```{r, out.width="150%", out.height="150%"}
spe$outlier_ctrl_tot <- spe$ctrl_total_ratio > 0.1 # Voyager
spe$polygons_whole_custom$outlier_ctrl_tot <- ifelse(spe$outlier_ctrl_tot == TRUE, "red", "grey80")

tm_shape(spe$polygons_whole_custom) + 
  tm_fill(col= "outlier_ctrl_tot") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = paste0("Outlier cells for control / total counts ratio > 0.1, \nflagged cells = ",
                                table(colData(spe)$outlier_ctrl_tot)["TRUE"], " out of ", dim(spe)[2], ", ",
                                round(table(colData(spe)$outlier_ctrl_tot)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"),
            main.title.fontface = 2,
            main.title.size = 1, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

```

```{r}
ctrltot_spe <- spe[, spe$outlier_ctrl_tot == FALSE]
dim(ctrltot_spe)
```

```{r}
summary(ctrltot_spe$ctrl_total_ratio)
```

```{r, out.width="150%", out.height="150%"}
hist(ctrltot_spe$ctrl_total_ratio)
```

```{r, out.width="150%", out.height="150%"}
boxplot(ctrltot_spe$ctrl_total_ratio)
```

# First cell skimming with skewness adjustment to boxplot {.tabset .tabset-fade .tabset-pills}

Introducing medcouple (MC) from Hubert, M. and Vandervieren, E. (2008). Boxplot whiskers are defined as:
1st quartile - 1.5 * hl(MC) * IQR 
3rd quartile + 1.5 * hu(MC) * IQR
If MC >= 0, hl = exp(-4 * MC), hu = exp(3 * MC) 
If MC < 0, hl = exp(-3 * MC), hu = exp(4 * MC)

```{r}
spe$polygons_whole_custom$control_counts <- spe$control_counts
spe$polygons_whole_custom$ctrl_total_ratio <- spe$ctrl_total_ratio
spe$polygons_whole_custom$um_area <- spe$um_area
spe$polygons_whole_custom$log2_AspectRatio <- spe$log2_AspectRatio
spe$polygons_whole_custom$Mean.DAPI <- spe$Mean.DAPI
```

## Um area

```{r}
summary(spe$um_area)
```

```{r, out.width="150%", out.height="150%"}
hist(spe$um_area)
```

Skewness

```{r}
skewness(spe$um_area)[1]
```

Medcouple

```{r}
mc(spe$um_area)[1]
```

How many outliers were found?

```{r}
names(spe$um_area) <- colnames(spe)
out <- adjbox(spe$um_area)
length(out$out)
area_fence <- out$fence
```

Distribution of outlier values

```{r}
summary(out$out)
```

```{r, out.width="150%", out.height="150%"}
outlier_color <- colnames(spe)
for(i in 1:out$n){
  outlier_color[i] <- ifelse(outlier_color[i] %in% names(out$out), "red", "grey80")
}

spe$polygons_whole_custom$umarea_outlier <- outlier_color

tm_shape(spe$polygons_whole_custom) + 
  tm_fill(col= "umarea_outlier") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = paste0("Outlier cells for um area", 
                                "\n fence = " , round(out$fence, 2),
                                "\nFlagged cells = ",
                                table(spe$polygons_whole_custom$umarea_outlier)["red"], " out of ", dim(spe)[2], ", ",
                                round(table(spe$polygons_whole_custom$umarea_outlier)["red"]/dim(spe)[2]*100, 2), "% of total cells"),
            main.title.fontface = 2,
            main.title.size = 1, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))
```

```{r}
area_poly <- spe$polygons_whole_custom[spe$polygons_whole_custom$umarea_outlier == "grey80",]
hist(area_poly$um_area)
```

```{r, out.width="150%", out.height="150%"}
boxplot(area_poly$um_area)
```

```{r}
summary(area_poly$um_area)
```

```{r}
area_spe <- spe[,spe$polygons_whole_custom$umarea_outlier == "grey80"]
dim(area_spe)
```

## Mean DAPI signal

```{r}
summary(spe$Mean.DAPI)
```

```{r, out.width="150%", out.height="150%"}
hist(spe$Mean.DAPI)
```

Skewness

```{r}
skewness(spe$Mean.DAPI)[1]
```

Medcouple

```{r}
mc(spe$Mean.DAPI)[1]
```

How many outliers were found?

```{r}
names(spe$Mean.DAPI) <- colnames(spe)
out <- adjbox(spe$Mean.DAPI)
length(out$out)
dapi_fence <- out$fence
```

Distribution of outlier values

```{r}
summary(out$out)
```

```{r, out.width="150%", out.height="150%"}
outlier_color <- colnames(spe)
for(i in 1:out$n)
  outlier_color[i] <- ifelse(outlier_color[i] %in% names(out$out), "red", "grey80")

spe$polygons_whole_custom$dapi_outlier <- outlier_color

tm_shape(spe$polygons_whole_custom) + 
  tm_fill(col= "dapi_outlier") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = paste0("Outlier cells for dapi", 
                                "\n fence = " , round(out$fence, 2),
                                "\nFlagged cells = ",
                                table(spe$polygons_whole_custom$dapi_outlier)["red"], " out of ", dim(spe)[2], ", ",
                                round(table(spe$polygons_whole_custom$dapi_outlier)["red"]/dim(spe)[2]*100, 2), "% of total cells"),
            main.title.fontface = 2,
            main.title.size = 1, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))
```

```{r}
dapi_poly <- spe$polygons_whole_custom[spe$polygons_whole_custom$dapi_outlier == "grey80",]
hist(dapi_poly$Mean.DAPI)
```

```{r, out.width="150%", out.height="150%"}
boxplot(dapi_poly$Mean.DAPI)
```

```{r}
summary(dapi_poly$Mean.DAPI)
```

```{r}
dapi_spe <- spe[, spe$polygons_whole_custom$dapi_outlier == "grey80"]
dim(dapi_spe)
```

```{r}
out_boundaries <- c(0, 0.1, 
               paste(round(area_fence[1,1], 2), round(area_fence[2,1], 2)),
               paste(round(dapi_fence[1,1], 2), round(dapi_fence[2,1], 2)))

flagged_cells <- c(sum(spe$polygons_whole_custom$flagged_0total=="red", na.rm = T),
                   sum(spe$polygons_whole_custom$outlier_ctrl_tot=="red", na.rm = T),
                   sum(spe$polygons_whole_custom$umarea_outlier=="red",na.rm = T),
                   sum(spe$polygons_whole_custom$dapi_outlier=="red",na.rm = T))

unflagged_cells <- c(sum(spe$polygons_whole_custom$flagged_0total=="grey80", na.rm = T),
                     sum(spe$polygons_whole_custom$outlier_ctrl_tot=="grey80", na.rm = T),
                     sum(spe$polygons_whole_custom$umarea_outlier=="grey80",na.rm = T),
                     sum(spe$polygons_whole_custom$dapi_outlier=="grey80",na.rm = T))

percent_flagged <- round(flagged_cells/dim(spe)[2]*100, 2)

union_flagged <- length(unique(c(spe[,spe$polygons_whole_custom$flagged_0total == "red"]$fov_id,
                   spe[,spe$polygons_whole_custom$outlier_ctrl_tot == "red"]$fov_id,
                   spe[,spe$polygons_whole_custom$umarea_outlier == "red"]$fov_id,
                   spe[,spe$polygons_whole_custom$dapi_outlier == "red"]$fov_id)))

stat_df <- data.frame(out_boundaries = out_boundaries, unflagged_cells = unflagged_cells, flagged_cells = flagged_cells, percent_flagged = percent_flagged, row.names = c("Total counts", "Control/total counts ratio", "Cell area in um", "Mean DAPI signal"))

kable(stat_df, format = "html", col.names = c("Outlier boundaries", "Unflagged cells", "Flagged cells", "% flagged cells"), align = "c", caption = paste0("<center>Flagged cells by different metrics, total cells in whole sample: ", dim(spe$polygons_whole_custom)[1], ", total unique flagged cells in whole sample: ", union_flagged, ", % unique flagged cells: ",
round(union_flagged/dim(spe)[2]*100, 2),"</center>"), padding = 1L)

```

```{r}
list_input <- list(flagged_0total = spe[,spe$polygons_whole_custom$flagged_0total == "red"]$fov_id, 
           outlier_ctrl_tot = spe[,spe$polygons_whole_custom$outlier_ctrl_tot == "red"]$fov_id,
           umarea_outlier = spe[,spe$polygons_whole_custom$umarea_outlier == "red"]$fov_id,
           dapi_outlier = spe[,spe$polygons_whole_custom$dapi_outlier == "red"]$fov_id)

upset(fromList(list_input))
```
