---
title: "05_08_merfish_mouse_liver1_slice1_first_flag_mc"
author: "Benedetta Banzi"
date: "2024-05-20"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

<STYLE TYPE="text/css">
<!--
  td{
    font-family: Arial; 
    font-size: 10 pt;
    padding:0px;
    cellpadding="0";
    cellspacing="0"
  }
  th {
    font-family: Arial; 
    font-size: 10 pt;
    height: 20px;
    font-weight: bold;
    text-align: "center";
    background-color: #ccccff;
    padding-left: 15px;
    padding-right: 15px;
  }
  table { 
    border-spacing: 4px;
    border-collapse: collapse;
  }
--->
</STYLE>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Preliminary steps

Library import

```{r}
library(SpatialExperiment)
library(data.table)
library(scater) # it imports scuttle
library(cowplot)
library(ggplot2)
theme_set(theme_bw())
library(matrixStats)
library(dplyr)
library(tidyr)
library(tibble)
library(arrow)
library(scales)
library(sf)
library(tmap)
library(knitr)
library(grid)
library(ggpointdensity)
library(ggpubr)
library(spdep)
library(ggExtra)
library(DT)
library(e1071)
library(robustbase)
library(UpSetR)
```

Read-in function

```{r}
readMerfishSPE <- function(dirname = dirname, 
                            countmatfpattern = "cell_by_gene.csv", 
                            metadatafpattern = "cell_metadata.csv", 
                            coord_names = c("center_x", "center_y")){

  countmat_file <- file.path(dirname, list.files(dirname, countmatfpattern))
  metadata_file <- file.path(dirname, list.files(dirname, metadatafpattern))

  # Read in 
  countmat <- data.table::fread(countmat_file)
  names(countmat)[names(countmat) == "V1"] <- "cell_id"
  countmat <- countmat[base::order(countmat$cell_id), ]
  rownames(countmat) <- countmat$cell_id
  
  metadata <- data.table::fread(metadata_file)
  names(metadata)[names(metadata) == "V1"] <- "cell_id"
  metadata <- metadata[order(metadata$cell_id), ]
  rownames(metadata) <- metadata$cell_id
  
  if(dim(metadata)[1] < dim(countmat)[1]){
    countmat <- countmat[rownames(countmat) %in% rownames(metadata),]
  }
  
  if(dim(countmat)[1] < dim(metadata)[1]){
    metadata <- metadata[rownames(metadata) %in% rownames(countmat),]
  }

  # Count matrix 
  counts <- t(countmat)
  colnames(counts) <- counts["cell_id",]
  counts <- counts[!rownames(counts)=="cell_id",]
  count_matrix <- apply(counts, 2, as.integer)
  rownames(count_matrix) <- rownames(counts)
  

  # rowData (does not exist)

  # colData
  colData <- metadata

  spe <- SpatialExperiment::SpatialExperiment(
    assays = list(counts = count_matrix),
    # rowData = rowData,
    colData = colData,
    spatialCoordsNames = coord_names
  )

  return(spe)

}
```

Setting directories

```{r}
dirname <- "/data01/Shared/Benedetta/spatial_transcriptomics/Vizgen_MERSCOPE/Datasets/LiverMouse1Slice1"
sample_name <- "Merfish_mouse_liver1_slice1"
```

# Merscope dataset exploratory analysis for preprocessing {.tabset .tabset-fade .tabset-pills}

SpatialExperiment object creation and exploration

```{r}
spe <- readMerfishSPE(dirname = dirname)
spe <- scuttle::addPerCellQC(spe, subsets=list(Blank = grep("^Blank-", rownames(spe))))
names(colData(spe)@listData)
spe$control_counts <- spe$subsets_Blank_sum
spe$control_detected <- spe$subsets_Blank_detected
```

Dataset dimensions

```{r}
dim(spe)
```

Added variables

```{r, class.source = 'fold-show'}
# Importing global coordinates from spatialCoords slot since they are not available in colData
colData(spe)$CenterX_global_um <- spatialCoords(spe)[,1]
colData(spe)$CenterY_global_um <- spatialCoords(spe)[,2]

# fov start from 0
spe$fov <- spe$fov + 1

# Logged total counts
spe$logtot <- log10(spe$total)

# Keeping only target counts by subtracting counts referred to negative control probes
spe$target_counts <- spe$total - spe$control_counts

# Keeping only detected features by subtracting negative control probes
spe$target_detected <- spe$detected - spe$control_detected

# Proportion between counts referred only to target probes and detected features
spe$tot_det_ratio <- spe$target_counts/spe$target_detected

# Control probe counts/total counts
ctrl_total_ratio <- spe$control_counts/spe$total
ctrl_total_ratio <- replace_na(ctrl_total_ratio, 0)
spe$ctrl_total_ratio <- ctrl_total_ratio
```

Loading custom polygon file and setting the correct format as sf object

Creating custom polygon file to add 

```{r}
spat_obj <- readRDS(file.path(dirname, "polygons.rds"))
names(spat_obj)[names(spat_obj) == "cellID"] <- "cell_id"

polygons_whole <- spat_obj

polygons_whole <- polygons_whole[base::order(polygons_whole$cell_id),]
rownames(polygons_whole) <- polygons_whole$cell_id

if(attr(table(rownames(polygons_whole) == colnames(spe)), "dimnames")[[1]][1] == FALSE){
  spe <- spe[, base::order(spe$fov, spe$cell_id)]
  colnames(spe) <- paste(spe$fov, spe$cell_id, sep = "_")
  if(dim(spe)[2] > dim(polygons_whole)[1]){
    spe <- spe[, colnames(spe) %in% rownames(polygons_whole)]
  }
  else{
    polygons_whole <- polygons_whole[rownames(polygons_whole) %in% colnames(spe),]
  }
}

spe$polygons_whole <- polygons_whole
```

# Adding fov information for future local plotting

```{r}
spe$polygons_whole$fov <- spe$fov
```

Checking if the computed polygons are valid and removing elements that could result in geometry errors (Voyager)

```{r}
valid_poly <- st_is_valid(spe$polygons_whole)
table(valid_poly)
```

```{r}
if(attr(table(valid_poly), "dimnames")[[1]][1] == FALSE){
  spe$polygons_whole <- st_buffer(spe$polygons_whole, dist = 0)
}
valid_poly <- st_is_valid(spe$polygons_whole)
table(valid_poly)
```

Checking if all the geometries are polygons

```{r}
table(st_geometry_type(spe$polygons_whole$geometry)) # 0 multipolygons
```

Subsetting to remove non-polygons

```{r}
if (table(st_geometry_type(spe$polygons_whole$geometry))[7] > 0){
  cellids <- c() 
  for(i in 1:dim(spe$polygons_whole)[1]){
	cellids[i] <- attr(spe$polygons_whole$geometry[[i]], "class")[2]
	cellids[i] <- cellids[i] == "MULTIPOLYGON"
	}
 spe <- spe[, cellids == FALSE]
 return(spe)
}

table(st_geometry_type(spe$polygons_whole))
```

```{r}
custom_area <- st_area(spe$polygons_whole)
spe$um_area <- custom_area*0.0324
spe$polygons_whole$um_area <- spe$um_area 

AspectRatio <- c()
for(i in 1:dim(spe$polygons_whole)[1]){
 AspectRatio[i] <- (max(spe$polygons_whole$geometry[[i]][[1]][,2]) -     min(spe$polygons_whole$geometry[[i]][[1]][,2]))/(max(spe$polygons_whole$geometry[[i]][[1]][,1]) - min(spe$polygons_whole$geometry[[i]][[1]][,1]))
}

spe$polygons_whole$log2_AspectRatio <- log2(AspectRatio) 
spe$log2_AspectRatio <- log2(AspectRatio) 
```

# Large-scale to small-scale {.tabset .tabset-fade .tabset-pills}

Approaching dataset preprocessing starting from global-scale, then fov-focused exploratory analysis

## Global polygon map

To view the spatial organization of both the sample as it is and FOVs

```{r, out.width="900%", out.height="600%"}
palette <- c("gray11", "gray47","blue","royalblue1","dodgerblue4","deepskyblue","cyan","lightblue1","darkseagreen1"
             ,"darkolivegreen1","limegreen","green3","darkgreen","yellow4","yellow","greenyellow","olivedrab1","seagreen2"
             ,"lightyellow","goldenrod4","goldenrod1","lightgoldenrod1","lightgoldenrod4","darkorange1","plum1","orchid4"
             ,"lightpink","lightcoral","orangered1","deeppink1","darkmagenta","red4")

color <- rep(palette, length.out = 1791)

fov_color <- c()
for(i in 1:dim(spe)[2]){
  fov_color[i] <- color[spe$fov[i]]  
}


spe$polygons_whole$fov_color <- fov_color

tm_shape(spe$polygons_whole)+
  tm_borders(lwd = 0.1, col = "grey50")+
  tm_fill(col = "fov_color") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = sample_name,
            main.title.size = 0.5,
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))
```

## Global feature maps

To spot areas with anomalous feature values and spatial patterns on a global scale

```{r, out.width="150%", out.height="150%"}
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "total", point_size = 0.1) + ggtitle("Total counts per cell") + theme(aspect.ratio = 1)
```

```{r, out.width="150%", out.height="150%"}
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "logtot", point_size = 0.1) + ggtitle("Log-total counts per cell") + theme(aspect.ratio = 1)
```

```{r, out.width="150%", out.height="150%"}
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "detected", point_size = 0.1) + ggtitle("Detected features per cell") + theme(aspect.ratio = 1)
```

```{r, out.width="150%", out.height="150%"}
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "um_area", point_size = 0.1) + ggtitle("Area in μm per cell") + theme(aspect.ratio = 1)
```

```{r, out.width="150%", out.height="150%"}
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "log2_AspectRatio", point_size = 0.1) + ggtitle("Logged width/height ratio per cell") + theme(aspect.ratio = 1)
```

```{r, out.width="150%", out.height="150%"}
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "tot_det_ratio", point_size = 0.1) + ggtitle("Target counts/detected features ratio per cell") + theme(aspect.ratio = 1)
```

```{r, out.width="150%", out.height="150%"}
plotColData(spe, "CenterY_global_um", "CenterX_global_um", order_by = "control_counts", colour_by = "control_counts", point_size = 0.1) +
  scale_color_gradient(low = "white", high = "red", name = "control_counts") + ggtitle("Negative control probe counts per cell") +
  theme(aspect.ratio = 1,
        panel.background = element_rect(fill = "black",color = NA),
        plot.background = element_rect(fill = "black",color = NA),
        axis.title.x = element_text(color = "white"),
        axis.title.y = element_text(color = "white"),
        axis.ticks = element_line(color = "white"),
        axis.text.y = element_text(color = "white"),
        axis.text.x = element_text(color = "white"),
        axis.line = element_line(color = "white"),
        legend.title = element_text(color = "white"),
        legend.text = element_text(color = "white"),
        plot.title = element_text(color = "white"))
```

```{r, out.width="150%", out.height="150%"}
plotColData(spe, "CenterY_global_um", "CenterX_global_um", colour_by = "ctrl_total_ratio", point_size = 0.1) + ggtitle("Control counts/total counts ratio per cell") + theme(aspect.ratio = 1)
```

# Feature histograms 

To check feature distribution before highlighting where extreme values are located in the global sample map

```{r}
variables <- c("total", "detected", "logtot", "um_area", "log2_AspectRatio", "tot_det_ratio", "control_counts", "ctrl_total_ratio")
for (var in 1:length(variables)){
  n <- grep(paste0("^", variables[var], "$"), colnames(spe@colData))
  histo_var <- names(colData(spe))[n]
  p <- ggplot(data = data.frame(colData(spe))) +
    geom_histogram(aes(x = .data[[histo_var]]), fill = "#69b3a2") +
    ggtitle(paste0(histo_var))
  print(p)
}  
```

# First cell skimming with constant values {.tabset .tabset-fade .tabset-pills}

Coarse-grain cell filtering to remove gross segmentation errors.

## Total counts

```{r}
summary(spe$total)
```

```{r, out.width="150%", out.height="150%"}
hist(spe$total)
```

```{r, out.width="150%", out.height="150%"}
boxplot(spe$total)
```

```{r, out.width="150%", out.height="150%"}
threshold <- 0
spe$flagged_0total <- as.factor(ifelse(colData(spe)$total == 0, TRUE, FALSE))
spe$polygons_whole$flagged_0total <- ifelse(spe$flagged_0total == TRUE, "red", "grey80")

tm_shape(spe$polygons_whole) + 
  tm_fill(col = "flagged_0total") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = paste0("Cells having 0 total counts", "\nFlagged cells = ",
                                table(colData(spe)$flagged_0total)["TRUE"], " out of ", dim(spe)[2], ", ",
                                round(table(colData(spe)$flagged_0total)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"),
            main.title.fontface = 2,
            main.title.size = 1, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))
```

```{r}
total0_spe <- spe[,spe$flagged_0total == FALSE]
dim(total0_spe)
```

## Control count / total counts

```{r}
summary(spe$ctrl_total_ratio)
```

```{r, out.width="150%", out.height="150%"}
hist(spe$ctrl_total_ratio)
```

```{r, out.width="150%", out.height="150%"}
boxplot(spe$ctrl_total_ratio)
```

```{r, out.width="150%", out.height="150%"}
spe$outlier_ctrl_tot <- spe$ctrl_total_ratio > 0.1 # Voyager
spe$polygons_whole$outlier_ctrl_tot <- ifelse(spe$outlier_ctrl_tot == TRUE, "red", "grey80")

tm_shape(spe$polygons_whole) + 
  tm_fill(col= "outlier_ctrl_tot") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = paste0("Outlier cells for control / total counts ratio > 0.1, \nflagged cells = ",
                                table(colData(spe)$outlier_ctrl_tot)["TRUE"], " out of ", dim(spe)[2], ", ",
                                round(table(colData(spe)$outlier_ctrl_tot)["TRUE"]/dim(spe)[2]*100, 2), "% of total cells"),
            main.title.fontface = 2,
            main.title.size = 1, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

```

```{r}
ctrltot_spe <- spe[, spe$outlier_ctrl_tot == FALSE]
dim(ctrltot_spe)
```

```{r}
summary(ctrltot_spe$ctrl_total_ratio)
```

```{r, out.width="150%", out.height="150%"}
hist(ctrltot_spe$ctrl_total_ratio)
```

```{r, out.width="150%", out.height="150%"}
boxplot(ctrltot_spe$ctrl_total_ratio)
```

# First cell skimming with skewness adjustment to boxplot {.tabset .tabset-fade .tabset-pills}

Introducing medcouple (MC) from Hubert, M. and Vandervieren, E. (2008). Boxplot whiskers are defined as:
1st quartile - 1.5 * hl(MC) * IQR 
3rd quartile + 1.5 * hu(MC) * IQR
If MC >= 0, hl = exp(-4 * MC), hu = exp(3 * MC) 
If MC < 0, hl = exp(-3 * MC), hu = exp(4 * MC)

```{r}
spe$polygons_whole$control_counts <- spe$control_counts
spe$polygons_whole$ctrl_total_ratio <- spe$ctrl_total_ratio
spe$polygons_whole$um_area <- spe$um_area
spe$polygons_whole$log2_AspectRatio <- spe$log2_AspectRatio
spe$polygons_whole$Mean.DAPI <- spe$Mean.DAPI
```

## Um area

```{r}
summary(spe$um_area)
```

```{r, out.width="150%", out.height="150%"}
hist(spe$um_area)
```

Skewness

```{r}
skewness(spe$um_area)[1]
```

Medcouple

```{r}
mc(spe$um_area)[1]
```

How many outliers were found?

```{r}
names(spe$um_area) <- colnames(spe)
out <- adjbox(spe$um_area)
length(out$out)
area_fence <- out$fence
```

Distribution of outlier values

```{r}
summary(out$out)
```

```{r, out.width="150%", out.height="150%"}
outlier_color <- colnames(spe)
for(i in 1:out$n){
  outlier_color[i] <- ifelse(outlier_color[i] %in% names(out$out), "red", "grey80")
}

spe$polygons_whole$umarea_outlier <- outlier_color

tm_shape(spe$polygons_whole) + 
  tm_fill(col= "umarea_outlier") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = paste0("Outlier cells for um area", 
                                "\n fence = " , round(out$fence, 2),
                                "\nFlagged cells = ",
                                table(spe$polygons_whole$umarea_outlier)["red"], " out of ", dim(spe)[2], ", ",
                                round(table(spe$polygons_whole$umarea_outlier)["red"]/dim(spe)[2]*100, 2), "% of total cells"),
            main.title.fontface = 2,
            main.title.size = 1, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))
```

```{r}
area_poly <- spe$polygons_whole[spe$polygons_whole$umarea_outlier == "grey80",]
hist(area_poly$um_area)
```

```{r, out.width="150%", out.height="150%"}
boxplot(area_poly$um_area)
```

```{r}
summary(area_poly$um_area)
```

```{r}
area_spe <- spe[,spe$polygons_whole$umarea_outlier == "grey80"]
dim(area_spe)
```


```{r}
out_boundaries <- c(0, 0.1, 
               paste(round(area_fence[1,1], 2), round(area_fence[2,1], 2)))

flagged_cells <- c(sum(spe$polygons_whole$flagged_0total=="red", na.rm = T),
                   sum(spe$polygons_whole$outlier_ctrl_tot=="red", na.rm = T),
                   sum(spe$polygons_whole$umarea_outlier=="red",na.rm = T))

unflagged_cells <- c(sum(spe$polygons_whole$flagged_0total=="grey80", na.rm = T),
                     sum(spe$polygons_whole$outlier_ctrl_tot=="grey80", na.rm = T),
                     sum(spe$polygons_whole$umarea_outlier=="grey80",na.rm = T))

percent_flagged <- round(flagged_cells/dim(spe)[2]*100, 2)

union_flagged <- length(unique(c(spe[,spe$polygons_whole$flagged_0total == "red"]$cell_id,
                   spe[,spe$polygons_whole$outlier_ctrl_tot == "red"]$cell_id,
                   spe[,spe$polygons_whole$umarea_outlier == "red"]$cell_id)))

stat_df <- data.frame(out_boundaries = out_boundaries, unflagged_cells = unflagged_cells, flagged_cells = flagged_cells, percent_flagged = percent_flagged, row.names = c("Total counts", "Control/total counts ratio", "Cell area in um"))

kable(stat_df, format = "html", col.names = c("Outlier boundaries", "Unflagged cells", "Flagged cells", "% flagged cells"), align = "c", caption = paste0("<center>Flagged cells by different metrics, total cells in whole sample: ", dim(spe$polygons_whole)[1], ", total unique flagged cells in whole sample: ", union_flagged, ", % unique flagged cells: ",
round(union_flagged/dim(spe)[2]*100, 2), "</center>"), padding = 1L)

```


```{r}
list_input <- list(flagged_0total = spe[,spe$polygons_whole$flagged_0total == "red"]$cell_id, 
           outlier_ctrl_tot = spe[,spe$polygons_whole$outlier_ctrl_tot == "red"]$cell_id,
           umarea_outlier = spe[,spe$polygons_whole$umarea_outlier == "red"]$cell_id)

upset(fromList(list_input))
```
