---
title: "Spatial_preprocessing_dev"
author: "Benedetta Banzi"
date: "2024-01-25"
output: html_document
---

<STYLE TYPE="text/css">
<!--
  td{
    font-family: Arial; 
    font-size: 10 pt;
    padding:0px;
    cellpadding="0";
    cellspacing="0"
  }
  th {
    font-family: Arial; 
    font-size: 10 pt;
    height: 20px;
    font-weight: bold;
    text-align: "center";
    background-color: #ccccff;
  }
  table { 
    border-spacing: 4px;
    border-collapse: collapse;
  }
--->
</STYLE>


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Library import

```{r}
library(SpatialExperiment)
library(data.table)
library(scater) # it imports scuttle
library(cowplot)
library(ggplot2)
theme_set(theme_bw())
library(matrixStats)
library(dplyr)
library(tidyr)
library(tibble)
library(arrow)
library(scales)
library(sf)
library(tmap)
library(knitr)
library(grid)
```

Read-in function

```{r}
readCosmxSPE <- function(dirname = dirname, 
                         countmatfpattern = "exprMat_file.csv", 
                         metadatafpattern = "metadata_file.csv", 
                         coord_names = c("CenterX_global_px",
                                         "CenterY_global_px")){
  
  countmat_file <- file.path(dirname, list.files(dirname, countmatfpattern))
  metadata_file <- file.path(dirname, list.files(dirname, metadatafpattern))
  
  # Read in 
  countmat <- data.table::fread(countmat_file)
  metadata <- data.table::fread(metadata_file)
  
  # Count matrix   
  counts <- merge(countmat, metadata[, c("fov", "cell_ID")])
  counts <- subset(counts, select = -c(fov, cell_ID))
  cell_codes <- rownames(counts)
  features <- colnames(counts)
  counts <- t(counts)
  
  rownames(counts) <- features
  colnames(counts) <- cell_codes
  
  # rowData (does not exist)
  
  # colData
  colData <- merge(metadata, countmat[, c("fov", "cell_ID")])
  
  spe <- SpatialExperiment::SpatialExperiment(
    assays = list(counts = counts),
    # rowData = rowData,
    colData = colData,
    spatialCoordsNames = coord_names
  )
  
  return(spe)
}
```

Setting directories

```{r}
dirname <- "/NAS06/work/Spatial_omics/DBKero/CosMx_Breast/CosMX_data_Case2"
count_pattern <- "Run5810_Case2_exprMat_file.csv"
meta_pattern <- "Run5810_Case2_metadata_file.csv"
sample_name <- "CosMx_DBKero_BC"
```

# CosMx dataset exploratory analysis for preprocessing

SpatialExperiment object creation and exploration

```{r}
spe <- readCosmxSPE(dirname = dirname, countmatfpattern = count_pattern, metadatafpattern = meta_pattern)
spe <- scuttle::addPerCellQC(spe, subsets=list(NegProb = grep("^NegPrb", rownames(spe))))
names(colData(spe)@listData)
```

```{r}
dim(spe)
```

# Fov position file loading

```{r}
fov_pattern <- "Run5810_Case2_fov_positions_file.csv"
fovpos_file <- file.path(dirname, list.files(dirname, fov_pattern))

metadata <- data.table::fread(file.path(dirname, list.files(dirname, meta_pattern)))
fov_positions <- data.table::fread(fovpos_file, header = T)
fov_positions <- fov_positions[fov_positions$fov%in%unique(metadata$fov),]
```

# Added variables

```{r}
fov_size <- 4256

# Importing global coordinates from spatialCoords slot since they are not available in colData
colData(spe)$CenterX_global_px <- spatialCoords(spe)[,1]
colData(spe)$CenterY_global_px <- spatialCoords(spe)[,2]

# Converting global coordinates to um 
spe$CenterX_global_um <- spe$CenterX_global_px*0.18
spe$CenterY_global_um <- spe$CenterY_global_px*0.18 

# Logged total counts
spe$logtot <- log10(spe$total)

# Converting area in um
spe$um_area <- spe$Area*0.0324

# Logged width/height ratio so there are no holes in the distribution in the middle and also negative values can be appreciated
spe$log2_AspectRatio <- log2(spe$AspectRatio)

# Keeping only target counts by subtracting counts referred to negative control probes
spe$target_counts <- spe$total - spe$subsets_NegProb_sum

# Keeping only detected features by subtracting negative control probes
spe$target_detected <- spe$detected - spe$subsets_NegProb_detected

# Proportion between counts referred only to target probes and detected features
spe$tot_det_ratio <- spe$target_counts/spe$target_detected

spe$ctrl_total_ratio <- spe$subsets_NegProb_sum/spe$total
```

# Sf object creation from metadata

```{r}
meta_df <- data.frame(colData(spe))
meta_df <- group_by(meta_df, fov) |> summarize(n_cells = n(), mean_total = mean(total), mean_detected = mean(detected), mean_umarea = mean(um_area), mean_log2_AspectRatio = mean(log2_AspectRatio), mean_tot_det_ratio = mean(tot_det_ratio), mean_subsets_NegProb_sum = mean(subsets_NegProb_sum), median_total = median(total), median_detected = median(detected), median_umarea = median(um_area),median_log2_AspectRatio = median(log2_AspectRatio), median_tot_det_ratio = median(tot_det_ratio), median_subsets_NegProb_sum = median(subsets_NegProb_sum))

fov_positions <- left_join(fov_positions, meta_df, by = join_by(fov))

fov_size <- 4256
poly_df <- transmute(fov_positions, x_1 = x_global_px, y_1 = y_global_px, 
                     x_2 = x_global_px + fov_size, y_2 = y_global_px, 
                     x_3 = x_global_px + fov_size, y_3 = y_global_px + fov_size, 
                     x_4 = x_global_px, y_4 = y_global_px + fov_size)

str(poly_df)

ls_list <- list()
for (k in 1:dim(fov_positions)[1]){
  ls_list[[k]] <- st_linestring(matrix(as.numeric(poly_df[k]), ncol = 2, byrow = T))
}

sf_data <- st_as_sf(fov_positions, coords = c("x_global_px", "y_global_px"), geometry = st_sfc(ls_list))

sf_data <- st_cast(sf_data, "POLYGON")

# plot(sf_data)
names(sf_data)
sf_data$fov <- as.factor(sf_data$fov)

```

# Flagging score development

Starting from feature scatterplots to see the relationship between possible candidate features

Total counts ~ logged width/height ratio

```{r}
cell_df <- data.frame(colData(spe))

ggplot(data = cell_df, aes(x = log2_AspectRatio, y = total)) +
  geom_point() + 
  ggtitle("Counts ~ log2_AspectRatio")
```

Detected features ~ logged width/height ratio

```{r}
ggplot(data = cell_df, aes(x = log2_AspectRatio, y = detected)) +
  geom_point() + 
  ggtitle("Detected features ~ log2_AspectRatio")
```

Total counts ~ area in um

```{r}
ggplot(data = cell_df, aes(x = um_area, y = total)) +
  geom_point() + 
  ggtitle("Total target counts ~ μm area")
```

Total counts ~ area in px

```{r}
ggplot(data = cell_df, aes(x = Area, y = total)) +
  geom_point() + 
  ggtitle("Total target counts ~ px area")
```

Logged width/height ratio ~ negative control probe counts

```{r}
ggplot(data = cell_df, aes(x = subsets_NegProb_sum, y = log2_AspectRatio)) +
  geom_point() + 
  ggtitle("Logged width/height ratio ~ negative control probe counts")
```

Logged width/height ratio ~ local coordinates on X axis in px

```{r}
ggplot(data = cell_df, aes(x = CenterX_local_px, y = log2_AspectRatio)) +
  geom_point() + 
  ggtitle("Logged width/height ratio ~ local coordinates on X axis in px")
```

Logged width/height ratio ~ local coordinates on X axis in px

```{r}
ggplot(data = cell_df, aes(x = CenterY_local_px, y = log2_AspectRatio)) +
  geom_point() + 
  ggtitle("Logged width/height ratio ~ local coordinates on Y axis in px")
```

# Flag score formula

```{r}
spe$flag_score <- 1/(1 + exp(-0.3*spe$target_counts/spe$um_area + 0.8*abs(spe$log2_AspectRatio)))
summary(spe$flag_score)
plot(spe$flag_score)
```

```{r}
hist(spe$flag_score)
```

## Cell boundary file loading

Loading polygon file and setting the correct format as sf object

```{r}
spat_obj <- fread("/NAS06/work/Spatial_omics/DBKero/CosMx_Breast/CosMX_data_Case2/Run5810_Case2-polygons.csv")
str(spat_obj)
```

```{r}
sf_data <- st_as_sf(spat_obj, coords = c("x_local_px", "y_local_px"))

polygons <- sf_data %>%
  group_by(fov, cellID) %>%
  summarize(geometry = st_combine(geometry))#st_combine = combine several feature geometries into one, without unioning or resolving internal boundaries

polygons <- st_cast(polygons, "POLYGON")
```


```{r, out.width="150%", out.height="150%"}
polygons$total <- spe$total
polygons$target_counts <- spe$target_counts
polygons$um_area <- spe$um_area
polygons$log2_AspectRatio <- spe$log2_AspectRatio
polygons$flag_score <- spe$flag_score
```

Local feature map of flag score, fov 21

```{r, out.width="150%", out.height="150%"}
polygons21 <- filter(polygons, fov==21) 

tm_shape(polygons21) + 
  tm_fill(col= "flag_score", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE)
```

Local threshold map of flag score, target counts, um_area and logged width/height ratio, 20th percentile

```{r, out.width="300%", out.height="600%"}
metric_color <- ifelse(polygons21$flag_score < quantile(spe$flag_score, probs = 0.2), "red", "white")
target_color <- ifelse(polygons21$target_counts < quantile(spe$target_counts, probs = 0.2), "red", "white")
area_color <- ifelse(polygons21$um_area < quantile(spe$um_area, probs = 0.2), "red", "white")
log_hw_color_lo <- ifelse(polygons21$log2_AspectRatio < quantile(spe$log2_AspectRatio, probs = 0.2), "red", "white")
log_hw_color_hi <- ifelse(polygons21$log2_AspectRatio > quantile(spe$log2_AspectRatio, probs = 0.85), "red", "white")

polygons21 <- cbind(polygons21, metric_color = metric_color, target_color = target_color, area_color = area_color, log_hw_color_lo = log_hw_color_lo, 
                    log_hw_color_hi = log_hw_color_hi)

p1 <- tm_shape(polygons21) + 
  tm_fill(col= "metric_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Flag score",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))


p2 <- tm_shape(polygons21) + 
  tm_fill(col= "target_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Target counts",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p3 <- tm_shape(polygons21) + 
  tm_fill(col= "area_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "um area",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))


p4 <- tm_shape(polygons21) + 
  tm_fill(col= "log_hw_color_lo", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Logged AspectRatio below",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p5 <- tm_shape(polygons21) + 
  tm_fill(col= "log_hw_color_hi", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Logged AspectRatio above",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

tmap_arrange(p1, p2, p3, p4, p5, ncol = 2)
```

```{r, out.width="150%", out.height="150%"}
plot_list <- list(p1, p2, p3, p4, p5)
for(i in 1:length(plot_list)){
  print(plot_list[[i]])
}
```


```{r, out.width="150%", out.height="150%"}

flag_value <- c(round(quantile(polygons21$flag_score, probs = 0.2), 2),
                round(quantile(polygons21$target_counts, probs = 0.2), 2),
                round(quantile(polygons21$um_area, probs = 0.2), 2),
                round(quantile(polygons21$log2_AspectRatio, probs = 0.2), 2),
                round(quantile(polygons21$log2_AspectRatio, probs = 0.85), 2))

flagged_cells <- c(sum(polygons21$metric_color=="red", na.rm = T),
                   sum(polygons21$target_color=="red", na.rm = T),
                   sum(polygons21$area_color=="red", na.rm = T),
                   sum(polygons21$log_hw_color_lo=="red", na.rm = T),
                   sum(polygons21$log_hw_color_hi=="red", na.rm = T))

unflagged_cells <- c(sum(polygons21$metric_color=="white", na.rm = T),
                     sum(polygons21$target_color=="white", na.rm = T),
                     sum(polygons21$area_color=="white", na.rm = T),
                     sum(polygons21$log_hw_color_lo=="white", na.rm = T),
                     sum(polygons21$log_hw_color_hi=="white", na.rm = T))

poly21 <- data.frame(threshold = c(0.2, 0.2, 0.2, 0.2, 0.85), flag_value = flag_value, unflagged_cells = unflagged_cells, flagged_cells = flagged_cells,
                     row.names = c("Flag score", "Target counts", "Area in um", "Logged aspect ratio low", "Logged aspect ratio high"))

kable(poly21, format = "html", col.names = c("Threshold", "Flag value", "Flagged cells", "Unflagged cells"), align = "c", caption = paste0("<center>Flagged cells by different metrics, total cells in fov: ", dim(polygons21)[1], ", total cells in whole sample: ", dim(polygons)[1], "</center>"),
      padding = 1L)
```

#
#

Local threshold map of flag score, 10th percentile

```{r, out.width="300%", out.height="600%"}
metric_color <- ifelse(polygons21$flag_score < quantile(spe$flag_score, probs = 0.1), "red", "white")
target_color <- ifelse(polygons21$target_counts < quantile(spe$target_counts, probs = 0.1), "red", "white")
area_color <- ifelse(polygons21$um_area < quantile(spe$um_area, probs = 0.1), "red", "white")
log_hw_color_lo <- ifelse(polygons21$log2_AspectRatio < quantile(spe$log2_AspectRatio, probs = 0.1), "red", "white")
log_hw_color_hi <- ifelse(polygons21$log2_AspectRatio > quantile(spe$log2_AspectRatio, probs = 0.95), "red", "white")

polygons21 <- cbind(polygons21, metric_color = metric_color, target_color = target_color, area_color = area_color, log_hw_color_lo = log_hw_color_lo, 
                    log_hw_color_hi = log_hw_color_hi)

p1 <- tm_shape(polygons21) + 
  tm_fill(col= "metric_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Flag score",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p2 <- tm_shape(polygons21) + 
  tm_fill(col= "target_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Target counts",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p3 <- tm_shape(polygons21) + 
  tm_fill(col= "area_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "um area",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p4 <- tm_shape(polygons21) + 
  tm_fill(col= "log_hw_color_lo", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Logged AspectRatio below",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p5 <- tm_shape(polygons21) + 
  tm_fill(col= "log_hw_color_hi", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Logged AspectRatio above",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

tmap_arrange(p1, p2, p3, p4, p5, ncol = 2)

```

```{r, out.width="150%", out.height="150%"}
plot_list <- list(p1, p2, p3, p4, p5)
for(i in 1:length(plot_list)){
  print(plot_list[[i]])
}
```

```{r, out.width="300%", out.height="300%"}

flag_value <- c(round(quantile(polygons21$flag_score, probs = 0.1), 2),
                round(quantile(polygons21$target_counts, probs = 0.1), 2),
                round(quantile(polygons21$um_area, probs = 0.1), 2),
                round(quantile(polygons21$log2_AspectRatio, probs = 0.1), 2),
                round(quantile(polygons21$log2_AspectRatio, probs = 0.95), 2))

flagged_cells <- c(sum(polygons21$metric_color=="red", na.rm = T),
                   sum(polygons21$target_color=="red", na.rm = T),
                   sum(polygons21$area_color=="red", na.rm = T),
                   sum(polygons21$log_hw_color_lo=="red", na.rm = T),
                   sum(polygons21$log_hw_color_hi=="red", na.rm = T))

unflagged_cells <- c(sum(polygons21$metric_color=="white", na.rm = T),
                     sum(polygons21$target_color=="white", na.rm = T),
                     sum(polygons21$area_color=="white", na.rm = T),
                     sum(polygons21$log_hw_color_lo=="white", na.rm = T),
                     sum(polygons21$log_hw_color_hi=="white", na.rm = T))

poly21 <- data.frame(threshold = c(0.1, 0.1, 0.1, 0.1, 0.95), flag_value = flag_value, unflagged_cells = unflagged_cells, flagged_cells = flagged_cells,
                     row.names = c("Flag score", "Target counts", "Area in um", "Logged aspect ratio low", "Logged aspect ratio high"))

kable(poly21, format = "html", col.names = c("Threshold", "Flag value", "Flagged cells", "Unflagged cells"), align = "c", caption = paste0("<center>Flagged cells by different metrics, total cells in fov: ", dim(polygons21)[1], ", total cells in whole sample: ", dim(polygons)[1], "</center>"),
      padding = 1L)
```

Local feature map of flag score, fov 25

```{r, out.width="150%", out.height="150%"}
polygons25 <- filter(polygons, fov==25)

tm_shape(polygons25) + 
  tm_fill(col= "flag_score", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE)

```

Local threshold map of flag score, 15th percentile

```{r, out.width="300%", out.height="600%"}
metric_color <- ifelse(polygons25$flag_score < quantile(spe$flag_score, probs = 0.15), "red", "white")
target_color <- ifelse(polygons25$target_counts < quantile(spe$target_counts, probs = 0.15), "red", "white")
area_color <- ifelse(polygons25$um_area < quantile(spe$um_area, probs = 0.15), "red", "white")
log_hw_color_lo <- ifelse(polygons25$log2_AspectRatio < quantile(spe$log2_AspectRatio, probs = 0.15), "red", "white")
log_hw_color_hi <- ifelse(polygons25$log2_AspectRatio > quantile(spe$log2_AspectRatio, probs = 0.90), "red", "white")

polygons25 <- cbind(polygons25, metric_color = metric_color, target_color = target_color, area_color = area_color, log_hw_color_lo = log_hw_color_lo, 
                    log_hw_color_hi = log_hw_color_hi)

p1 <- tm_shape(polygons25) + 
  tm_fill(col= "metric_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Flag score",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p2 <- tm_shape(polygons25) + 
  tm_fill(col= "target_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Target counts",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p3 <- tm_shape(polygons25) + 
  tm_fill(col= "area_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "um area",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p4 <- tm_shape(polygons25) + 
  tm_fill(col= "log_hw_color_lo", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Logged AspectRatio below",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p5 <- tm_shape(polygons25) + 
  tm_fill(col= "log_hw_color_hi", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Logged AspectRatio above",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))


tmap_arrange(p1, p2, p3, p4, p5, ncol = 2)

```

```{r, out.width="150%", out.height="150%"}
plot_list <- list(p1, p2, p3, p4, p5)
for(i in 1:length(plot_list)){
  print(plot_list[[i]])
}
```

```{r, out.width="150%", out.height="150%"}

flag_value <- c(round(quantile(polygons25$flag_score, probs = 0.15), 2),
                round(quantile(polygons25$target_counts, probs = 0.15), 2),
                round(quantile(polygons25$um_area, probs = 0.15), 2),
                round(quantile(polygons25$log2_AspectRatio, probs = 0.15), 2),
                round(quantile(polygons25$log2_AspectRatio, probs = 0.90), 2))

flagged_cells <- c(sum(polygons25$metric_color=="red", na.rm = T),
                   sum(polygons25$target_color=="red", na.rm = T),
                   sum(polygons25$area_color=="red", na.rm = T),
                   sum(polygons25$log_hw_color_lo=="red", na.rm = T),
                   sum(polygons25$log_hw_color_hi=="red", na.rm = T))

unflagged_cells <- c(sum(polygons25$metric_color=="white", na.rm = T),
                     sum(polygons25$target_color=="white", na.rm = T),
                     sum(polygons25$area_color=="white", na.rm = T),
                     sum(polygons25$log_hw_color_lo=="white", na.rm = T),
                     sum(polygons25$log_hw_color_hi=="white", na.rm = T))

poly25 <- data.frame(threshold = c(0.15, 0.15, 0.15, 0.15, 0.90), flag_value = flag_value, unflagged_cells = unflagged_cells, flagged_cells = flagged_cells,
                     row.names = c("Flag score", "Target counts", "Area in um", "Logged aspect ratio low", "Logged aspect ratio high"))

kable(poly25, format = "html", col.names = c("Threshold", "Flag value", "Flagged cells", "Unflagged cells"), align = "c", caption = paste0("<center>Flagged cells by different metrics, total cells in fov: ", dim(polygons25)[1], ", total cells in whole sample: ", dim(polygons)[1], "</center>"),
      padding = 1L)
```

Local feature map of flag score, fov 16

```{r, out.width="150%", out.height="150%"}
polygons16 <- filter(polygons, fov==16)

tm_shape(polygons16) + 
  tm_fill(col= "flag_score", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE)

```

Local threshold map of flag score, 15th percentile

```{r, out.width="300%", out.height="600%"}
metric_color <- ifelse(polygons16$flag_score < quantile(spe$flag_score, probs = 0.15), "red", "white")
target_color <- ifelse(polygons16$target_counts < quantile(spe$target_counts, probs = 0.15), "red", "white")
area_color <- ifelse(polygons16$um_area < quantile(spe$um_area, probs = 0.15), "red", "white")
log_hw_color_lo <- ifelse(polygons16$log2_AspectRatio < quantile(spe$log2_AspectRatio, probs = 0.15), "red", "white")
log_hw_color_hi <- ifelse(polygons16$log2_AspectRatio > quantile(spe$log2_AspectRatio, probs = 0.90), "red", "white")

polygons16 <- cbind(polygons16, metric_color = metric_color, target_color = target_color, area_color = area_color, log_hw_color_lo = log_hw_color_lo, 
                    log_hw_color_hi = log_hw_color_hi)

p1 <- tm_shape(polygons16) + 
  tm_fill(col= "metric_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Flag score",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p2 <- tm_shape(polygons16) + 
  tm_fill(col= "target_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Target counts",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p3 <- tm_shape(polygons16) + 
  tm_fill(col= "area_color", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "um area",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p4 <- tm_shape(polygons16) + 
  tm_fill(col= "log_hw_color_lo", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Logged AspectRatio below",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

p5 <- tm_shape(polygons16) + 
  tm_fill(col= "log_hw_color_hi", alpha = 0.5) +
  tm_borders(lwd = 0.1, col = "grey50") +
  tm_layout(legend.outside = TRUE,
            main.title.position = c("center", "top"),
            main.title = "Logged AspectRatio above",
            main.title.size = 0.5, 
            inner.margins = c(0, 0, 0, 0),
            outer.margins = c(0, 0, 0, 0))

tmap_arrange(p1, p2, p3, p4, p5, ncol = 2)

```

```{r, out.width="150%", out.height="150%"}
plot_list <- list(p1, p2, p3, p4, p5)
for(i in 1:length(plot_list)){
  print(plot_list[[i]])
}
```

```{r, out.width="150%", out.height="150%"}

flag_value <- c(round(quantile(polygons16$flag_score, probs = 0.15), 2),
                round(quantile(polygons16$target_counts, probs = 0.15), 2),
                round(quantile(polygons16$um_area, probs = 0.15), 2),
                round(quantile(polygons16$log2_AspectRatio, probs = 0.15), 2),
                round(quantile(polygons16$log2_AspectRatio, probs = 0.90), 2))

flagged_cells <- c(sum(polygons16$metric_color=="red", na.rm = T),
                   sum(polygons16$target_color=="red", na.rm = T),
                   sum(polygons16$area_color=="red", na.rm = T),
                   sum(polygons16$log_hw_color_lo=="red", na.rm = T),
                   sum(polygons16$log_hw_color_hi=="red", na.rm = T))

unflagged_cells <- c(sum(polygons16$metric_color=="white", na.rm = T),
                     sum(polygons16$target_color=="white", na.rm = T),
                     sum(polygons16$area_color=="white", na.rm = T),
                     sum(polygons16$log_hw_color_lo=="white", na.rm = T),
                     sum(polygons16$log_hw_color_hi=="white", na.rm = T))

poly16 <- data.frame(threshold = c(0.15, 0.15, 0.15, 0.15, 0.90), flag_value = flag_value, unflagged_cells = unflagged_cells, flagged_cells = flagged_cells,
                     row.names = c("Flag score", "Target counts", "Area in um", "Logged aspect ratio low", "Logged aspect ratio high"))

kable(poly16, format = "html", col.names = c("Threshold", "Flag value", "Flagged cells", "Unflagged cells"), align = "c", caption = paste0("<center>Flagged cells by different metrics, total cells in fov: ", dim(polygons16)[1], ", total cells in whole sample: ", dim(polygons)[1], "</center>"),
      padding = 1L)
```

