---
title: "Spatial Data Analysis with the SpaceTrooper Package"
author: "Dario Righelli"
date: "`r BiocStyle::doc_date()`"
output:
  BiocStyle::html_document:
    toc: true
    toc_float: true
vignette: >
  %\VignetteIndexEntry{Spatial Data Analysis with the SpaceTrooper Package}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
# Set chunk options: suppress echo, messages, and warnings in code output
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Introduction

This vignette introduces the `SpaceTrooper` package for spatial data analysis
from platforms like **Xenium**, **Merfish**, and **CosMx**. We cover data
loading, quality control, and result visualization.

# Installation

To install `SpaceTrooper`, use the following commands:

```{r install, eval=FALSE}
# Install BiocManager if not already installed, then install SpaceTrooper
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install("drighelli/SpaceTrooper")
```

# Data Loading

In this section, we load data from various platforms using the package's
functions. The goal is to provide a uniform `SpatialExperiment` object across
all technologies, allowing for consistent QC analysis.

The functions in `SpaceTrooper` compute missing metrics as needed and allow
for the inclusion of polygons with the `keep_polygons` argument. This stores
polygons in the `colData` of the `SpatialExperiment`.
We suggest to use the `keep_polygons` argument for technologies like 
**Xenium** and **Merfish/Merscope** because we already load the polygons to 
compute missing metrics in these cases.

Eventually, I/O functions will be Pull Requested into the `SpatialExperimentIO`
Bioconductor package, to discuss with the authors.

Additionally, we are developing compatibility with the
`SpatialFeatureExperiment` class and Python's `spatialData` class for
cross-language support.

```{r load-data, eval=FALSE}
# Load the SpaceTrooper library
library(SpaceTrooper)

# Load Xenium data into a Spatial Experiment object (SPE)
xeniumFolder <- "~/Downloads/Xenium_data/pancreas/"
spe <- readXeniumSPE(xeniumFolder, compute_missing_metrics=TRUE,
                     keep_polygons=TRUE)

# Load Merfish data into an SPE with parquet boundaries
merscopeFolder <- "~/Downloads/Merfish_data/human_uterine_cancer_patient2/"
spe <- readMerfishSPE(merscopeFolder, boundaries_type="parquet",
                      compute_missing_metrics=TRUE, keep_polygons=TRUE)

# Load CosMx data into an SPE without polygons
cosmxFolder <- "~/Downloads/CosMx_data/DBKero/CosMx_Breast/CosMx_data_Case2/"
spe <- readCosmxSPE(cosmxFolder)
```

# Data Analysis based on CosMx

The package offers several functions for spatial data analysis, including
quality control and visualization.

This tutorial focuses on CosMx data, which provides Fields of View (FoVs)
with cell identifiers. Note that FoVs are unique to CosMx and are not
available for other technologies like Xenium and Merfish/Merscope.

For CosMx data, you don't need to `keep_polygons` initially because the
metrics can be computed without them. Polygons can be loaded later if needed.

```{r cosmx-analysis}
library(SpaceTrooper)

# Reload CosMx data with sample name and without polygons
cosmxFolder <- "~/Downloads/CosMx_data/DBKero/CosMx_Breast/CosMx_data_Case2/"
spe <- readCosmxSPE(cosmxFolder, sample_name="DBKero_CosMx",
                    keep_polygons=FALSE)
```

The `spatialPerCellQC` function, inspired by `scater::addPerCellQC`, computes
additional metrics for each cell in the `SpatialExperiment`. It also allows for
the detection of negative control probes, which is crucial for QC.

```{r cosmx-analysis-qc}
# Perform per-cell quality control checks
spe <- spatialPerCellQC(spe, negProbList=c("NegPrb", "Negative",
                                           "SystemControl"))
# Display the first few rows of the SPE's column data
head(colData(spe))
```

Next, we use `computeQCScore` to calculate a flag score based on previously
computed metrics. The flag score combines transcript counts related to cell
area, the aspect ratio of each cell, and its distance from the FoV border
(only for CosMx, this last one is not used otherwise).

Logical filters can then be computed using `computeFilterFlags`, which requires
thresholds for various metrics. Currently, the function considers:

- **Flag Score (`fs_threshold`)**: Cells with scores below this threshold
(default 0.5) are flagged for exclusion.
- **Flag Score Quantiles (`use_fs_quantiles`)**: Option to filter based on
quantiles (default `FALSE`).
- **Total Counts (`total_threshold`)**: Minimum count threshold (default 0).
- **Negative Probe Ratio (`ctrl_tot_ratio_threshold`)**: Minimum ratio of
negative probes to total counts (default 0.1).

```{r cosmx-analysis-score}
# Calculate quality scores for each cell
spe <- computeQCScore(spe)
# Display the updated column data
head(colData(spe))

# Compute flags to identify cells for filtering
spe <- computeFilterFlags(spe, fs_threshold=0.7)
# Display the updated column data
head(colData(spe))
```

Spatial outlier detection is another critical step in QC. While the flag
score addresses some metrics, other outlier detection methods may be needed.

The `computeSpatialOutlier` function allows the computation of medcouple
statistics on a specified metric (`compute_by` argument). It can also use
`scuttle::isOutlier` for asymmetric distributions. The `method` argument
supports "mc", "scuttle", or "both".

```{r cosmx-analysis-outlier}
# Identify spatial outliers based on cell area (Area_um)
spe <- computeSpatialOutlier(spe, compute_by="Area_um", method="both")
# Display the updated column data
head(colData(spe))

# Identify spatial outliers based on mean DAPI intensity
spe <- computeSpatialOutlier(spe, compute_by="Mean.DAPI", method="both")
# Display the updated column data
head(colData(spe))
```

# Results Visualization

The following functions can be used to visualize the analysis results.

## Field of Views (FOVs) Visualization

The `plotCellsFovs` function shows a map of the FoVs within an experiment. 
This plot sis specific to CosMx data and uses cell centroids.

```{r plot-fovs}
# Plot the cells within their respective Field of Views (FOVs)
plotCellsFovs(spe)
```

## Centroid Visualization

To focus on cell centroids, use the `plotCentroidsSPE` function. It provides
a `colour_by` argument, similar to `scater`.

```{r plot-centroids}
# Plot the centroids of the cells in the SPE
plotCentroidsSPE(spe)

# Plot centroids, colored by total intensity across all channels
plotCentroidsSPE(spe, colour_by="total")

# Plot centroids, highlighting cells flagged as outliers by QC scores
plotCentroidsSPE(spe, colour_by="is_fscore_outlier")
```

## Metric Histograms

You can investigate individual metrics by viewing their histograms. For
outliers, use the `use_fences` argument to display the fences computed by
`computeSpatialOutlier`.

```{r plot-metrics}
# Plot a histogram of cell areas (Area_um)
plotMetricHist(spe, metric="Area_um")

# Plot a histogram with fences to identify outliers
plotMetricHist(spe, metric="Area_um", use_fences="Area_um_outlier_mc")

plotMetricHist(spe, metric="Area_um", use_fences="Area_um_outlier_sc")
```

## Polygon Addition and Visualization

Next, load and add polygons to your SPE object using the following functions.
Each technology has its own `readPolygons` function to standardize the
loaded `sf` object and handle different file types.

```{r plot-polygons}
# Read polygon data associated with cells in the SPE
pols <- readPolygonsCosmx(metadata(spe)$polygons)

# Add the polygon data to the SPE object
spe <- addPolygonsToSPE(spe, pols)
```

Once the polygons are stored in an `sf` object within `colData`, you can
visualize them using functions based on the `tmap` library.

The `plotZoomFovsMap` function allows you to visualize a map of the FoVs with
a zoom-in of selected FoVs, colored by the `colour_by` argument. You can also
toggle between `tmap` and `ggplot2` for visualization (this is a
work-in-progress).



```{r, eval=FALSE}
## not working in vignette compilation
plotZoomFovsMap(spe, fovs=c(16:19), colour_by="Area_um")
plotZoomFovsMap(spe, fovs=c(16:19), colour_by="Area_um", useggplot=FALSE)
```

To focus on specific FoVs, subset the original object and visualize only the
polygons of this subset using `plotPolygonsSPE`.

```{r plot-polygons-fov}
# Subset the SPE to include only specified FOVs
spe1 <- spe[,spe$fov %in% c(16:19)]

# Plot the polygons of the selected cells
plotPolygonsSPE(spe1)

# Plot polygons colored by cell area
plotPolygonsSPE(spe1, colour_by="Area_um", palette="viridis")

# Plot polygons highlighting cells identified as outliers by area
plotPolygonsSPE(spe1, colour_by="Area_um_outlier_mc", palette="viridis")

# Plot polygons highlighting cells flagged as outliers by QC scores
plotPolygonsSPE(spe1, colour_by="is_fscore_outlier", palette="viridis")
```

# Conclusion

In this vignette, we explored the main functionalities of the `SpaceTrooper`
package for spatial data analysis. We covered data loading, quality control,
and result visualization.

# Other Features

We released several functions for polygon handling, `sf` object manipulation,
and palette creation. These will be improved and generalized in future versions.

More vignettes to come...

```{r}

# Polygon handling
# Accessories for sf object
# Palette creation -> to improve and generalize
```

# Session Information

```{r}
sessionInfo()
```
