---
title: "Spatial Data Quality Control with the SpaceTrooper Package"
author: "Dario Righelli"
date: "`r BiocStyle::doc_date()`"
output:
  BiocStyle::html_document:
    toc: true
    toc_float: true
vignette: >
  %\VignetteIndexEntry{Spatial Data Quality Control with the SpaceTrooper Package}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
# Set chunk options: suppress echo, messages, and warnings in code output
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Introduction

This vignette introduces the `SpaceTrooper` package for spatial data analysis
from platforms like **Xenium**, **Merfish**, and **CosMx**. We cover data
loading, quality control, and result visualization.

# Installation

To install `SpaceTrooper`, use the following commands:

```{r install, eval=FALSE}
# Install BiocManager if not already installed, then install SpaceTrooper
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install("drighelli/SpaceTrooper")
```

# Data Loading

In this section, we load data from various platforms using the package's
functions. The goal is to provide a uniform `SpatialExperiment` object across
all technologies, allowing for consistent QC analysis.

The functions in `SpaceTrooper` compute missing metrics as needed and allow
for the inclusion of polygons with the `keep_polygons` argument. This stores
polygons in the `colData` of the `SpatialExperiment`.
We suggest to use the `keep_polygons` argument for technologies like 
**Xenium** and **Merfish/Merscope** because we already load the polygons to 
compute missing metrics in these cases.

Eventually, I/O functions will be Pull Requested into the `SpatialExperimentIO`
Bioconductor package, to discuss with the authors.

Additionally, we are developing compatibility with the
`SpatialFeatureExperiment` class and Python's `spatialData` class for
cross-language support.

```{r load-data, eval=FALSE}
# Load the SpaceTrooper library
library(SpaceTrooper)

# Load Xenium data into a Spatial Experiment object (SPE)
xeniumFolder <- "~/Downloads/Xenium_data/pancreas/"
spe <- readXeniumSPE(xeniumFolder, compute_missing_metrics=TRUE,
                     keep_polygons=TRUE)

# Load Merfish data into an SPE with parquet boundaries
merscopeFolder <- "~/Downloads/Merfish_data/human_uterine_cancer_patient2/"
spe <- readMerfishSPE(merscopeFolder, boundaries_type="parquet",
                      compute_missing_metrics=TRUE, keep_polygons=TRUE)

# Load CosMx data into an SPE without polygons
cosmxFolder <- "~/Downloads/CosMx_data/DBKero/CosMx_Breast/CosMx_data_Case2/"
spe <- readCosmxSPE(cosmxFolder)
```

# Data Analysis based on CosMx

The package offers several functions for spatial data analysis, including
quality control and visualization.

This tutorial focuses on CosMx data, which provides Fields of View (FoVs)
with cell identifiers. Note that FoVs are unique to CosMx and are not
available for other technologies like Xenium and Merfish/Merscope.

For CosMx data, you don't need to `keep_polygons` initially because the
metrics can be computed without them. Polygons can be loaded later if needed.

```{r cosmx-analysis}
library(SpaceTrooper)

# Reload CosMx data with sample name and without polygons
cosmxFolder <- "~/Downloads/CosMx_data/DBKero/CosMx_Breast/CosMx_data_Case2/"
spe <- readCosmxSPE(cosmxFolder, sample_name="DBKero_CosMx",
                    keep_polygons=FALSE)
## a custom function for the workshop
source(system.file("scripts/labelsfunct.R", package="SpaceTrooper"))
spe <- addLabelsDBKero(spe)

spe
```

# Field of Views (FOVs) Visualization

The `plotCellsFovs` function shows a map of the FoVs within an experiment. 
This plot sis specific to CosMx data and uses cell centroids.

```{r plot-fovs}
# Plot the cells within their respective Field of Views (FOVs)
plotCellsFovs(spe)
```

# Visualizing cell types

To focus on cell centroids, use the `plotCentroidsSPE` function. It provides
a `colour_by` argument, similar to `scater`.
Additionally, if you have a column with a color palette it automatically maps it
to the values passed in the `colour_by` argument.

```{r plot-centroids-labels}
# Plot the centroids of the cells in the SPE
plotCentroidsSPE(spe, colour_by="labels")
# Plot the centroids of the cells in the SPE with personalized labels
plotCentroidsSPE(spe, colour_by="labels", palette="labels_colors")
```



# Quality control 

The `spatialPerCellQC` function, inspired by `scater::addPerCellQC`, computes
additional metrics for each cell in the `SpatialExperiment`. It also allows for
the detection of negative control probes, which is crucial for QC.

```{r cosmx-analysis-qc}
# Perform per-cell quality control checks
spe <- spatialPerCellQC(spe, negProbList=c("NegPrb", "Negative",
                                           "SystemControl"))
```


# Metrics Histograms

You can investigate individual metrics by viewing their histograms. For
outliers, use the `use_fences` argument to display the fences computed by
`computeSpatialOutlier`.

```{r plot-metrics}
# Plot a histogram of counts (sum)
plotMetricHist(spe, metric="sum")
# Plot a histogram of cell areas (Area_um)
plotMetricHist(spe, metric="Area_um")
# Plot a histogram of proportion of counts respect to the cell area in micron 
plotMetricHist(spe, metric="log2CountArea")
```


# Spatial Outlier Detection 

Spatial outlier detection is another critical step in QC. While the flag
score addresses some metrics, other outlier detection methods may be needed.

The `computeSpatialOutlier` function allows the computation of medcouple
statistics on a specified metric (`compute_by` argument). It can also use
`scuttle::isOutlier` for asymmetric distributions. The `method` argument
supports "mc", "scuttle", or "both".

This outlier detection approach can be used to decide if and which cells can 
be discarded on a singular metric.

```{r cosmx-analysis-outlier}
# Identify spatial outliers based on cell area (Area_um)
spe <- computeSpatialOutlier(spe, compute_by="Area_um", method="both")

# Identify spatial outliers based on mean DAPI intensity
spe <- computeSpatialOutlier(spe, compute_by="Mean.DAPI", method="both")
```

If we computed outliers with the computeSpatialOutlier function we can also 
visualize which fences have been used to create the filter on the cells.

```{r plot-metrics-2}
# Plot a histogram with fences to identify outliers using the medcouple
plotMetricHist(spe, metric="Area_um", use_fences="Area_um_outlier_mc")

# Plot a histogram with fences to identify outliers using scuttle
plotMetricHist(spe, metric="Area_um", use_fences="Area_um_outlier_sc")

# Plot a histogram with fences to identify outliers using the medcouple
plotMetricHist(spe, metric="Mean.DAPI", use_fences="Mean.DAPI_outlier_mc")

# Plot a histogram with fences to identify outliers using scuttle
plotMetricHist(spe, metric="Mean.DAPI", use_fences="Mean.DAPI_outlier_sc")
```

# The Flag Score

Next, we use `computeQCScore` to calculate a flag score based on previously
computed metrics. 
The flag score combines **transcript counts** related to 
**cell area**, the **aspect ratio** of each cell, and its 
**distance from the FoV border** (only for CosMx, this last one is not used 
otherwise).

```{r cosmx-analysis-score}
# Calculate quality scores for each cell
spe <- computeQCScore(spe)
```

Logical filters can then be computed using `computeFilterFlags`, which requires
thresholds for various metrics. Currently, the function considers:

* **Flag Score (`fs_threshold`)**: Cells with scores below this threshold
    (default 0.5) are flagged for exclusion. This value can be used to indicate 
    the quantile for the filtering when setting the `use_fs_quantiles` argument to `TRUE`.
    
* **Flag Score Quantiles (`use_fs_quantiles`)**: Option to filter based on
    quantiles (default `FALSE`).
  
* **Total Counts (`total_threshold`)**: Minimum count threshold (default 0).

* **Negative Probe Ratio (`ctrl_tot_ratio_threshold`)**: Minimum ratio of
    negative probes to total counts (default 0.1).


```{r cosmx-analysis-score2}
# Compute flags to identify cells for filtering
spe <- computeFilterFlags(spe, fs_threshold=0.7)
```

# Polygon Addition and Visualization

To better understand the flag score values we start to load the polygons, 
giving us a better overview of the cells characteristics.

We can load and add polygons to your SPE object using the following functions.
Each technology has its own `readPolygons` function to standardize the
loaded `sf` object and handle different file types.

```{r plot-polygons}
# Read polygon data associated with cells in the SPE
# the polygon file path is stored in the spe metadata
pols <- readPolygonsCosmx(metadata(spe)$polygons)

# Add the polygon data to the SPE object
spe <- addPolygonsToSPE(spe, pols)
```

Once the polygons are stored in an `sf` object within `colData`, you can
visualize them using functions based on the `tmap` library.

```{r plot-polygons-fov-1}
# Subset the SPE to include only specified FOVs
spe1112 <- spe[,spe$fov %in% c(11:12)]

# Plot the polygons of the selected cells
plotPolygonsSPE(spe1112, bg_color="white")

# Plot polygons colored by cell area
plotPolygonsSPE(spe1112, colour_by="log2AspectRatio")
plotPolygonsSPE(spe1112, colour_by="Area_um")

plotPolygonsSPE(spe1112, colour_by="flag_score", palette="viridis")
plotPolygonsSPE(spe1112, colour_by="is_fscore_outlier")
```

# Fov Zoom and Map

The `plotZoomFovsMap` function allows you to visualize a map of the FoVs with
a zoom-in of selected FoVs, colored by the `colour_by` argument. You can also
toggle between `tmap` and `ggplot2` for visualization (this is a
work-in-progress).

```{r, eval=FALSE}
## not working in vignette compilation
plotZoomFovsMap(spe, fovs=c(16:19), colour_by="flag_score")
plotZoomFovsMap(spe, fovs=c(16:19), colour_by="is_fscore_outlier", useggplot=FALSE)
```

# Conclusion

In this vignette, we explored the main functionalities of the `SpaceTrooper`
package for spatial data analysis. 
Main steps shown are:
* data loading: Xenium, Merfish/Merscope, CosMx 
* polygons loading: Xenium, Merfish/Merscope, CosMx 
* quality control: 
    + outlier detection: medcouple and scuttle MAD
    + flag score: a score combining **transcript counts**, **cell area**, 
    **aspect ratio** and  **distance from the FoV border**
* visualization: 
    + centroids: with ggplot2
    + polygons: sf + tmap

# Pending

## Other Features 

We released several functions for polygon handling, `sf` object manipulation,
and palette creation. These will be improved and generalized in future versions.

More vignettes to come:

* Polygons handling

* sf object accessory functions

* palette creation from colData (work-in-progress)

## More to come

Interoperability:

* within R: SpatialFeatureExperiment support 

* outside R: spatialData Python class I/O


Visualization:

* more plots:

    + interactive plots with polygons
    
    + metrics violin plot
    
    + metrics scatter plot
    
    + multiple metrics polygons visualization

# Session Information

```{r}
sessionInfo()
```
